### 2019-10-07 Moral_Communities_and_Crime_v1.Rmd
Created first version of paper

---
title: Do moral communities have a spatial dimension? A spatial exploratory analysis of places of worship and violent crime in the city of Recife, Brazil
subtitle: 
titlerunning: Moral communities and crime
authorrunning: 
thanks: | 
    Grants or other notes about the article that should go on the front 
    page should be placed here. General acknowledgments should be placed at the
    end of the article.

authors: 
- name: Äüthör 1
  address: Department of YYY, University of XXX
  email: abc@def

- name: Âuthóř 2
  address: Department of ZZZ, University of WWW
  email: djf@wef

keywords:
- Moral communities
- Crime
- Places of worship
- Point pattern analysis
- Intensity

#PACS: 
#- PAC1
#- superPAC

MSC:
- MSC code 1
- MSC code 2

abstract: |
  A wealth of research exists to suggest that environmental attributes can facilitate or deter criminal activities, as for instance the presence of alcohol outlets or the design of defensible spaces. In this article we present an exploratory spatial analysis of the relationship between violent crime and places of worship in Recife, Brazil. This analysis aims to explore whether there is a spatial dimension to “moral communities” around places of worship. Use of a cluster detection methodology, the results suggest the presence of a strong geographical association between churches and crime. The evidence obtained for the year 2010 indicates a process of clustering of homicide events in the vicinity of places of worship. In addition, it was also possible to verify a pattern of concentration between evangelical churches and violent crimes.

bibliography: bibliography.bib
output: rticles::springer_article
---

# Introduction {#intro}

Your text comes here. Separate text sections with

# Section title {#sec:1}

Text with citations by @Galyardt14mmm, [@Mislevy06Cog].

## Subsection title {#sec:2}

as required. Don't forget to give each section
and subsection a unique label (see Sect. \ref{sec:1}).

#### Paragraph headings

Use paragraph headings as needed.

\begin{align}
a^2+b^2=c^2
\end{align}

# References

### 2019-10-09 Moral_Communities_and_Crime_v1.Rmd
Using the initial draft from Edivaldo et al., first round of revisions to the text and the analysis. Worked on the introduction, structure of the paper, and started to put the analysis in.

@@
 
-#PACS: 
-#- PAC1
-#- superPAC
-
-MSC:
-- MSC code 1
-- MSC code 2
-
 abstract: |
@@
 
+<!-- This paper is an R markdown document that includes self-contained, reproducible analysis.
+To reproduce the paper the following file is needed: `Moral_Communities_and_Crime.RData`.
+For reference: this data file was created using R notebook `00.
+Data Preprocessing.Rmd`.
+The initial analysis was done in R notebook `01.
+Analysis.Rmd`.
+Neither of these two files are needed for this document.
+-->
+
+```{r load-packages, include=FALSE}
+
+# Load packages used in this document
+
+library(tidyverse)
+library(kableExtra)
+library(sf)
+library(spatstat)
+library(maptools)
+library(ggthemes)
+library(raster)
+library(fasterize)
+```
+
+```{r load-data, include=FALSE}
+
+# To run this notebook, a file `Moral_Communities_and_Crime.RData` is needed. This file contains the following objects:
+# 
+# - events_recife.sf: a simple features object with points for violent crimes, churches (Catholic, Evangelical, Spiritist) and commercial establishments (ice cream shops, pharmacies, supermarkets) in the city of Recife.
+# 
+# - recife.sf: a simple features object with the boundaries of the city of Recife
+# 
+# - setores.sf: a simple features object with setores (census tracts) and selected socio-economic and demographic information for the city of Recife, including:
+#   - Population density in persons per sq.km
+#   - Median income in setor
+#   - Proportion of population in setor who are unemployed
+#   - Proportion of households with a single mother
+#   - Proportion of households with a young single mother, a woman between the ages of 15 and 25
+#   - Proportion of population in poverty
+# 
+# Read the RData file:
+
+load("Moral_Communities_and_Crime.RData")
+```
+
+```{r data-preparation, include=FALSE, cache=TRUE}
+
+# In preparation for doing point pattern analysis, convert the `recife.sf` object to an `owin` object for use in `spatstat`:
+
+recife.owin <- recife.sf %>% as("Spatial") %>% as("owin")
+plot(recife.owin)
+
+# Convert the events to a `ppp` object:
+
+events_recife.ppp <- events_recife.sf %>% 
+  st_coordinates() %>% 
+  data.frame() %>%
+  mutate(Type = events_recife.sf$Type) %>%
+  as.ppp(W = recife.owin)
+
+#Split the point pattern into each type of event: crime, Catholic place of worship, Evangelical place of worship, Spiritist place of worship, ice cream shop, pharmacy, and supermarket
+
+events_recife_list <- split(events_recife.ppp)
+```
+
+```{r socio-economic-deprivation, include=FALSE, cache=TRUE}
+
+# Here an indicator of socio-economic deprivation is obtained using principal component analysis
+
+# PCA is implemented in R as `prcomp`:
+setores.pca <- setores.sf %>% 
+  st_drop_geometry() %>%
+  dplyr::select(median_income, prop_unemployed, prop_single_mother, prop_young_single_mother, prop_poverty) %>%
+  prcomp(center = TRUE, scale = TRUE)
+
+# The loadings of the pca are as follows:
+setores.pca$rotation
+
+# Socio-economic deprivation is a combination of high poverty, high proportion of single mothers, especially young women between the ages of 15-25, high proportion of unemployed people, and low median income.
+
+# Various diagnostics of the results are possible, such as the variance:
+
+std_dev <- setores.pca$sdev
+
+# Compute variance
+
+pr_var <- std_dev^2
+
+# Check variance of the components
+
+pr_var[1:5]
+
+# And the proportion of the variance:
+
+pr_var/sum(pr_var)
+
+# Once that there is principal component (call it SED for socio-economic deprivation), it can be joined to the `sf` object before it can be rasterized and converted to an `im` object:
+
+setores.sf$SED <- setores.pca$x[,1]
+
+# Normalize the SDE factor (with the addition of a small correction; the `rhohat` function does not like values <= 0):
+
+setores.sf <- setores.sf %>% mutate(SED = (SED - min(SED) + 0.1)/(max(SED) + 0.1 - min(SED)))
+```
+
 # Introduction {#intro}
 
-Your text comes here. Separate text sections with
+Violent crime is a widespread phenomenon with negative impacts on many spheres of life and society.
+Although the rate of homicides worldwide has grown at a slower rate than the population, the number of people killed in homicides still increased from 362,000 in 1990 to 464,000 in 2017 [@Unodc2019executive].
+Variations in violent crime also tend to be extremely uneven.
+The Americas, with a population of approximately 793.8 million people in 2019 (or about 10.3% of the world population), accounted for approximately 173,000 homicides in 2017, or approximately 37.3% of all homicides in the world.
+Within the Americas, Brazil (along with Venezuela, Colombia, and Mexico) is one of the largest countries in the region with high homicide rates.
+The relevance of variations in the prevalence of violent crime is in fact recognized by the United Nations Office on Drug and Crime  [-@Unodc2019executive] as being key to achieving policy goals:
 
-# Section title {#sec:1}
+>> High levels of homicidal violence are concentrated in geographic and demographic “pockets”, so
+achieving target 16.1 of the Sustainable Development Goals requires interventions within the specific
+regions, countries, communities and population groups that are most at risk (p. 35)
 
-Text with citations by @Galyardt14mmm, [@Mislevy06Cog].
+The endemic malady of violent crime  represents in fact an important factor that threatens to derail sustainable development goals.
+Violent crime, furthermore, tends to be more accute in those regions most in need of development.
+On the one hand, violent crime represents an economic and social drag [@Becker2017analise] that affects people's well-being, either through the loss of human life, mental health, limitations in the right to public spaces [@Doran2011putting], or through disturbances in schooling and academic achievement [@Unodc2019development].
+In turn, these negative effects combine to create the unfortunate conditions that exacerbate criminal behavior, thus creating a vicious cycle of economic disadvantage and crime [@Unodc2019development].
+Added to this scenario, social control institutions in Brazil, which in practice should inhibit criminal practice, have deep deficiencies that end up reducing the deterrent power of the criminal justice system.
+Among these serious problems are the inefficiency of the police, the lack of national legislation, the glacial pace of judicial processes.
+and the weak situation of the prison system in the country [@Menezes2013spatial].
+It is not surprising, given the above, that researchers have joined calls for more research to increase our understanding of the patterns of violent crime in Low and Middle Income Countries [LMIC; @Murray2013crime].
 
-## Subsection title {#sec:2}
+In order to counteract any criminological factors, it is important to identify empirical regularities.
+Criminological factors include concrete elements, such as the presence of arms or drugs, as well as environmental factors, such as the built environment, in addition to figurative factors, which include the moral costs of deviant behavior as well as family supervision.
+Accordingly, a number of studies have investigated various aspects of the environment and neighborhood design [e.g., @Foster2010neighbourhood; @He2017built; @Loukaitou2001measuring], whereas other studies have focused on exposure to environmental attributes that signal weakened norms, such as liquor and tobacco outlets associated with geographical hotspots of crime [@Brower2007spatial; @Deryol2016crime; @Lipton2008spatial; @Quick2017influence].
+Yet another fruitful avenue for research, and one that has only recently begun to attract attention, is the presence of environmental attributes that can help reinforce moral norms, such as schools and churches [e.g., @Abdullah2018amenities; @Davignon2015christian; @Furr2010metric; @Traunmuller2011moral].
+It is thus that in a recent paper, Warner and Konkel [-@Warner2019neighborhood] note that the role of places of worship, as distinct entities from the members of the congregations, have received less attention in empirical and theoretical research for their potential to prevent crime.
+The role of these institutions might be particularly important in places where the State lacks the means or the will to enforce norms - as is the case in Brazil.
 
-as required. Don't forget to give each section
-and subsection a unique label (see Sect. \ref{sec:1}).
+With the above considerations in mind, the objective of this paper is to investigate whether moral communities have a spatial dimension discernible from the presence of places of worship.
+The case study is the city of Recife, in the state of Pernambuco in Brazil's Northeast.
+Recife is a large metropolitan area in a historically poor region, and afflicted by high levels of violent crime.
+The empirical strategy is to use spatial analysis to explore the potential geographical relationships between violent criminal events, on the one hand, and places of worship and a selection of commercial establishments, on the other.
+Dissagregated data allows us to analyze the phenomena of interest at a very high level of resolution as spatial point patterns.
 
-#### Paragraph headings
+After this introductory section, the rest of this paper is structured as follows: Sect.
+\ref{background} reviews the literature on religion and crime; then Sect.
+\ref{methods} presents the empirical strategy used in the work.
+Sect.
+\ref{case} presents the context as well as the data used ub the study; Sect.
+\ref{results} systematizes and discusses the main results; and finally, the conclusions are presented in section \ref{conclusions}.
 
-Use paragraph headings as needed.
+# Background: Theoretical Perspectives {#background}
 
-\begin{align}
-a^2+b^2=c^2
-\end{align}
+Words go here.
+
+# Empirical Strategy {#methods}
+
+Two different techniques are deployed in this paper to study the spatial distribution of criminal events, and their possible relationship to the location of places of worship.
+Firstly, the F-function is a nearest neighbor technique that is used to summarize the distance between a set of points and their nearest events.
+This technique has found extensive use in spatial epidemiology (see Gatrell et al., 1996) and has also been used in spatial criminology (see, inter alia, Craglia et al., 2000; Kiani et al., 2015; Levine, 2006; and Rogerson and Sun, 2001).
+The second technique is the scan statistic.
+This method has also been extensively used in spatial epidemiology (e.g., Hjalmars et al., 1996) and has found use in spatial criminology as well (see He et al., 2017a; Nakaya and Yano, 2010; Malleson and Andresen, 2015; and Shiode, 2011).
+Use of these two techniques is meant to provide insights with respect to 1) the location of violent criminal events with respect to various types of facilities; and 2) geographically identify places of worship associated with clusters of criminal events.
+These two techniques are briefly reviewed next.
+
+## Quadrat Analysis
+
+Words go here.
+
+## Relative Intensity
+
+Words go here.
+
+# Case Study {#case}
+
+## Context {#context}
+
+The study is of the city of Recife, the capital of the state of Pernambuco in the Northeast region of Brazil.
+With a population of 1,550,390 million inhabitants in 2010, and with an area of 7,133.2 km2, Recife is one of the main Brazilian metropolises, exerting a great economic influence in neighboring regions.
+However, the city experiences a serious problem with violent crime, and has the dubious honor of being one of the five capital cities in Brazil with the highest homicide rates in the period under study [@Menezes2013spatial].
+
+In the current context, the term "violent crime" is an umbrella for several forms of infractions to the penal code.
+Following recommendations of the National Secretariat of Public Security of the Ministry of Justice of 2006, these are Violent Lethal and Intentional Crimes (CLVI; which includes intentional homicide), theft followed by death (robbery), and corporal injury followed by death.
+The data on CVLI were extracted from the Police Information System of the Secretariat of Social Defense of Pernambuco (INFOPOL / SDS-PE), which is the most reliable, detailed, and comprehensive information on violent deaths in the region.
+
+```{r descriptive-statistics, echo=FALSE}
+descriptive_stats <- data.frame(Attribute = c("Man",
+                                              "Woman",
+                                              "Murder",
+                                              "Robbery",
+                                              "Body injury followed by death",
+                                              "Black and White",
+                                              "Yellow and White",
+                                              "Not Reported",
+                                              "1 to 12 years",
+                                              "13 to 17 years",
+                                              "18 to 30 years",
+                                              "31 to 65 years",
+                                              "65 years and older",
+                                              "Not Reported",
+                                              "Firearm",
+                                              "Other"),
+                                Percentage = c(91.35,
+                                               8.65,
+                                               97.82,
+                                               2.00,
+                                               0.18,
+                                               95.11,
+                                               1.27,
+                                               3.62,
+                                               0.18,
+                                               11.31,
+                                               61.70,
+                                               25.84,
+                                               0.73,
+                                               0.24,
+                                               87.51,
+                                               12.49))
+
+kable(descriptive_stats, 
+      "latex", 
+      booktabs = TRUE,
+      digits = 3,
+      caption = "\\label{tab:descriptive-statistics}Characteristics of Violent Crime in Recife, July, 2008 - June, 2009",
+      col.names = c("", "Percentage")) %>%
+  kable_styling() %>%
+  pack_rows("Gender of Victim", 1, 2) %>%
+  pack_rows("Type of Crime", 3, 5) %>%
+  pack_rows("Ethnicity of Victim", 6, 8) %>%
+  pack_rows("Age of Victim", 9, 14) %>%
+  pack_rows("Weapon Used", 15, 16)
+```
+
+## Data {#data}
+
+The data are organized at the individual level and it is possible to obtain information about the location, day of the week, day of the month, period of the day, as well as gender, age, and race of the victim.
+The database used in this study comprises the period from July 1, 2008 to June 30, 2010.
+Some descriptive statistics regarding this dataset are reported in Table 1, where it can be seen that overall, about 91% of the victims of CVLI in the period analyzed were men, while approximately 98% of violent crimes were homicides.
+In addition, most of the victims were black or brown, and the youth population between the ages of 18 and 30 is the most affected by violent crime.
+Lastly, it should be noted that about 88% of the criminal events under analysis involved firearms.
+Information about places of interest was obtained from the National Register of Addresses for Statistical Purposes (CNEFE - Census 2010), which lists 78,056,411 urban and rural addresses, distributed among the 316,574 census tracts.
+This is the first database of its kind produced by IBGE, and the first version was produced at the time of the 2000 Census.
+The way addresses are described in the National Register is very rich, and it is possible to identify the names of the places of worship including their denomination.
+Georeferencing was used to geolocate each place of worship.
+In this way, a total of 1,719 places of worship were geolocated in the city of Recife.
+Figure 1 (Panel a) shows the spatial distribution of the 1,657 CVLI crimes that occurred in the city of Recife between 2008.2 and 2010.1.
+Panel b) of the figure shows the spatial distribution of places of worship in the region.
+
+```{r plot-crime, echo=FALSE, fig.align = "left", fig.cap="\\label{fig:plot-crime}Lethal and Intentional Violent Crime in Recife"}
+
+#We can experiment with plotting the events:
+
+ggplot(data = events_recife.sf %>% filter(Type == "Violent Crime")) + 
+  geom_sf(color = "red") +
+  geom_sf(data = recife.sf, fill = NA) + 
+  theme_tufte()
+```
+
+In addition to places of worship, the National Register of Addresses for Statistical Purposes was queried to extract facilities other than places of worship.
+As discussed above, the idea is to identify points of reference that can be used as controls, having a neutral morality profile.
+For the sake of the present study, we selected pharmacies, ice cream shops, and bakeries to construct our control group.
+These three types of establishments comply with the criteria of being morally neutral and having a spatial distribution commensurate with places of worship.
+Figure 2 shows the spatial distribution of the control establishments, namely pharmacies, bakeries, and ice cream shops in the city of Recife.
+Note the similarity between the maps.
+As expected, there are differences in the number of points, but the locations of cases and controls are quite similar.
+
+```{r plot-events, echo=FALSE, fig.align = "left", fig.cap="\\label{fig:plot-events}Places of Worship and Commercial Establishments in Recife"}
+
+#We can experiment with plotting the events:
+
+ggplot(data = events_recife.sf %>% filter(Type != "Violent Crime")) + 
+  geom_sf(aes(color = Type, fill = Type)) +
+  geom_sf(data = recife.sf, fill = NA) + 
+  facet_wrap(~ Type, nrow = 2) +
+  scale_fill_viridis_d(option = "viridis", direction = 1) +
+  scale_color_viridis_d(option = "viridis", direction = 1) +
+  theme_tufte()  +
+  theme(axis.text.x = element_text(angle = 90))
+
+```
+
+```{r summary-point-patterns, include=FALSE}
+
+# Extract the intensity of each point pattern for the study area
+
+summary_crime <- summary(events_recife_list$`Violent Crime`)
+summary_catholic <- summary(events_recife_list$Catholic)
+summary_evangelic <- summary(events_recife_list$Evangelical)
+summary_spiritist <- summary(events_recife_list$Spiritist)
+summary_ice_cream <- summary(events_recife_list$`Ice Cream Shop`)
+summary_pharmacy <- summary(events_recife_list$Pharmacy)
+summary_supermarket <- summary(events_recife_list$Supermarket)
+```
+
+```{r plot-sed, echo=FALSE, fig.align = "left", fig.cap="\\label{fig:plot-sed}Socio-Economic Deprivation in Recife"}
+ggplot(data = setores.sf) + 
+  geom_sf(aes(fill = SED),
+          color = NA) +
+  scale_fill_viridis_c(option = "plasma", direction = 1, values = c(0, 0.05, 0.2, 0.3, 0.4, 1)) +
+  theme_tufte()
+```
+
+# Analysis and results {#results}
+
+## Quadrat Analysis
+
+```{r rasterize-sed, include=FALSE}
+
+# Rasterize and convert to image:
+
+# First, we need a base raster; this will be the boundaries of the study area.
+After some trial and error with different values of number of rows and columns, these values create pixels that are as close to squares as possible, while also being relatively small
+
+recife.r <- raster(recife.sf, nrows = 400, ncols = 276)
+
+# The raster `recife.r` is the template for rasterizing other items, in this case socio-economic deprivation. Use package `fasterize` to rasterize an `sf` object.
+
+sed.r <- fasterize(setores.sf, recife.r, field = "SED")
+
+# Once that the object is a raster, it can be converted to a `spatstat` `img` class object:
+sed.img <- as.im(sed.r)
+```
+
+```{r quadrat-analysis-preliminaries, include=FALSE}
+
+# To do quadrat analysis by different levels of SED we need the quintiles of this variable (other criteria could be used to slice the distribution, e.g., standard deviations, but this one is intuitive and simple)
+
+sed5 <- quantile(sed.img, c(0, 0.2, 0.4, 0.6, 0.8, 1))
+
+# Next we can cut the `sed.im` using the quintiles:
+
+sed_cut5.img <- cut(sed.img, breaks = sed5, include.lowest = TRUE, labels = c(1:5))
+```
+
+```{r tesselate-sed5, include=FALSE}
+
+# To do quadrat analysis we need to tesellate the cut image:
+
+V <- tess(image = sed_cut5.img)
+
+# This object is used for the analysis, but it is not very neat for plotting purposes. Lets convert instead to an `sf` object. First, extract and expand the coordinates:
+
+coords_V <- expand.grid(X = V$image$xcol, Y = V$image$yrow)
+
+# Next, extract the values of the cells and convert to a vector:
+
+sed5_V <- as.vector(t(V$image$v))
+
+# Join the coordinates with the values in a dataframe and drop all NAs, before converting to `sf` and adding coordinate referencing system information (crs) to ensure that its projection information is there and compatible with the rest of `sf` objects (i.e., recife.sf, setores.sf, events_recife.sf):
+
+sed5.sf <- data.frame(coords_V, sed5_V) %>% 
+  drop_na() %>% transmute(X, Y, SED = factor(sed5_V, 
+                                              levels = c(1:5),
+                                              labels = c("Low", 
+                                                         "Medium Low", 
+                                                         "Medium", 
+                                                         "Medium High", 
+                                                         "High"))) %>%
+  st_as_sf(coords = c("X", "Y")) %>%
+  st_set_crs(value = st_crs(recife.sf))
+```
+
+```{r plot-sed-as-quintiles, echo=FALSE, cache = TRUE, fig.cap="\\label{fig:plot-sed-as-quintiles}Socio-Economic Deprivation as Quintiles" }
+ggplot(data = sed5.sf) + 
+  geom_sf(aes(fill = SED, color = SED)) +
+  geom_sf(data = recife.sf, fill = NA) +
+  scale_fill_viridis_d(option = "plasma", direction = 1) +
+  scale_color_viridis_d(option = "plasma", direction = 1) +
+  guides(fill = guide_legend(reverse = TRUE),
+         color = guide_legend(reverse = TRUE)) +
+  theme_tufte()
+```
+
+Calculate the intensity of other ppp by quadrat:
+```{r quadrat-analysis, include=FALSE}
+
+# Calculate the intensity of the various point patterns by quadrat:
+
+# Crime
+
+crime5 <- quadratcount(events_recife_list$`Violent Crime`, tess = V)
+crime5_intensity <- intensity(crime5)
+
+# Catholic
+
+catholic5 <- quadratcount(events_recife_list$Catholic, tess = V)
+catholic5_intensity <- intensity(catholic5)
+
+# Evangelical
+
+evangelic5 <- quadratcount(events_recife_list$Evangelical, tess = V)
+evangelic5_intensity <- intensity(evangelic5)
+
+# Spiritist
+
+spiritist5 <- quadratcount(events_recife_list$Spiritist, tess = V)
+spiritist5_intensity <- intensity(spiritist5)
+
+# Ice cream
+
+ice_cream5 <- quadratcount(events_recife_list$`Ice Cream Shop`, tess = V)
+ice_cream5_intensity <- intensity(ice_cream5)
+
+# Pharmacies
+
+pharmacies5 <- quadratcount(events_recife_list$Pharmacy, tess = V)
+pharmacies5_intensity <- intensity(pharmacies5)
+
+# Supermarkets
+
+super_markets5 <- quadratcount(events_recife_list$Supermarket, tess = V)
+supermarkets5_intensity <- intensity(super_markets5)
+```
+
+```{r instensity-dataframe, include=FALSE}
+
+# Collect in a data.frame for convenience of plotting:
+
+intensity_all <- data.frame(SED = rep(c("Low", 
+                                        "Middle Low", 
+                                        "Middle", 
+                                        "Middle High", 
+                                        "High"), 
+                                      times = 7),
+                            Type = rep(c("Violent Crime", 
+                                         "Catholic",
+                                         "Evangelical",
+                                         "Spiritist",
+                                         "Ice Cream",
+                                         "Pharmacies",
+                                         "Super Markets"), 
+                                       each = 5),
+                            Intensity = c(crime5_intensity,
+                                          catholic5_intensity,
+                                          evangelic5_intensity,
+                                          spiritist5_intensity,
+                                          ice_cream5_intensity,
+                                          pharmacies5_intensity,
+                                          supermarkets5_intensity),
+                            Global_Intensity = rep(c(summary_crime$intensity,
+                                                     summary_catholic$intensity,
+                                                     summary_evangelic$intensity,
+                                                     summary_spiritist$intensity,
+                                                     summary_ice_cream$intensity,
+                                                     summary_pharmacy$intensity,
+                                                     summary_supermarket$intensity),
+                                                   each = 5)) %>%
+  mutate(SED = factor(SED, levels = c("Low", 
+                                        "Middle Low", 
+                                        "Middle", 
+                                        "Middle High", 
+                                        "High"),
+                      ordered = TRUE),
+         Type = factor(Type, levels = c("Violent Crime", 
+                                        "Catholic",
+                                        "Evangelical",
+                                        "Spiritist",
+                                        "Ice Cream",
+                                        "Pharmacies",
+                                        "Super Markets")))
+```
+
+```{r plot-crime-quadrat, echo=FALSE, fig.width=3, fig.asp=1 , cache = TRUE, fig.cap="\\label{fig:plot-crime-quadrat}Intensity of crime by level of Socio-Economic Deprivation; the dotted line indicates the global intensity of crime"}
+
+#Plot the intensity of crime:
+
+ggplot(data = intensity_all %>% filter(Type == "Violent Crime"), 
+       aes(x = SED, y = Intensity, fill = Type)) + 
+  geom_col(fill = "red") +
+  xlab("Socio-Economic Deprivation") +
+  geom_hline(aes(yintercept = Global_Intensity), linetype = "dotted", size = 1) +
+  #facet_wrap(Type ~.) +
+  theme_tufte() +
+  theme(axis.text.x = element_text(angle = 90))
+```
+
+```{r plot-events-quadrat, echo=FALSE, cache = TRUE, fig.cap="\\label{fig:plot-events-quadrat}Intensity of places of worship and commercial establishments by level of Socio-Economic Deprivation; the dotted lines indicates the respective global intensities"}
+
+#Plot the intensity of point patterns:
+
+ggplot(data = intensity_all %>% filter(Type != "Violent Crime"), 
+       aes(x = SED, y = Intensity, fill = Type)) + 
+  geom_col() +
+  xlab("Socio-Economic Deprivation") +
+  geom_hline(aes(yintercept = Global_Intensity), linetype = "dotted", size = 1) +
+  scale_fill_viridis_d(option = "viridis", direction = 1) +
+  scale_color_viridis_d(option = "viridis", direction = 1) +
+  facet_wrap(Type ~.) +
+  theme_tufte() +
+  theme(axis.text.x = element_text(angle = 90))
+```
+
+## Relative Intensity
+
+```{r distance-functions, include=FALSE, cache = TRUE}
+
+# Estimation of the relative intensity (see Baddeley et al. (2015) Spatial Point Pattern, pp. 179-183).
+
+# This needs that we convert the point patterns for places of worship to distance functions.
+This way, the covariates for the relative distribution estimate are distance estimates to places of worship.
+The function `distfun` is used for this:
+
+z_catholic <- distfun(events_recife_list$Catholic)
+z_evangelical <- distfun(events_recife_list$Evangelical)
+z_spiritist <- distfun(events_recife_list$Spiritist)
+z_ice_cream <- distfun(events_recife_list$`Ice Cream Shop`)
+z_pharmacies <- distfun(events_recife_list$Pharmacy)
+z_super_markets <- distfun(events_recife_list$Supermarket)
+```
+
+# Conclusions {#conclusions}
+
+Words go here.
 

### 2019-10-09 Moral_Communities_and_Crime_v1.tex
Using the initial draft from Edivaldo et al., first round of revisions to the text and the analysis. Worked on the introduction, structure of the paper, and started to put the analysis in.

% !TeX program = pdfLaTeX
\documentclass[smallextended]{svjour3}       % onecolumn (second format)
%\documentclass[twocolumn]{svjour3}          % twocolumn
%
\smartqed  % flush right qed marks, e.g. at end of proof
%
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage[utf8]{inputenc}

\usepackage[hyphens]{url} % not crucial - just used below for the URL
\usepackage{hyperref}
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

%
% \usepackage{mathptmx}      % use Times fonts if available on your TeX system
%
% insert here the call for the packages your document requires
%\usepackage{latexsym}
% etc.
%
% please place your own definitions here and don't use \def but
% \newcommand{}{}
%
% Insert the name of "your journal" with
% \journalname{myjournal}
%

%% load any required packages here

\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{multirow}
\usepackage{wrapfig}
\usepackage{float}
\usepackage{colortbl}
\usepackage{pdflscape}
\usepackage{tabu}
\usepackage{threeparttable}
\usepackage{threeparttablex}
\usepackage[normalem]{ulem}
\usepackage{makecell}
\usepackage{xcolor}

\begin{document}

\title{Do moral communities have a spatial dimension? A spatial exploratory
analysis of places of worship and violent crime in the city of Recife,
Brazil \thanks{Grants or other notes about the article that should go on the front page
should be placed here. General acknowledgments should be placed at the
end of the article.} }

    \titlerunning{Moral communities and crime}

\author{  Äüthör 1 \and  Âuthóř 2 \and  }

\institute{
        Äüthör 1 \at
     Department of YYY, University of XXX \\
     \email{\href{mailto:abc@def}{\nolinkurl{abc@def}}}  %  \\
%             \emph{Present address:} of F. Author  %  if needed
    \and
        Âuthóř 2 \at
     Department of ZZZ, University of WWW \\
     \email{\href{mailto:djf@wef}{\nolinkurl{djf@wef}}}  %  \\
%             \emph{Present address:} of F. Author  %  if needed
    \and
    }

\date{Received: date / Accepted: date}
% The correct dates will be entered by the editor

\maketitle

\begin{abstract}
A wealth of research exists to suggest that environmental attributes can
facilitate or deter criminal activities, as for instance the presence of
alcohol outlets or the design of defensible spaces. In this article we
present an exploratory spatial analysis of the relationship between
violent crime and places of worship in Recife, Brazil. This analysis
aims to explore whether there is a spatial dimension to ``moral
communities'' around places of worship. Use of a cluster detection
methodology, the results suggest the presence of a strong geographical
association between churches and crime. The evidence obtained for the
year 2010 indicates a process of clustering of homicide events in the
vicinity of places of worship. In addition, it was also possible to
verify a pattern of concentration between evangelical churches and
violent crimes.
\\
\keywords{
        Moral communities \and
        Crime \and
        Places of worship \and
        Point pattern analysis \and
        Intensity \and
    }

\end{abstract}

\def\spacingset#1{\renewcommand{\baselinestretch}%
{#1}\small\normalsize} \spacingset{1}

\hypertarget{intro}{%
\section{Introduction}\label{intro}}

Violent crime is a widespread phenomenon with negative impacts on many
spheres of life and society. Although the rate of homicides worldwide
has grown at a slower rate than the population, the number of people
killed in homicides still increased from 362,000 in 1990 to 464,000 in
2017 (United Nations Office on Drugs and Crime 2019a). Variations in
violent crime also tend to be extremely uneven. The Americas, with a
population of approximately 793.8 million people in 2019 (or about
10.3\% of the world population), accounted for approximately 173,000
homicides in 2017, or approximately 37.3\% of all homicides in the
world. Within the Americas, Brazil (along with Venezuela, Colombia, and
Mexico) is one of the largest countries in the region with high homicide
rates. The relevance of variations in the prevalence of violent crime is
in fact recognized by the United Nations Office on Drug and Crime
(2019a) as being key to achieving policy goals:

\begin{quote}
\begin{quote}
High levels of homicidal violence are concentrated in geographic and
demographic ``pockets'', so achieving target 16.1 of the Sustainable
Development Goals requires interventions within the specific regions,
countries, communities and population groups that are most at risk
(p.~35)
\end{quote}
\end{quote}

The endemic malady of violent crime represents in fact an important
factor that threatens to derail sustainable development goals. Violent
crime, furthermore, tends to be more accute in those regions most in
need of development. On the one hand, violent crime represents an
economic and social drag (Becker and Kassouf 2017) that affects people's
well-being, either through the loss of human life, mental health,
limitations in the right to public spaces (Doran and Burgess 2011), or
through disturbances in schooling and academic achievement (United
Nations Office on Drugs and Crime 2019b). In turn, these negative
effects combine to create the unfortunate conditions that exacerbate
criminal behavior, thus creating a vicious cycle of economic
disadvantage and crime (United Nations Office on Drugs and Crime 2019b).
Added to this scenario, social control institutions in Brazil, which in
practice should inhibit criminal practice, have deep deficiencies that
end up reducing the deterrent power of the criminal justice system.
Among these serious problems are the inefficiency of the police, the
lack of national legislation, the glacial pace of judicial processes.
and the weak situation of the prison system in the country (Menezes et
al. 2013). It is not surprising, given the above, that researchers have
joined calls for more research to increase our understanding of the
patterns of violent crime in Low and Middle Income Countries (LMIC;
Murray, Cerqueira, and Kahn 2013).

In order to counteract any criminological factors, it is important to
identify empirical regularities. Criminological factors include concrete
elements, such as the presence of arms or drugs, as well as
environmental factors, such as the built environment, in addition to
figurative factors, which include the moral costs of deviant behavior as
well as family supervision. Accordingly, a number of studies have
investigated various aspects of the environment and neighborhood design
(e.g., Foster, Giles-Corti, and Knuiman 2010; He, Páez, and Liu 2017;
Loukaitou-Sideris et al. 2001), whereas other studies have focused on
exposure to environmental attributes that signal weakened norms, such as
liquor and tobacco outlets associated with geographical hotspots of
crime (Brower and Carroll 2007; Deryol et al. 2016; Lipton et al. 2008;
Quick, Law, and Luan 2017). Yet another fruitful avenue for research,
and one that has only recently begun to attract attention, is the
presence of environmental attributes that can help reinforce moral
norms, such as schools and churches (e.g., Abdullah et al. 2018;
Davignon and Thomson 2015; Furr-Holden et al. 2010; Traunmuller 2011).
It is thus that in a recent paper, Warner and Konkel (n.d.) note that
the role of places of worship, as distinct entities from the members of
the congregations, have received less attention in empirical and
theoretical research for their potential to prevent crime. The role of
these institutions might be particularly important in places where the
State lacks the means or the will to enforce norms - as is the case in
Brazil.

With the above considerations in mind, the objective of this paper is to
investigate whether moral communities have a spatial dimension
discernible from the presence of places of worship. The case study is
the city of Recife, in the state of Pernambuco in Brazil's Northeast.
Recife is a large metropolitan area in a historically poor region, and
afflicted by high levels of violent crime. The empirical strategy is to
use spatial analysis to explore the potential geographical relationships
between violent criminal events, on the one hand, and places of worship
and a selection of commercial establishments, on the other.
Dissagregated data allows us to analyze the phenomena of interest at a
very high level of resolution as spatial point patterns.

After this introductory section, the rest of this paper is structured as
follows: Sect. \ref{background} reviews the literature on religion and
crime; then Sect. \ref{methods} presents the empirical strategy used in
the work. Sect. \ref{case} presents the context as well as the data used
ub the study; Sect. \ref{results} systematizes and discusses the main
results; and finally, the conclusions are presented in section
\ref{conclusions}.

\hypertarget{background}{%
\section{Background: Theoretical Perspectives}\label{background}}

Words go here.

\hypertarget{methods}{%
\section{Empirical Strategy}\label{methods}}

Two different techniques are deployed in this paper to study the spatial
distribution of criminal events, and their possible relationship to the
location of places of worship. Firstly, the F-function is a nearest
neighbor technique that is used to summarize the distance between a set
of points and their nearest events. This technique has found extensive
use in spatial epidemiology (see Gatrell et al., 1996) and has also been
used in spatial criminology (see, inter alia, Craglia et al., 2000;
Kiani et al., 2015; Levine, 2006; and Rogerson and Sun, 2001). The
second technique is the scan statistic. This method has also been
extensively used in spatial epidemiology (e.g., Hjalmars et al., 1996)
and has found use in spatial criminology as well (see He et al., 2017a;
Nakaya and Yano, 2010; Malleson and Andresen, 2015; and Shiode, 2011).
Use of these two techniques is meant to provide insights with respect to
1) the location of violent criminal events with respect to various types
of facilities; and 2) geographically identify places of worship
associated with clusters of criminal events. These two techniques are
briefly reviewed next.

\hypertarget{quadrat-analysis}{%
\subsection{Quadrat Analysis}\label{quadrat-analysis}}

Words go here.

\hypertarget{relative-intensity}{%
\subsection{Relative Intensity}\label{relative-intensity}}

Words go here.

\hypertarget{case}{%
\section{Case Study}\label{case}}

\hypertarget{context}{%
\subsection{Context}\label{context}}

The study is of the city of Recife, the capital of the state of
Pernambuco in the Northeast region of Brazil. With a population of
1,550,390 million inhabitants in 2010, and with an area of 7,133.2 km2,
Recife is one of the main Brazilian metropolises, exerting a great
economic influence in neighboring regions. However, the city experiences
a serious problem with violent crime, and has the dubious honor of being
one of the five capital cities in Brazil with the highest homicide rates
in the period under study (Menezes et al. 2013).

In the current context, the term ``violent crime'' is an umbrella for
several forms of infractions to the penal code. Following
recommendations of the National Secretariat of Public Security of the
Ministry of Justice of 2006, these are Violent Lethal and Intentional
Crimes (CLVI; which includes intentional homicide), theft followed by
death (robbery), and corporal injury followed by death. The data on CVLI
were extracted from the Police Information System of the Secretariat of
Social Defense of Pernambuco (INFOPOL / SDS-PE), which is the most
reliable, detailed, and comprehensive information on violent deaths in
the region.

\begin{table}

\caption{\label{tab:descriptive-statistics}\label{tab:descriptive-statistics}Characteristics of Violent Crime in Recife, July, 2008 - June, 2009}
\centering
\begin{tabular}[t]{lr}
\toprule
 & Percentage\\
\midrule
\addlinespace[0.3em]
\multicolumn{2}{l}{\textbf{Gender of Victim}}\\
\hspace{1em}Man & 91.35\\
\hspace{1em}Woman & 8.65\\
\addlinespace[0.3em]
\multicolumn{2}{l}{\textbf{Type of Crime}}\\
\hspace{1em}Murder & 97.82\\
\hspace{1em}Robbery & 2.00\\
\hspace{1em}Body injury followed by death & 0.18\\
\addlinespace[0.3em]
\multicolumn{2}{l}{\textbf{Ethnicity of Victim}}\\
\hspace{1em}Black and White & 95.11\\
\hspace{1em}Yellow and White & 1.27\\
\hspace{1em}Not Reported & 3.62\\
\addlinespace[0.3em]
\multicolumn{2}{l}{\textbf{Age of Victim}}\\
\hspace{1em}1 to 12 years & 0.18\\
\hspace{1em}13 to 17 years & 11.31\\
\hspace{1em}18 to 30 years & 61.70\\
\hspace{1em}31 to 65 years & 25.84\\
\hspace{1em}65 years and older & 0.73\\
\hspace{1em}Not Reported & 0.24\\
\addlinespace[0.3em]
\multicolumn{2}{l}{\textbf{Weapon Used}}\\
\hspace{1em}Firearm & 87.51\\
\hspace{1em}Other & 12.49\\
\bottomrule
\end{tabular}
\end{table}

\hypertarget{data}{%
\subsection{Data}\label{data}}

The data are organized at the individual level and it is possible to
obtain information about the location, day of the week, day of the
month, period of the day, as well as gender, age, and race of the
victim. The database used in this study comprises the period from July
1, 2008 to June 30, 2010. Some descriptive statistics regarding this
dataset are reported in Table 1, where it can be seen that overall,
about 91\% of the victims of CVLI in the period analyzed were men, while
approximately 98\% of violent crimes were homicides. In addition, most
of the victims were black or brown, and the youth population between the
ages of 18 and 30 is the most affected by violent crime. Lastly, it
should be noted that about 88\% of the criminal events under analysis
involved firearms. Information about places of interest was obtained
from the National Register of Addresses for Statistical Purposes (CNEFE
- Census 2010), which lists 78,056,411 urban and rural addresses,
distributed among the 316,574 census tracts. This is the first database
of its kind produced by IBGE, and the first version was produced at the
time of the 2000 Census. The way addresses are described in the National
Register is very rich, and it is possible to identify the names of the
places of worship including their denomination. Georeferencing was used
to geolocate each place of worship. In this way, a total of 1,719 places
of worship were geolocated in the city of Recife. Figure 1 (Panel a)
shows the spatial distribution of the 1,657 CVLI crimes that occurred in
the city of Recife between 2008.2 and 2010.1. Panel b) of the figure
shows the spatial distribution of places of worship in the region.

\begin{figure}

\includegraphics{Moral_Communities_and_Crime_v1_files/figure-latex/plot-crime-1} \hfill{}

\caption{\label{fig:plot-crime}Lethal and Intentional Violent Crime in Recife}\label{fig:plot-crime}
\end{figure}

In addition to places of worship, the National Register of Addresses for
Statistical Purposes was queried to extract facilities other than places
of worship. As discussed above, the idea is to identify points of
reference that can be used as controls, having a neutral morality
profile. For the sake of the present study, we selected pharmacies, ice
cream shops, and bakeries to construct our control group. These three
types of establishments comply with the criteria of being morally
neutral and having a spatial distribution commensurate with places of
worship. Figure 2 shows the spatial distribution of the control
establishments, namely pharmacies, bakeries, and ice cream shops in the
city of Recife. Note the similarity between the maps. As expected, there
are differences in the number of points, but the locations of cases and
controls are quite similar.

\begin{figure}

\includegraphics{Moral_Communities_and_Crime_v1_files/figure-latex/plot-events-1} \hfill{}

\caption{\label{fig:plot-events}Places of Worship and Commercial Establishments in Recife}\label{fig:plot-events}
\end{figure}

\begin{figure}

\includegraphics{Moral_Communities_and_Crime_v1_files/figure-latex/plot-sed-1} \hfill{}

\caption{\label{fig:plot-sed}Socio-Economic Deprivation in Recife}\label{fig:plot-sed}
\end{figure}

\hypertarget{results}{%
\section{Analysis and results}\label{results}}

\hypertarget{quadrat-analysis-1}{%
\subsection{Quadrat Analysis}\label{quadrat-analysis-1}}

\begin{figure}
\centering
\includegraphics{Moral_Communities_and_Crime_v1_files/figure-latex/plot-sed-as-quintiles-1.pdf}
\caption{\label{fig:plot-sed-as-quintiles}Socio-Economic Deprivation as
Quintiles}
\end{figure}

Calculate the intensity of other ppp by quadrat:

\begin{figure}
\centering
\includegraphics{Moral_Communities_and_Crime_v1_files/figure-latex/plot-crime-quadrat-1.pdf}
\caption{\label{fig:plot-crime-quadrat}Intensity of crime by level of
Socio-Economic Deprivation; the dotted line indicates the global
intensity of crime}
\end{figure}

\begin{figure}
\centering
\includegraphics{Moral_Communities_and_Crime_v1_files/figure-latex/plot-events-quadrat-1.pdf}
\caption{\label{fig:plot-events-quadrat}Intensity of places of worship
and commercial establishments by level of Socio-Economic Deprivation;
the dotted lines indicates the respective global intensities}
\end{figure}

\hypertarget{relative-intensity-1}{%
\subsection{Relative Intensity}\label{relative-intensity-1}}

\hypertarget{conclusions}{%
\section{Conclusions}\label{conclusions}}

Words go here.

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\leavevmode\hypertarget{ref-Abdullah2018amenities}{}%
Abdullah, Snhs, F. A. Bohani, Z. A. Nazri, Y. Jeffry, M. A. Abdullah, M.
N. Junoh, and Z. A. Kasim. 2018. ``Amenities Surrounding Commercial
Serial Crime Prediction at Greater Valley and Kuala Lumpur Using K-Means
Clustering.'' Journal Article. \emph{Jurnal Teknologi} 80 (4): 43--53.
\href{\%3CGo\%20to\%20ISI\%3E://WOS:000437011000005}{\textless{}Go to ISI\textgreater{}://WOS:000437011000005}.

\leavevmode\hypertarget{ref-Becker2017analise}{}%
Becker, Kalinca Léia, and Ana Lúcia Kassouf. 2017. ``Uma análise Do
Efeito Dos Gastos Públicos Em Educação Sobre a Criminalidade No
Brasil.'' \emph{Economia E Sociedade} 26 (1): 215--42.

\leavevmode\hypertarget{ref-Brower2007spatial}{}%
Brower, A. M., and L. Carroll. 2007. ``Spatial and Temporal Aspects of
Alcohol-Related Crime in a College Town.'' Journal Article.
\emph{Journal of American College Health} 55 (5): 267--75.
\url{https://doi.org/10.3200/jach.55.5.267-276}.

\leavevmode\hypertarget{ref-Davignon2015christian}{}%
Davignon, P., and R. A. Thomson. 2015. ``Christian Colleges and
Universities as Moral Communities: The Effects of Institutional
Characteristics on Student Religiosity.'' Journal Article. \emph{Review
of Religious Research} 57 (4): 531--54.
\url{https://doi.org/10.1007/s13644-015-0214-5}.

\leavevmode\hypertarget{ref-Deryol2016crime}{}%
Deryol, R., P. Wilcox, M. Logan, and J. Wooldredge. 2016. ``Crime Places
in Context: An Illustration of the Multilevel Nature of Hot Spot
Development.'' Journal Article. \emph{Journal of Quantitative
Criminology} 32 (2): 305--25.
\url{https://doi.org/10.1007/s10940-015-9278-1}.

\leavevmode\hypertarget{ref-Doran2011putting}{}%
Doran, Bruce J, and Melissa B Burgess. 2011. \emph{Putting Fear of Crime
on the Map: Investigating Perceptions of Crime Using Geographic
Information Systems}. Springer Science \& Business Media.

\leavevmode\hypertarget{ref-Foster2010neighbourhood}{}%
Foster, S., B. Giles-Corti, and M. Knuiman. 2010. ``Neighbourhood Design
and Fear of Crime: A Social-Ecological Examination of the Correlates of
Residents' Fear in New Suburban Housing Developments.'' Journal Article.
\emph{Health \& Place} 16 (6): 1156--65.
\url{https://doi.org/10.1016/j.healthplace.2010.07.007}.

\leavevmode\hypertarget{ref-Furr2010metric}{}%
Furr-Holden, C. D. M., K. D. M. Campbell, A. J. Milam, M. J. Smart, N.
A. Ialongo, and P. J. Leaf. 2010. ``Metric Properties of the
Neighborhood Inventory for Environmental Typology (Nifety): An
Environmental Assessment Tool for Measuring Indicators of Violence,
Alcohol, Tobacco, and Other Drug Exposures.'' Journal Article.
\emph{Evaluation Review} 34 (3): 159--84.
\url{https://doi.org/10.1177/0193841x10368493}.

\leavevmode\hypertarget{ref-He2017built}{}%
He, Li, Antonio Páez, and Desheng Liu. 2017. ``Built Environment and
Violent Crime: An Environmental Audit Approach Using Google Street
View.'' Journal Article. \emph{Computers, Environment and Urban Systems}
66: 83--95.
\url{https://doi.org/http://dx.doi.org/10.1016/j.compenvurbsys.2017.08.001}.

\leavevmode\hypertarget{ref-Lipton2008spatial}{}%
Lipton, R., A. Banerjee, D. Levy, N. Manzanilla, and M. Cochrane. 2008.
``The Spatial Distribution of Underage Tobacco Sales in Los Angeles.''
Journal Article. \emph{Substance Use \& Misuse} 43 (11): 1597--1617.
\url{https://doi.org/10.1080/10826080802241110}.

\leavevmode\hypertarget{ref-Loukaitou2001measuring}{}%
Loukaitou-Sideris, A., R. Liggett, H. Iseki, and W. Thurlow. 2001.
``Measuring the Effects of Built Environment on Bus Stop Crime.''
Journal Article. \emph{Environment and Planning B-Planning \& Design} 28
(2): 255--80. \url{https://doi.org/10.1068/b2642r}.

\leavevmode\hypertarget{ref-Menezes2013spatial}{}%
Menezes, T., R. Silveira-Neto, C. Monteiro, and J. L. Ratton. 2013.
``Spatial Correlation Between Homicide Rates and Inequality: Evidence
from Urban Neighborhoods.'' Journal Article. \emph{Economics Letters}
120 (1): 97--99. \url{https://doi.org/10.1016/j.econlet.2013.03.040}.

\leavevmode\hypertarget{ref-Murray2013crime}{}%
Murray, J., D. R. D. Cerqueira, and T. Kahn. 2013. ``Crime and Violence
in Brazil: Systematic Review of Time Trends, Prevalence Rates and Risk
Factors.'' Journal Article. \emph{Aggression and Violent Behavior} 18
(5): 471--83. \url{https://doi.org/10.1016/j.avb.2013.07.003}.

\leavevmode\hypertarget{ref-Quick2017influence}{}%
Quick, M., J. Law, and H. Luan. 2017. ``The Influence of on-Premise and
Off-Premise Alcohol Outlets on Reported Violent Crime in the Region of
Waterloo, Ontario: Applying Bayesian Spatial Modeling to Inform Land Use
Planning and Policy.'' Journal Article. \emph{Applied Spatial Analysis
and Policy} 10 (3): 435--54.
\url{https://doi.org/10.1007/s12061-016-9191-5}.

\leavevmode\hypertarget{ref-Traunmuller2011moral}{}%
Traunmuller, R. 2011. ``Moral Communities? Religion as a Source of
Social Trust in a Multilevel Analysis of 97 German Regions.'' Journal
Article. \emph{European Sociological Review} 27 (3): 346--63.
\url{https://doi.org/10.1093/esr/jcq011}.

\leavevmode\hypertarget{ref-Unodc2019executive}{}%
United Nations Office on Drugs and Crime. 2019a. ``Global Study on
Homicide 2019: Executive Summary.'' UNODC Vienna.

\leavevmode\hypertarget{ref-Unodc2019development}{}%
---------. 2019b. ``Global Study on Homicide 2019: Homicide, Development
and the Sustainable Development Goals.'' UNODC Vienna.

\leavevmode\hypertarget{ref-Warner2019neighborhood}{}%
Warner, B. D., and R. H. Konkel. n.d. ``Neighborhood Churches and Their
Relationship to Neighborhood Processes Important for Crime Prevention.''
Journal Article. \emph{Journal of Urban Affairs}.
\url{https://doi.org/10.1080/07352166.2019.1581030}.

\bibliographystyle{spbasic}
\bibliography{bibliography.bib}

\end{document}

### 2019-10-13 Moral_Communities_and_Crime_v1.Rmd
Completed revisions to Background section and abstract.

@@
 authors: 
-- name: Äüthör 1
-  address: Department of YYY, University of XXX
-  email: abc@def
+- name: Aye B. Cee
+  address: A Department, University of XXX
+  email: abc@example.edu
 
-- name: Âuthóř 2
-  address: Department of ZZZ, University of WWW
-  email: djf@wef
+- name: Dee E. Eff
+  address: A Department, University of XXX
+  email: def@example.edu
+
+- name: Gee H. Eye
+  address: A Department, University of XXX
+  email: ghi@example.edu
+
+- name: Jay K. Elle
+  address: Some School, University of WWW
+  email: jkl@school.edu
 
@@
 abstract: |
-  A wealth of research exists to suggest that environmental attributes can facilitate or deter criminal activities, as for instance the presence of alcohol outlets or the design of defensible spaces.
-In this article we present an exploratory spatial analysis of the relationship between violent crime and places of worship in Recife, Brazil.
-This analysis aims to explore whether there is a spatial dimension to “moral communities” around places of worship.
-Use of a cluster detection methodology, the results suggest the presence of a strong geographical association between churches and crime.
-The evidence obtained for the year 2010 indicates a process of clustering of homicide events in the vicinity of places of worship.
-In addition, it was also possible to verify a pattern of concentration between evangelical churches and violent crimes.
+  Religious tenets of the type "thou shalt not kill" and their equivalents in many world religions have functioned as _de facto_ social policy for determining appropriate and acceptable behavior over the centuries.
+With the advent of the scientific study of religion, there has been a growth of interest in the role of religions to operate as moral communities.
+Moral communities, a concept closely related to informal social controls, are of interest in countries and regions where formal controls are weak and ineffective.
+The objective of this paper is to present a spatial analysis of Violent and Intentional Crime in the city of Recife in Brazil, with a focus on the possible interactions between criminal events and places of worship.
+Previous research into moral communities has advocated the need for analysis at different scales, and this analysis contributes to the literature by using micro-level data and appropriate spatial analytical tools for spatial point patterns.
+Analysis is condcuted using three different types of places of worship (Catholic, Evangelical, and Spiritist) and three types of business establishments as controls (ice cream shops, pharmacies, and supermarkets).
+The results suggest that Catholic and Spiritist places of worship do not project moral communities geographically more than, say, ice cream shops.
+The intensity of criminal events in the proximity of Evangelical places of worship, in contrast, is markedly higher than for any of our other referential events.
 
@@
 Violent crime is a widespread phenomenon with negative impacts on many spheres of life and society.
-Although the rate of homicides worldwide has grown at a slower rate than the population, the number of people killed in homicides still increased from 362,000 in 1990 to 464,000 in 2017 [@Unodc2019executive].
-Variations in violent crime also tend to be extremely uneven.
+Although the rate of homicides worldwide has grown at a slower rate than the population in the past few decades, the number of people killed in homicides still increased from 362,000 in 1990 to 464,000 in 2017 [@Unodc2019executive].
+Furthermore, variations in the prevalence of violent crime tend to be extremely uneven.
 The Americas, with a population of approximately 793.8 million people in 2019 (or about 10.3% of the world population), accounted for approximately 173,000 homicides in 2017, or approximately 37.3% of all homicides in the world.
 Within the Americas, Brazil (along with Venezuela, Colombia, and Mexico) is one of the largest countries in the region with high homicide rates.
-The relevance of variations in the prevalence of violent crime is in fact recognized by the United Nations Office on Drug and Crime  [-@Unodc2019executive] as being key to achieving policy goals:
+And there are notable variations at the sub-national and intra-metropolitan levels as well.
+In fact, understanding the variations in the prevalence of violent crime is recognized as key to achieving policy goals:
 
@@
 achieving target 16.1 of the Sustainable Development Goals requires interventions within the specific
-regions, countries, communities and population groups that are most at risk (p. 35)
+regions, countries, communities and population groups that are most at risk [@Unodc2019executive, p. 35]
 
-The endemic malady of violent crime  represents in fact an important factor that threatens to derail sustainable development goals.
+This concern is not iddle: the endemic malady of violent crime represents an important factor that threatens to derail sustainable development goals.
 Violent crime, furthermore, tends to be more accute in those regions most in need of development.
-On the one hand, violent crime represents an economic and social drag [@Becker2017analise] that affects people's well-being, either through the loss of human life, mental health, limitations in the right to public spaces [@Doran2011putting], or through disturbances in schooling and academic achievement [@Unodc2019development].
-In turn, these negative effects combine to create the unfortunate conditions that exacerbate criminal behavior, thus creating a vicious cycle of economic disadvantage and crime [@Unodc2019development].
-Added to this scenario, social control institutions in Brazil, which in practice should inhibit criminal practice, have deep deficiencies that end up reducing the deterrent power of the criminal justice system.
-Among these serious problems are the inefficiency of the police, the lack of national legislation, the glacial pace of judicial processes.
-and the weak situation of the prison system in the country [@Menezes2013spatial].
-It is not surprising, given the above, that researchers have joined calls for more research to increase our understanding of the patterns of violent crime in Low and Middle Income Countries [LMIC; @Murray2013crime].
+On the one hand, violent crime represents an economic and social drag [@Becker2017analise].
+The impacts are multifaceted, since they affect people's well-being, either through the loss of human life, mental health, and limitations in the right to public spaces [@Doran2011putting], or through disturbances in schooling and academic achievement [@Unodc2019development].
+In turn, these negative effects combine to create the unfortunate conditions that tend to breed criminal behavior, thus creating a vicious cycle of economic disadvantage and violence [@Unodc2019development].
+It is not surprising, given these pecularities, that researchers have joined calls for increased attention to the study of the patterns of violent crime in Low and Middle Income Countries [LMIC; @Murray2013crime].
 
-In order to counteract any criminological factors, it is important to identify empirical regularities.
-Criminological factors include concrete elements, such as the presence of arms or drugs, as well as environmental factors, such as the built environment, in addition to figurative factors, which include the moral costs of deviant behavior as well as family supervision.
-Accordingly, a number of studies have investigated various aspects of the environment and neighborhood design [e.g., @Foster2010neighbourhood; @He2017built; @Loukaitou2001measuring], whereas other studies have focused on exposure to environmental attributes that signal weakened norms, such as liquor and tobacco outlets associated with geographical hotspots of crime [@Brower2007spatial; @Deryol2016crime; @Lipton2008spatial; @Quick2017influence].
-Yet another fruitful avenue for research, and one that has only recently begun to attract attention, is the presence of environmental attributes that can help reinforce moral norms, such as schools and churches [e.g., @Abdullah2018amenities; @Davignon2015christian; @Furr2010metric; @Traunmuller2011moral].
-It is thus that in a recent paper, Warner and Konkel [-@Warner2019neighborhood] note that the role of places of worship, as distinct entities from the members of the congregations, have received less attention in empirical and theoretical research for their potential to prevent crime.
-The role of these institutions might be particularly important in places where the State lacks the means or the will to enforce norms - as is the case in Brazil.
+Added to the scenario, formal social control institutions in countries like Brazil leave something (or much) to be desired: while in practice they should inhibit criminal behavior, they suffer from deep deficiencies that end up reducing the deterrent power of the justice system.
+Serious institutional problems include the inefficiency of the police, the lack of national legislation, the glacial pace of judicial processes, and the weak situation of the prison system in the country [@Menezes2013spatial].
+This makes it even more urgent to understand the role of _informal social controls_, i.e., the ability of community organizations such as schools, clubs, and neighborbood associations to suppress crime by strengthening the capabilities of neighbors to control inappropriate behavior [@Groff2015informal].
 
-With the above considerations in mind, the objective of this paper is to investigate whether moral communities have a spatial dimension discernible from the presence of places of worship.
-The case study is the city of Recife, in the state of Pernambuco in Brazil's Northeast.
-Recife is a large metropolitan area in a historically poor region, and afflicted by high levels of violent crime.
-The empirical strategy is to use spatial analysis to explore the potential geographical relationships between violent criminal events, on the one hand, and places of worship and a selection of commercial establishments, on the other.
-Dissagregated data allows us to analyze the phenomena of interest at a very high level of resolution as spatial point patterns.
+In order to understand the factors that can potentially help to deter crime, beyond formal control, it is important to identify empirical regularities.
+Criminological factors include concrete elements (such as te presence of arms or drugs, or elements of the built environment), and also figurative factors, which include the soci costs of deviant behavior as well as family supervision.
+Accordingly, a number of studies have investigated various aspects of the environment and neighborhood design [e.g., @Foster2010neighbourhood; @He2017built; @Loukaitou2001measuring], whereas other studies have focused on exposure to environmental attributes that signal weakened norms, such as liquor and tobacco outlets [e.g., @Brower2007spatial; @Deryol2016crime; @Lipton2008spatial; @Quick2017influence].
 
-After this introductory section, the rest of this paper is structured as follows: Sect.
-\ref{background} reviews the literature on religion and crime; then Sect.
-\ref{methods} presents the empirical strategy used in the work.
-Sect.
-\ref{case} presents the context as well as the data used ub the study; Sect.
-\ref{results} systematizes and discusses the main results; and finally, the conclusions are presented in section \ref{conclusions}.
+Yet another fruitful avenue for research, and one that has only recently begun to be explored, is the presence of environmental attributes that can help to reinforce moral norms, such as schools and churches [e.g., @Abdullah2018amenities; @Davignon2015christian; @Furr2010metric; @Traunmuller2011moral].
+It is thus that in a recent paper, Warner and Konkel [-@Warner2019neighborhood] note that the role of places of worship, as distinct entities from the members of the congregations, have received less attention in empirical and theoretical research for their deterrence potential.
+The role of these institutions might be particularly important in places where formal state institutions lack the means or the will to enforce norms - as is the case in Brazil.
 
-# Background: Theoretical Perspectives {#background}
+With the above considerations in mind, the objective of this paper is to investigate whether and how places of workship correlate spatially with criminal events, or in other words, to investigate whether their signals as moral communities have a discernible spatial dimension.
+The case study is the city of Recife in the state of Pernambuco, in Brazil's Northeast.
+Recife is a large and important metropolitan area in a historically poor region, and afflicted by high levels of violent crime, having the dubious honor of being one of five state capitals with the highest rates of homicide in the period under study [@Menezes2013spatial].
+The empirical strategy is to use spatial analysis to explore the potential geographical relationships between violent criminal events, on the one hand, and places of worship and a selection of commercial establishments that serve as controls, on the other.
+Dissagregated data allows us to analyze the relevant point patterns at a very high level of resolution.
 
-Words go here.
+After this introductory section, the rest of this paper is structured as follows: Section \ref{background} reviews some relevant theoretical perspectives on religion and crime; then Section \ref{methods} presents the empirical strategy used in the work.
+Section \ref{case} presents the context as well as the data used in the study; Section \ref{results} presents the analysis, whereas the results are discussed in Section \ref{discussion}.
+Finally, some concluding remarks are offered in Section \ref{conclusions}.
+
+# Background: Theoretical Perspectives on Religion and Crime {#background}
+
+Why do humans behave morally?
+For millennia, it has been the role of religion to provide the basic tenets of morally acceptable behavior: thou shalt not kill et al., and their equivalent in many world religions [e.g., @Donovan1986different].
+Enforcement of such tenets implies different mechanisms, including _sin_, _haram_, _karma_, and _tapu_, with punishment delivered by hellfire and exile to _Jahannam_ or _Gehenna_, to mention just some choice places of torment.
+In addition to acting as _de facto_ social policy for much of history, many of these religious tenets still have the force of law in many cultures and regions - and even where they do not, they are held by some researchers and policy experts to be helpful complements to reduce crime in any case [e.g., @Durrant2017religion; @Johnson2011more].
+The hypothesis that religion can act as a factor that deters and reduces criminal behavior, therefore, has prompted the scientific study of the effectiveness of religion on moral behavior [@Hoffmann2015religion].
+
+An early effort to theorize the effect of religion on criminal behavior was the _hellfire hypothesis_ of Hirschi and Stark [-@Hirschi1969hellfire].
+The focus of this hypothesis is the threat of extratemporal (and possibly eternal) punishment, and how this threat can deter believers from commandment-breaking (also see Pascal's wager).
+Accordingly, Hirsch and Stark [-@Hirschi1969hellfire] posit that negative correlations between crime and religion are a consequence of a sense of commitment with normative values, a commitment that is ritually reinforced in a regular fashion (e.g., the liturgical rite of peace in Catholic Mass).
+Rohrbaugh and Jessor [-@Rohrbaugh1975religiosity], for instance, note that by attributing to divinity (or some other supra-terrenal entity) the supreme force of punishment, religion helps to build a sense of obduracy, or "hardening" against temptation.
+
+Despite its intuitive appeal, research has not provided much direct evidence for the hypothesis of hellfire [@Hoffmann2015religion, p.
+1].
+In a bizarre twist, even mainstream doctrines (such as absolution) may in fact encourage beliefs that neutralize the fear of _terrenal_ punishment, and thus turn out to be criminogenic [e.g., @Topalli2013god].
+More deviant cases can even recast criminal behavior as a form of spiritual insurgency [@Chesnut2017devoted].
+Counterexamples like this notwithstanding, researchers remain open to the possibility that "religion may still serve as a social control mechanism by encouraging conventional beliefs, monitoring behaviors, enhancing family attachments, or providing conventional activities" [@Hoffmann2015religion, p.
+1].
+In this way, the hypothesis of _moral communities_ [@Stark1996religion] recognizes that social integration is essential for increasing social control, thus reducing the practice of behavior that is not in accordance with current norms.
+Rohrbaugh and Jessor [-@Rohrbaugh1975religiosity] also emphasize that religion acts as social control since it defines what is an appropriate attitude according to moral values, thus making moral communities a close relative of the concept of _informal social control_, defined as “the ability of social groups or institutions to make norms or rules effective” [@Reiss1951delinquency, p.
+196; cited in @Groff2015informal, p.
+91].
+
+The hypothesis of moral communities has over the years been used to examine a variety of outcomes of interest.
+Recent examples include Stroope and Baker's [-@Stroope2018moral] exploration of religiosity and self-rated health; Davignon and Thomson [-@Davignon2015christian] with their research on institutional context and the religiosity of students; and the study of religion as a source of trust of Traunmuller [-@Traunmuller2011moral].
+This is in addition, of course, to numerous studies on criminal behavior such as Eitle [-@Eitle2011religion], an author who explored the deterrence power of religion on gambling; Lee and Bartkowski's [-@Lee2004love] investigation of juvenile homicide in rural areas; and the research of Regnerus [-@Regnerus2003moral] on adolescent delinquency.
+The hypothesis of moral communities as a form of informal social control, on the other hand, has been less studied from a geographical perspective, and it is only recently that has attracted the attention of researchers.
+Groff [-@Groff2015informal], for instance, discusses informal social control as a phenomenon that can plaussibly operate at different geographical scales, from the level of the home and family, through the street block and neighborhood, possibly to other scales such as the county.
+In this way, Nie and Yang [-@Nie2019smoking] remark on a recent paper on the lack of research conducted to study how the religious context of a geographical area (e.g., a county) may influence (youth smoking) behavior (p.
+2).
+This point is echoed by Warner and Konkel [-@Warner2019neighborhood], who moreover recommend the use of even smaller units of analysis (e.g., Census Block Groups), in an effort to reduce aggregation bias [see @Hipp2007block], but more importantly because the processes defined by social disorganization theory are thought to occur at relatively small geographical scales.
+
+The preceding review makes clear the need for research at various geographical scales.
+From a theoretical perspective, the social mechanisms that underpin the hypothesis of moral communities and informal social control can happen at various geographical scales, some of which previous research has already addressed: the work of Nie and Yang [-@Nie2019smoking], for one, highlights the role of processes detectable at a relatively high level of aggregation, whereas Warner and Konkel [-@Warner2019neighborhood], looking specifically at the effect of places of worship, bring the study closer to the level of the neighborhood.
+In what follows, we propose to adopt a much more disaggregated approach by considering the intensity of criminal events with respect to places of worship.
+This requires micro-level data both of crimes and places of worship, and the use of appropriate spatial analytical tools for point patterns, as discussed next.
 
@@
 In the current context, the term "violent crime" is an umbrella for several forms of infractions to the penal code.
-Following recommendations of the National Secretariat of Public Security of the Ministry of Justice of 2006, these are Violent Lethal and Intentional Crimes (CLVI; which includes intentional homicide), theft followed by death (robbery), and corporal injury followed by death.
+Following recommendations of the National Secretariat of Public Security of the Ministry of Justice of 2006, these are Violent Lethal and Intentional Crimes (VLIC; which includes intentional homicide), theft followed by death (robbery), and corporal injury followed by death.
 The data on CVLI were extracted from the Police Information System of the Secretariat of Social Defense of Pernambuco (INFOPOL / SDS-PE), which is the most reliable, detailed, and comprehensive information on violent deaths in the region.
@@
 
+# Discussion {#discussion}
+
+An opposing viewpoint is offered by the theory of conservative Protestants [@Ellison2003enduring].
+This theory argues that religion is not always beneficial in order to guarantee social order, and consequently to result in lower crime rates.
+This theory suggested that, conservative Protestants are more tolerate with violent behavior when associated with defense of honor, family, property, or woman.
+However, high rates are not expected for all types of crimes in densely populated areas of conservative Protestants.
+But what justifies this behavior?
+Ellison et al.
+[-@Ellison2003enduring] argue that this tolerance stems from a religious view of the world.
+More clearly, punishment is seen as a legitimate form of retribution whose explanation can be attributed to the literal interpretation of the Bible.
+In the specific context of Brazil, it is important to consider that the interpretations and meanings attributed to urban violence during the last two decades in Brazil often had religious figures as protagonists [@Birman2012violencia].
+
 # Conclusions {#conclusions}
 
-Words go here.
+The theory of conservative Protestants, argues that conservative Protestants are more likely to tolerate violent behavior of a defensive or punitive character, especially when associated to defense of honor, family, property, or woman.
+Our objective in the study is to explore the question whether moral communities have a spatial dimension.
+To this end we deploy an array of spatial analytical tools to investigate potential patterns of clustering of violent crime and places of worship, using as well other kinds of urban facilities to act as controls.
+The case study is the city of Recife, in the northeast of Brazil, and analysis is based on geographically detailed datasets of criminal events as well as several types of urban facilities.
 

### 2019-10-13 Moral_Communities_and_Crime_v1.tex
Completed revisions to Background section and abstract.

@@
 
-\author{  Äüthör 1 \and  Âuthóř 2 \and  }
+\author{  Aye B. Cee \and  Dee E. Eff \and  Gee H. Eye \and  Jay K. Elle \and  }
 
 \institute{
-        Äüthör 1 \at
-     Department of YYY, University of XXX \\
-     \email{\href{mailto:abc@def}{\nolinkurl{abc@def}}}  %  \\
+        Aye B. Cee \at
+     A Department, University of XXX \\
+     \email{\href{mailto:abc@example.edu}{\nolinkurl{abc@example.edu}}}  %  \\
 %             \emph{Present address:} of F. Author  %  if needed
     \and
-        Âuthóř 2 \at
-     Department of ZZZ, University of WWW \\
-     \email{\href{mailto:djf@wef}{\nolinkurl{djf@wef}}}  %  \\
+        Dee E. Eff \at
+     A Department, University of XXX \\
+     \email{\href{mailto:def@example.edu}{\nolinkurl{def@example.edu}}}  %  \\
+%             \emph{Present address:} of F. Author  %  if needed
+    \and
+        Gee H. Eye \at
+     A Department, University of XXX \\
+     \email{\href{mailto:ghi@example.edu}{\nolinkurl{ghi@example.edu}}}  %  \\
+%             \emph{Present address:} of F. Author  %  if needed
+    \and
+        Jay K. Elle \at
+     Some School, University of WWW \\
+     \email{\href{mailto:jkl@school.edu}{\nolinkurl{jkl@school.edu}}}  %  \\
 %             \emph{Present address:} of F. Author  %  if needed
@@
 \begin{abstract}
-A wealth of research exists to suggest that environmental attributes can
-facilitate or deter criminal activities, as for instance the presence of
-alcohol outlets or the design of defensible spaces. In this article we
-present an exploratory spatial analysis of the relationship between
-violent crime and places of worship in Recife, Brazil. This analysis
-aims to explore whether there is a spatial dimension to ``moral
-communities'' around places of worship. Use of a cluster detection
-methodology, the results suggest the presence of a strong geographical
-association between churches and crime. The evidence obtained for the
-year 2010 indicates a process of clustering of homicide events in the
-vicinity of places of worship. In addition, it was also possible to
-verify a pattern of concentration between evangelical churches and
-violent crimes.
+Religious tenets of the type ``thou shalt not kill'' and their
+equivalents in many world religions have functioned as \emph{de facto}
+social policy for determining appropriate and acceptable behavior over
+the centuries. With the advent of the scientific study of religion,
+there has been a growth of interest in the role of religions to operate
+as moral communities. Moral communities, a concept closely related to
+informal social controls, are of interest in countries and regions where
+formal controls are weak and ineffective. The objective of this paper is
+to present a spatial analysis of Violent and Intentional Crime in the
+city of Recife in Brazil, with a focus on the possible interactions
+between criminal events and places of worship. Previous research into
+moral communities has advocated the need for analysis at different
+scales, and this analysis contributes to the literature by using
+micro-level data and appropriate spatial analytical tools for spatial
+point patterns. Analysis is condcuted using three different types of
+places of worship (Catholic, Evangelical, and Spiritist) and three types
+of business establishments as controls (ice cream shops, pharmacies, and
+supermarkets). The results suggest that Catholic and Spiritist places of
+worship do not project moral communities geographically more than, say,
+ice cream shops. The intensity of criminal events in the proximity of
+Evangelical places of worship, in contrast, is markedly higher than for
+any of our other referential events.
 \\
@@
 spheres of life and society. Although the rate of homicides worldwide
-has grown at a slower rate than the population, the number of people
-killed in homicides still increased from 362,000 in 1990 to 464,000 in
-2017 (United Nations Office on Drugs and Crime 2019a). Variations in
-violent crime also tend to be extremely uneven. The Americas, with a
-population of approximately 793.8 million people in 2019 (or about
-10.3\% of the world population), accounted for approximately 173,000
-homicides in 2017, or approximately 37.3\% of all homicides in the
-world. Within the Americas, Brazil (along with Venezuela, Colombia, and
-Mexico) is one of the largest countries in the region with high homicide
-rates. The relevance of variations in the prevalence of violent crime is
-in fact recognized by the United Nations Office on Drug and Crime
-(2019a) as being key to achieving policy goals:
+has grown at a slower rate than the population in the past few decades,
+the number of people killed in homicides still increased from 362,000 in
+1990 to 464,000 in 2017 (United Nations Office on Drugs and Crime
+2019a). Furthermore, variations in the prevalence of violent crime tend
+to be extremely uneven. The Americas, with a population of approximately
+793.8 million people in 2019 (or about 10.3\% of the world population),
+accounted for approximately 173,000 homicides in 2017, or approximately
+37.3\% of all homicides in the world. Within the Americas, Brazil (along
+with Venezuela, Colombia, and Mexico) is one of the largest countries in
+the region with high homicide rates. And there are notable variations at
+the sub-national and intra-metropolitan levels as well. In fact,
+understanding the variations in the prevalence of violent crime is
+recognized as key to achieving policy goals:
 
@@
 countries, communities and population groups that are most at risk
-(p.~35)
+(United Nations Office on Drugs and Crime 2019a, 35)
 \end{quote}
@@
 
-The endemic malady of violent crime represents in fact an important
-factor that threatens to derail sustainable development goals. Violent
-crime, furthermore, tends to be more accute in those regions most in
-need of development. On the one hand, violent crime represents an
-economic and social drag (Becker and Kassouf 2017) that affects people's
-well-being, either through the loss of human life, mental health,
-limitations in the right to public spaces (Doran and Burgess 2011), or
-through disturbances in schooling and academic achievement (United
-Nations Office on Drugs and Crime 2019b). In turn, these negative
-effects combine to create the unfortunate conditions that exacerbate
-criminal behavior, thus creating a vicious cycle of economic
-disadvantage and crime (United Nations Office on Drugs and Crime 2019b).
-Added to this scenario, social control institutions in Brazil, which in
-practice should inhibit criminal practice, have deep deficiencies that
-end up reducing the deterrent power of the criminal justice system.
-Among these serious problems are the inefficiency of the police, the
-lack of national legislation, the glacial pace of judicial processes.
-and the weak situation of the prison system in the country (Menezes et
-al. 2013). It is not surprising, given the above, that researchers have
-joined calls for more research to increase our understanding of the
-patterns of violent crime in Low and Middle Income Countries (LMIC;
-Murray, Cerqueira, and Kahn 2013).
+This concern is not iddle: the endemic malady of violent crime
+represents an important factor that threatens to derail sustainable
+development goals. Violent crime, furthermore, tends to be more accute
+in those regions most in need of development. On the one hand, violent
+crime represents an economic and social drag (Becker and Kassouf 2017).
+The impacts are multifaceted, since they affect people's well-being,
+either through the loss of human life, mental health, and limitations in
+the right to public spaces (Doran and Burgess 2011), or through
+disturbances in schooling and academic achievement (United Nations
+Office on Drugs and Crime 2019b). In turn, these negative effects
+combine to create the unfortunate conditions that tend to breed criminal
+behavior, thus creating a vicious cycle of economic disadvantage and
+violence (United Nations Office on Drugs and Crime 2019b). It is not
+surprising, given these pecularities, that researchers have joined calls
+for increased attention to the study of the patterns of violent crime in
+Low and Middle Income Countries (LMIC; Murray, Cerqueira, and Kahn
+2013).
 
-In order to counteract any criminological factors, it is important to
-identify empirical regularities. Criminological factors include concrete
-elements, such as the presence of arms or drugs, as well as
-environmental factors, such as the built environment, in addition to
-figurative factors, which include the moral costs of deviant behavior as
-well as family supervision. Accordingly, a number of studies have
-investigated various aspects of the environment and neighborhood design
-(e.g., Foster, Giles-Corti, and Knuiman 2010; He, Páez, and Liu 2017;
-Loukaitou-Sideris et al. 2001), whereas other studies have focused on
-exposure to environmental attributes that signal weakened norms, such as
-liquor and tobacco outlets associated with geographical hotspots of
-crime (Brower and Carroll 2007; Deryol et al. 2016; Lipton et al. 2008;
-Quick, Law, and Luan 2017). Yet another fruitful avenue for research,
-and one that has only recently begun to attract attention, is the
-presence of environmental attributes that can help reinforce moral
-norms, such as schools and churches (e.g., Abdullah et al. 2018;
-Davignon and Thomson 2015; Furr-Holden et al. 2010; Traunmuller 2011).
-It is thus that in a recent paper, Warner and Konkel (n.d.) note that
-the role of places of worship, as distinct entities from the members of
-the congregations, have received less attention in empirical and
-theoretical research for their potential to prevent crime. The role of
-these institutions might be particularly important in places where the
-State lacks the means or the will to enforce norms - as is the case in
-Brazil.
+Added to the scenario, formal social control institutions in countries
+like Brazil leave something (or much) to be desired: while in practice
+they should inhibit criminal behavior, they suffer from deep
+deficiencies that end up reducing the deterrent power of the justice
+system. Serious institutional problems include the inefficiency of the
+police, the lack of national legislation, the glacial pace of judicial
+processes, and the weak situation of the prison system in the country
+(Menezes et al. 2013). This makes it even more urgent to understand the
+role of \emph{informal social controls}, i.e., the ability of community
+organizations such as schools, clubs, and neighborbood associations to
+suppress crime by strengthening the capabilities of neighbors to control
+inappropriate behavior (Groff 2015).
+
+In order to understand the factors that can potentially help to deter
+crime, beyond formal control, it is important to identify empirical
+regularities. Criminological factors include concrete elements (such as
+te presence of arms or drugs, or elements of the built environment), and
+also figurative factors, which include the soci costs of deviant
+behavior as well as family supervision. Accordingly, a number of studies
+have investigated various aspects of the environment and neighborhood
+design (e.g., Foster, Giles-Corti, and Knuiman 2010; He, Páez, and Liu
+2017; Loukaitou-Sideris et al. 2001), whereas other studies have focused
+on exposure to environmental attributes that signal weakened norms, such
+as liquor and tobacco outlets (e.g., Brower and Carroll 2007; Deryol et
+al. 2016; Lipton et al. 2008; Quick, Law, and Luan 2017).
+
+Yet another fruitful avenue for research, and one that has only recently
+begun to be explored, is the presence of environmental attributes that
+can help to reinforce moral norms, such as schools and churches (e.g.,
+Abdullah et al. 2018; Davignon and Thomson 2015; Furr-Holden et al.
+2010; Traunmuller 2011). It is thus that in a recent paper, Warner and
+Konkel (2019) note that the role of places of worship, as distinct
+entities from the members of the congregations, have received less
+attention in empirical and theoretical research for their deterrence
+potential. The role of these institutions might be particularly
+important in places where formal state institutions lack the means or
+the will to enforce norms - as is the case in Brazil.
 
 With the above considerations in mind, the objective of this paper is to
-investigate whether moral communities have a spatial dimension
-discernible from the presence of places of worship. The case study is
-the city of Recife, in the state of Pernambuco in Brazil's Northeast.
-Recife is a large metropolitan area in a historically poor region, and
-afflicted by high levels of violent crime. The empirical strategy is to
-use spatial analysis to explore the potential geographical relationships
-between violent criminal events, on the one hand, and places of worship
-and a selection of commercial establishments, on the other.
-Dissagregated data allows us to analyze the phenomena of interest at a
-very high level of resolution as spatial point patterns.
+investigate whether and how places of workship correlate spatially with
+criminal events, or in other words, to investigate whether their signals
+as moral communities have a discernible spatial dimension. The case
+study is the city of Recife in the state of Pernambuco, in Brazil's
+Northeast. Recife is a large and important metropolitan area in a
+historically poor region, and afflicted by high levels of violent crime,
+having the dubious honor of being one of five state capitals with the
+highest rates of homicide in the period under study (Menezes et al.
+2013). The empirical strategy is to use spatial analysis to explore the
+potential geographical relationships between violent criminal events, on
+the one hand, and places of worship and a selection of commercial
+establishments that serve as controls, on the other. Dissagregated data
+allows us to analyze the relevant point patterns at a very high level of
+resolution.
 
 After this introductory section, the rest of this paper is structured as
-follows: Sect. \ref{background} reviews the literature on religion and
-crime; then Sect. \ref{methods} presents the empirical strategy used in
-the work. Sect. \ref{case} presents the context as well as the data used
-ub the study; Sect. \ref{results} systematizes and discusses the main
-results; and finally, the conclusions are presented in section
-\ref{conclusions}.
+follows: Section \ref{background} reviews some relevant theoretical
+perspectives on religion and crime; then Section \ref{methods} presents
+the empirical strategy used in the work. Section \ref{case} presents the
+context as well as the data used in the study; Section \ref{results}
+presents the analysis, whereas the results are discussed in Section
+\ref{discussion}. Finally, some concluding remarks are offered in
+Section \ref{conclusions}.
 
 \hypertarget{background}{%
-\section{Background: Theoretical Perspectives}\label{background}}
+\section{Background: Theoretical Perspectives on Religion and
+Crime}\label{background}}
 
-Words go here.
+Why do humans behave morally? For millennia, it has been the role of
+religion to provide the basic tenets of morally acceptable behavior:
+thou shalt not kill et al., and their equivalent in many world religions
+(e.g., Donovan 1986). Enforcement of such tenets implies different
+mechanisms, including \emph{sin}, \emph{haram}, \emph{karma}, and
+\emph{tapu}, with punishment delivered by hellfire and exile to
+\emph{Jahannam} or \emph{Gehenna}, to mention just some choice places of
+torment. In addition to acting as \emph{de facto} social policy for much
+of history, many of these religious tenets still have the force of law
+in many cultures and regions - and even where they do not, they are held
+by some researchers and policy experts to be helpful complements to
+reduce crime in any case (e.g., Durrant and Poppelwell 2017; Johnson
+2011). The hypothesis that religion can act as a factor that deters and
+reduces criminal behavior, therefore, has prompted the scientific study
+of the effectiveness of religion on moral behavior (Hoffmann 2015).
+
+An early effort to theorize the effect of religion on criminal behavior
+was the \emph{hellfire hypothesis} of Hirschi and Stark (1969). The
+focus of this hypothesis is the threat of extratemporal (and possibly
+eternal) punishment, and how this threat can deter believers from
+commandment-breaking (also see Pascal's wager). Accordingly, Hirsch and
+Stark (1969) posit that negative correlations between crime and religion
+are a consequence of a sense of commitment with normative values, a
+commitment that is ritually reinforced in a regular fashion (e.g., the
+liturgical rite of peace in Catholic Mass). Rohrbaugh and Jessor (1975),
+for instance, note that by attributing to divinity (or some other
+supra-terrenal entity) the supreme force of punishment, religion helps
+to build a sense of obduracy, or ``hardening'' against temptation.
+
+Despite its intuitive appeal, research has not provided much direct
+evidence for the hypothesis of hellfire (Hoffmann 2015, 1). In a bizarre
+twist, even mainstream doctrines (such as absolution) may in fact
+encourage beliefs that neutralize the fear of \emph{terrenal}
+punishment, and thus turn out to be criminogenic (e.g., Topalli,
+Brezina, and Bernhardt 2013). More deviant cases can even recast
+criminal behavior as a form of spiritual insurgency (Chesnut 2017).
+Counterexamples like this notwithstanding, researchers remain open to
+the possibility that ``religion may still serve as a social control
+mechanism by encouraging conventional beliefs, monitoring behaviors,
+enhancing family attachments, or providing conventional activities''
+(Hoffmann 2015, 1). In this way, the hypothesis of \emph{moral
+communities} (Stark 1996) recognizes that social integration is
+essential for increasing social control, thus reducing the practice of
+behavior that is not in accordance with current norms. Rohrbaugh and
+Jessor (1975) also emphasize that religion acts as social control since
+it defines what is an appropriate attitude according to moral values,
+thus making moral communities a close relative of the concept of
+\emph{informal social control}, defined as ``the ability of social
+groups or institutions to make norms or rules effective'' (Reiss 1951,
+196; cited in Groff 2015, 91).
+
+The hypothesis of moral communities has over the years been used to
+examine a variety of outcomes of interest. Recent examples include
+Stroope and Baker's (2018) exploration of religiosity and self-rated
+health; Davignon and Thomson (2015) with their research on institutional
+context and the religiosity of students; and the study of religion as a
+source of trust of Traunmuller (2011). This is in addition, of course,
+to numerous studies on criminal behavior such as Eitle (2011), an author
+who explored the deterrence power of religion on gambling; Lee and
+Bartkowski's (2004) investigation of juvenile homicide in rural areas;
+and the research of Regnerus (2003) on adolescent delinquency. The
+hypothesis of moral communities as a form of informal social control, on
+the other hand, has been less studied from a geographical perspective,
+and it is only recently that has attracted the attention of researchers.
+Groff (2015), for instance, discusses informal social control as a
+phenomenon that can plaussibly operate at different geographical scales,
+from the level of the home and family, through the street block and
+neighborhood, possibly to other scales such as the county. In this way,
+Nie and Yang (2019) remark on a recent paper on the lack of research
+conducted to study how the religious context of a geographical area
+(e.g., a county) may influence (youth smoking) behavior (p.~2). This
+point is echoed by Warner and Konkel (2019), who moreover recommend the
+use of even smaller units of analysis (e.g., Census Block Groups), in an
+effort to reduce aggregation bias (see Hipp 2007), but more importantly
+because the processes defined by social disorganization theory are
+thought to occur at relatively small geographical scales.
+
+The preceding review makes clear the need for research at various
+geographical scales. From a theoretical perspective, the social
+mechanisms that underpin the hypothesis of moral communities and
+informal social control can happen at various geographical scales, some
+of which previous research has already addressed: the work of Nie and
+Yang (2019), for one, highlights the role of processes detectable at a
+relatively high level of aggregation, whereas Warner and Konkel (2019),
+looking specifically at the effect of places of worship, bring the study
+closer to the level of the neighborhood. In what follows, we propose to
+adopt a much more disaggregated approach by considering the intensity of
+criminal events with respect to places of worship. This requires
+micro-level data both of crimes and places of worship, and the use of
+appropriate spatial analytical tools for point patterns, as discussed
+next.
 
@@
 Ministry of Justice of 2006, these are Violent Lethal and Intentional
-Crimes (CLVI; which includes intentional homicide), theft followed by
+Crimes (VLIC; which includes intentional homicide), theft followed by
 death (robbery), and corporal injury followed by death. The data on CVLI
@@
 
+\hypertarget{discussion}{%
+\section{Discussion}\label{discussion}}
+
+An opposing viewpoint is offered by the theory of conservative
+Protestants (Ellison, Burr, and McCall 2003). This theory argues that
+religion is not always beneficial in order to guarantee social order,
+and consequently to result in lower crime rates. This theory suggested
+that, conservative Protestants are more tolerate with violent behavior
+when associated with defense of honor, family, property, or woman.
+However, high rates are not expected for all types of crimes in densely
+populated areas of conservative Protestants. But what justifies this
+behavior? Ellison et al.~(2003) argue that this tolerance stems from a
+religious view of the world. More clearly, punishment is seen as a
+legitimate form of retribution whose explanation can be attributed to
+the literal interpretation of the Bible. In the specific context of
+Brazil, it is important to consider that the interpretations and
+meanings attributed to urban violence during the last two decades in
+Brazil often had religious figures as protagonists (Birman and Machado
+2012).
+
 \hypertarget{conclusions}{%
@@
 
-Words go here.
+The theory of conservative Protestants, argues that conservative
+Protestants are more likely to tolerate violent behavior of a defensive
+or punitive character, especially when associated to defense of honor,
+family, property, or woman. Our objective in the study is to explore the
+question whether moral communities have a spatial dimension. To this end
+we deploy an array of spatial analytical tools to investigate potential
+patterns of clustering of violent crime and places of worship, using as
+well other kinds of urban facilities to act as controls. The case study
+is the city of Recife, in the northeast of Brazil, and analysis is based
+on geographically detailed datasets of criminal events as well as
+several types of urban facilities.
 
@@
 
+\leavevmode\hypertarget{ref-Birman2012violencia}{}%
+Birman, Patricia, and Carly Machado. 2012. ``A Violencia Dos Justos:
+Evangelicos, Midia E Periferias Da Metropole.'' \emph{Revista Brasileira
+de Ciências Sociais} 27 (80): 55--69.
+
 \leavevmode\hypertarget{ref-Brower2007spatial}{}%
@@
 \url{https://doi.org/10.3200/jach.55.5.267-276}.
+
+\leavevmode\hypertarget{ref-Chesnut2017devoted}{}%
+Chesnut, R Andrew. 2017. \emph{Devoted to Death: Santa Muerte, the
+Skeleton Saint}. Oxford University Press.
 
@@
 
+\leavevmode\hypertarget{ref-Donovan1986different}{}%
+Donovan, Peter. 1986. ``Do Different Religions Share Moral Common
+Ground?'' Journal Article. \emph{Religious Studies} 22 (3/4): 367--75.
+\url{http://www.jstor.org.libaccess.lib.mcmaster.ca/stable/20006295}.
+
 \leavevmode\hypertarget{ref-Doran2011putting}{}%
@@
 Information Systems}. Springer Science \& Business Media.
+
+\leavevmode\hypertarget{ref-Durrant2017religion}{}%
+Durrant, Russil, and Zoe Poppelwell. 2017. ``Why Religion Matters.'' In
+\emph{Religion, Crime and Punishment}, 1--17. Springer.
+
+\leavevmode\hypertarget{ref-Eitle2011religion}{}%
+Eitle, D. 2011. ``Religion and Gambling Among Young Adults in the United
+States: Moral Communities and the Deterrence Hypothesis.'' Journal
+Article. \emph{Journal for the Scientific Study of Religion} 50 (1):
+61--81. \url{https://doi.org/10.1111/j.1468-5906.2010.01552.x}.
+
+\leavevmode\hypertarget{ref-Ellison2003enduring}{}%
+Ellison, Christopher G, Jeffrey A Burr, and Patricia L McCall. 2003.
+``The Enduring Puzzle of Southern Homicide: Is Regional Religious
+Culture the Missing Piece?'' \emph{Homicide Studies} 7 (4): 326--52.
 
@@
 
+\leavevmode\hypertarget{ref-Groff2015informal}{}%
+Groff, E. R. 2015. ``Informal Social Control and Crime Events.'' Journal
+Article. \emph{Journal of Contemporary Criminal Justice} 31 (1):
+90--106. \url{https://doi.org/10.1177/1043986214552619}.
+
 \leavevmode\hypertarget{ref-He2017built}{}%
@@
 \url{https://doi.org/http://dx.doi.org/10.1016/j.compenvurbsys.2017.08.001}.
+
+\leavevmode\hypertarget{ref-Hipp2007block}{}%
+Hipp, J. R. 2007. ``Block, Tract, and Levels of Aggregation:
+Neighborhood Structure and Crime and Disorder as a Case in Point.''
+Journal Article. \emph{American Sociological Review} 72 (5): 659--80.
+\url{https://doi.org/10.1177/000312240707200501}.
+
+\leavevmode\hypertarget{ref-Hirschi1969hellfire}{}%
+Hirschi, Travis, and Rodney Stark. 1969. ``Hellfire and Delinquency.''
+\emph{Social Problems} 17 (2): 202--13.
+
+\leavevmode\hypertarget{ref-Hoffmann2015religion}{}%
+Hoffmann, John P. 2015. ``Religion and Crime.'' \emph{The Encyclopedia
+of Crime and Punishment}, 1--5.
+
+\leavevmode\hypertarget{ref-Johnson2011more}{}%
+Johnson, Byron. 2011. \emph{More God, Less Crime: Why Faith Matters and
+How It Could Matter More}. Templeton Foundation Press.
+
+\leavevmode\hypertarget{ref-Lee2004love}{}%
+Lee, M. R., and J. P. Bartkowski. 2004. ``Love Thy Neighbor? Moral
+Communities, Civic Engagement, and Juvenile Homicide in Rural Areas.''
+Journal Article. \emph{Social Forces} 82 (3): 1001--35.
+\url{https://doi.org/10.1353/sof.2004.0044}.
 
@@
 
+\leavevmode\hypertarget{ref-Nie2019smoking}{}%
+Nie, F. H., and X. Z. Y. Yang. 2019. ``Smoking in the Temple of the Holy
+Spirit? Geographic Location Matters.'' Journal Article. \emph{Health \&
+Place} 58. \url{https://doi.org/10.1016/j.healthplace.2019.05.017}.
+
 \leavevmode\hypertarget{ref-Quick2017influence}{}%
@@
 \url{https://doi.org/10.1007/s12061-016-9191-5}.
+
+\leavevmode\hypertarget{ref-Regnerus2003moral}{}%
+Regnerus, M. D. 2003. ``Moral Communities and Adolescent Delinquency:
+Religious Contexts and Community Social Control.'' Journal Article.
+\emph{Sociological Quarterly} 44 (4): 523--54.
+\href{\%3CGo\%20to\%20ISI\%3E://WOS:000188319800001}{\textless{}Go to ISI\textgreater{}://WOS:000188319800001}.
+
+\leavevmode\hypertarget{ref-Reiss1951delinquency}{}%
+Reiss, Albert J. 1951. ``Delinquency as the Failure of Personal and
+Social Controls.'' Journal Article. \emph{American Sociological Review}
+16 (2): 196--207. \url{https://doi.org/10.2307/2087693}.
+
+\leavevmode\hypertarget{ref-Rohrbaugh1975religiosity}{}%
+Rohrbaugh, John, and Richard Jessor. 1975. ``Religiosity in Youth: A
+Personal Control Against Deviant Behavior.'' \emph{Journal of
+Personality} 43 (1): 136--55.
+
+\leavevmode\hypertarget{ref-Stark1996religion}{}%
+Stark, R. 1996. ``Religion as Context: Hellfire and Delinquency One More
+Time.'' Journal Article. \emph{Sociology of Religion} 57 (2): 163--73.
+\url{https://doi.org/10.2307/3711948}.
+
+\leavevmode\hypertarget{ref-Stroope2018moral}{}%
+Stroope, S., and J. O. Baker. 2018. ``Whose Moral Community?
+Religiosity, Secularity, and Self-Rated Health Across Communal Religious
+Contexts.'' Journal Article. \emph{Journal of Health and Social
+Behavior} 59 (2): 185--99.
+\url{https://doi.org/10.1177/0022146518755698}.
+
+\leavevmode\hypertarget{ref-Topalli2013god}{}%
+Topalli, V., T. Brezina, and M. Bernhardt. 2013. ``With God on My Side:
+The Paradoxical Relationship Between Religious Belief and Criminality
+Among Hardcore Street Offenders.'' Journal Article. \emph{Theoretical
+Criminology} 17 (1): 49--69.
+\url{https://doi.org/10.1177/1362480612463114}.
 
@@
 \leavevmode\hypertarget{ref-Warner2019neighborhood}{}%
-Warner, B. D., and R. H. Konkel. n.d. ``Neighborhood Churches and Their
+Warner, B. D., and R. H. Konkel. 2019. ``Neighborhood Churches and Their
 Relationship to Neighborhood Processes Important for Crime Prevention.''

### 2019-10-14 Moral_Communities_and_Crime_v1.Rmd
Completed first draft of methods and started working on the sections on context/data and analysis and results. Changed the name of the RData file to Moral_Communities_Crime_Recife.RData

@@
 
-load("Moral_Communities_and_Crime.RData")
+load("Moral_Communities_and_Crime_Recife.RData")
 ```
@@
 
-The preceding review makes clear the need for research at various geographical scales.
+The preceding review, as well as other research [-@Malleson2019identifying], makes clear the need for research at various geographical scales.
 From a theoretical perspective, the social mechanisms that underpin the hypothesis of moral communities and informal social control can happen at various geographical scales, some of which previous research has already addressed: the work of Nie and Yang [-@Nie2019smoking], for one, highlights the role of processes detectable at a relatively high level of aggregation, whereas Warner and Konkel [-@Warner2019neighborhood], looking specifically at the effect of places of worship, bring the study closer to the level of the neighborhood.
@@
 
-Two different techniques are deployed in this paper to study the spatial distribution of criminal events, and their possible relationship to the location of places of worship.
-Firstly, the F-function is a nearest neighbor technique that is used to summarize the distance between a set of points and their nearest events.
-This technique has found extensive use in spatial epidemiology (see Gatrell et al., 1996) and has also been used in spatial criminology (see, inter alia, Craglia et al., 2000; Kiani et al., 2015; Levine, 2006; and Rogerson and Sun, 2001).
-The second technique is the scan statistic.
-This method has also been extensively used in spatial epidemiology (e.g., Hjalmars et al., 1996) and has found use in spatial criminology as well (see He et al., 2017a; Nakaya and Yano, 2010; Malleson and Andresen, 2015; and Shiode, 2011).
-Use of these two techniques is meant to provide insights with respect to 1) the location of violent criminal events with respect to various types of facilities; and 2) geographically identify places of worship associated with clusters of criminal events.
-These two techniques are briefly reviewed next.
+As noted above, previous research that has investigated the presence of churches from a geographical perspective has used data aggregated at different scales.
+In the case of Nie and Yang [-@Nie2019smoking], the unit of analysis was the county, whereas Warner and Konkel [-@Warner2019neighborhood] used Census Block Groups, a much smaller unit of analysis.
+Research in spatial criminology, on the other hand, has studied criminal events as point patterns since at least the work of Levine et al.
+[-@Levine1986crime] investigated the concentration of criminal events in the proximity of bus stops.
+Since then, many other works have applied tools of spatial point pattern analysis to investigate the empirical properties of the distribution of crime.
+This includes Craglia et al.
+[-@Craglia2000comparative] who used the Ripley's K-function [@Ripley1976second] to investigate clustering processes, and Rogerson and Sun [-@Rogerson2001spatial] who applied nearest neighbor techniques to the study of arson.
+More recent studies of crime as point patterns are Nakaya and Yano [-@Nakaya2010visualizing], Kiani et al.
+[-@Kiani2015analysis], and Malleson and Andresen [-@Malleson2015spatio].
 
-## Quadrat Analysis
+Readers interested in point pattern analysis techniques are urged to consult the still valuable Bailey and Gatrell [-@Bailey1995interactive] or for an up-to-date and in-depth coverage of the topic, the excellent text by Baddeley et al.
+[-@Baddeley2015spatial].
+In this paper we will concentrate on one property of point processes, namely the intensity.
+Suppose that the outcome of interest is the number of events per unit area $Y(\mathcal{A})$.
+This could be, for instance, the number of criminal events observed in an arbitrary area.
+The intensity of a point pattern then is as follows:
+$$
+\lambda(s)=\lim_{\text{d}s \to 0}\frac{E[Y(\text{d}s)]}{\text{d}s}
+$$
 
-Words go here.
+where $\lambda(s)$ is the intensity of the process at point $s$, given by the expected number of events in a small area $\text{d}s$ around $s$, as the area becomes arbitrarily small.
+If the process is homogeneous (i.e., the intensity is a constant over space), an apt estimator of the intensity is the global intensity, simply the number of events divided by the area of the region under analysis:
+$$
+\hat{\lambda} = \frac{n}{|\mathcal{A}|}
+$$
 
-## Relative Intensity
+When the process is not homogeneous, other estimators are more appropriate.
+In the analysis to follow we use two estimators of intensity: conditional quadrat counts and relative distribution estimate.
+These techniques are briefly discussed next.
 
-Words go here.
+## Conditional Quadrat Counts
 
-# Case Study {#case}
+Quadrat counts is a relatively simple technique to analyze spatial variations in the intensity of a point process.
+It operates by partitioning the region under study $\mathcal{A}$ into subregions $\mathcal{A}_1,\cdots,\mathcal{A}_m$.
+These subregions are mutually exclusive, and their union is identical to $\mathcal{A}$.
+In the simplest case, the subregions have equal area.
+The number of events within each subregion (i.e., $n_{\mathcal{A}}$) divided by the area of the subregion (i.e., $|\mathcal{A}|$) is a local estimate of the intensity:
+$$
+\lambda(\mathcal{A}_i)=\frac{n_{\mathcal{A}_i}}{|\mathcal{A}_i|}
+$$
 
-## Context {#context}
+A test for homogeneity consists of assessing whether the intensity of the point process at the quadrats is uniform:
+$$
+\lambda(\mathcal{A}_1)=\cdots=\lambda(\mathcal{A}_m)
+$$
+
+Using quadrats of equal size is a convenient simplication, but in principle the areas could be different - in which case the count of events would be proportional to the area of the quadrat under homogeneity.
+An interesting variation of this technique is conditional quadrat counts, whereby the partition of regions $\mathcal{A}_1,\cdots,\mathcal{A}_m$ is done to reflect an underlying covariate of interest, say $Z$.
+By introducing a covariate as a partitioning criterion, it becomes possible to calculate estimators of the intensity of quadrats at different levels of the value of the covariate.
+For exploratory purposes, we can plot average intensity by quadrat, and compare to the global intensity.
+
+## Relative Distribution Estimate
+
+Conditional quadrat counts allow us to explore whether the intensity of the process depends on a covariate $Z$. A different way of expressing this is as follows:
+$$
+\lambda(s)=\rho\big(Z(s)\big)
+$$
+
+In this case, the intensity is a function $\rho$ that maps how the intensity depends on covariate $Z$.
+Non-parametric estimation of $\rho$ uses the ratio of the density of covariate values at the locations of the point process, relative to the _spatial distribution function_ $G$, the density of covariate values at random locations [see @Baddeley2015spatial, p.
+179].
+
+The density of covariate values at the locations of the is obtained by means of a kernel density estimator, for example:
+$$
+\lambda_0(s) = \frac{1}{|\mathcal{A}|}\sum_{i = 1}^n\kappa(Z(s_i) - z)
+$$
+where $s_i$ are the locations of the points.
+
+On the other hand, $G$ which is the cumulative distribution function of $Z$ at random point $S$ uniformly distributed in $\mathcal{A}$  :
+$$
+G(z) = \frac{1}{|\mathcal{A}|} \int_{\mathcal{A}}\text{I}\{Z(s)\leq z\}\text{d}s
+$$
+where $\text{I}$ is an indicator function that takes the value of $1$ if the argument is true, and the value of $0$ otherwise.
+In practice, the spatial distribution function is approximated based on a discretization of space using a fine grid of pixels as follows:
+$$
+G(z) = \frac{\#\{\text{pixels }s:Z(s)\leq z\}}{\# \text{pixels}}
+$$
+
+Therefore, an estimator of $\rho$ is as follows:
+$$
+\rho(z) = \frac{1}{|\mathcal{A}|G'(z)}\sum_{i=1}^n\kappa(Z(s_i) - z)
+$$
+where the derivative $G'$ can be approximated by differentiating a smoothed estimated of $G$. Other estimators are discussed by Baddeley et al. [@Baddeley2015spatial, p. 180].
+
+It is possible to adjust the relative distribution estimate by means of a baseline; a baseline in this case can be a function of other covariates that might confound the estimates of the relative distribution, so that the relative intensity $\lambda(s)/B(s)$ can be assumed to depend only on covariate $Z$.
+Therefore:
+$$
+\lambda(s) = \rho(Z(s))B(s)
+$$
+where $B(s)$ is a baseline function.
+It can be seen that $\rho(z)=1$ is the baseline intensity; values of $\rho(z)>1$ correspond to intensities higher than the baseline _as a function of_ $Z(u)$, whereas values of $\rho(z)<1$ correspond to intensities lower than the baseline, again as a function of $Z(u)$.
+
+# Case Study: Context and Data {#case}
 
 The study is of the city of Recife, the capital of the state of Pernambuco in the Northeast region of Brazil.
-With a population of 1,550,390 million inhabitants in 2010, and with an area of 7,133.2 km2, Recife is one of the main Brazilian metropolises, exerting a great economic influence in neighboring regions.
+With a population of 1,550,390 million inhabitants in 2010, Recife is one of the main Brazilian metropolises, exerting a great economic influence in neighboring regions.
 However, the city experiences a serious problem with violent crime, and has the dubious honor of being one of the five capital cities in Brazil with the highest homicide rates in the period under study [@Menezes2013spatial].
@@
 Following recommendations of the National Secretariat of Public Security of the Ministry of Justice of 2006, these are Violent Lethal and Intentional Crimes (VLIC; which includes intentional homicide), theft followed by death (robbery), and corporal injury followed by death.
-The data on CVLI were extracted from the Police Information System of the Secretariat of Social Defense of Pernambuco (INFOPOL / SDS-PE), which is the most reliable, detailed, and comprehensive information on violent deaths in the region.
+The data on LIVC were extracted from the Police Information System of the Secretariat of Social Defense of Pernambuco (INFOPOL / SDS-PE), which is the most reliable, detailed, and comprehensive information on violent deaths in the region.
 
-```{r descriptive-statistics, echo=FALSE}
+```{r table-descriptive-statistics, echo=FALSE}
 descriptive_stats <- data.frame(Attribute = c("Man",
@@
 
-## Data {#data}
-
 The data are organized at the individual level and it is possible to obtain information about the location, day of the week, day of the month, period of the day, as well as gender, age, and race of the victim.
 The database used in this study comprises the period from July 1, 2008 to June 30, 2010.
-Some descriptive statistics regarding this dataset are reported in Table 1, where it can be seen that overall, about 91% of the victims of CVLI in the period analyzed were men, while approximately 98% of violent crimes were homicides.
+Some descriptive statistics regarding this dataset are reported in Table 1, where it can be seen that overall, about 91% of the victims of LIVC in the period analyzed were men, while approximately 98% of violent crimes were homicides.
 In addition, most of the victims were black or brown, and the youth population between the ages of 18 and 30 is the most affected by violent crime.
 Lastly, it should be noted that about 88% of the criminal events under analysis involved firearms.
-Information about places of interest was obtained from the National Register of Addresses for Statistical Purposes (CNEFE - Census 2010), which lists 78,056,411 urban and rural addresses, distributed among the 316,574 census tracts.
-This is the first database of its kind produced by IBGE, and the first version was produced at the time of the 2000 Census.
-The way addresses are described in the National Register is very rich, and it is possible to identify the names of the places of worship including their denomination.
-Georeferencing was used to geolocate each place of worship.
-In this way, a total of 1,719 places of worship were geolocated in the city of Recife.
-Figure 1 (Panel a) shows the spatial distribution of the 1,657 CVLI crimes that occurred in the city of Recife between 2008.2 and 2010.1.
-Panel b) of the figure shows the spatial distribution of places of worship in the region.
+Figure \ref{fig:plot-crime} shows the spatial distribution of the 1,657 LIVC crimes that occurred in the city of Recife between July, 2008 and June, 2010.
 
-```{r plot-crime, echo=FALSE, fig.align = "left", fig.cap="\\label{fig:plot-crime}Lethal and Intentional Violent Crime in Recife"}
+```{r fig-plot-crime, echo=FALSE, fig.align = "left", fig.cap="\\label{fig:plot-crime}Location of Lethal and Intentional Violent Crime in Recife, July, 2008 - June, 2009"}
 
@@
 
+Information about places of interest was obtained from the National Register of Addresses for Statistical Purposes (CNEFE - Census 2010), which lists 78,056,411 urban and rural addresses, distributed among the 316,574 census tracts.
+This is the first database of its kind produced by IBGE, and the first version was produced at the time of the 2000 Census.
+The way addresses are described in the National Register is very rich, and it is possible to identify the names of the places of worship including their denomination.
+Georeferencing was used to geolocate each place of worship.
+In this way, a total of 1,719 places of worship were geolocated in the city of Recife.
+
 In addition to places of worship, the National Register of Addresses for Statistical Purposes was queried to extract facilities other than places of worship.
@@
 
-```{r plot-events, echo=FALSE, fig.align = "left", fig.cap="\\label{fig:plot-events}Places of Worship and Commercial Establishments in Recife"}
+```{r fig-plot-events, echo=FALSE, fig.align = "left", fig.cap="\\label{fig:plot-events}Location of Places of Worship and Commercial Establishments in Recife"}
 
@@
 
-```{r plot-sed, echo=FALSE, fig.align = "left", fig.cap="\\label{fig:plot-sed}Socio-Economic Deprivation in Recife"}
+# Analysis and results {#results}
+
+The analysis is implemented using the `R` statistical computing language, and documents to replicate the analysis, in addition to all data necessary, can be obtained from this anonymous shared folder:
+
+https://drive.google.com/open?id=1tuJM4Mhi0Ftq3ZEwjv6RP9vO3veGjIR_
+
+## Socio-Economic Deprivation
+
+```{r sed-descriptive-statistics, echo=FALSE}
+
+# Create a dataframe with the descriptive statistics of the socio-economic and demographic variables used for the creation of the Socio-Economic Deprivation index.
+The descriptive statistics are the min, 2nd quartile, median, mean, 3rd quartile, max, and include as well the loadings of the first principal component for discussion
+
+sed_descriptives <- data.frame(variable = c("Median Income", 
+                    "Proportion Unemployed", 
+                    "Proportion Poverty", 
+                    "Proportion Single Mother", 
+                    "Proportion Young Single Mother"),
+           min = c(min(setores.sf$median_income),
+                   min(setores.sf$prop_unemployed),
+                   min(setores.sf$prop_poverty),
+                   min(setores.sf$prop_single_mother),
+                   min(setores.sf$prop_young_single_mother)),
+           second_q = c(quantile(setores.sf$median_income, c(0, 0.25, 0.5, 0.75, 1))[2],
+                        quantile(setores.sf$prop_unemployed, c(0, 0.25, 0.5, 0.75, 1))[2],
+                        quantile(setores.sf$prop_poverty, c(0, 0.25, 0.5, 0.75, 1))[2],
+                        quantile(setores.sf$prop_single_mother, c(0, 0.25, 0.5, 0.75, 1))[2],
+                        quantile(setores.sf$prop_young_single_mother, c(0, 0.25, 0.5, 0.75, 1))[2]),
+           median = c(median(setores.sf$median_income),
+                      median(setores.sf$prop_unemployed),
+                      median(setores.sf$prop_poverty),
+                      median(setores.sf$prop_single_mother),
+                      median(setores.sf$prop_young_single_mother)),
+           mean = c(mean(setores.sf$median_income),
+                    mean(setores.sf$prop_unemployed),
+                    mean(setores.sf$prop_poverty),
+                    mean(setores.sf$prop_single_mother),
+                    mean(setores.sf$prop_young_single_mother)),
+           third_q = c(quantile(setores.sf$median_income, c(0, 0.25, 0.5, 0.75, 1))[4],
+                       quantile(setores.sf$prop_unemployed, c(0, 0.25, 0.5, 0.75, 1))[4],
+                       quantile(setores.sf$prop_poverty, c(0, 0.25, 0.5, 0.75, 1))[4],
+                       quantile(setores.sf$prop_single_mother, c(0, 0.25, 0.5, 0.75, 1))[4],
+                       quantile(setores.sf$prop_young_single_mother, c(0, 0.25, 0.5, 0.75, 1))[4]),
+           max = c(max(setores.sf$median_income),
+                   max(setores.sf$prop_unemployed),
+                   max(setores.sf$prop_poverty),
+                   max(setores.sf$prop_single_mother),
+                   max(setores.sf$prop_young_single_mother)),
+           pca_loadings = as.numeric(setores.pca$rotation[c(1, 2, 5, 3, 4), 1]))
+```
+
+```{r fig-plot-sed, echo=FALSE, include = FALSE, fig.align = "left", fig.cap="\\label{fig:plot-sed}Socio-Economic Deprivation in Recife"}
 ggplot(data = setores.sf) + 
@@
 
-# Analysis and results {#results}
+The first step in our analysis is to obtain an indicator of Socio-Economic Deprivation (SED).
+Socio-economic deprivation is known to correlated positively with crime, as it often reflects relevant criminogenic factors such as poverty and family disruption [@He2015temporal].
+In the present case, we have for the city of Recife information about the variables shown in Table \ref{tab:sed-descriptive-statistics} at the level of _setores_, a small Brazilian census geography.
 
-## Quadrat Analysis
+As seen in the table, Recife is a city with large socio-economic and demographic disparities; for example, the _setor_ with the highest median income has a median income that is `r prettyNum(filter(sed_descriptives, variable == "Median Income")["max"]/filter(sed_descriptives, variable == "Median Income")["min"] * 100, digits = 2, big.mark = ",")`% higher than the median income in the _setor_ with the lowest median income.
+Whereas there are _setores_ where the proportion of population who are unemployed is zero, there are _setores_ where almost 60% of the population are unemployed.
+Likewise, there are _setores_ where almost 70% of the population lives in poverty.
+In addition to these economic indicators, two variables are used to represent family disruption, the proportion of families whose head is a single mother, and the proportion of families whose head is a _young_ single mother, that is, a woman between the ages of 15 and 25.
+As can be seen, there are _setores_ with approximately 70% of households led by single mothers, and of these, over 21% are led by younger women, indicating a high degree of family disruption.
+
+For the analysis, we use Principal Component Analysis, a data reduction technique, to obtain an indicator of Socio-Economic Deprivation.
+The loadings of the first factor, which accounts for almost 54% of the variance, are shown in Table \ref{tab:sed-descriptive-statistics}.
+The loadings of this factor indicate that high Socio-Economic Deprivation is a combination of (in terms of importance): high levels of poverty, high levels of unemployment, and low median income, followed by high levels of family disruption, in particular proportion of families led by young single mothers.
+
+```{r table-sed-descriptive-statistics, echo=FALSE}
+
+# Render table with descriptives of socio-economic and demographic variables
+
+kable(sed_descriptives,
+      "latex",
+      booktabs = TRUE,
+      digits = 2,
+      escape = FALSE,
+      caption = "\\label{tab:sed-descriptive-statistics}Descriptive statistics of some key socio-economic and demographic variables in Recife",
+      col.names = linebreak(c("Variable", 
+                              "Min", 
+                              "2nd Quartile", 
+                              "Median", 
+                              "Mean", 
+                              "3rd Quartile", 
+                              "Max", "PC Factor 1\n Loadings")),
+      row.names = FALSE) %>%
+  kable_styling(bootstrap_options = c("striped", "scaled_down")) %>%
+  footnote(general = c("The variable for young single mothers is for women aged 15-25",
+                       "The first principal component accounts for 53.91% of the variance"))
+
+```
+
+Figure \ref{fig:plot-sed-as-quintiles} displays the geography of Socio-Economic Deprivation in Recife, after classifying _setores_ by quintiles, whereby the "High" class corresponds to _setores_ in the top 20% of the Socio-Economic Deprivation indicator, and the "Low" class corresponds to _setores_ in the bottom 20% of the Socio-Economic Deprivation indicator.
+The figure shows a veritable mosaic of affluence and deprivation, with high deprivation areas directly in contact, and in some cases even completely surrounded, by low deprivation areas.
+This geographical pattern of unequality, on the other hand, seems to be characteristic of Brazilian metropolitan regions, where enclaves of wealth and _favelas_ (i.e., urban slums) can be found in close proximity [see for example @Feitosa2007global].
 
@@
 
-# Rasterize and convert to image:
+# To use the SED indicator for conditional quadrat counts and relative distribution, we need to rasterize and convert this variable to an image:
 
@@
 
-```{r plot-sed-as-quintiles, echo=FALSE, cache = TRUE, fig.cap="\\label{fig:plot-sed-as-quintiles}Socio-Economic Deprivation as Quintiles" }
+```{r plot-sed-as-quintiles, echo=FALSE, cache = TRUE, fig.cap="\\label{fig:plot-sed-as-quintiles}Socio-Economic Deprivation in Recife classified by quintiles" }
 ggplot(data = sed5.sf) + 
@@
 ```
+
+## Conditional Quadrat Analysis
 
@@
 
-## Relative Intensity
+## Relative Distribution
 
@@
 ```
+
+## Relative Distribution with a Baseline Function
 

### 2019-10-14 Moral_Communities_and_Crime_v1.tex
Completed first draft of methods and started working on the sections on context/data and analysis and results. Changed the name of the RData file to Moral_Communities_Crime_Recife.RData

@@
 
-The preceding review makes clear the need for research at various
-geographical scales. From a theoretical perspective, the social
-mechanisms that underpin the hypothesis of moral communities and
-informal social control can happen at various geographical scales, some
-of which previous research has already addressed: the work of Nie and
-Yang (2019), for one, highlights the role of processes detectable at a
-relatively high level of aggregation, whereas Warner and Konkel (2019),
-looking specifically at the effect of places of worship, bring the study
-closer to the level of the neighborhood. In what follows, we propose to
-adopt a much more disaggregated approach by considering the intensity of
-criminal events with respect to places of worship. This requires
-micro-level data both of crimes and places of worship, and the use of
-appropriate spatial analytical tools for point patterns, as discussed
-next.
+The preceding review, as well as other research (2019), makes clear the
+need for research at various geographical scales. From a theoretical
+perspective, the social mechanisms that underpin the hypothesis of moral
+communities and informal social control can happen at various
+geographical scales, some of which previous research has already
+addressed: the work of Nie and Yang (2019), for one, highlights the role
+of processes detectable at a relatively high level of aggregation,
+whereas Warner and Konkel (2019), looking specifically at the effect of
+places of worship, bring the study closer to the level of the
+neighborhood. In what follows, we propose to adopt a much more
+disaggregated approach by considering the intensity of criminal events
+with respect to places of worship. This requires micro-level data both
+of crimes and places of worship, and the use of appropriate spatial
+analytical tools for point patterns, as discussed next.
 
@@
 
-Two different techniques are deployed in this paper to study the spatial
-distribution of criminal events, and their possible relationship to the
-location of places of worship. Firstly, the F-function is a nearest
-neighbor technique that is used to summarize the distance between a set
-of points and their nearest events. This technique has found extensive
-use in spatial epidemiology (see Gatrell et al., 1996) and has also been
-used in spatial criminology (see, inter alia, Craglia et al., 2000;
-Kiani et al., 2015; Levine, 2006; and Rogerson and Sun, 2001). The
-second technique is the scan statistic. This method has also been
-extensively used in spatial epidemiology (e.g., Hjalmars et al., 1996)
-and has found use in spatial criminology as well (see He et al., 2017a;
-Nakaya and Yano, 2010; Malleson and Andresen, 2015; and Shiode, 2011).
-Use of these two techniques is meant to provide insights with respect to
-1) the location of violent criminal events with respect to various types
-of facilities; and 2) geographically identify places of worship
-associated with clusters of criminal events. These two techniques are
-briefly reviewed next.
+As noted above, previous research that has investigated the presence of
+churches from a geographical perspective has used data aggregated at
+different scales. In the case of Nie and Yang (2019), the unit of
+analysis was the county, whereas Warner and Konkel (2019) used Census
+Block Groups, a much smaller unit of analysis. Research in spatial
+criminology, on the other hand, has studied criminal events as point
+patterns since at least the work of Levine et al.~(1986) investigated
+the concentration of criminal events in the proximity of bus stops.
+Since then, many other works have applied tools of spatial point pattern
+analysis to investigate the empirical properties of the distribution of
+crime. This includes Craglia et al.~(2000) who used the Ripley's
+K-function (Ripley 1976) to investigate clustering processes, and
+Rogerson and Sun (2001) who applied nearest neighbor techniques to the
+study of arson. More recent studies of crime as point patterns are
+Nakaya and Yano (2010), Kiani et al.~(2015), and Malleson and Andresen
+(2015).
 
-\hypertarget{quadrat-analysis}{%
-\subsection{Quadrat Analysis}\label{quadrat-analysis}}
+Readers interested in point pattern analysis techniques are urged to
+consult the still valuable Bailey and Gatrell (1995) or for an
+up-to-date and in-depth coverage of the topic, the excellent text by
+Baddeley et al.~(2015). In this paper we will concentrate on one
+property of point processes, namely the intensity. Suppose that the
+outcome of interest is the number of events per unit area
+\(Y(\mathcal{A})\). This could be, for instance, the number of criminal
+events observed in an arbitrary area. The intensity of a point pattern
+then is as follows: \[
+\lambda(s)=\lim_{\text{d}s \to 0}\frac{E[Y(\text{d}s)]}{\text{d}s}
+\]
 
-Words go here.
+where \(\lambda(s)\) is the intensity of the process at point \(s\),
+given by the expected number of events in a small area \(\text{d}s\)
+around \(s\), as the area becomes arbitrarily small. If the process is
+homogeneous (i.e., the intensity is a constant over space), an apt
+estimator of the intensity is the global intensity, simply the number of
+events divided by the area of the region under analysis: \[
+\hat{\lambda} = \frac{n}{|\mathcal{A}|}
+\]
 
-\hypertarget{relative-intensity}{%
-\subsection{Relative Intensity}\label{relative-intensity}}
+When the process is not homogeneous, other estimators are more
+appropriate. In the analysis to follow we use two estimators of
+intensity: conditional quadrat counts and relative distribution
+estimate. These techniques are briefly discussed next.
 
-Words go here.
+\hypertarget{conditional-quadrat-counts}{%
+\subsection{Conditional Quadrat
+Counts}\label{conditional-quadrat-counts}}
+
+Quadrat counts is a relatively simple technique to analyze spatial
+variations in the intensity of a point process. It operates by
+partitioning the region under study \(\mathcal{A}\) into subregions
+\(\mathcal{A}_1,\cdots,\mathcal{A}_m\). These subregions are mutually
+exclusive, and their union is identical to \(\mathcal{A}\). In the
+simplest case, the subregions have equal area. The number of events
+within each subregion (i.e., \(n_{\mathcal{A}}\)) divided by the area of
+the subregion (i.e., \(|\mathcal{A}|\)) is a local estimate of the
+intensity: \[
+\lambda(\mathcal{A}_i)=\frac{n_{\mathcal{A}_i}}{|\mathcal{A}_i|}
+\]
+
+A test for homogeneity consists of assessing whether the intensity of
+the point process at the quadrats is uniform: \[
+\lambda(\mathcal{A}_1)=\cdots=\lambda(\mathcal{A}_m)
+\]
+
+Using quadrats of equal size is a convenient simplication, but in
+principle the areas could be different - in which case the count of
+events would be proportional to the area of the quadrat under
+homogeneity. An interesting variation of this technique is conditional
+quadrat counts, whereby the partition of regions
+\(\mathcal{A}_1,\cdots,\mathcal{A}_m\) is done to reflect an underlying
+covariate of interest, say \(Z\). By introducing a covariate as a
+partitioning criterion, it becomes possible to calculate estimators of
+the intensity of quadrats at different levels of the value of the
+covariate. For exploratory purposes, we can plot average intensity by
+quadrat, and compare to the global intensity.
+
+\hypertarget{relative-distribution-estimate}{%
+\subsection{Relative Distribution
+Estimate}\label{relative-distribution-estimate}}
+
+Conditional quadrat counts allow us to explore whether the intensity of
+the process depends on a covariate \(Z\). A different way of expressing
+this is as follows: \[
+\lambda(s)=\rho\big(Z(s)\big)
+\]
+
+In this case, the intensity is a function \(\rho\) that maps how the
+intensity depends on covariate \(Z\). Non-parametric estimation of
+\(\rho\) uses the ratio of the density of covariate values at the
+locations of the point process, relative to the \emph{spatial
+distribution function} \(G\), the density of covariate values at random
+locations (see Baddeley, Rubak, and Turner 2015, 179).
+
+The density of covariate values at the locations of the is obtained by
+means of a kernel density estimator, for example: \[
+\lambda_0(s) = \frac{1}{|\mathcal{A}|}\sum_{i = 1}^n\kappa(Z(s_i) - z)
+\] where \(s_i\) are the locations of the points.
+
+On the other hand, \(G\) which is the cumulative distribution function
+of \(Z\) at random point \(S\) uniformly distributed in \(\mathcal{A}\)
+: \[
+G(z) = \frac{1}{|\mathcal{A}|} \int_{\mathcal{A}}\text{I}\{Z(s)\leq z\}\text{d}s
+\] where \(\text{I}\) is an indicator function that takes the value of
+\(1\) if the argument is true, and the value of \(0\) otherwise. In
+practice, the spatial distribution function is approximated based on a
+discretization of space using a fine grid of pixels as follows: \[
+G(z) = \frac{\#\{\text{pixels }s:Z(s)\leq z\}}{\# \text{pixels}}
+\]
+
+Therefore, an estimator of \(\rho\) is as follows: \[
+\rho(z) = \frac{1}{|\mathcal{A}|G'(z)}\sum_{i=1}^n\kappa(Z(s_i) - z)
+\] where the derivative \(G'\) can be approximated by differentiating a
+smoothed estimated of \(G\). Other estimators are discussed by Baddeley
+et al.~(Baddeley, Rubak, and Turner 2015, 180).
+
+It is possible to adjust the relative distribution estimate by means of
+a baseline; a baseline in this case can be a function of other
+covariates that might confound the estimates of the relative
+distribution, so that the relative intensity \(\lambda(s)/B(s)\) can be
+assumed to depend only on covariate \(Z\). Therefore: \[
+\lambda(s) = \rho(Z(s))B(s)
+\] where \(B(s)\) is a baseline function. It can be seen that
+\(\rho(z)=1\) is the baseline intensity; values of \(\rho(z)>1\)
+correspond to intensities higher than the baseline \emph{as a function
+of} \(Z(u)\), whereas values of \(\rho(z)<1\) correspond to intensities
+lower than the baseline, again as a function of \(Z(u)\).
 
 \hypertarget{case}{%
-\section{Case Study}\label{case}}
-
-\hypertarget{context}{%
-\subsection{Context}\label{context}}
+\section{Case Study: Context and Data}\label{case}}
 
@@
 Pernambuco in the Northeast region of Brazil. With a population of
-1,550,390 million inhabitants in 2010, and with an area of 7,133.2 km2,
-Recife is one of the main Brazilian metropolises, exerting a great
-economic influence in neighboring regions. However, the city experiences
-a serious problem with violent crime, and has the dubious honor of being
-one of the five capital cities in Brazil with the highest homicide rates
-in the period under study (Menezes et al. 2013).
+1,550,390 million inhabitants in 2010, Recife is one of the main
+Brazilian metropolises, exerting a great economic influence in
+neighboring regions. However, the city experiences a serious problem
+with violent crime, and has the dubious honor of being one of the five
+capital cities in Brazil with the highest homicide rates in the period
+under study (Menezes et al. 2013).
 
@@
 Crimes (VLIC; which includes intentional homicide), theft followed by
-death (robbery), and corporal injury followed by death. The data on CVLI
+death (robbery), and corporal injury followed by death. The data on LIVC
 were extracted from the Police Information System of the Secretariat of
@@
 
-\caption{\label{tab:descriptive-statistics}\label{tab:descriptive-statistics}Characteristics of Violent Crime in Recife, July, 2008 - June, 2009}
+\caption{\label{tab:table-descriptive-statistics}\label{tab:descriptive-statistics}Characteristics of Violent Crime in Recife, July, 2008 - June, 2009}
 \centering
@@
 
-\hypertarget{data}{%
-\subsection{Data}\label{data}}
-
 The data are organized at the individual level and it is possible to
@@
 dataset are reported in Table 1, where it can be seen that overall,
-about 91\% of the victims of CVLI in the period analyzed were men, while
+about 91\% of the victims of LIVC in the period analyzed were men, while
 approximately 98\% of violent crimes were homicides. In addition, most
@@
 should be noted that about 88\% of the criminal events under analysis
-involved firearms. Information about places of interest was obtained
-from the National Register of Addresses for Statistical Purposes (CNEFE
-- Census 2010), which lists 78,056,411 urban and rural addresses,
-distributed among the 316,574 census tracts. This is the first database
-of its kind produced by IBGE, and the first version was produced at the
-time of the 2000 Census. The way addresses are described in the National
-Register is very rich, and it is possible to identify the names of the
-places of worship including their denomination. Georeferencing was used
-to geolocate each place of worship. In this way, a total of 1,719 places
-of worship were geolocated in the city of Recife. Figure 1 (Panel a)
-shows the spatial distribution of the 1,657 CVLI crimes that occurred in
-the city of Recife between 2008.2 and 2010.1. Panel b) of the figure
-shows the spatial distribution of places of worship in the region.
+involved firearms. Figure \ref{fig:plot-crime} shows the spatial
+distribution of the 1,657 LIVC crimes that occurred in the city of
+Recife between July, 2008 and June, 2010.
 
@@
 
-\includegraphics{Moral_Communities_and_Crime_v1_files/figure-latex/plot-crime-1} \hfill{}
+\includegraphics{Moral_Communities_and_Crime_v1_files/figure-latex/fig-plot-crime-1} \hfill{}
 
-\caption{\label{fig:plot-crime}Lethal and Intentional Violent Crime in Recife}\label{fig:plot-crime}
+\caption{\label{fig:plot-crime}Location of Lethal and Intentional Violent Crime in Recife, July, 2008 - June, 2009}\label{fig:fig-plot-crime}
 \end{figure}
+
+Information about places of interest was obtained from the National
+Register of Addresses for Statistical Purposes (CNEFE - Census 2010),
+which lists 78,056,411 urban and rural addresses, distributed among the
+316,574 census tracts. This is the first database of its kind produced
+by IBGE, and the first version was produced at the time of the 2000
+Census. The way addresses are described in the National Register is very
+rich, and it is possible to identify the names of the places of worship
+including their denomination. Georeferencing was used to geolocate each
+place of worship. In this way, a total of 1,719 places of worship were
+geolocated in the city of Recife.
 
@@
 
-\includegraphics{Moral_Communities_and_Crime_v1_files/figure-latex/plot-events-1} \hfill{}
+\includegraphics{Moral_Communities_and_Crime_v1_files/figure-latex/fig-plot-events-1} \hfill{}
 
-\caption{\label{fig:plot-events}Places of Worship and Commercial Establishments in Recife}\label{fig:plot-events}
-\end{figure}
-
-\begin{figure}
-
-\includegraphics{Moral_Communities_and_Crime_v1_files/figure-latex/plot-sed-1} \hfill{}
-
-\caption{\label{fig:plot-sed}Socio-Economic Deprivation in Recife}\label{fig:plot-sed}
+\caption{\label{fig:plot-events}Location of Places of Worship and Commercial Establishments in Recife}\label{fig:fig-plot-events}
 \end{figure}
@@
 
-\hypertarget{quadrat-analysis-1}{%
-\subsection{Quadrat Analysis}\label{quadrat-analysis-1}}
+The analysis is implemented using the \texttt{R} statistical computing
+language, and documents to replicate the analysis, in addition to all
+data necessary, can be obtained from this anonymous shared folder:
+
+\url{https://drive.google.com/open?id=1tuJM4Mhi0Ftq3ZEwjv6RP9vO3veGjIR_}
+
+\hypertarget{socio-economic-deprivation}{%
+\subsection{Socio-Economic
+Deprivation}\label{socio-economic-deprivation}}
+
+The first step in our analysis is to obtain an indicator of
+Socio-Economic Deprivation (SED). Socio-economic deprivation is known to
+correlated positively with crime, as it often reflects relevant
+criminogenic factors such as poverty and family disruption (He et al.
+2015). In the present case, we have for the city of Recife information
+about the variables shown in Table \ref{tab:sed-descriptive-statistics}
+at the level of \emph{setores}, a small Brazilian census geography.
+
+As seen in the table, Recife is a city with large socio-economic and
+demographic disparities; for example, the \emph{setor} with the highest
+median income has a median income that is 6,517\% higher than the median
+income in the \emph{setor} with the lowest median income. Whereas there
+are \emph{setores} where the proportion of population who are unemployed
+is zero, there are \emph{setores} where almost 60\% of the population
+are unemployed. Likewise, there are \emph{setores} where almost 70\% of
+the population lives in poverty. In addition to these economic
+indicators, two variables are used to represent family disruption, the
+proportion of families whose head is a single mother, and the proportion
+of families whose head is a \emph{young} single mother, that is, a woman
+between the ages of 15 and 25. As can be seen, there are \emph{setores}
+with approximately 70\% of households led by single mothers, and of
+these, over 21\% are led by younger women, indicating a high degree of
+family disruption.
+
+For the analysis, we use Principal Component Analysis, a data reduction
+technique, to obtain an indicator of Socio-Economic Deprivation. The
+loadings of the first factor, which accounts for almost 54\% of the
+variance, are shown in Table \ref{tab:sed-descriptive-statistics}. The
+loadings of this factor indicate that high Socio-Economic Deprivation is
+a combination of (in terms of importance): high levels of poverty, high
+levels of unemployment, and low median income, followed by high levels
+of family disruption, in particular proportion of families led by young
+single mothers.
+
+\begin{table}
+
+\caption{\label{tab:table-sed-descriptive-statistics}\label{tab:sed-descriptive-statistics}Descriptive statistics of some key socio-economic and demographic variables in Recife}
+\centering
+\begin{tabular}[t]{lrrrrrrr}
+\toprule
+Variable & Min & 2nd Quartile & Median & Mean & 3rd Quartile & Max & \makecell[l]{PC Factor 1\\ Loadings}\\
+\midrule
+Median Income & 122.05 & 337.40 & 581.67 & 1031.09 & 1283.79 & 7954.19 & -0.42\\
+Proportion Unemployed & 0.00 & 0.05 & 0.10 & 0.10 & 0.14 & 0.59 & 0.56\\
+Proportion Poverty & 0.00 & 0.05 & 0.11 & 0.12 & 0.17 & 0.63 & 0.58\\
+Proportion Single Mother & 0.00 & 0.33 & 0.37 & 0.37 & 0.41 & 0.70 & 0.17\\
+Proportion Young Single Mother & 0.00 & 0.01 & 0.02 & 0.02 & 0.02 & 0.21 & 0.38\\
+\bottomrule
+\multicolumn{8}{l}{\textit{Note: }}\\
+\multicolumn{8}{l}{The variable for young single mothers is for women aged 15-25}\\
+\multicolumn{8}{l}{The first principal component accounts for 53.91\% of the variance}\\
+\end{tabular}
+\end{table}
+
+Figure \ref{fig:plot-sed-as-quintiles} displays the geography of
+Socio-Economic Deprivation in Recife, after classifying \emph{setores}
+by quintiles, whereby the ``High'' class corresponds to \emph{setores}
+in the top 20\% of the Socio-Economic Deprivation indicator, and the
+``Low'' class corresponds to \emph{setores} in the bottom 20\% of the
+Socio-Economic Deprivation indicator. The figure shows a veritable
+mosaic of affluence and deprivation, with high deprivation areas
+directly in contact, and in some cases even completely surrounded, by
+low deprivation areas. This geographical pattern of unequality, on the
+other hand, seems to be characteristic of Brazilian metropolitan
+regions, where enclaves of wealth and \emph{favelas} (i.e., urban slums)
+can be found in close proximity (see for example Feitosa et al. 2007).
 
@@
 \includegraphics{Moral_Communities_and_Crime_v1_files/figure-latex/plot-sed-as-quintiles-1.pdf}
-\caption{\label{fig:plot-sed-as-quintiles}Socio-Economic Deprivation as
-Quintiles}
+\caption{\label{fig:plot-sed-as-quintiles}Socio-Economic Deprivation in
+Recife classified by quintiles}
 \end{figure}
+
+\hypertarget{conditional-quadrat-analysis}{%
+\subsection{Conditional Quadrat
+Analysis}\label{conditional-quadrat-analysis}}
 
@@
 
-\hypertarget{relative-intensity-1}{%
-\subsection{Relative Intensity}\label{relative-intensity-1}}
+\hypertarget{relative-distribution}{%
+\subsection{Relative Distribution}\label{relative-distribution}}
+
+\hypertarget{relative-distribution-with-a-baseline-function}{%
+\subsection{Relative Distribution with a Baseline
+Function}\label{relative-distribution-with-a-baseline-function}}
 
@@
 
+\leavevmode\hypertarget{ref-Baddeley2015spatial}{}%
+Baddeley, Adrian, Ege Rubak, and Rolf Turner. 2015. \emph{Spatial Point
+Patterns: Methodology and Applications with R}. Book. Chapman; Hall/CRC.
+
+\leavevmode\hypertarget{ref-Bailey1995interactive}{}%
+Bailey, T. C., and A. C. Gatrell. 1995. \emph{Interactive Spatial Data
+Analysis}. Book. Essex: Addison Wesley Longman.
+
 \leavevmode\hypertarget{ref-Becker2017analise}{}%
@@
 Skeleton Saint}. Oxford University Press.
+
+\leavevmode\hypertarget{ref-Craglia2000comparative}{}%
+Craglia, M., R. Haining, and P. Wiles. 2000. ``A Comparative Evaluation
+of Approaches to Urban Crime Pattern Analysis.'' Journal Article.
+\emph{Urban Studies} 37 (4): 711--29. \url{ISI:000086520800005}.
 
@@
 
+\leavevmode\hypertarget{ref-Feitosa2007global}{}%
+Feitosa, F. F., G. Camara, A. M. V. Monteiro, T. Koschitzki, and M. P.
+S. Silva. 2007. ``Global and Local Spatial Indices of Urban
+Segregation.'' Journal Article. \emph{International Journal of
+Geographical Information Science} 21 (3): 299--323.
+\href{ISI:000244944300004\%0AC:/Papers/International\%20Journal\%20of\%20Geographical\%20Information\%20Science/IJGIS\%20(2007)\%2021\%20(3)\%20299-323.pdf}{ISI:000244944300004
+C:/Papers/International Journal of Geographical Information Science/IJGIS (2007) 21 (3) 299-323.pdf}.
+
 \leavevmode\hypertarget{ref-Foster2010neighbourhood}{}%
@@
 
+\leavevmode\hypertarget{ref-He2015temporal}{}%
+He, L., A. Paez, D. S. Liu, and S. G. Jiang. 2015. ``Temporal Stability
+of Model Parameters in Crime Rate Analysis: An Empirical Examination.''
+Journal Article. \emph{Applied Geography} 58: 141--52.
+\url{https://doi.org/10.1016/j.apgeog.2015.02.002}.
+
 \leavevmode\hypertarget{ref-Hipp2007block}{}%
@@
 
+\leavevmode\hypertarget{ref-Kiani2015analysis}{}%
+Kiani, Rasoul, Siamak Mahdavi, and Amin Keshavarzi. 2015. ``Analysis and
+Prediction of Crimes by Clustering and Classification.''
+\emph{International Journal of Advanced Research in Artificial
+Intelligence} 4 (8): 11--17.
+
 \leavevmode\hypertarget{ref-Lee2004love}{}%
@@
 \url{https://doi.org/10.1353/sof.2004.0044}.
+
+\leavevmode\hypertarget{ref-Levine1986crime}{}%
+Levine, Ned, Martin Wachs, and Elham Shirazi. 1986. ``Crime at Bus
+Stops: A Study of Environmental Factors.'' \emph{Journal of
+Architectural and Planning Research}, 339--61.
 
@@
 
+\leavevmode\hypertarget{ref-Malleson2015spatio}{}%
+Malleson, Nick, and Martin A Andresen. 2015. ``Spatio-Temporal Crime
+Hotspots and the Ambient Population.'' \emph{Crime Science} 4 (1): 10.
+
+\leavevmode\hypertarget{ref-Malleson2019identifying}{}%
+Malleson, N., W. Steenbeek, and M. A. Andresen. 2019. ``Identifying the
+Appropriate Spatial Resolution for the Analysis of Crime Patterns.''
+Journal Article. \emph{Plos One} 14 (6).
+\url{https://doi.org/10.1371/journal.pone.0218324}.
+
 \leavevmode\hypertarget{ref-Menezes2013spatial}{}%
@@
 (5): 471--83. \url{https://doi.org/10.1016/j.avb.2013.07.003}.
+
+\leavevmode\hypertarget{ref-Nakaya2010visualizing}{}%
+Nakaya, T., and K. J. Yano. 2010. ``Visualising Crime Clusters in a
+Space-Time Cube: An Exploratory Data-Analysis Approach Using Space-Time
+Kernel Density Estimation and Scan Statistics.'' Journal Article.
+\emph{Transactions in Gis} 14 (3): 223--39.
+\url{https://doi.org/10.1111/j.1467-9671.2010.01194.x}.
 
@@
 16 (2): 196--207. \url{https://doi.org/10.2307/2087693}.
+
+\leavevmode\hypertarget{ref-Ripley1976second}{}%
+Ripley, B. D. 1976. ``The Second-Order Analysis of Stationary Point
+Processes.'' Journal Article. \emph{Journal of Applied Probability} 13
+(2): 255--66. \url{ISI:A1976CA37400007}.
+
+\leavevmode\hypertarget{ref-Rogerson2001spatial}{}%
+Rogerson, Peter, and Y Sun. 2001. ``Spatial Monitoring of Geographic
+Patterns: An Application to Crime Analysis.'' \emph{Computers,
+Environment and Urban Systems} 25 (6): 539--56.
 

### 2019-10-15 Moral_Communities_and_Crime_v1.Rmd
Completed first draft of analysis/results and conclusions.

@@
 It is thus that in a recent paper, Warner and Konkel [-@Warner2019neighborhood] note that the role of places of worship, as distinct entities from the members of the congregations, have received less attention in empirical and theoretical research for their deterrence potential.
-The role of these institutions might be particularly important in places where formal state institutions lack the means or the will to enforce norms - as is the case in Brazil.
+The role of these institutions might be particularly important in places where formal state institutions lack the means or the will to enforce norms in a consistent way - as is the case in Brazil [@Garmany2014space].
 
@@
 
-For the analysis, we use Principal Component Analysis, a data reduction technique, to obtain an indicator of Socio-Economic Deprivation.
-The loadings of the first factor, which accounts for almost 54% of the variance, are shown in Table \ref{tab:sed-descriptive-statistics}.
-The loadings of this factor indicate that high Socio-Economic Deprivation is a combination of (in terms of importance): high levels of poverty, high levels of unemployment, and low median income, followed by high levels of family disruption, in particular proportion of families led by young single mothers.
+For the analysis, we use Principal Component Analysis, a data reduction technique, to obtain an indicator of Socio-Economic Deprivation, essentially the first principal component in the output.
+The loadings of this factor (Table \ref{tab:sed-descriptive-statistics}) indicate that high Socio-Economic Deprivation is a combination of (in terms of importance): high levels of poverty, high levels of unemployment, and low median income, followed by high levels of family disruption, in particular proportion of families led by young single mothers.
+This factor accounts for almost 54% of the variance..
 
@@
       escape = FALSE,
+      align = c("l", "c", "c", "c", "c", "c", "c", "c"),
       caption = "\\label{tab:sed-descriptive-statistics}Descriptive statistics of some key socio-economic and demographic variables in Recife",
@@
                               "3rd Quartile", 
-                              "Max", "PC Factor 1\n Loadings")),
+                              "Max", 
+                              "PC Factor 1\n Loadings")),
       row.names = FALSE) %>%
@@
 Figure \ref{fig:plot-sed-as-quintiles} displays the geography of Socio-Economic Deprivation in Recife, after classifying _setores_ by quintiles, whereby the "High" class corresponds to _setores_ in the top 20% of the Socio-Economic Deprivation indicator, and the "Low" class corresponds to _setores_ in the bottom 20% of the Socio-Economic Deprivation indicator.
-The figure shows a veritable mosaic of affluence and deprivation, with high deprivation areas directly in contact, and in some cases even completely surrounded, by low deprivation areas.
+The figure shows a veritable mosaic of affluence and deprivation, with high deprivation areas directly in contact, and in some cases almost completely surrounded, by low deprivation areas.
 This geographical pattern of unequality, on the other hand, seems to be characteristic of Brazilian metropolitan regions, where enclaves of wealth and _favelas_ (i.e., urban slums) can be found in close proximity [see for example @Feitosa2007global].
@@
 
-Calculate the intensity of other ppp by quadrat:
+Once the indicator of Socio-Economic Deprivation has been obtained, as outlined above, it can be used for conditional quadrat analysis.
+As previously described, the quadrats are irregular areas that are defined based on a covariate, in this case the quintiles of the Socio-Economic Deprivation indicator.
+In essence, this entails counting the number of events at each level of the covariate, and then calculating an estimator of the intensity.
+
 ```{r quadrat-analysis, include=FALSE}
 
-# Calculate the intensity of the various point patterns by quadrat:
+# Calculate the intensity of the various point patterns by quadrat, conditional on Socio-Economic Deprivation:
 
@@
 
+Figure \ref{fig:plot-crime-quadrat} shows the results of conditional quadrat counts for Lethal and Intentional Violent Crime.
+The dotted line indicates the global intensity, simply the total number of criminal events divided by the area of the region.
+It can be seen that the intensity of crime clearly increases with Socio-Economic Deprivation, and the intensity in areas with the highest levels of Socio-Economic Deprivation is `r crime5_intensity[5]/crime5_intensity[1] * 100` % higher than in places with the lowest levels of Socio-Economic Deprivation.
+This is as expected, as research has consistently found correlations between Socio-Economic Deprivation and a variety of negative health outcomes, including homicide [@Ichihara2018area].
+A $\chi^2$ test of independence on the quadrat counts yields a value of $p\leq 0.0001$ (four degrees of freedom), which comfortably rejects the null hypothesis of homogeneity of intensity of crime by level of Socio-Economic Deprivation.
+
 ```{r plot-crime-quadrat, echo=FALSE, fig.width=3, fig.asp=1 , cache = TRUE, fig.cap="\\label{fig:plot-crime-quadrat}Intensity of crime by level of Socio-Economic Deprivation; the dotted line indicates the global intensity of crime"}
@@
 ```
+
+Next, we repeat the analysis of conditional quadrat counts, but now for the six types of events of reference.
+As discussed above, this included three types of places of worship which are of interest from the perspective of moral communities, and three types of commercial establishments that we posit are morally neutral.
+The results of estimating the intensity of the process by means of conditional quadrat counts are shown in Figure \ref{fig:plot-events-quadrat}.
+
+The first thing that we note are the variations in global intensity of places of worship, with Evangelical places of worship being the most intense of the three.
+There are no clear variations in the locational pattern of Catholic places of worship by level of Socio-Economic Deprivation.
+Evangelical places of worship, on the other hand, tend to be found more frequently in places with middle-low and middle levels of Socio-Economic Deprivation, a pattern also discernible, albeit barely, for Spiritist churches.
+Interestingly (see Table \ref{tab:quadrat-test-statistics}) $\chi^2$ tests of independence on the conditional quadrat counts fail to reject the hypothesis of homogeneity for Catholic and Spiritist churches, and it is only in the case of Evangelical places of worship that we detect the possibility of locational patterns that vary by level of Socio-Economic Deprivation.
+
+Three types of commercial establishments also appear to display inhomogeneous locational patterns (Table \ref{tab:quadrat-test-statistics}), with the $\chi^2$ test of independence emphaticaly rejecting the null hypothesis.
+As seen in Figure \ref{fig:plot-events-quadrat}, ice cream shops tend to be found more frequently in places with middle-low Socio-Economic Deprivation, and less frequently in places with middle-high Socio-Economic Deprivation.
+Pharmacies show a clearer trend, with locational patterns that tend to favor places with low Socio-Economic Deprivation.
+Finally, supermarkets are found less frequently in places with middle-high and high Socio-Economic Deprivation.
 
@@
 
+```{r table-quadrat-test-statistics, echo=FALSE}
+
+quad.tests <- data.frame(Event = c("Catholic",
+                                   "Evangelical",
+                                   "Spiritist",
+                                   "Ice Crem Shop",
+                                   "Pharmacy",
+                                   "Supermarket"),
+                         rbind(quadrat.test(catholic5), 
+                               quadrat.test(evangelic5), 
+                               quadrat.test(spiritist5),
+                               quadrat.test(ice_cream5),
+                               quadrat.test(pharmacies5),
+                               quadrat.test(super_markets5))) %>%
+  transmute(Event,
+            Statistic = as.numeric(statistic),
+            degrees_of_freedom = as.numeric(parameter),
+            p_value = as.numeric(p.value)) %>%
+  mutate(p_value = ifelse(p_value >= 0.001, paste0(round(p_value, 4)), "$<0.001$"))
+
+# Render table with descriptives of socio-economic and demographic variables
+
+kable(quad.tests,
+      "latex",
+      booktabs = TRUE,
+      digits = 3,
+      escape = FALSE,
+      align = c("l", "c", "c", "c"),
+      caption = "\\label{tab:quadrat-test-statistics}$\\chi^2$ tests of independence of conditional quadrat counts",
+      col.names = linebreak(c("Event", 
+                              "Statistic", 
+                              "Degrees of Freedom", 
+                              "p-value"))) %>%
+  kable_styling(bootstrap_options = c("striped", "scaled_down"))
+```
+
 ## Relative Distribution
+
+The preceding analysis provides a valuable backdrop.
+As seen there, there is a clear distribution of criminal events, increasing on par with the level of Socio-Economic deprivation.
+Of the three different classes of places of worship, only Evangelical churches display a locational pattern that is commensurate, with Socio-Economic Deprivation, albeit the pattern is distinct from criminal events, with Evangelical churches found less frequently in both high and low Socio-Economic Deprivation areas.
+
+The analysis does not answer the question, yet, of possible covariations between crime, on the one hand, and places of worship and commercial establishments, on the other.
+In this subsection we implement the relative distribution estimator of intensity of Lethal and Intentional Violent Crime.
+The covariate $Z(u)$ in this analysis is distance to one of a place of worship or a commercial establishment:
+$$
+\lambda_k(s)=\rho(Z_k(s))
+$$
+with:
+$$
+Z_k(s) = d_{sk}
+$$
+In the expression above, $d_{s_k}$ is distance from location $s$ to an event of class $k$, and $k=\{\text{Catholic, Evangelical, Spiritist, Ice Cream Shop, Pharmacy, Supermarket}\}$.
 
@@
 
+```{r relative-distribution, include=FALSE, cache=TRUE}
+
+# The estimates of intensity are calculated with `rhohat`:
+
+rho_catholic <- rhohat(events_recife_list$`Violent Crime`, z_catholic)
+rho_evangelic <- rhohat(events_recife_list$`Violent Crime`, z_evangelical)
+rho_spiritist <- rhohat(events_recife_list$`Violent Crime`, z_spiritist)
+rho_ice_cream <- rhohat(events_recife_list$`Violent Crime`, z_ice_cream)
+rho_pharmacies <- rhohat(events_recife_list$`Violent Crime`, z_pharmacies)
+rho_supermarkets <- rhohat(events_recife_list$`Violent Crime`, z_super_markets)
+```
+
+```{r prepare-relative-distribution-data, include=FALSE}
+
+# Rename the variables in the results to collect in a single dataframe:
+
+rho_catholic <- rho_catholic %>% as.data.frame() %>% dplyr::rename(z = z_catholic)
+rho_evangelic <- rho_evangelic %>% as.data.frame() %>% rename(z = z_evangelica)
+rho_spiritist <- rho_spiritist %>% as.data.frame() %>% dplyr::rename(z = z_spiritist)
+rho_ice_cream <- rho_ice_cream %>% as.data.frame() %>% rename(z = z_ice_cream)
+rho_pharmacies <- rho_pharmacies %>% as.data.frame() %>% rename(z = z_pharmacies)
+rho_supermarkets <- rho_supermarkets %>% as.data.frame() %>% rename(z = z_super_mark)
+
+# Collect into dataframe for convenience of plotting:
+
+rho_all <- rbind(data.frame(rho_catholic, Type = "Catholic", Class = "Place of Worship"), 
+                 data.frame(rho_evangelic, Type = "Evangelical", Class = "Place of Worship"),
+                 data.frame(rho_spiritist, Type = "Spiritist", Class = "Place of Worship"),
+                 data.frame(rho_ice_cream, Type = "Ice Cream Shops", Class = "Business"), 
+                 data.frame(rho_pharmacies, Type = "Pharmacies", Class = "Business"), 
+                 data.frame(rho_supermarkets, Type = "Supermarkets", Class = "Business")) 
+```
+
+```{r analysis-relative-distribution, include=FALSE}
+
+# Extract some useful information from the relative distribution functions
+
+# First, find the max intensity for Evangelical:
+
+max_int_evangelical <- max(filter(rho_all, Type == "Evangelical")$rho)
+
+# Next, the distance at which the intensity of crime with respect to distance to Evangelical is max:
+
+dist_max_int_evangelical <- filter(rho_all, Type == "Evangelical")[filter(rho_all, Type == "Evangelical")$rho == max_int_evangelical,"z"]
+
+# Next, find the approximate intensity at that distance for the other types of events. First, what is the nearest distance to the distance at which the intensity with respect to Evangelical is max?
+
+near_dist_catholic_i <- which.min(abs(filter(rho_all, Type == "Catholic")$z - dist_max_int_evangelical))
+
+# If the nearest distance is greater than the one for evangelical, pick the preceding row in the dataframe, if it is less than then the following row.
+Use the two rows to obtain a linear approximation of the intensity at the desired distance
+
+if(filter(rho_all, Type == "Catholic")[near_dist_catholic_i, "z"] > dist_max_int_evangelical){
+  interv <- filter(rho_all, Type == "Catholic")[(near_dist_catholic_i-1):near_dist_catholic_i,] %>%
+    dplyr::select(z, rho)
+  slope <- (interv$rho[2] - interv$rho[1]) / (interv$z[2] - interv$z[1])
+  catholic_int_at_evangelical_max_int <- interv$rho[1] + (dist_max_int_evangelical - interv$z[1]) * slope
+}else{
+  interv <- filter(rho_all, Type == "Catholic")[near_dist_catholic_i:(near_dist_catholic_i + 1),] %>%
+    dplyr::select(z, rho)
+  slope <- (interv$rho[2] - interv$rho[1]) / (interv$z[2] - interv$z[1])
+  catholic_int_at_evangelical_max_int <- interv$rho[1] + (dist_max_int_evangelical - interv$z[1]) * slope
+}
+
+# Repeat for Spiritist
+
+near_dist_spiritist_i <- which.min(abs(filter(rho_all, Type == "Spiritist")$z - dist_max_int_evangelical))
+
+# If the nearest distance is greater than the one for evangelical, pick the preceding row in the dataframe, if it is less than then the following row.
+Use the two rows to obtain a linear approximation of the intensity at the desired distance
+
+if(filter(rho_all, Type == "Spiritist")[near_dist_spiritist_i, "z"] > dist_max_int_evangelical){
+  interv <- filter(rho_all, Type == "Spiritist")[(near_dist_spiritist_i-1):near_dist_spiritist_i,] %>%
+    dplyr::select(z, rho)
+  slope <- (interv$rho[2] - interv$rho[1]) / (interv$z[2] - interv$z[1])
+  spiritist_int_at_evangelical_max_int <- interv$rho[1] + (dist_max_int_evangelical - interv$z[1]) * slope
+}else{
+  interv <- filter(rho_all, Type == "Spiritist")[near_dist_spiritist_i:(near_dist_spiritist_i + 1),] %>%
+    dplyr::select(z, rho)
+  slope <- (interv$rho[2] - interv$rho[1]) / (interv$z[2] - interv$z[1])
+  spiritist_int_at_evangelical_max_int <- interv$rho[1] + (dist_max_int_evangelical - interv$z[1]) * slope
+}
+
+# Repeat for ice cream shop
+
+near_dist_ice_cream_i <- which.min(abs(filter(rho_all, Type == "Ice Cream Shops")$z - dist_max_int_evangelical))
+
+# If the nearest distance is greater than the one for evangelical, pick the preceding row in the dataframe, if it is less than then the following row.
+Use the two rows to obtain a linear approximation of the intensity at the desired distance
+
+if(filter(rho_all, Type == "Ice Cream Shops")[near_dist_ice_cream_i, "z"] > dist_max_int_evangelical){
+  interv <- filter(rho_all, Type == "Ice Cream Shops")[(near_dist_ice_cream_i-1):near_dist_ice_cream_i,] %>%
+    dplyr::select(z, rho)
+  slope <- (interv$rho[2] - interv$rho[1]) / (interv$z[2] - interv$z[1])
+  ice_cream_int_at_evangelical_max_int <- interv$rho[1] + (dist_max_int_evangelical - interv$z[1]) * slope
+}else{
+  interv <- filter(rho_all, Type == "Ice Cream Shops")[near_dist_ice_cream_i:(near_dist_ice_cream_i + 1),] %>%
+    dplyr::select(z, rho)
+  slope <- (interv$rho[2] - interv$rho[1]) / (interv$z[2] - interv$z[1])
+  ice_cream_int_at_evangelical_max_int <- interv$rho[1] + (dist_max_int_evangelical - interv$z[1]) * slope
+}
+
+# Repeat for pharmacy
+
+near_dist_pharmacies_i <- which.min(abs(filter(rho_all, Type == "Pharmacies")$z - dist_max_int_evangelical))
+
+# If the nearest distance is greater than the one for evangelical, pick the preceding row in the dataframe, if it is less than then the following row.
+Use the two rows to obtain a linear approximation of the intensity at the desired distance
+
+if(filter(rho_all, Type == "Pharmacies")[near_dist_pharmacies_i, "z"] > dist_max_int_evangelical){
+  interv <- filter(rho_all, Type == "Pharmacies")[(near_dist_pharmacies_i - 1):near_dist_pharmacies_i,] %>%
+    dplyr::select(z, rho)
+  slope <- (interv$rho[2] - interv$rho[1]) / (interv$z[2] - interv$z[1])
+  pharmacies_int_at_evangelical_max_int <- interv$rho[1] + (dist_max_int_evangelical - interv$z[1]) * slope
+}else{
+  interv <- filter(rho_all, Type == "Pharmacies")[near_dist_pharmacies_i:(near_dist_pharmacies_i + 1),] %>%
+    dplyr::select(z, rho)
+  slope <- (interv$rho[2] - interv$rho[1]) / (interv$z[2] - interv$z[1])
+  pharmacies_int_at_evangelical_max_int <- interv$rho[1] + (dist_max_int_evangelical - interv$z[1]) * slope
+}
+
+# Repeat for supermarkets
+
+near_dist_supermarkets_i <- which.min(abs(filter(rho_all, Type == "Supermarkets")$z - dist_max_int_evangelical))
+
+# If the nearest distance is greater than the one for evangelical, pick the preceding row in the dataframe, if it is less than then the following row.
+Use the two rows to obtain a linear approximation of the intensity at the desired distance
+
+if(filter(rho_all, Type == "Supermarkets")[near_dist_supermarkets_i, "z"] > dist_max_int_evangelical){
+  interv <- filter(rho_all, Type == "Supermarkets")[(near_dist_supermarkets_i - 1):near_dist_supermarkets_i,] %>%
+    dplyr::select(z, rho)
+  slope <- (interv$rho[2] - interv$rho[1]) / (interv$z[2] - interv$z[1])
+  supermarkets_int_at_evangelical_max_int <- interv$rho[1] + (dist_max_int_evangelical - interv$z[1]) * slope
+}else{
+  interv <- filter(rho_all, Type == "Supermarkets")[near_dist_supermarkets_i:(near_dist_supermarkets_i + 1),] %>%
+    dplyr::select(z, rho)
+  slope <- (interv$rho[2] - interv$rho[1]) / (interv$z[2] - interv$z[1])
+  supermarkets_int_at_evangelical_max_int <- interv$rho[1] + (dist_max_int_evangelical - interv$z[1]) * slope
+}
+```
+
+The results of this analysis are shown in Figure \ref{fig:plot-relative-distribution}.
+The global estimator for the intensity of crime is shown in this figure as a dotted line.
+Three types of places of worship are shown in solid lines, and three types of commercial establishments are shown in dashed lines.
+Each relative distribution function is shown with its corresponding 95% confidence bands.
+
+Several interesting things emerge from inspection of this figure.
+First, we notice that with the exception of Spiritist churches (which has an intensity significantly lower than the global intensity at a distance of zero), the intensity of crime at for other churches and commercial establishments is close, but higher than, the global intensity of crime.
+In general, the intensity tends to increase at short distances; however, this effect is much more marked for Evangelical places of worship.
+The intensity of crime in the proximity of these places of worship grows rapidly within a short distance, reaching a peak of `r signif(max_int_evangelical, 5)` at a distance of `r round(dist_max_int_evangelical, 2)` m.
+At that distance, the intensity of crime with respect to Evangelical places of worship is `r round(max_int_evangelical/catholic_int_at_evangelical_max_int * 100, 2)`% higher than the intensity of Catholic places of worship; `r round(max_int_evangelical/spiritist_int_at_evangelical_max_int * 100, 2)`% higher than the intensity of crime with respect to Spiritist churches; `r round(max_int_evangelical/ice_cream_int_at_evangelical_max_int * 100, 2)`% higher than the intensity with respect to ice cream shops; `r round(max_int_evangelical/pharmacies_int_at_evangelical_max_int * 100, 2)`% higher than the intensity with respect to pharmacies; and `r round(max_int_evangelical/supermarkets_int_at_evangelical_max_int * 100, 2)`% higher than the intensity with respect to supermarkets.
+
+Intriguingly, after reaching its peak intensity, the intensity of crime with respect to Evangelical places of worship declines sharply, until at distances of approximately 200 m it is lower than for other places of worship and commercial establishments, and at distances of approximately 270 m it is lower than the global intensity of crime.
+This does not happen with other places of worship or commercial establishments: the intensity of crime with respect to these locations remains higher than the global intensity of crime for the interval of distances examined.
+Another interesting observation is that the intensity of crime with respect to Catholic places of worship is not substantially different when compared to the intensity of crime with respect to any of the commercial establishments considered.
+
+These results, while intriguing, beg the question of whether variations in Socio-Economic Deprivation may have a confounding effect on the relative distributions; for instance, if as seen in the case of conditional quadrat counts, there is some overlap in the locational patterns of, say, Evangelical places of worship and crime with respect to Socio-Economic Deprivation.
+Next section explores this question.
+
+```{r figure-plot-relative-distribution, echo=FALSE, cache = TRUE, fig.cap="\\label{fig:plot-relative-distribution}Intensity of crime as a function of distance to a selected place of worship or commercial establishment; bands are 95% confidence interval and the dotted line indicates the global intensity of crime"}
+# Plot all estimates. The dotted line is the intensity if the intensity was constant (this is a summary measure of the `crime.ppp` object). Select only distance less than 500 m:
+
+ggplot(data = filter(rho_all, z <= 500)) +
+  geom_ribbon(aes(x = z, ymin = lo, ymax = hi, fill = Type), alpha = 0.2) + 
+  geom_line(aes(x = z, y = rho, color = Type, linetype = Class), size = 1) +
+  geom_abline(intercept = 7.575416e-06, slope = 0, linetype = "dotted") +
+  #xlim(c(0, 500)) +
+  theme_tufte()
+```
+
 ## Relative Distribution with a Baseline Function
+
+In this section, the preceding analysis regarding the relative distribution of crime using distance to places of worship and commercial establishments is repeated, after introducing a baseline function.
+The interpretation of $\rho(Z(s))$ in this case is as follows: instead of the raw intensity at the value of $z$, it is the intensity _relative to the intensity according to the level of Socio-Economic Deprivation_.
+Therefore, $rho(z)=1$ matches the intensity according to Socio-Economic Deprivation; values of $rho(z)>1$ indicate a higher intensity than explained by Socio-Economic Deprivation; and values of $rho(z)<1$ correspond to lower intensities than explained by Socio-Economic Deprivation.
+
+```{r relative-distribution-with-baseline, include=FALSE}
+
+#The estimates of intensity are calculated with `rhohat`, but now with a baseline intensity function that depends on median income:
+
+rho_catholic_bl <- rhohat(events_recife_list$`Violent Crime`, z_catholic, baseline = sed.img)
+rho_evangelic_bl <- rhohat(events_recife_list$`Violent Crime`, z_evangelical, baseline = sed.img)
+rho_spiritist_bl <- rhohat(events_recife_list$`Violent Crime`, z_spiritist, baseline = sed.img)
+rho_ice_cream_bl <- rhohat(events_recife_list$`Violent Crime`, z_ice_cream, baseline = sed.img)
+rho_pharmacies_bl <- rhohat(events_recife_list$`Violent Crime`, z_pharmacies, baseline = sed.img)
+rho_supermarkets_bl <- rhohat(events_recife_list$`Violent Crime`, z_super_markets, baseline = sed.img)
+```
+
+```{r prepare-relative-distribution-with-baseline-data, include=FALSE}
+
+# Rename variables to combine dataframes:
+
+rho_catholic_bl <- rho_catholic_bl %>% 
+  as.data.frame() %>% 
+  dplyr::rename(z = z_catholic)
+rho_evangelic_bl <- rho_evangelic_bl %>% 
+  as.data.frame() %>% 
+  rename(z = z_evangelica)
+rho_spiritist_bl <- rho_spiritist_bl %>% 
+  as.data.frame() %>% 
+  dplyr::rename(z = z_spiritist)
+rho_ice_cream_bl <- rho_ice_cream_bl %>% 
+  as.data.frame() %>% 
+  rename(z = z_ice_cream)
+rho_pharmacies_bl <- rho_pharmacies_bl %>% 
+  as.data.frame() %>% 
+  rename(z = z_pharmacies)
+rho_supermarkets_bl <- rho_supermarkets_bl %>% 
+  as.data.frame() %>% 
+  rename(z = z_super_mark)
+
+# Collect into dataframe for convenience of plotting:
+
+rho_all_bl <- rbind(data.frame(rho_catholic_bl, Type = "Catholic", Class = "Place of Worship"), 
+                    data.frame(rho_evangelic_bl, Type = "Evangelical", Class = "Place of Worship"),
+                    data.frame(rho_spiritist_bl, Type = "Spiritist", Class = "Place of Worship"),  
+                    data.frame(rho_ice_cream_bl, Type = "Ice Cream Shops", Class = "Business"), 
+                    data.frame(rho_pharmacies_bl, Type = "Pharmacies", Class = "Business"), 
+                    data.frame(rho_supermarkets_bl, Type = "Supermarkets", Class = "Business")) 
+```
+
+```{r analysis-relative-distribution-with-baseline, include=FALSE}
+
+# Extract some useful information from the relative distribution functions
+
+# First, find the max intensity for each type of event:
+
+max_int_evangelical_bl <- max(filter(rho_all_bl, Type == "Evangelical")$rho)
+max_int_catholic_bl <- max(filter(rho_all_bl, Type == "Catholic")$rho)
+max_int_spiritist_bl <- max(filter(rho_all_bl, Type == "Spiritist")$rho)
+max_int_ice_cream_bl <- max(filter(rho_all_bl, Type == "Ice Cream Shops")$rho)
+max_int_pharmacy_bl <- max(filter(rho_all_bl, Type == "Pharmacies")$rho)
+max_int_supermarket_bl <- max(filter(rho_all_bl, Type == "Supermarkets")$rho)
+
+# Next, the distance at which the intensity of crime with respect to distance to Evangelical is max:
+
+dist_max_int_evangelical_bl <- filter(rho_all_bl, Type == "Evangelical")[filter(rho_all_bl, Type == "Evangelical")$rho == max_int_evangelical_bl,"z"]
+
+# Next, find the approximate intensity at that distance for the other types of events. First, what is the nearest distance to the distance at which the intensity with respect to Evangelical is max?
+
+near_dist_catholic_i <- which.min(abs(filter(rho_all_bl, Type == "Catholic")$z - dist_max_int_evangelical_bl))
+
+# If the nearest distance is greater than the one for evangelical, pick the preceding row in the dataframe, if it is less than then the following row.
+Use the two rows to obtain a linear approximation of the intensity at the desired distance
+
+if(filter(rho_all_bl, Type == "Catholic")[near_dist_catholic_i, "z"] > dist_max_int_evangelical_bl){
+  interv <- filter(rho_all_bl, Type == "Catholic")[(near_dist_catholic_i-1):near_dist_catholic_i,] %>%
+    dplyr::select(z, rho)
+  slope <- (interv$rho[2] - interv$rho[1]) / (interv$z[2] - interv$z[1])
+  catholic_int_at_evangelical_max_int_bl <- interv$rho[1] + (dist_max_int_evangelical_bl - interv$z[1]) * slope
+}else{
+  interv <- filter(rho_all_bl, Type == "Catholic")[near_dist_catholic_i:(near_dist_catholic_i + 1),] %>%
+    dplyr::select(z, rho)
+  slope <- (interv$rho[2] - interv$rho[1]) / (interv$z[2] - interv$z[1])
+  catholic_int_at_evangelical_max_int_bl <- interv$rho[1] + (dist_max_int_evangelical_bl - interv$z[1]) * slope
+}
+
+# Repeat for Spiritist
+
+near_dist_spiritist_i <- which.min(abs(filter(rho_all_bl, Type == "Spiritist")$z - dist_max_int_evangelical_bl))
+
+# If the nearest distance is greater than the one for evangelical, pick the preceding row in the dataframe, if it is less than then the following row.
+Use the two rows to obtain a linear approximation of the intensity at the desired distance
+
+if(filter(rho_all_bl, Type == "Spiritist")[near_dist_spiritist_i, "z"] > dist_max_int_evangelical_bl){
+  interv <- filter(rho_all_bl, Type == "Spiritist")[(near_dist_spiritist_i-1):near_dist_spiritist_i,] %>%
+    dplyr::select(z, rho)
+  slope <- (interv$rho[2] - interv$rho[1]) / (interv$z[2] - interv$z[1])
+  spiritist_int_at_evangelical_max_int_bl <- interv$rho[1] + (dist_max_int_evangelical_bl - interv$z[1]) * slope
+}else{
+  interv <- filter(rho_all_bl, Type == "Spiritist")[near_dist_spiritist_i:(near_dist_spiritist_i + 1),] %>%
+    dplyr::select(z, rho)
+  slope <- (interv$rho[2] - interv$rho[1]) / (interv$z[2] - interv$z[1])
+  spiritist_int_at_evangelical_max_int_bl <- interv$rho[1] + (dist_max_int_evangelical_bl - interv$z[1]) * slope
+}
+
+# Repeat for ice cream shop
+
+near_dist_ice_cream_i <- which.min(abs(filter(rho_all_bl, Type == "Ice Cream Shops")$z - dist_max_int_evangelical_bl))
+
+# If the nearest distance is greater than the one for evangelical, pick the preceding row in the dataframe, if it is less than then the following row.
+Use the two rows to obtain a linear approximation of the intensity at the desired distance
+
+if(filter(rho_all_bl, Type == "Ice Cream Shops")[near_dist_ice_cream_i, "z"] > dist_max_int_evangelical_bl){
+  interv <- filter(rho_all_bl, Type == "Ice Cream Shops")[(near_dist_ice_cream_i-1):near_dist_ice_cream_i,] %>%
+    dplyr::select(z, rho)
+  slope <- (interv$rho[2] - interv$rho[1]) / (interv$z[2] - interv$z[1])
+  ice_cream_int_at_evangelical_max_int_bl <- interv$rho[1] + (dist_max_int_evangelical_bl - interv$z[1]) * slope
+}else{
+  interv <- filter(rho_all_bl, Type == "Ice Cream Shops")[near_dist_ice_cream_i:(near_dist_ice_cream_i + 1),] %>%
+    dplyr::select(z, rho)
+  slope <- (interv$rho[2] - interv$rho[1]) / (interv$z[2] - interv$z[1])
+  ice_cream_int_at_evangelical_max_int_bl <- interv$rho[1] + (dist_max_int_evangelical_bl - interv$z[1]) * slope
+}
+
+# Repeat for pharmacy
+
+near_dist_pharmacies_i <- which.min(abs(filter(rho_all_bl, Type == "Pharmacies")$z - dist_max_int_evangelical_bl))
+
+# If the nearest distance is greater than the one for evangelical, pick the preceding row in the dataframe, if it is less than then the following row.
+Use the two rows to obtain a linear approximation of the intensity at the desired distance
+
+if(filter(rho_all_bl, Type == "Pharmacies")[near_dist_pharmacies_i, "z"] > dist_max_int_evangelical_bl){
+  interv <- filter(rho_all_bl, Type == "Pharmacies")[(near_dist_pharmacies_i - 1):near_dist_pharmacies_i,] %>%
+    dplyr::select(z, rho)
+  slope <- (interv$rho[2] - interv$rho[1]) / (interv$z[2] - interv$z[1])
+  pharmacies_int_at_evangelical_max_int_bl <- interv$rho[1] + (dist_max_int_evangelical_bl - interv$z[1]) * slope
+}else{
+  interv <- filter(rho_all_bl, Type == "Pharmacies")[near_dist_pharmacies_i:(near_dist_pharmacies_i + 1),] %>%
+    dplyr::select(z, rho)
+  slope <- (interv$rho[2] - interv$rho[1]) / (interv$z[2] - interv$z[1])
+  pharmacies_int_at_evangelical_max_int_bl <- interv$rho[1] + (dist_max_int_evangelical_bl - interv$z[1]) * slope
+}
+
+# Repeat for supermarkets
+
+near_dist_supermarkets_i <- which.min(abs(filter(rho_all_bl, Type == "Supermarkets")$z - dist_max_int_evangelical_bl))
+
+# If the nearest distance is greater than the one for evangelical, pick the preceding row in the dataframe, if it is less than then the following row.
+Use the two rows to obtain a linear approximation of the intensity at the desired distance
+
+if(filter(rho_all_bl, Type == "Supermarkets")[near_dist_supermarkets_i, "z"] > dist_max_int_evangelical_bl){
+  interv <- filter(rho_all_bl, Type == "Supermarkets")[(near_dist_supermarkets_i - 1):near_dist_supermarkets_i,] %>%
+    dplyr::select(z, rho)
+  slope <- (interv$rho[2] - interv$rho[1]) / (interv$z[2] - interv$z[1])
+  supermarkets_int_at_evangelical_max_int_bl <- interv$rho[1] + (dist_max_int_evangelical_bl - interv$z[1]) * slope
+}else{
+  interv <- filter(rho_all_bl, Type == "Supermarkets")[near_dist_supermarkets_i:(near_dist_supermarkets_i + 1),] %>%
+    dplyr::select(z, rho)
+  slope <- (interv$rho[2] - interv$rho[1]) / (interv$z[2] - interv$z[1])
+  supermarkets_int_at_evangelical_max_int_bl <- interv$rho[1] + (dist_max_int_evangelical_bl - interv$z[1]) * slope
+}
+```
+
+The results of the analysis are shown in Figure \ref{fig:plot-relative-distribution-with-baseline}.
+The general trends shown by the relative distribution functions are similar to those seen above, before the introduction of a baseline function.
+We notice that at short distances the intensity of crime as a function of distance to the various places of worship and commercial establishments is higher than the baseline intesity (i.e., the intensity according to Socio-Ecomomic Deprivation).
+The only exception is intensity as a function of distance to a Spiritist church, which at short distances is not significantly different from the baseline intensity (see the confidence bands).
+
+The peak intensity as a function of distance to Evangelical churches is `r round(max_int_evangelical_bl, 2)` times the baseline intensity; for Catholic places of worship, the highest intesity is `r round(max_int_catholic_bl, 2)` times the baseline intensity; and the intensity reaches a peak of `r round(max_int_spiritist_bl, 2)` times the baseline intensity when the covariate is distance to Spiritist churches.
+Turning now to inspect the three types of commercial establishments, we see that the highest intensities are `r round(max_int_ice_cream_bl, 2)`, `r round(max_int_pharmacy_bl, 2)`, and `r round(max_int_supermarket_bl, 2)` times the intensity according to Socio-Economic Deprivation for ice cream shops, pharmacies, and supermarkets, respectively.
+
+Some differences are observed as well.
+Whereas the peak intensity as a function of distance to Evangelical places of worship was `r round(max_int_evangelical/catholic_int_at_evangelical_max_int * 100, 2)`% higher than the intensity of Catholic places of worship _without a baseline function_, once that we account for Socio-Economic Deprivation by means of a baseline function, the peak relative intensity as a function of distance to Evangelical churches declines to `r round(max_int_evangelical_bl/catholic_int_at_evangelical_max_int_bl * 100, 2)`% higher than the intensity as a function of distance to Catholic places of worship at the same distance.
+Accounting for Socio-Economic Deprivation in this analysis of intensity has in general a moderating effect: whereas the peak intensity of crime was was `r round(max_int_evangelical/spiritist_int_at_evangelical_max_int * 100, 2)`% higher when comparing Evangelical and Spiritist churches at the distance of the peak intensity of Evangelical churches, this difference declines to  `r round(max_int_evangelical_bl/spiritist_int_at_evangelical_max_int_bl * 100, 2)`% after accounting for Socio-Economic Deprivation.
+Similarly, this changes from `r round(max_int_evangelical/ice_cream_int_at_evangelical_max_int * 100, 2)`% to `r round(max_int_evangelical_bl/ice_cream_int_at_evangelical_max_int_bl * 100, 2)`% in the case of ice cream shops, from `r round(max_int_evangelical/pharmacies_int_at_evangelical_max_int * 100, 2)`% to `r round(max_int_evangelical_bl/pharmacies_int_at_evangelical_max_int_bl * 100, 2)`% in the case of pharmacies, and finally from `r round(max_int_evangelical/supermarkets_int_at_evangelical_max_int * 100, 2)`% to `r round(max_int_evangelical_bl/supermarkets_int_at_evangelical_max_int_bl * 100, 2)`% in the case of supermarkets.
+
+Again, it is interesting to note that while the intensity with respect to distance to Catholic and Spiritist places of worship, as well as ice cream shops, pharmacies, and supermarkets remains higher than the background intensity (given by the baseline according to Socio-Economic Deprivation), the intensity of crime at distances of approximately 300 m of an Evangelical church are significantly lower than the background intensity of crime.
+In the following section we discuss the possible implications of these findings.
+
+```{r figure-plot-relative-distribution-with-baseline, echo=FALSE, cache = TRUE, fig.cap="\\label{fig:plot-relative-distribution-with-baseline}Intensity of crime as a function of distance to a selected place of worship or commercial establishment after introducing Socio-Economic Deprivation as a baseline function; bands are 95% confidence intervals and the dotted line indicates the baseline intensity with respect to Socio-Economic Deprivation"}
+
+# Plot relative distribution with baseline function
+
+ggplot(data = filter(rho_all_bl, z <= 500)) +
+  geom_ribbon(aes(x = z, ymin = lo, ymax = hi, fill = Type), alpha = 0.2) + 
+  geom_line(aes(x = z, y = rho, color = Type, linetype = Class), size = 1) +
+  geom_abline(intercept = 1, slope = 0, linetype = "dotted") +
+  #xlim(c(0, 500)) +
+  theme_tufte()
+```
 
@@
 
-An opposing viewpoint is offered by the theory of conservative Protestants [@Ellison2003enduring].
-This theory argues that religion is not always beneficial in order to guarantee social order, and consequently to result in lower crime rates.
-This theory suggested that, conservative Protestants are more tolerate with violent behavior when associated with defense of honor, family, property, or woman.
-However, high rates are not expected for all types of crimes in densely populated areas of conservative Protestants.
-But what justifies this behavior?
-Ellison et al.
-[-@Ellison2003enduring] argue that this tolerance stems from a religious view of the world.
-More clearly, punishment is seen as a legitimate form of retribution whose explanation can be attributed to the literal interpretation of the Bible.
-In the specific context of Brazil, it is important to consider that the interpretations and meanings attributed to urban violence during the last two decades in Brazil often had religious figures as protagonists [@Birman2012violencia].
+Words go here.
 
@@
 
-The theory of conservative Protestants, argues that conservative Protestants are more likely to tolerate violent behavior of a defensive or punitive character, especially when associated to defense of honor, family, property, or woman.
-Our objective in the study is to explore the question whether moral communities have a spatial dimension.
-To this end we deploy an array of spatial analytical tools to investigate potential patterns of clustering of violent crime and places of worship, using as well other kinds of urban facilities to act as controls.
-The case study is the city of Recife, in the northeast of Brazil, and analysis is based on geographically detailed datasets of criminal events as well as several types of urban facilities.
+The objective of this research was to investigate the covariation of incidence of crime to proximity of places of worship.
+The hypothesis of moral communities (closely related to social disorganization theory) posits that a negative correlation between the presence of churches and crime.
+Recent papers have argued that it is important to study this phenomenon from a geographical perspective, paying attention to the level of aggregation.
+Accordingly, our research for the city of Recife in Brazil makes the following contributions to the literature:
+
+1.
+By using disaggregated data for Lethal and Intentional Violent Crime and various types of places of worship, we were able to analyze the intensity of crime with respect to proximity to churches as a point pattern.
+
+2.
+Following the suggestion of Warner and Konkel [-@Warner2019neighborhood], we also disaggregate our data by type of place of worship, and consider the three following denominations: Catholic, Evangelical, and Spiritist.
+
+2. Use of relative distribution functions is effectively a multi-scale technique, since the intensity of crime can be estimated at different distances from various types of places of worship.
+
+3. We also used a set of putatively morally neutral commercial establishments to serve as controls in our analysis.
+
+Our key findings can be summarized as follows:
+
+1.
+Catholic places of worhsip do not seem to geographically project more or less of a moral community effect than, say, ice cream shops, pharmacies, or supermarkets: the intensity of crime as a function of distance to Catholic churches is not significantly different that what we observe when we consider distance to any of these types of commercial establishments, at any distance.
+
+2.
+The intensity of crime with respect to distance to Evangelical churches is more complex than previously known: whereas the intensity of crime is higher at relativey short distance from Evangelical places of worship, at longer distances it tends to decline even below the global intensity of crime and the background level of crime according to Socio-Economic Deprivation.
+
+3.
+Spiritist churches are associated with lower intensity of crime at very short distance; however, at longer distances(approximately between 200 m and 500 m), the intensity of crime is in fact higher than the intensity of crime at comparable distances from Catholic and Evangelical places of worship, or any of the three kinds of commercial establishments.
+
+These results are interesting of and in themselves as they help to clarify the potential of various types of moral communities to act as restraints on crime or, on the contrary, as criminogenic factors.
+They also suggest several avenues for further research.
+
+1.
+The analysis presented in this paper is to a large extent exploratory.
+For example, to introduce a baseline function in the analysis of relative distribution, we used a data reduction technique to obtain an indicator of Socio-Economic Deprivation.
+This indicator was useful to determine a general spatial pattern of crime conditional on this variable, and to account for a baseline (or background) intensity of crime.
+However, it would be interesting to investigate the effect of the individual socio-economic and demographic variables on the intensity of crime, as opposed to the aggregated effect of Socio-Economic Deprivation.
+This implies the use of multivariate analytical techniques.
+
+2.
+It would be interesting to explore the effect of aggregation on the results.
+On the one hand, this would be informative with respect to the well-known issues with aggregation bias in spatial analysis [related to the Modifiable Areal Unit Problem in geography; see @Fotheringham1991modifiable; @Openshaw1979million; and @Tagashira2002modifiable].
+On the other hand, as the results of the research presented here, it is possible that analysis at different levels of aggregation may tend to capture different spatial and social processes of interest.
+
+3.
+And finally, in relation to the latter point, more in-depth research, perhaps observational, ethnographic, or participatory studies, could help to develop a more refined understanding of the kinds of processes that operate at different scales and how they can influence moral behavior and crime.
+
+In summary, the research presented on this paper provides information about the potential of different types of moral communities to reduce crime, and should be of interest to policy makers as they assess whether formal or informal forms of social control can be effective to deter criminal behavior.
 

### 2019-10-15 Moral_Communities_and_Crime_v1.tex
Completed first draft of analysis/results and conclusions.

@@
 important in places where formal state institutions lack the means or
-the will to enforce norms - as is the case in Brazil.
+the will to enforce norms in a consistent way - as is the case in Brazil
+(Garmany 2014).
 
@@
 For the analysis, we use Principal Component Analysis, a data reduction
-technique, to obtain an indicator of Socio-Economic Deprivation. The
-loadings of the first factor, which accounts for almost 54\% of the
-variance, are shown in Table \ref{tab:sed-descriptive-statistics}. The
-loadings of this factor indicate that high Socio-Economic Deprivation is
-a combination of (in terms of importance): high levels of poverty, high
-levels of unemployment, and low median income, followed by high levels
-of family disruption, in particular proportion of families led by young
-single mothers.
+technique, to obtain an indicator of Socio-Economic Deprivation,
+essentially the first principal component in the output. The loadings of
+this factor (Table \ref{tab:sed-descriptive-statistics}) indicate that
+high Socio-Economic Deprivation is a combination of (in terms of
+importance): high levels of poverty, high levels of unemployment, and
+low median income, followed by high levels of family disruption, in
+particular proportion of families led by young single mothers. This
+factor accounts for almost 54\% of the variance..
 
@@
 \centering
-\begin{tabular}[t]{lrrrrrrr}
+\begin{tabular}[t]{lccccccc}
 \toprule
@@
 mosaic of affluence and deprivation, with high deprivation areas
-directly in contact, and in some cases even completely surrounded, by
+directly in contact, and in some cases almost completely surrounded, by
 low deprivation areas. This geographical pattern of unequality, on the
@@
 
-Calculate the intensity of other ppp by quadrat:
+Once the indicator of Socio-Economic Deprivation has been obtained, as
+outlined above, it can be used for conditional quadrat analysis. As
+previously described, the quadrats are irregular areas that are defined
+based on a covariate, in this case the quintiles of the Socio-Economic
+Deprivation indicator. In essence, this entails counting the number of
+events at each level of the covariate, and then calculating an estimator
+of the intensity.
+
+Figure \ref{fig:plot-crime-quadrat} shows the results of conditional
+quadrat counts for Lethal and Intentional Violent Crime. The dotted line
+indicates the global intensity, simply the total number of criminal
+events divided by the area of the region. It can be seen that the
+intensity of crime clearly increases with Socio-Economic Deprivation,
+and the intensity in areas with the highest levels of Socio-Economic
+Deprivation is 208.7065292 \% higher than in places with the lowest
+levels of Socio-Economic Deprivation. This is as expected, as research
+has consistently found correlations between Socio-Economic Deprivation
+and a variety of negative health outcomes, including homicide (Ichihara
+et al. 2018). A \(\chi^2\) test of independence on the quadrat counts
+yields a value of \(p\leq 0.0001\) (four degrees of freedom), which
+comfortably rejects the null hypothesis of homogeneity of intensity of
+crime by level of Socio-Economic Deprivation.
 
@@
 
+Next, we repeat the analysis of conditional quadrat counts, but now for
+the six types of events of reference. As discussed above, this included
+three types of places of worship which are of interest from the
+perspective of moral communities, and three types of commercial
+establishments that we posit are morally neutral. The results of
+estimating the intensity of the process by means of conditional quadrat
+counts are shown in Figure \ref{fig:plot-events-quadrat}.
+
+The first thing that we note are the variations in global intensity of
+places of worship, with Evangelical places of worship being the most
+intense of the three. There are no clear variations in the locational
+pattern of Catholic places of worship by level of Socio-Economic
+Deprivation. Evangelical places of worship, on the other hand, tend to
+be found more frequently in places with middle-low and middle levels of
+Socio-Economic Deprivation, a pattern also discernible, albeit barely,
+for Spiritist churches. Interestingly (see Table
+\ref{tab:quadrat-test-statistics}) \(\chi^2\) tests of independence on
+the conditional quadrat counts fail to reject the hypothesis of
+homogeneity for Catholic and Spiritist churches, and it is only in the
+case of Evangelical places of worship that we detect the possibility of
+locational patterns that vary by level of Socio-Economic Deprivation.
+
+Three types of commercial establishments also appear to display
+inhomogeneous locational patterns (Table
+\ref{tab:quadrat-test-statistics}), with the \(\chi^2\) test of
+independence emphaticaly rejecting the null hypothesis. As seen in
+Figure \ref{fig:plot-events-quadrat}, ice cream shops tend to be found
+more frequently in places with middle-low Socio-Economic Deprivation,
+and less frequently in places with middle-high Socio-Economic
+Deprivation. Pharmacies show a clearer trend, with locational patterns
+that tend to favor places with low Socio-Economic Deprivation. Finally,
+supermarkets are found less frequently in places with middle-high and
+high Socio-Economic Deprivation.
+
 \begin{figure}
@@
 
+\begin{table}
+
+\caption{\label{tab:table-quadrat-test-statistics}\label{tab:quadrat-test-statistics}$\chi^2$ tests of independence of conditional quadrat counts}
+\centering
+\begin{tabular}[t]{lccc}
+\toprule
+Event & Statistic & Degrees of Freedom & p-value\\
+\midrule
+Catholic & 6.913 & 4 & 0.2811\\
+Evangelical & 38.669 & 4 & $<0.001$\\
+Spiritist & 9.108 & 4 & 0.1169\\
+Ice Crem Shop & 15.108 & 4 & 0.009\\
+Pharmacy & 51.869 & 4 & $<0.001$\\
+\addlinespace
+Supermarket & 23.826 & 4 & $<0.001$\\
+\bottomrule
+\end{tabular}
+\end{table}
+
 \hypertarget{relative-distribution}{%
 \subsection{Relative Distribution}\label{relative-distribution}}
+
+The preceding analysis provides a valuable backdrop. As seen there,
+there is a clear distribution of criminal events, increasing on par with
+the level of Socio-Economic deprivation. Of the three different classes
+of places of worship, only Evangelical churches display a locational
+pattern that is commensurate, with Socio-Economic Deprivation, albeit
+the pattern is distinct from criminal events, with Evangelical churches
+found less frequently in both high and low Socio-Economic Deprivation
+areas.
+
+The analysis does not answer the question, yet, of possible covariations
+between crime, on the one hand, and places of worship and commercial
+establishments, on the other. In this subsection we implement the
+relative distribution estimator of intensity of Lethal and Intentional
+Violent Crime. The covariate \(Z(u)\) in this analysis is distance to
+one of a place of worship or a commercial establishment: \[
+\lambda_k(s)=\rho(Z_k(s))
+\] with: \[
+Z_k(s) = d_{sk}
+\] In the expression above, \(d_{s_k}\) is distance from location \(s\)
+to an event of class \(k\), and
+\(k=\{\text{Catholic, Evangelical, Spiritist, Ice Cream Shop, Pharmacy, Supermarket}\}\).
+
+The results of this analysis are shown in Figure
+\ref{fig:plot-relative-distribution}. The global estimator for the
+intensity of crime is shown in this figure as a dotted line. Three types
+of places of worship are shown in solid lines, and three types of
+commercial establishments are shown in dashed lines. Each relative
+distribution function is shown with its corresponding 95\% confidence
+bands.
+
+Several interesting things emerge from inspection of this figure. First,
+we notice that with the exception of Spiritist churches (which has an
+intensity significantly lower than the global intensity at a distance of
+zero), the intensity of crime at for other churches and commercial
+establishments is close, but higher than, the global intensity of crime.
+In general, the intensity tends to increase at short distances; however,
+this effect is much more marked for Evangelical places of worship. The
+intensity of crime in the proximity of these places of worship grows
+rapidly within a short distance, reaching a peak of
+\ensuremath{2.4916\times 10^{-5}} at a distance of 64.92 m. At that
+distance, the intensity of crime with respect to Evangelical places of
+worship is 183.78\% higher than the intensity of Catholic places of
+worship; 194.54\% higher than the intensity of crime with respect to
+Spiritist churches; 174.19\% higher than the intensity with respect to
+ice cream shops; 151.52\% higher than the intensity with respect to
+pharmacies; and 177.63\% higher than the intensity with respect to
+supermarkets.
+
+Intriguingly, after reaching its peak intensity, the intensity of crime
+with respect to Evangelical places of worship declines sharply, until at
+distances of approximately 200 m it is lower than for other places of
+worship and commercial establishments, and at distances of approximately
+270 m it is lower than the global intensity of crime. This does not
+happen with other places of worship or commercial establishments: the
+intensity of crime with respect to these locations remains higher than
+the global intensity of crime for the interval of distances examined.
+Another interesting observation is that the intensity of crime with
+respect to Catholic places of worship is not substantially different
+when compared to the intensity of crime with respect to any of the
+commercial establishments considered.
+
+These results, while intriguing, beg the question of whether variations
+in Socio-Economic Deprivation may have a confounding effect on the
+relative distributions; for instance, if as seen in the case of
+conditional quadrat counts, there is some overlap in the locational
+patterns of, say, Evangelical places of worship and crime with respect
+to Socio-Economic Deprivation. Next section explores this question.
+
+\begin{figure}
+\centering
+\includegraphics{Moral_Communities_and_Crime_v1_files/figure-latex/figure-plot-relative-distribution-1.pdf}
+\caption{\label{fig:plot-relative-distribution}Intensity of crime as a
+function of distance to a selected place of worship or commercial
+establishment; bands are 95\% confidence interval and the dotted line
+indicates the global intensity of crime}
+\end{figure}
 
@@
 
+In this section, the preceding analysis regarding the relative
+distribution of crime using distance to places of worship and commercial
+establishments is repeated, after introducing a baseline function. The
+interpretation of \(\rho(Z(s))\) in this case is as follows: instead of
+the raw intensity at the value of \(z\), it is the intensity
+\emph{relative to the intensity according to the level of Socio-Economic
+Deprivation}. Therefore, \(rho(z)=1\) matches the intensity according to
+Socio-Economic Deprivation; values of \(rho(z)>1\) indicate a higher
+intensity than explained by Socio-Economic Deprivation; and values of
+\(rho(z)<1\) correspond to lower intensities than explained by
+Socio-Economic Deprivation.
+
+The results of the analysis are shown in Figure
+\ref{fig:plot-relative-distribution-with-baseline}. The general trends
+shown by the relative distribution functions are similar to those seen
+above, before the introduction of a baseline function. We notice that at
+short distances the intensity of crime as a function of distance to the
+various places of worship and commercial establishments is higher than
+the baseline intesity (i.e., the intensity according to Socio-Ecomomic
+Deprivation). The only exception is intensity as a function of distance
+to a Spiritist church, which at short distances is not significantly
+different from the baseline intensity (see the confidence bands).
+
+The peak intensity as a function of distance to Evangelical churches is
+3.33 times the baseline intensity; for Catholic places of worship, the
+highest intesity is 2.12 times the baseline intensity; and the intensity
+reaches a peak of 2.67 times the baseline intensity when the covariate
+is distance to Spiritist churches. Turning now to inspect the three
+types of commercial establishments, we see that the highest intensities
+are 2.28, 2.46, and 2.09 times the intensity according to Socio-Economic
+Deprivation for ice cream shops, pharmacies, and supermarkets,
+respectively.
+
+Some differences are observed as well. Whereas the peak intensity as a
+function of distance to Evangelical places of worship was 183.78\%
+higher than the intensity of Catholic places of worship \emph{without a
+baseline function}, once that we account for Socio-Economic Deprivation
+by means of a baseline function, the peak relative intensity as a
+function of distance to Evangelical churches declines to 176.53\% higher
+than the intensity as a function of distance to Catholic places of
+worship at the same distance. Accounting for Socio-Economic Deprivation
+in this analysis of intensity has in general a moderating effect:
+whereas the peak intensity of crime was was 194.54\% higher when
+comparing Evangelical and Spiritist churches at the distance of the peak
+intensity of Evangelical churches, this difference declines to 186.45\%
+after accounting for Socio-Economic Deprivation. Similarly, this changes
+from 174.19\% to 168.33\% in the case of ice cream shops, from 151.52\%
+to 140.97\% in the case of pharmacies, and finally from 177.63\% to
+164.33\% in the case of supermarkets.
+
+Again, it is interesting to note that while the intensity with respect
+to distance to Catholic and Spiritist places of worship, as well as ice
+cream shops, pharmacies, and supermarkets remains higher than the
+background intensity (given by the baseline according to Socio-Economic
+Deprivation), the intensity of crime at distances of approximately 300 m
+of an Evangelical church are significantly lower than the background
+intensity of crime. In the following section we discuss the possible
+implications of these findings.
+
+\begin{figure}
+\centering
+\includegraphics{Moral_Communities_and_Crime_v1_files/figure-latex/figure-plot-relative-distribution-with-baseline-1.pdf}
+\caption{\label{fig:plot-relative-distribution-with-baseline}Intensity
+of crime as a function of distance to a selected place of worship or
+commercial establishment after introducing Socio-Economic Deprivation as
+a baseline function; bands are 95\% confidence intervals and the dotted
+line indicates the baseline intensity with respect to Socio-Economic
+Deprivation}
+\end{figure}
+
 \hypertarget{discussion}{%
@@
 
-An opposing viewpoint is offered by the theory of conservative
-Protestants (Ellison, Burr, and McCall 2003). This theory argues that
-religion is not always beneficial in order to guarantee social order,
-and consequently to result in lower crime rates. This theory suggested
-that, conservative Protestants are more tolerate with violent behavior
-when associated with defense of honor, family, property, or woman.
-However, high rates are not expected for all types of crimes in densely
-populated areas of conservative Protestants. But what justifies this
-behavior? Ellison et al.~(2003) argue that this tolerance stems from a
-religious view of the world. More clearly, punishment is seen as a
-legitimate form of retribution whose explanation can be attributed to
-the literal interpretation of the Bible. In the specific context of
-Brazil, it is important to consider that the interpretations and
-meanings attributed to urban violence during the last two decades in
-Brazil often had religious figures as protagonists (Birman and Machado
-2012).
+Words go here.
 
@@
 
-The theory of conservative Protestants, argues that conservative
-Protestants are more likely to tolerate violent behavior of a defensive
-or punitive character, especially when associated to defense of honor,
-family, property, or woman. Our objective in the study is to explore the
-question whether moral communities have a spatial dimension. To this end
-we deploy an array of spatial analytical tools to investigate potential
-patterns of clustering of violent crime and places of worship, using as
-well other kinds of urban facilities to act as controls. The case study
-is the city of Recife, in the northeast of Brazil, and analysis is based
-on geographically detailed datasets of criminal events as well as
-several types of urban facilities.
+The objective of this research was to investigate the covariation of
+incidence of crime to proximity of places of worship. The hypothesis of
+moral communities (closely related to social disorganization theory)
+posits that a negative correlation between the presence of churches and
+crime. Recent papers have argued that it is important to study this
+phenomenon from a geographical perspective, paying attention to the
+level of aggregation. Accordingly, our research for the city of Recife
+in Brazil makes the following contributions to the literature:
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\item
+  By using disaggregated data for Lethal and Intentional Violent Crime
+  and various types of places of worship, we were able to analyze the
+  intensity of crime with respect to proximity to churches as a point
+  pattern.
+\item
+  Following the suggestion of Warner and Konkel (2019), we also
+  disaggregate our data by type of place of worship, and consider the
+  three following denominations: Catholic, Evangelical, and Spiritist.
+\item
+  Use of relative distribution functions is effectively a multi-scale
+  technique, since the intensity of crime can be estimated at different
+  distances from various types of places of worship.
+\item
+  We also used a set of putatively morally neutral commercial
+  establishments to serve as controls in our analysis.
+\end{enumerate}
+
+Our key findings can be summarized as follows:
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\item
+  Catholic places of worhsip do not seem to geographically project more
+  or less of a moral community effect than, say, ice cream shops,
+  pharmacies, or supermarkets: the intensity of crime as a function of
+  distance to Catholic churches is not significantly different that what
+  we observe when we consider distance to any of these types of
+  commercial establishments, at any distance.
+\item
+  The intensity of crime with respect to distance to Evangelical
+  churches is more complex than previously known: whereas the intensity
+  of crime is higher at relativey short distance from Evangelical places
+  of worship, at longer distances it tends to decline even below the
+  global intensity of crime and the background level of crime according
+  to Socio-Economic Deprivation.
+\item
+  Spiritist churches are associated with lower intensity of crime at
+  very short distance; however, at longer distances(approximately
+  between 200 m and 500 m), the intensity of crime is in fact higher
+  than the intensity of crime at comparable distances from Catholic and
+  Evangelical places of worship, or any of the three kinds of commercial
+  establishments.
+\end{enumerate}
+
+These results are interesting of and in themselves as they help to
+clarify the potential of various types of moral communities to act as
+restraints on crime or, on the contrary, as criminogenic factors. They
+also suggest several avenues for further research.
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\item
+  The analysis presented in this paper is to a large extent exploratory.
+  For example, to introduce a baseline function in the analysis of
+  relative distribution, we used a data reduction technique to obtain an
+  indicator of Socio-Economic Deprivation. This indicator was useful to
+  determine a general spatial pattern of crime conditional on this
+  variable, and to account for a baseline (or background) intensity of
+  crime. However, it would be interesting to investigate the effect of
+  the individual socio-economic and demographic variables on the
+  intensity of crime, as opposed to the aggregated effect of
+  Socio-Economic Deprivation. This implies the use of multivariate
+  analytical techniques.
+\item
+  It would be interesting to explore the effect of aggregation on the
+  results. On the one hand, this would be informative with respect to
+  the well-known issues with aggregation bias in spatial analysis
+  (related to the Modifiable Areal Unit Problem in geography; see
+  Fotheringham and Wong 1991; Openshaw and Taylor 1979; and Tagashira
+  and Okabe 2002). On the other hand, as the results of the research
+  presented here, it is possible that analysis at different levels of
+  aggregation may tend to capture different spatial and social processes
+  of interest.
+\item
+  And finally, in relation to the latter point, more in-depth research,
+  perhaps observational, ethnographic, or participatory studies, could
+  help to develop a more refined understanding of the kinds of processes
+  that operate at different scales and how they can influence moral
+  behavior and crime.
+\end{enumerate}
+
+In summary, the research presented on this paper provides information
+about the potential of different types of moral communities to reduce
+crime, and should be of interest to policy makers as they assess whether
+formal or informal forms of social control can be effective to deter
+criminal behavior.
 
@@
 
-\leavevmode\hypertarget{ref-Birman2012violencia}{}%
-Birman, Patricia, and Carly Machado. 2012. ``A Violencia Dos Justos:
-Evangelicos, Midia E Periferias Da Metropole.'' \emph{Revista Brasileira
-de Ciências Sociais} 27 (80): 55--69.
-
 \leavevmode\hypertarget{ref-Brower2007spatial}{}%
@@
 61--81. \url{https://doi.org/10.1111/j.1468-5906.2010.01552.x}.
-
-\leavevmode\hypertarget{ref-Ellison2003enduring}{}%
-Ellison, Christopher G, Jeffrey A Burr, and Patricia L McCall. 2003.
-``The Enduring Puzzle of Southern Homicide: Is Regional Religious
-Culture the Missing Piece?'' \emph{Homicide Studies} 7 (4): 326--52.
 
@@
 \url{https://doi.org/10.1016/j.healthplace.2010.07.007}.
+
+\leavevmode\hypertarget{ref-Fotheringham1991modifiable}{}%
+Fotheringham, A. S., and D. W. S. Wong. 1991. ``The Modifiable Areal
+Unit Problem in Multivariate Statistical-Analysis.'' Journal Article.
+\emph{Environment and Planning A} 23 (7): 1025--44.
+\url{ISI:A1991GA61200008}.
 
@@
 
+\leavevmode\hypertarget{ref-Garmany2014space}{}%
+Garmany, J. 2014. ``Space for the State? Police, Violence, and Urban
+Poverty in Brazil.'' Journal Article. \emph{Annals of the Association of
+American Geographers} 104 (6): 1239--55.
+\url{https://doi.org/10.1080/00045608.2014.944456}.
+
 \leavevmode\hypertarget{ref-Groff2015informal}{}%
@@
 
+\leavevmode\hypertarget{ref-Ichihara2018area}{}%
+Ichihara, M. Y. T., D. Ramos, P. Reboucas, F. J. Oliveira, A. J. F.
+Ferreira, C. Teixeira, M. Allik, et al. 2018. ``Area Deprivation
+Measures Used in Brazil: A Scoping Review.'' Journal Article.
+\emph{Revista de Saude Publica} 52 (83): 1--14.
+\url{https://doi.org/10.11606/s1518-8787.2018052000933}.
+
 \leavevmode\hypertarget{ref-Johnson2011more}{}%
@@
 Place} 58. \url{https://doi.org/10.1016/j.healthplace.2019.05.017}.
+
+\leavevmode\hypertarget{ref-Openshaw1979million}{}%
+Openshaw, S., and P. J. Taylor. 1979. ``A Million or so Correlation
+Coefficients: Three Experiments on the Modifiable Areal Unit Problem.''
+Book Section. In \emph{Statistical Applications in the Spatial
+Sciences}, edited by N. Wrigley, 127--44. London: Pion.
 
@@
 
+\leavevmode\hypertarget{ref-Tagashira2002modifiable}{}%
+Tagashira, N., and A. Okabe. 2002. ``The Modifiable Areal Unit Problem
+in a Regression Model Whose Independent Variable Is a Distance from a
+Predetermined Point.'' Journal Article. \emph{Geographical Analysis} 34
+(1): 1--20. \url{ISI:000173383000001}.
+
 \leavevmode\hypertarget{ref-Topalli2013god}{}%

### 2019-10-15 Moral_Communities_and_Crime_Title_Page_v1.Rmd
First complete draft of paper and title page

--- 2019-10-13 Moral_Communities_and_Crime_v1.Rmd
@@
 authors: 
-- name: Aye B. Cee
+- name: Edivaldo Constantino das Neves Júnior 
   address: A Department, University of XXX
-  email: abc@example.edu
+  email: edivaldoconstantino@gmail.com
 
-- name: Dee E. Eff
+- name: Breno Caldas de Araujo
   address: A Department, University of XXX
@@
 
-- name: Gee H. Eye
-  address: A Department, University of XXX
-  email: ghi@example.edu
+- name: Tatiane Almeida de Menezes
+  address: Programa de Pós-Graduação em Gestão e Economia da Saúde, Universidade Federal de Pernambuco
+  email: tatianedemenezes@gmail.com
 
-- name: Jay K. Elle
-  address: Some School, University of WWW
-  email: jkl@school.edu
+- name: Antonio Paez
+  address: School of Geography and Earth Sciences, McMaster University
+  email: paezha@mcmaster.ca
 
@@
 ---
-
-<!-- This paper is an R markdown document that includes self-contained, reproducible analysis.
-To reproduce the paper the following file is needed: `Moral_Communities_and_Crime.RData`.
-For reference: this data file was created using R notebook `00.
-Data Preprocessing.Rmd`.
-The initial analysis was done in R notebook `01.
-Analysis.Rmd`.
-Neither of these two files are needed for this document.
--->
-
-```{r load-packages, include=FALSE}
-
-# Load packages used in this document
-
-library(tidyverse)
-library(kableExtra)
-library(sf)
-library(spatstat)
-library(maptools)
-library(ggthemes)
-library(raster)
-library(fasterize)
-```
-
-```{r load-data, include=FALSE}
-
-# To run this notebook, a file `Moral_Communities_and_Crime.RData` is needed. This file contains the following objects:
-# 
-# - events_recife.sf: a simple features object with points for violent crimes, churches (Catholic, Evangelical, Spiritist) and commercial establishments (ice cream shops, pharmacies, supermarkets) in the city of Recife.
-# 
-# - recife.sf: a simple features object with the boundaries of the city of Recife
-# 
-# - setores.sf: a simple features object with setores (census tracts) and selected socio-economic and demographic information for the city of Recife, including:
-#   - Population density in persons per sq.km
-#   - Median income in setor
-#   - Proportion of population in setor who are unemployed
-#   - Proportion of households with a single mother
-#   - Proportion of households with a young single mother, a woman between the ages of 15 and 25
-#   - Proportion of population in poverty
-# 
-# Read the RData file:
-
-load("Moral_Communities_and_Crime.RData")
-```
-
-```{r data-preparation, include=FALSE, cache=TRUE}
-
-# In preparation for doing point pattern analysis, convert the `recife.sf` object to an `owin` object for use in `spatstat`:
-
-recife.owin <- recife.sf %>% as("Spatial") %>% as("owin")
-plot(recife.owin)
-
-# Convert the events to a `ppp` object:
-
-events_recife.ppp <- events_recife.sf %>% 
-  st_coordinates() %>% 
-  data.frame() %>%
-  mutate(Type = events_recife.sf$Type) %>%
-  as.ppp(W = recife.owin)
-
-#Split the point pattern into each type of event: crime, Catholic place of worship, Evangelical place of worship, Spiritist place of worship, ice cream shop, pharmacy, and supermarket
-
-events_recife_list <- split(events_recife.ppp)
-```
-
-```{r socio-economic-deprivation, include=FALSE, cache=TRUE}
-
-# Here an indicator of socio-economic deprivation is obtained using principal component analysis
-
-# PCA is implemented in R as `prcomp`:
-setores.pca <- setores.sf %>% 
-  st_drop_geometry() %>%
-  dplyr::select(median_income, prop_unemployed, prop_single_mother, prop_young_single_mother, prop_poverty) %>%
-  prcomp(center = TRUE, scale = TRUE)
-
-# The loadings of the pca are as follows:
-setores.pca$rotation
-
-# Socio-economic deprivation is a combination of high poverty, high proportion of single mothers, especially young women between the ages of 15-25, high proportion of unemployed people, and low median income.
-
-# Various diagnostics of the results are possible, such as the variance:
-
-std_dev <- setores.pca$sdev
-
-# Compute variance
-
-pr_var <- std_dev^2
-
-# Check variance of the components
-
-pr_var[1:5]
-
-# And the proportion of the variance:
-
-pr_var/sum(pr_var)
-
-# Once that there is principal component (call it SED for socio-economic deprivation), it can be joined to the `sf` object before it can be rasterized and converted to an `im` object:
-
-setores.sf$SED <- setores.pca$x[,1]
-
-# Normalize the SDE factor (with the addition of a small correction; the `rhohat` function does not like values <= 0):
-
-setores.sf <- setores.sf %>% mutate(SED = (SED - min(SED) + 0.1)/(max(SED) + 0.1 - min(SED)))
-```
-
-# Introduction {#intro}
-
-Violent crime is a widespread phenomenon with negative impacts on many spheres of life and society.
-Although the rate of homicides worldwide has grown at a slower rate than the population in the past few decades, the number of people killed in homicides still increased from 362,000 in 1990 to 464,000 in 2017 [@Unodc2019executive].
-Furthermore, variations in the prevalence of violent crime tend to be extremely uneven.
-The Americas, with a population of approximately 793.8 million people in 2019 (or about 10.3% of the world population), accounted for approximately 173,000 homicides in 2017, or approximately 37.3% of all homicides in the world.
-Within the Americas, Brazil (along with Venezuela, Colombia, and Mexico) is one of the largest countries in the region with high homicide rates.
-And there are notable variations at the sub-national and intra-metropolitan levels as well.
-In fact, understanding the variations in the prevalence of violent crime is recognized as key to achieving policy goals:
-
->> High levels of homicidal violence are concentrated in geographic and demographic “pockets”, so
-achieving target 16.1 of the Sustainable Development Goals requires interventions within the specific
-regions, countries, communities and population groups that are most at risk [@Unodc2019executive, p. 35]
-
-This concern is not iddle: the endemic malady of violent crime represents an important factor that threatens to derail sustainable development goals.
-Violent crime, furthermore, tends to be more accute in those regions most in need of development.
-On the one hand, violent crime represents an economic and social drag [@Becker2017analise].
-The impacts are multifaceted, since they affect people's well-being, either through the loss of human life, mental health, and limitations in the right to public spaces [@Doran2011putting], or through disturbances in schooling and academic achievement [@Unodc2019development].
-In turn, these negative effects combine to create the unfortunate conditions that tend to breed criminal behavior, thus creating a vicious cycle of economic disadvantage and violence [@Unodc2019development].
-It is not surprising, given these pecularities, that researchers have joined calls for increased attention to the study of the patterns of violent crime in Low and Middle Income Countries [LMIC; @Murray2013crime].
-
-Added to the scenario, formal social control institutions in countries like Brazil leave something (or much) to be desired: while in practice they should inhibit criminal behavior, they suffer from deep deficiencies that end up reducing the deterrent power of the justice system.
-Serious institutional problems include the inefficiency of the police, the lack of national legislation, the glacial pace of judicial processes, and the weak situation of the prison system in the country [@Menezes2013spatial].
-This makes it even more urgent to understand the role of _informal social controls_, i.e., the ability of community organizations such as schools, clubs, and neighborbood associations to suppress crime by strengthening the capabilities of neighbors to control inappropriate behavior [@Groff2015informal].
-
-In order to understand the factors that can potentially help to deter crime, beyond formal control, it is important to identify empirical regularities.
-Criminological factors include concrete elements (such as te presence of arms or drugs, or elements of the built environment), and also figurative factors, which include the soci costs of deviant behavior as well as family supervision.
-Accordingly, a number of studies have investigated various aspects of the environment and neighborhood design [e.g., @Foster2010neighbourhood; @He2017built; @Loukaitou2001measuring], whereas other studies have focused on exposure to environmental attributes that signal weakened norms, such as liquor and tobacco outlets [e.g., @Brower2007spatial; @Deryol2016crime; @Lipton2008spatial; @Quick2017influence].
-
-Yet another fruitful avenue for research, and one that has only recently begun to be explored, is the presence of environmental attributes that can help to reinforce moral norms, such as schools and churches [e.g., @Abdullah2018amenities; @Davignon2015christian; @Furr2010metric; @Traunmuller2011moral].
-It is thus that in a recent paper, Warner and Konkel [-@Warner2019neighborhood] note that the role of places of worship, as distinct entities from the members of the congregations, have received less attention in empirical and theoretical research for their deterrence potential.
-The role of these institutions might be particularly important in places where formal state institutions lack the means or the will to enforce norms - as is the case in Brazil.
-
-With the above considerations in mind, the objective of this paper is to investigate whether and how places of workship correlate spatially with criminal events, or in other words, to investigate whether their signals as moral communities have a discernible spatial dimension.
-The case study is the city of Recife in the state of Pernambuco, in Brazil's Northeast.
-Recife is a large and important metropolitan area in a historically poor region, and afflicted by high levels of violent crime, having the dubious honor of being one of five state capitals with the highest rates of homicide in the period under study [@Menezes2013spatial].
-The empirical strategy is to use spatial analysis to explore the potential geographical relationships between violent criminal events, on the one hand, and places of worship and a selection of commercial establishments that serve as controls, on the other.
-Dissagregated data allows us to analyze the relevant point patterns at a very high level of resolution.
-
-After this introductory section, the rest of this paper is structured as follows: Section \ref{background} reviews some relevant theoretical perspectives on religion and crime; then Section \ref{methods} presents the empirical strategy used in the work.
-Section \ref{case} presents the context as well as the data used in the study; Section \ref{results} presents the analysis, whereas the results are discussed in Section \ref{discussion}.
-Finally, some concluding remarks are offered in Section \ref{conclusions}.
-
-# Background: Theoretical Perspectives on Religion and Crime {#background}
-
-Why do humans behave morally?
-For millennia, it has been the role of religion to provide the basic tenets of morally acceptable behavior: thou shalt not kill et al., and their equivalent in many world religions [e.g., @Donovan1986different].
-Enforcement of such tenets implies different mechanisms, including _sin_, _haram_, _karma_, and _tapu_, with punishment delivered by hellfire and exile to _Jahannam_ or _Gehenna_, to mention just some choice places of torment.
-In addition to acting as _de facto_ social policy for much of history, many of these religious tenets still have the force of law in many cultures and regions - and even where they do not, they are held by some researchers and policy experts to be helpful complements to reduce crime in any case [e.g., @Durrant2017religion; @Johnson2011more].
-The hypothesis that religion can act as a factor that deters and reduces criminal behavior, therefore, has prompted the scientific study of the effectiveness of religion on moral behavior [@Hoffmann2015religion].
-
-An early effort to theorize the effect of religion on criminal behavior was the _hellfire hypothesis_ of Hirschi and Stark [-@Hirschi1969hellfire].
-The focus of this hypothesis is the threat of extratemporal (and possibly eternal) punishment, and how this threat can deter believers from commandment-breaking (also see Pascal's wager).
-Accordingly, Hirsch and Stark [-@Hirschi1969hellfire] posit that negative correlations between crime and religion are a consequence of a sense of commitment with normative values, a commitment that is ritually reinforced in a regular fashion (e.g., the liturgical rite of peace in Catholic Mass).
-Rohrbaugh and Jessor [-@Rohrbaugh1975religiosity], for instance, note that by attributing to divinity (or some other supra-terrenal entity) the supreme force of punishment, religion helps to build a sense of obduracy, or "hardening" against temptation.
-
-Despite its intuitive appeal, research has not provided much direct evidence for the hypothesis of hellfire [@Hoffmann2015religion, p.
-1].
-In a bizarre twist, even mainstream doctrines (such as absolution) may in fact encourage beliefs that neutralize the fear of _terrenal_ punishment, and thus turn out to be criminogenic [e.g., @Topalli2013god].
-More deviant cases can even recast criminal behavior as a form of spiritual insurgency [@Chesnut2017devoted].
-Counterexamples like this notwithstanding, researchers remain open to the possibility that "religion may still serve as a social control mechanism by encouraging conventional beliefs, monitoring behaviors, enhancing family attachments, or providing conventional activities" [@Hoffmann2015religion, p.
-1].
-In this way, the hypothesis of _moral communities_ [@Stark1996religion] recognizes that social integration is essential for increasing social control, thus reducing the practice of behavior that is not in accordance with current norms.
-Rohrbaugh and Jessor [-@Rohrbaugh1975religiosity] also emphasize that religion acts as social control since it defines what is an appropriate attitude according to moral values, thus making moral communities a close relative of the concept of _informal social control_, defined as “the ability of social groups or institutions to make norms or rules effective” [@Reiss1951delinquency, p.
-196; cited in @Groff2015informal, p.
-91].
-
-The hypothesis of moral communities has over the years been used to examine a variety of outcomes of interest.
-Recent examples include Stroope and Baker's [-@Stroope2018moral] exploration of religiosity and self-rated health; Davignon and Thomson [-@Davignon2015christian] with their research on institutional context and the religiosity of students; and the study of religion as a source of trust of Traunmuller [-@Traunmuller2011moral].
-This is in addition, of course, to numerous studies on criminal behavior such as Eitle [-@Eitle2011religion], an author who explored the deterrence power of religion on gambling; Lee and Bartkowski's [-@Lee2004love] investigation of juvenile homicide in rural areas; and the research of Regnerus [-@Regnerus2003moral] on adolescent delinquency.
-The hypothesis of moral communities as a form of informal social control, on the other hand, has been less studied from a geographical perspective, and it is only recently that has attracted the attention of researchers.
-Groff [-@Groff2015informal], for instance, discusses informal social control as a phenomenon that can plaussibly operate at different geographical scales, from the level of the home and family, through the street block and neighborhood, possibly to other scales such as the county.
-In this way, Nie and Yang [-@Nie2019smoking] remark on a recent paper on the lack of research conducted to study how the religious context of a geographical area (e.g., a county) may influence (youth smoking) behavior (p.
-2).
-This point is echoed by Warner and Konkel [-@Warner2019neighborhood], who moreover recommend the use of even smaller units of analysis (e.g., Census Block Groups), in an effort to reduce aggregation bias [see @Hipp2007block], but more importantly because the processes defined by social disorganization theory are thought to occur at relatively small geographical scales.
-
-The preceding review makes clear the need for research at various geographical scales.
-From a theoretical perspective, the social mechanisms that underpin the hypothesis of moral communities and informal social control can happen at various geographical scales, some of which previous research has already addressed: the work of Nie and Yang [-@Nie2019smoking], for one, highlights the role of processes detectable at a relatively high level of aggregation, whereas Warner and Konkel [-@Warner2019neighborhood], looking specifically at the effect of places of worship, bring the study closer to the level of the neighborhood.
-In what follows, we propose to adopt a much more disaggregated approach by considering the intensity of criminal events with respect to places of worship.
-This requires micro-level data both of crimes and places of worship, and the use of appropriate spatial analytical tools for point patterns, as discussed next.
-
-# Empirical Strategy {#methods}
-
-Two different techniques are deployed in this paper to study the spatial distribution of criminal events, and their possible relationship to the location of places of worship.
-Firstly, the F-function is a nearest neighbor technique that is used to summarize the distance between a set of points and their nearest events.
-This technique has found extensive use in spatial epidemiology (see Gatrell et al., 1996) and has also been used in spatial criminology (see, inter alia, Craglia et al., 2000; Kiani et al., 2015; Levine, 2006; and Rogerson and Sun, 2001).
-The second technique is the scan statistic.
-This method has also been extensively used in spatial epidemiology (e.g., Hjalmars et al., 1996) and has found use in spatial criminology as well (see He et al., 2017a; Nakaya and Yano, 2010; Malleson and Andresen, 2015; and Shiode, 2011).
-Use of these two techniques is meant to provide insights with respect to 1) the location of violent criminal events with respect to various types of facilities; and 2) geographically identify places of worship associated with clusters of criminal events.
-These two techniques are briefly reviewed next.
-
-## Quadrat Analysis
-
-Words go here.
-
-## Relative Intensity
-
-Words go here.
-
-# Case Study {#case}
-
-## Context {#context}
-
-The study is of the city of Recife, the capital of the state of Pernambuco in the Northeast region of Brazil.
-With a population of 1,550,390 million inhabitants in 2010, and with an area of 7,133.2 km2, Recife is one of the main Brazilian metropolises, exerting a great economic influence in neighboring regions.
-However, the city experiences a serious problem with violent crime, and has the dubious honor of being one of the five capital cities in Brazil with the highest homicide rates in the period under study [@Menezes2013spatial].
-
-In the current context, the term "violent crime" is an umbrella for several forms of infractions to the penal code.
-Following recommendations of the National Secretariat of Public Security of the Ministry of Justice of 2006, these are Violent Lethal and Intentional Crimes (VLIC; which includes intentional homicide), theft followed by death (robbery), and corporal injury followed by death.
-The data on CVLI were extracted from the Police Information System of the Secretariat of Social Defense of Pernambuco (INFOPOL / SDS-PE), which is the most reliable, detailed, and comprehensive information on violent deaths in the region.
-
-```{r descriptive-statistics, echo=FALSE}
-descriptive_stats <- data.frame(Attribute = c("Man",
-                                              "Woman",
-                                              "Murder",
-                                              "Robbery",
-                                              "Body injury followed by death",
-                                              "Black and White",
-                                              "Yellow and White",
-                                              "Not Reported",
-                                              "1 to 12 years",
-                                              "13 to 17 years",
-                                              "18 to 30 years",
-                                              "31 to 65 years",
-                                              "65 years and older",
-                                              "Not Reported",
-                                              "Firearm",
-                                              "Other"),
-                                Percentage = c(91.35,
-                                               8.65,
-                                               97.82,
-                                               2.00,
-                                               0.18,
-                                               95.11,
-                                               1.27,
-                                               3.62,
-                                               0.18,
-                                               11.31,
-                                               61.70,
-                                               25.84,
-                                               0.73,
-                                               0.24,
-                                               87.51,
-                                               12.49))
-
-kable(descriptive_stats, 
-      "latex", 
-      booktabs = TRUE,
-      digits = 3,
-      caption = "\\label{tab:descriptive-statistics}Characteristics of Violent Crime in Recife, July, 2008 - June, 2009",
-      col.names = c("", "Percentage")) %>%
-  kable_styling() %>%
-  pack_rows("Gender of Victim", 1, 2) %>%
-  pack_rows("Type of Crime", 3, 5) %>%
-  pack_rows("Ethnicity of Victim", 6, 8) %>%
-  pack_rows("Age of Victim", 9, 14) %>%
-  pack_rows("Weapon Used", 15, 16)
-```
-
-## Data {#data}
-
-The data are organized at the individual level and it is possible to obtain information about the location, day of the week, day of the month, period of the day, as well as gender, age, and race of the victim.
-The database used in this study comprises the period from July 1, 2008 to June 30, 2010.
-Some descriptive statistics regarding this dataset are reported in Table 1, where it can be seen that overall, about 91% of the victims of CVLI in the period analyzed were men, while approximately 98% of violent crimes were homicides.
-In addition, most of the victims were black or brown, and the youth population between the ages of 18 and 30 is the most affected by violent crime.
-Lastly, it should be noted that about 88% of the criminal events under analysis involved firearms.
-Information about places of interest was obtained from the National Register of Addresses for Statistical Purposes (CNEFE - Census 2010), which lists 78,056,411 urban and rural addresses, distributed among the 316,574 census tracts.
-This is the first database of its kind produced by IBGE, and the first version was produced at the time of the 2000 Census.
-The way addresses are described in the National Register is very rich, and it is possible to identify the names of the places of worship including their denomination.
-Georeferencing was used to geolocate each place of worship.
-In this way, a total of 1,719 places of worship were geolocated in the city of Recife.
-Figure 1 (Panel a) shows the spatial distribution of the 1,657 CVLI crimes that occurred in the city of Recife between 2008.2 and 2010.1.
-Panel b) of the figure shows the spatial distribution of places of worship in the region.
-
-```{r plot-crime, echo=FALSE, fig.align = "left", fig.cap="\\label{fig:plot-crime}Lethal and Intentional Violent Crime in Recife"}
-
-#We can experiment with plotting the events:
-
-ggplot(data = events_recife.sf %>% filter(Type == "Violent Crime")) + 
-  geom_sf(color = "red") +
-  geom_sf(data = recife.sf, fill = NA) + 
-  theme_tufte()
-```
-
-In addition to places of worship, the National Register of Addresses for Statistical Purposes was queried to extract facilities other than places of worship.
-As discussed above, the idea is to identify points of reference that can be used as controls, having a neutral morality profile.
-For the sake of the present study, we selected pharmacies, ice cream shops, and bakeries to construct our control group.
-These three types of establishments comply with the criteria of being morally neutral and having a spatial distribution commensurate with places of worship.
-Figure 2 shows the spatial distribution of the control establishments, namely pharmacies, bakeries, and ice cream shops in the city of Recife.
-Note the similarity between the maps.
-As expected, there are differences in the number of points, but the locations of cases and controls are quite similar.
-
-```{r plot-events, echo=FALSE, fig.align = "left", fig.cap="\\label{fig:plot-events}Places of Worship and Commercial Establishments in Recife"}
-
-#We can experiment with plotting the events:
-
-ggplot(data = events_recife.sf %>% filter(Type != "Violent Crime")) + 
-  geom_sf(aes(color = Type, fill = Type)) +
-  geom_sf(data = recife.sf, fill = NA) + 
-  facet_wrap(~ Type, nrow = 2) +
-  scale_fill_viridis_d(option = "viridis", direction = 1) +
-  scale_color_viridis_d(option = "viridis", direction = 1) +
-  theme_tufte()  +
-  theme(axis.text.x = element_text(angle = 90))
-
-```
-
-```{r summary-point-patterns, include=FALSE}
-
-# Extract the intensity of each point pattern for the study area
-
-summary_crime <- summary(events_recife_list$`Violent Crime`)
-summary_catholic <- summary(events_recife_list$Catholic)
-summary_evangelic <- summary(events_recife_list$Evangelical)
-summary_spiritist <- summary(events_recife_list$Spiritist)
-summary_ice_cream <- summary(events_recife_list$`Ice Cream Shop`)
-summary_pharmacy <- summary(events_recife_list$Pharmacy)
-summary_supermarket <- summary(events_recife_list$Supermarket)
-```
-
-```{r plot-sed, echo=FALSE, fig.align = "left", fig.cap="\\label{fig:plot-sed}Socio-Economic Deprivation in Recife"}
-ggplot(data = setores.sf) + 
-  geom_sf(aes(fill = SED),
-          color = NA) +
-  scale_fill_viridis_c(option = "plasma", direction = 1, values = c(0, 0.05, 0.2, 0.3, 0.4, 1)) +
-  theme_tufte()
-```
-
-# Analysis and results {#results}
-
-## Quadrat Analysis
-
-```{r rasterize-sed, include=FALSE}
-
-# Rasterize and convert to image:
-
-# First, we need a base raster; this will be the boundaries of the study area.
-After some trial and error with different values of number of rows and columns, these values create pixels that are as close to squares as possible, while also being relatively small
-
-recife.r <- raster(recife.sf, nrows = 400, ncols = 276)
-
-# The raster `recife.r` is the template for rasterizing other items, in this case socio-economic deprivation. Use package `fasterize` to rasterize an `sf` object.
-
-sed.r <- fasterize(setores.sf, recife.r, field = "SED")
-
-# Once that the object is a raster, it can be converted to a `spatstat` `img` class object:
-sed.img <- as.im(sed.r)
-```
-
-```{r quadrat-analysis-preliminaries, include=FALSE}
-
-# To do quadrat analysis by different levels of SED we need the quintiles of this variable (other criteria could be used to slice the distribution, e.g., standard deviations, but this one is intuitive and simple)
-
-sed5 <- quantile(sed.img, c(0, 0.2, 0.4, 0.6, 0.8, 1))
-
-# Next we can cut the `sed.im` using the quintiles:
-
-sed_cut5.img <- cut(sed.img, breaks = sed5, include.lowest = TRUE, labels = c(1:5))
-```
-
-```{r tesselate-sed5, include=FALSE}
-
-# To do quadrat analysis we need to tesellate the cut image:
-
-V <- tess(image = sed_cut5.img)
-
-# This object is used for the analysis, but it is not very neat for plotting purposes. Lets convert instead to an `sf` object. First, extract and expand the coordinates:
-
-coords_V <- expand.grid(X = V$image$xcol, Y = V$image$yrow)
-
-# Next, extract the values of the cells and convert to a vector:
-
-sed5_V <- as.vector(t(V$image$v))
-
-# Join the coordinates with the values in a dataframe and drop all NAs, before converting to `sf` and adding coordinate referencing system information (crs) to ensure that its projection information is there and compatible with the rest of `sf` objects (i.e., recife.sf, setores.sf, events_recife.sf):
-
-sed5.sf <- data.frame(coords_V, sed5_V) %>% 
-  drop_na() %>% transmute(X, Y, SED = factor(sed5_V, 
-                                              levels = c(1:5),
-                                              labels = c("Low", 
-                                                         "Medium Low", 
-                                                         "Medium", 
-                                                         "Medium High", 
-                                                         "High"))) %>%
-  st_as_sf(coords = c("X", "Y")) %>%
-  st_set_crs(value = st_crs(recife.sf))
-```
-
-```{r plot-sed-as-quintiles, echo=FALSE, cache = TRUE, fig.cap="\\label{fig:plot-sed-as-quintiles}Socio-Economic Deprivation as Quintiles" }
-ggplot(data = sed5.sf) + 
-  geom_sf(aes(fill = SED, color = SED)) +
-  geom_sf(data = recife.sf, fill = NA) +
-  scale_fill_viridis_d(option = "plasma", direction = 1) +
-  scale_color_viridis_d(option = "plasma", direction = 1) +
-  guides(fill = guide_legend(reverse = TRUE),
-         color = guide_legend(reverse = TRUE)) +
-  theme_tufte()
-```
-
-Calculate the intensity of other ppp by quadrat:
-```{r quadrat-analysis, include=FALSE}
-
-# Calculate the intensity of the various point patterns by quadrat:
-
-# Crime
-
-crime5 <- quadratcount(events_recife_list$`Violent Crime`, tess = V)
-crime5_intensity <- intensity(crime5)
-
-# Catholic
-
-catholic5 <- quadratcount(events_recife_list$Catholic, tess = V)
-catholic5_intensity <- intensity(catholic5)
-
-# Evangelical
-
-evangelic5 <- quadratcount(events_recife_list$Evangelical, tess = V)
-evangelic5_intensity <- intensity(evangelic5)
-
-# Spiritist
-
-spiritist5 <- quadratcount(events_recife_list$Spiritist, tess = V)
-spiritist5_intensity <- intensity(spiritist5)
-
-# Ice cream
-
-ice_cream5 <- quadratcount(events_recife_list$`Ice Cream Shop`, tess = V)
-ice_cream5_intensity <- intensity(ice_cream5)
-
-# Pharmacies
-
-pharmacies5 <- quadratcount(events_recife_list$Pharmacy, tess = V)
-pharmacies5_intensity <- intensity(pharmacies5)
-
-# Supermarkets
-
-super_markets5 <- quadratcount(events_recife_list$Supermarket, tess = V)
-supermarkets5_intensity <- intensity(super_markets5)
-```
-
-```{r instensity-dataframe, include=FALSE}
-
-# Collect in a data.frame for convenience of plotting:
-
-intensity_all <- data.frame(SED = rep(c("Low", 
-                                        "Middle Low", 
-                                        "Middle", 
-                                        "Middle High", 
-                                        "High"), 
-                                      times = 7),
-                            Type = rep(c("Violent Crime", 
-                                         "Catholic",
-                                         "Evangelical",
-                                         "Spiritist",
-                                         "Ice Cream",
-                                         "Pharmacies",
-                                         "Super Markets"), 
-                                       each = 5),
-                            Intensity = c(crime5_intensity,
-                                          catholic5_intensity,
-                                          evangelic5_intensity,
-                                          spiritist5_intensity,
-                                          ice_cream5_intensity,
-                                          pharmacies5_intensity,
-                                          supermarkets5_intensity),
-                            Global_Intensity = rep(c(summary_crime$intensity,
-                                                     summary_catholic$intensity,
-                                                     summary_evangelic$intensity,
-                                                     summary_spiritist$intensity,
-                                                     summary_ice_cream$intensity,
-                                                     summary_pharmacy$intensity,
-                                                     summary_supermarket$intensity),
-                                                   each = 5)) %>%
-  mutate(SED = factor(SED, levels = c("Low", 
-                                        "Middle Low", 
-                                        "Middle", 
-                                        "Middle High", 
-                                        "High"),
-                      ordered = TRUE),
-         Type = factor(Type, levels = c("Violent Crime", 
-                                        "Catholic",
-                                        "Evangelical",
-                                        "Spiritist",
-                                        "Ice Cream",
-                                        "Pharmacies",
-                                        "Super Markets")))
-```
-
-```{r plot-crime-quadrat, echo=FALSE, fig.width=3, fig.asp=1 , cache = TRUE, fig.cap="\\label{fig:plot-crime-quadrat}Intensity of crime by level of Socio-Economic Deprivation; the dotted line indicates the global intensity of crime"}
-
-#Plot the intensity of crime:
-
-ggplot(data = intensity_all %>% filter(Type == "Violent Crime"), 
-       aes(x = SED, y = Intensity, fill = Type)) + 
-  geom_col(fill = "red") +
-  xlab("Socio-Economic Deprivation") +
-  geom_hline(aes(yintercept = Global_Intensity), linetype = "dotted", size = 1) +
-  #facet_wrap(Type ~.) +
-  theme_tufte() +
-  theme(axis.text.x = element_text(angle = 90))
-```
-
-```{r plot-events-quadrat, echo=FALSE, cache = TRUE, fig.cap="\\label{fig:plot-events-quadrat}Intensity of places of worship and commercial establishments by level of Socio-Economic Deprivation; the dotted lines indicates the respective global intensities"}
-
-#Plot the intensity of point patterns:
-
-ggplot(data = intensity_all %>% filter(Type != "Violent Crime"), 
-       aes(x = SED, y = Intensity, fill = Type)) + 
-  geom_col() +
-  xlab("Socio-Economic Deprivation") +
-  geom_hline(aes(yintercept = Global_Intensity), linetype = "dotted", size = 1) +
-  scale_fill_viridis_d(option = "viridis", direction = 1) +
-  scale_color_viridis_d(option = "viridis", direction = 1) +
-  facet_wrap(Type ~.) +
-  theme_tufte() +
-  theme(axis.text.x = element_text(angle = 90))
-```
-
-## Relative Intensity
-
-```{r distance-functions, include=FALSE, cache = TRUE}
-
-# Estimation of the relative intensity (see Baddeley et al. (2015) Spatial Point Pattern, pp. 179-183).
-
-# This needs that we convert the point patterns for places of worship to distance functions.
-This way, the covariates for the relative distribution estimate are distance estimates to places of worship.
-The function `distfun` is used for this:
-
-z_catholic <- distfun(events_recife_list$Catholic)
-z_evangelical <- distfun(events_recife_list$Evangelical)
-z_spiritist <- distfun(events_recife_list$Spiritist)
-z_ice_cream <- distfun(events_recife_list$`Ice Cream Shop`)
-z_pharmacies <- distfun(events_recife_list$Pharmacy)
-z_super_markets <- distfun(events_recife_list$Supermarket)
-```
-
-# Discussion {#discussion}
-
-An opposing viewpoint is offered by the theory of conservative Protestants [@Ellison2003enduring].
-This theory argues that religion is not always beneficial in order to guarantee social order, and consequently to result in lower crime rates.
-This theory suggested that, conservative Protestants are more tolerate with violent behavior when associated with defense of honor, family, property, or woman.
-However, high rates are not expected for all types of crimes in densely populated areas of conservative Protestants.
-But what justifies this behavior?
-Ellison et al.
-[-@Ellison2003enduring] argue that this tolerance stems from a religious view of the world.
-More clearly, punishment is seen as a legitimate form of retribution whose explanation can be attributed to the literal interpretation of the Bible.
-In the specific context of Brazil, it is important to consider that the interpretations and meanings attributed to urban violence during the last two decades in Brazil often had religious figures as protagonists [@Birman2012violencia].
-
-# Conclusions {#conclusions}
-
-The theory of conservative Protestants, argues that conservative Protestants are more likely to tolerate violent behavior of a defensive or punitive character, especially when associated to defense of honor, family, property, or woman.
-Our objective in the study is to explore the question whether moral communities have a spatial dimension.
-To this end we deploy an array of spatial analytical tools to investigate potential patterns of clustering of violent crime and places of worship, using as well other kinds of urban facilities to act as controls.
-The case study is the city of Recife, in the northeast of Brazil, and analysis is based on geographically detailed datasets of criminal events as well as several types of urban facilities.
-
-# References

### 2019-10-15 Moral_Communities_and_Crime_Title_Page_v1.tex
First complete draft of paper and title page

--- 2019-10-13 Moral_Communities_and_Crime_v1.tex
@@
 
-\usepackage{booktabs}
-\usepackage{longtable}
-\usepackage{array}
-\usepackage{multirow}
-\usepackage{wrapfig}
-\usepackage{float}
-\usepackage{colortbl}
-\usepackage{pdflscape}
-\usepackage{tabu}
-\usepackage{threeparttable}
-\usepackage{threeparttablex}
-\usepackage[normalem]{ulem}
-\usepackage{makecell}
-\usepackage{xcolor}
-
 \begin{document}
@@
 
-\author{  Aye B. Cee \and  Dee E. Eff \and  Gee H. Eye \and  Jay K. Elle \and  }
+\author{  Edivaldo Constantino das Neves Júnior \and  Breno Caldas de Araujo \and  Tatiane Almeida de Menezes \and  Antonio Paez \and  }
 
 \institute{
-        Aye B. Cee \at
+        Edivaldo Constantino das Neves Júnior \at
      A Department, University of XXX \\
-     \email{\href{mailto:abc@example.edu}{\nolinkurl{abc@example.edu}}}  %  \\
+     \email{\href{mailto:edivaldoconstantino@gmail.com}{\nolinkurl{edivaldoconstantino@gmail.com}}}  %  \\
 %             \emph{Present address:} of F. Author  %  if needed
     \and
-        Dee E. Eff \at
+        Breno Caldas de Araujo \at
      A Department, University of XXX \\
@@
     \and
-        Gee H. Eye \at
-     A Department, University of XXX \\
-     \email{\href{mailto:ghi@example.edu}{\nolinkurl{ghi@example.edu}}}  %  \\
+        Tatiane Almeida de Menezes \at
+     Programa de Pós-Graduação em Gestão e Economia da Saúde, Universidade
+ Federal de Pernambuco \\
+     \email{\href{mailto:tatianedemenezes@gmail.com}{\nolinkurl{tatianedemenezes@gmail.com}}}  %  \\
 %             \emph{Present address:} of F. Author  %  if needed
     \and
-        Jay K. Elle \at
-     Some School, University of WWW \\
-     \email{\href{mailto:jkl@school.edu}{\nolinkurl{jkl@school.edu}}}  %  \\
+        Antonio Paez \at
+     School of Geography and Earth Sciences, McMaster University \\
+     \email{\href{mailto:paezha@mcmaster.ca}{\nolinkurl{paezha@mcmaster.ca}}}  %  \\
 %             \emph{Present address:} of F. Author  %  if needed
@@
 
-\hypertarget{intro}{%
-\section{Introduction}\label{intro}}
-
-Violent crime is a widespread phenomenon with negative impacts on many
-spheres of life and society. Although the rate of homicides worldwide
-has grown at a slower rate than the population in the past few decades,
-the number of people killed in homicides still increased from 362,000 in
-1990 to 464,000 in 2017 (United Nations Office on Drugs and Crime
-2019a). Furthermore, variations in the prevalence of violent crime tend
-to be extremely uneven. The Americas, with a population of approximately
-793.8 million people in 2019 (or about 10.3\% of the world population),
-accounted for approximately 173,000 homicides in 2017, or approximately
-37.3\% of all homicides in the world. Within the Americas, Brazil (along
-with Venezuela, Colombia, and Mexico) is one of the largest countries in
-the region with high homicide rates. And there are notable variations at
-the sub-national and intra-metropolitan levels as well. In fact,
-understanding the variations in the prevalence of violent crime is
-recognized as key to achieving policy goals:
-
-\begin{quote}
-\begin{quote}
-High levels of homicidal violence are concentrated in geographic and
-demographic ``pockets'', so achieving target 16.1 of the Sustainable
-Development Goals requires interventions within the specific regions,
-countries, communities and population groups that are most at risk
-(United Nations Office on Drugs and Crime 2019a, 35)
-\end{quote}
-\end{quote}
-
-This concern is not iddle: the endemic malady of violent crime
-represents an important factor that threatens to derail sustainable
-development goals. Violent crime, furthermore, tends to be more accute
-in those regions most in need of development. On the one hand, violent
-crime represents an economic and social drag (Becker and Kassouf 2017).
-The impacts are multifaceted, since they affect people's well-being,
-either through the loss of human life, mental health, and limitations in
-the right to public spaces (Doran and Burgess 2011), or through
-disturbances in schooling and academic achievement (United Nations
-Office on Drugs and Crime 2019b). In turn, these negative effects
-combine to create the unfortunate conditions that tend to breed criminal
-behavior, thus creating a vicious cycle of economic disadvantage and
-violence (United Nations Office on Drugs and Crime 2019b). It is not
-surprising, given these pecularities, that researchers have joined calls
-for increased attention to the study of the patterns of violent crime in
-Low and Middle Income Countries (LMIC; Murray, Cerqueira, and Kahn
-2013).
-
-Added to the scenario, formal social control institutions in countries
-like Brazil leave something (or much) to be desired: while in practice
-they should inhibit criminal behavior, they suffer from deep
-deficiencies that end up reducing the deterrent power of the justice
-system. Serious institutional problems include the inefficiency of the
-police, the lack of national legislation, the glacial pace of judicial
-processes, and the weak situation of the prison system in the country
-(Menezes et al. 2013). This makes it even more urgent to understand the
-role of \emph{informal social controls}, i.e., the ability of community
-organizations such as schools, clubs, and neighborbood associations to
-suppress crime by strengthening the capabilities of neighbors to control
-inappropriate behavior (Groff 2015).
-
-In order to understand the factors that can potentially help to deter
-crime, beyond formal control, it is important to identify empirical
-regularities. Criminological factors include concrete elements (such as
-te presence of arms or drugs, or elements of the built environment), and
-also figurative factors, which include the soci costs of deviant
-behavior as well as family supervision. Accordingly, a number of studies
-have investigated various aspects of the environment and neighborhood
-design (e.g., Foster, Giles-Corti, and Knuiman 2010; He, Páez, and Liu
-2017; Loukaitou-Sideris et al. 2001), whereas other studies have focused
-on exposure to environmental attributes that signal weakened norms, such
-as liquor and tobacco outlets (e.g., Brower and Carroll 2007; Deryol et
-al. 2016; Lipton et al. 2008; Quick, Law, and Luan 2017).
-
-Yet another fruitful avenue for research, and one that has only recently
-begun to be explored, is the presence of environmental attributes that
-can help to reinforce moral norms, such as schools and churches (e.g.,
-Abdullah et al. 2018; Davignon and Thomson 2015; Furr-Holden et al.
-2010; Traunmuller 2011). It is thus that in a recent paper, Warner and
-Konkel (2019) note that the role of places of worship, as distinct
-entities from the members of the congregations, have received less
-attention in empirical and theoretical research for their deterrence
-potential. The role of these institutions might be particularly
-important in places where formal state institutions lack the means or
-the will to enforce norms - as is the case in Brazil.
-
-With the above considerations in mind, the objective of this paper is to
-investigate whether and how places of workship correlate spatially with
-criminal events, or in other words, to investigate whether their signals
-as moral communities have a discernible spatial dimension. The case
-study is the city of Recife in the state of Pernambuco, in Brazil's
-Northeast. Recife is a large and important metropolitan area in a
-historically poor region, and afflicted by high levels of violent crime,
-having the dubious honor of being one of five state capitals with the
-highest rates of homicide in the period under study (Menezes et al.
-2013). The empirical strategy is to use spatial analysis to explore the
-potential geographical relationships between violent criminal events, on
-the one hand, and places of worship and a selection of commercial
-establishments that serve as controls, on the other. Dissagregated data
-allows us to analyze the relevant point patterns at a very high level of
-resolution.
-
-After this introductory section, the rest of this paper is structured as
-follows: Section \ref{background} reviews some relevant theoretical
-perspectives on religion and crime; then Section \ref{methods} presents
-the empirical strategy used in the work. Section \ref{case} presents the
-context as well as the data used in the study; Section \ref{results}
-presents the analysis, whereas the results are discussed in Section
-\ref{discussion}. Finally, some concluding remarks are offered in
-Section \ref{conclusions}.
-
-\hypertarget{background}{%
-\section{Background: Theoretical Perspectives on Religion and
-Crime}\label{background}}
-
-Why do humans behave morally? For millennia, it has been the role of
-religion to provide the basic tenets of morally acceptable behavior:
-thou shalt not kill et al., and their equivalent in many world religions
-(e.g., Donovan 1986). Enforcement of such tenets implies different
-mechanisms, including \emph{sin}, \emph{haram}, \emph{karma}, and
-\emph{tapu}, with punishment delivered by hellfire and exile to
-\emph{Jahannam} or \emph{Gehenna}, to mention just some choice places of
-torment. In addition to acting as \emph{de facto} social policy for much
-of history, many of these religious tenets still have the force of law
-in many cultures and regions - and even where they do not, they are held
-by some researchers and policy experts to be helpful complements to
-reduce crime in any case (e.g., Durrant and Poppelwell 2017; Johnson
-2011). The hypothesis that religion can act as a factor that deters and
-reduces criminal behavior, therefore, has prompted the scientific study
-of the effectiveness of religion on moral behavior (Hoffmann 2015).
-
-An early effort to theorize the effect of religion on criminal behavior
-was the \emph{hellfire hypothesis} of Hirschi and Stark (1969). The
-focus of this hypothesis is the threat of extratemporal (and possibly
-eternal) punishment, and how this threat can deter believers from
-commandment-breaking (also see Pascal's wager). Accordingly, Hirsch and
-Stark (1969) posit that negative correlations between crime and religion
-are a consequence of a sense of commitment with normative values, a
-commitment that is ritually reinforced in a regular fashion (e.g., the
-liturgical rite of peace in Catholic Mass). Rohrbaugh and Jessor (1975),
-for instance, note that by attributing to divinity (or some other
-supra-terrenal entity) the supreme force of punishment, religion helps
-to build a sense of obduracy, or ``hardening'' against temptation.
-
-Despite its intuitive appeal, research has not provided much direct
-evidence for the hypothesis of hellfire (Hoffmann 2015, 1). In a bizarre
-twist, even mainstream doctrines (such as absolution) may in fact
-encourage beliefs that neutralize the fear of \emph{terrenal}
-punishment, and thus turn out to be criminogenic (e.g., Topalli,
-Brezina, and Bernhardt 2013). More deviant cases can even recast
-criminal behavior as a form of spiritual insurgency (Chesnut 2017).
-Counterexamples like this notwithstanding, researchers remain open to
-the possibility that ``religion may still serve as a social control
-mechanism by encouraging conventional beliefs, monitoring behaviors,
-enhancing family attachments, or providing conventional activities''
-(Hoffmann 2015, 1). In this way, the hypothesis of \emph{moral
-communities} (Stark 1996) recognizes that social integration is
-essential for increasing social control, thus reducing the practice of
-behavior that is not in accordance with current norms. Rohrbaugh and
-Jessor (1975) also emphasize that religion acts as social control since
-it defines what is an appropriate attitude according to moral values,
-thus making moral communities a close relative of the concept of
-\emph{informal social control}, defined as ``the ability of social
-groups or institutions to make norms or rules effective'' (Reiss 1951,
-196; cited in Groff 2015, 91).
-
-The hypothesis of moral communities has over the years been used to
-examine a variety of outcomes of interest. Recent examples include
-Stroope and Baker's (2018) exploration of religiosity and self-rated
-health; Davignon and Thomson (2015) with their research on institutional
-context and the religiosity of students; and the study of religion as a
-source of trust of Traunmuller (2011). This is in addition, of course,
-to numerous studies on criminal behavior such as Eitle (2011), an author
-who explored the deterrence power of religion on gambling; Lee and
-Bartkowski's (2004) investigation of juvenile homicide in rural areas;
-and the research of Regnerus (2003) on adolescent delinquency. The
-hypothesis of moral communities as a form of informal social control, on
-the other hand, has been less studied from a geographical perspective,
-and it is only recently that has attracted the attention of researchers.
-Groff (2015), for instance, discusses informal social control as a
-phenomenon that can plaussibly operate at different geographical scales,
-from the level of the home and family, through the street block and
-neighborhood, possibly to other scales such as the county. In this way,
-Nie and Yang (2019) remark on a recent paper on the lack of research
-conducted to study how the religious context of a geographical area
-(e.g., a county) may influence (youth smoking) behavior (p.~2). This
-point is echoed by Warner and Konkel (2019), who moreover recommend the
-use of even smaller units of analysis (e.g., Census Block Groups), in an
-effort to reduce aggregation bias (see Hipp 2007), but more importantly
-because the processes defined by social disorganization theory are
-thought to occur at relatively small geographical scales.
-
-The preceding review makes clear the need for research at various
-geographical scales. From a theoretical perspective, the social
-mechanisms that underpin the hypothesis of moral communities and
-informal social control can happen at various geographical scales, some
-of which previous research has already addressed: the work of Nie and
-Yang (2019), for one, highlights the role of processes detectable at a
-relatively high level of aggregation, whereas Warner and Konkel (2019),
-looking specifically at the effect of places of worship, bring the study
-closer to the level of the neighborhood. In what follows, we propose to
-adopt a much more disaggregated approach by considering the intensity of
-criminal events with respect to places of worship. This requires
-micro-level data both of crimes and places of worship, and the use of
-appropriate spatial analytical tools for point patterns, as discussed
-next.
-
-\hypertarget{methods}{%
-\section{Empirical Strategy}\label{methods}}
-
-Two different techniques are deployed in this paper to study the spatial
-distribution of criminal events, and their possible relationship to the
-location of places of worship. Firstly, the F-function is a nearest
-neighbor technique that is used to summarize the distance between a set
-of points and their nearest events. This technique has found extensive
-use in spatial epidemiology (see Gatrell et al., 1996) and has also been
-used in spatial criminology (see, inter alia, Craglia et al., 2000;
-Kiani et al., 2015; Levine, 2006; and Rogerson and Sun, 2001). The
-second technique is the scan statistic. This method has also been
-extensively used in spatial epidemiology (e.g., Hjalmars et al., 1996)
-and has found use in spatial criminology as well (see He et al., 2017a;
-Nakaya and Yano, 2010; Malleson and Andresen, 2015; and Shiode, 2011).
-Use of these two techniques is meant to provide insights with respect to
-1) the location of violent criminal events with respect to various types
-of facilities; and 2) geographically identify places of worship
-associated with clusters of criminal events. These two techniques are
-briefly reviewed next.
-
-\hypertarget{quadrat-analysis}{%
-\subsection{Quadrat Analysis}\label{quadrat-analysis}}
-
-Words go here.
-
-\hypertarget{relative-intensity}{%
-\subsection{Relative Intensity}\label{relative-intensity}}
-
-Words go here.
-
-\hypertarget{case}{%
-\section{Case Study}\label{case}}
-
-\hypertarget{context}{%
-\subsection{Context}\label{context}}
-
-The study is of the city of Recife, the capital of the state of
-Pernambuco in the Northeast region of Brazil. With a population of
-1,550,390 million inhabitants in 2010, and with an area of 7,133.2 km2,
-Recife is one of the main Brazilian metropolises, exerting a great
-economic influence in neighboring regions. However, the city experiences
-a serious problem with violent crime, and has the dubious honor of being
-one of the five capital cities in Brazil with the highest homicide rates
-in the period under study (Menezes et al. 2013).
-
-In the current context, the term ``violent crime'' is an umbrella for
-several forms of infractions to the penal code. Following
-recommendations of the National Secretariat of Public Security of the
-Ministry of Justice of 2006, these are Violent Lethal and Intentional
-Crimes (VLIC; which includes intentional homicide), theft followed by
-death (robbery), and corporal injury followed by death. The data on CVLI
-were extracted from the Police Information System of the Secretariat of
-Social Defense of Pernambuco (INFOPOL / SDS-PE), which is the most
-reliable, detailed, and comprehensive information on violent deaths in
-the region.
-
-\begin{table}
-
-\caption{\label{tab:descriptive-statistics}\label{tab:descriptive-statistics}Characteristics of Violent Crime in Recife, July, 2008 - June, 2009}
-\centering
-\begin{tabular}[t]{lr}
-\toprule
- & Percentage\\
-\midrule
-\addlinespace[0.3em]
-\multicolumn{2}{l}{\textbf{Gender of Victim}}\\
-\hspace{1em}Man & 91.35\\
-\hspace{1em}Woman & 8.65\\
-\addlinespace[0.3em]
-\multicolumn{2}{l}{\textbf{Type of Crime}}\\
-\hspace{1em}Murder & 97.82\\
-\hspace{1em}Robbery & 2.00\\
-\hspace{1em}Body injury followed by death & 0.18\\
-\addlinespace[0.3em]
-\multicolumn{2}{l}{\textbf{Ethnicity of Victim}}\\
-\hspace{1em}Black and White & 95.11\\
-\hspace{1em}Yellow and White & 1.27\\
-\hspace{1em}Not Reported & 3.62\\
-\addlinespace[0.3em]
-\multicolumn{2}{l}{\textbf{Age of Victim}}\\
-\hspace{1em}1 to 12 years & 0.18\\
-\hspace{1em}13 to 17 years & 11.31\\
-\hspace{1em}18 to 30 years & 61.70\\
-\hspace{1em}31 to 65 years & 25.84\\
-\hspace{1em}65 years and older & 0.73\\
-\hspace{1em}Not Reported & 0.24\\
-\addlinespace[0.3em]
-\multicolumn{2}{l}{\textbf{Weapon Used}}\\
-\hspace{1em}Firearm & 87.51\\
-\hspace{1em}Other & 12.49\\
-\bottomrule
-\end{tabular}
-\end{table}
-
-\hypertarget{data}{%
-\subsection{Data}\label{data}}
-
-The data are organized at the individual level and it is possible to
-obtain information about the location, day of the week, day of the
-month, period of the day, as well as gender, age, and race of the
-victim. The database used in this study comprises the period from July
-1, 2008 to June 30, 2010. Some descriptive statistics regarding this
-dataset are reported in Table 1, where it can be seen that overall,
-about 91\% of the victims of CVLI in the period analyzed were men, while
-approximately 98\% of violent crimes were homicides. In addition, most
-of the victims were black or brown, and the youth population between the
-ages of 18 and 30 is the most affected by violent crime. Lastly, it
-should be noted that about 88\% of the criminal events under analysis
-involved firearms. Information about places of interest was obtained
-from the National Register of Addresses for Statistical Purposes (CNEFE
-- Census 2010), which lists 78,056,411 urban and rural addresses,
-distributed among the 316,574 census tracts. This is the first database
-of its kind produced by IBGE, and the first version was produced at the
-time of the 2000 Census. The way addresses are described in the National
-Register is very rich, and it is possible to identify the names of the
-places of worship including their denomination. Georeferencing was used
-to geolocate each place of worship. In this way, a total of 1,719 places
-of worship were geolocated in the city of Recife. Figure 1 (Panel a)
-shows the spatial distribution of the 1,657 CVLI crimes that occurred in
-the city of Recife between 2008.2 and 2010.1. Panel b) of the figure
-shows the spatial distribution of places of worship in the region.
-
-\begin{figure}
-
-\includegraphics{Moral_Communities_and_Crime_v1_files/figure-latex/plot-crime-1} \hfill{}
-
-\caption{\label{fig:plot-crime}Lethal and Intentional Violent Crime in Recife}\label{fig:plot-crime}
-\end{figure}
-
-In addition to places of worship, the National Register of Addresses for
-Statistical Purposes was queried to extract facilities other than places
-of worship. As discussed above, the idea is to identify points of
-reference that can be used as controls, having a neutral morality
-profile. For the sake of the present study, we selected pharmacies, ice
-cream shops, and bakeries to construct our control group. These three
-types of establishments comply with the criteria of being morally
-neutral and having a spatial distribution commensurate with places of
-worship. Figure 2 shows the spatial distribution of the control
-establishments, namely pharmacies, bakeries, and ice cream shops in the
-city of Recife. Note the similarity between the maps. As expected, there
-are differences in the number of points, but the locations of cases and
-controls are quite similar.
-
-\begin{figure}
-
-\includegraphics{Moral_Communities_and_Crime_v1_files/figure-latex/plot-events-1} \hfill{}
-
-\caption{\label{fig:plot-events}Places of Worship and Commercial Establishments in Recife}\label{fig:plot-events}
-\end{figure}
-
-\begin{figure}
-
-\includegraphics{Moral_Communities_and_Crime_v1_files/figure-latex/plot-sed-1} \hfill{}
-
-\caption{\label{fig:plot-sed}Socio-Economic Deprivation in Recife}\label{fig:plot-sed}
-\end{figure}
-
-\hypertarget{results}{%
-\section{Analysis and results}\label{results}}
-
-\hypertarget{quadrat-analysis-1}{%
-\subsection{Quadrat Analysis}\label{quadrat-analysis-1}}
-
-\begin{figure}
-\centering
-\includegraphics{Moral_Communities_and_Crime_v1_files/figure-latex/plot-sed-as-quintiles-1.pdf}
-\caption{\label{fig:plot-sed-as-quintiles}Socio-Economic Deprivation as
-Quintiles}
-\end{figure}
-
-Calculate the intensity of other ppp by quadrat:
-
-\begin{figure}
-\centering
-\includegraphics{Moral_Communities_and_Crime_v1_files/figure-latex/plot-crime-quadrat-1.pdf}
-\caption{\label{fig:plot-crime-quadrat}Intensity of crime by level of
-Socio-Economic Deprivation; the dotted line indicates the global
-intensity of crime}
-\end{figure}
-
-\begin{figure}
-\centering
-\includegraphics{Moral_Communities_and_Crime_v1_files/figure-latex/plot-events-quadrat-1.pdf}
-\caption{\label{fig:plot-events-quadrat}Intensity of places of worship
-and commercial establishments by level of Socio-Economic Deprivation;
-the dotted lines indicates the respective global intensities}
-\end{figure}
-
-\hypertarget{relative-intensity-1}{%
-\subsection{Relative Intensity}\label{relative-intensity-1}}
-
-\hypertarget{discussion}{%
-\section{Discussion}\label{discussion}}
-
-An opposing viewpoint is offered by the theory of conservative
-Protestants (Ellison, Burr, and McCall 2003). This theory argues that
-religion is not always beneficial in order to guarantee social order,
-and consequently to result in lower crime rates. This theory suggested
-that, conservative Protestants are more tolerate with violent behavior
-when associated with defense of honor, family, property, or woman.
-However, high rates are not expected for all types of crimes in densely
-populated areas of conservative Protestants. But what justifies this
-behavior? Ellison et al.~(2003) argue that this tolerance stems from a
-religious view of the world. More clearly, punishment is seen as a
-legitimate form of retribution whose explanation can be attributed to
-the literal interpretation of the Bible. In the specific context of
-Brazil, it is important to consider that the interpretations and
-meanings attributed to urban violence during the last two decades in
-Brazil often had religious figures as protagonists (Birman and Machado
-2012).
-
-\hypertarget{conclusions}{%
-\section{Conclusions}\label{conclusions}}
-
-The theory of conservative Protestants, argues that conservative
-Protestants are more likely to tolerate violent behavior of a defensive
-or punitive character, especially when associated to defense of honor,
-family, property, or woman. Our objective in the study is to explore the
-question whether moral communities have a spatial dimension. To this end
-we deploy an array of spatial analytical tools to investigate potential
-patterns of clustering of violent crime and places of worship, using as
-well other kinds of urban facilities to act as controls. The case study
-is the city of Recife, in the northeast of Brazil, and analysis is based
-on geographically detailed datasets of criminal events as well as
-several types of urban facilities.
-
-\hypertarget{references}{%
-\section*{References}\label{references}}
-\addcontentsline{toc}{section}{References}
-
-\hypertarget{refs}{}
-\leavevmode\hypertarget{ref-Abdullah2018amenities}{}%
-Abdullah, Snhs, F. A. Bohani, Z. A. Nazri, Y. Jeffry, M. A. Abdullah, M.
-N. Junoh, and Z. A. Kasim. 2018. ``Amenities Surrounding Commercial
-Serial Crime Prediction at Greater Valley and Kuala Lumpur Using K-Means
-Clustering.'' Journal Article. \emph{Jurnal Teknologi} 80 (4): 43--53.
-\href{\%3CGo\%20to\%20ISI\%3E://WOS:000437011000005}{\textless{}Go to ISI\textgreater{}://WOS:000437011000005}.
-
-\leavevmode\hypertarget{ref-Becker2017analise}{}%
-Becker, Kalinca Léia, and Ana Lúcia Kassouf. 2017. ``Uma análise Do
-Efeito Dos Gastos Públicos Em Educação Sobre a Criminalidade No
-Brasil.'' \emph{Economia E Sociedade} 26 (1): 215--42.
-
-\leavevmode\hypertarget{ref-Birman2012violencia}{}%
-Birman, Patricia, and Carly Machado. 2012. ``A Violencia Dos Justos:
-Evangelicos, Midia E Periferias Da Metropole.'' \emph{Revista Brasileira
-de Ciências Sociais} 27 (80): 55--69.
-
-\leavevmode\hypertarget{ref-Brower2007spatial}{}%
-Brower, A. M., and L. Carroll. 2007. ``Spatial and Temporal Aspects of
-Alcohol-Related Crime in a College Town.'' Journal Article.
-\emph{Journal of American College Health} 55 (5): 267--75.
-\url{https://doi.org/10.3200/jach.55.5.267-276}.
-
-\leavevmode\hypertarget{ref-Chesnut2017devoted}{}%
-Chesnut, R Andrew. 2017. \emph{Devoted to Death: Santa Muerte, the
-Skeleton Saint}. Oxford University Press.
-
-\leavevmode\hypertarget{ref-Davignon2015christian}{}%
-Davignon, P., and R. A. Thomson. 2015. ``Christian Colleges and
-Universities as Moral Communities: The Effects of Institutional
-Characteristics on Student Religiosity.'' Journal Article. \emph{Review
-of Religious Research} 57 (4): 531--54.
-\url{https://doi.org/10.1007/s13644-015-0214-5}.
-
-\leavevmode\hypertarget{ref-Deryol2016crime}{}%
-Deryol, R., P. Wilcox, M. Logan, and J. Wooldredge. 2016. ``Crime Places
-in Context: An Illustration of the Multilevel Nature of Hot Spot
-Development.'' Journal Article. \emph{Journal of Quantitative
-Criminology} 32 (2): 305--25.
-\url{https://doi.org/10.1007/s10940-015-9278-1}.
-
-\leavevmode\hypertarget{ref-Donovan1986different}{}%
-Donovan, Peter. 1986. ``Do Different Religions Share Moral Common
-Ground?'' Journal Article. \emph{Religious Studies} 22 (3/4): 367--75.
-\url{http://www.jstor.org.libaccess.lib.mcmaster.ca/stable/20006295}.
-
-\leavevmode\hypertarget{ref-Doran2011putting}{}%
-Doran, Bruce J, and Melissa B Burgess. 2011. \emph{Putting Fear of Crime
-on the Map: Investigating Perceptions of Crime Using Geographic
-Information Systems}. Springer Science \& Business Media.
-
-\leavevmode\hypertarget{ref-Durrant2017religion}{}%
-Durrant, Russil, and Zoe Poppelwell. 2017. ``Why Religion Matters.'' In
-\emph{Religion, Crime and Punishment}, 1--17. Springer.
-
-\leavevmode\hypertarget{ref-Eitle2011religion}{}%
-Eitle, D. 2011. ``Religion and Gambling Among Young Adults in the United
-States: Moral Communities and the Deterrence Hypothesis.'' Journal
-Article. \emph{Journal for the Scientific Study of Religion} 50 (1):
-61--81. \url{https://doi.org/10.1111/j.1468-5906.2010.01552.x}.
-
-\leavevmode\hypertarget{ref-Ellison2003enduring}{}%
-Ellison, Christopher G, Jeffrey A Burr, and Patricia L McCall. 2003.
-``The Enduring Puzzle of Southern Homicide: Is Regional Religious
-Culture the Missing Piece?'' \emph{Homicide Studies} 7 (4): 326--52.
-
-\leavevmode\hypertarget{ref-Foster2010neighbourhood}{}%
-Foster, S., B. Giles-Corti, and M. Knuiman. 2010. ``Neighbourhood Design
-and Fear of Crime: A Social-Ecological Examination of the Correlates of
-Residents' Fear in New Suburban Housing Developments.'' Journal Article.
-\emph{Health \& Place} 16 (6): 1156--65.
-\url{https://doi.org/10.1016/j.healthplace.2010.07.007}.
-
-\leavevmode\hypertarget{ref-Furr2010metric}{}%
-Furr-Holden, C. D. M., K. D. M. Campbell, A. J. Milam, M. J. Smart, N.
-A. Ialongo, and P. J. Leaf. 2010. ``Metric Properties of the
-Neighborhood Inventory for Environmental Typology (Nifety): An
-Environmental Assessment Tool for Measuring Indicators of Violence,
-Alcohol, Tobacco, and Other Drug Exposures.'' Journal Article.
-\emph{Evaluation Review} 34 (3): 159--84.
-\url{https://doi.org/10.1177/0193841x10368493}.
-
-\leavevmode\hypertarget{ref-Groff2015informal}{}%
-Groff, E. R. 2015. ``Informal Social Control and Crime Events.'' Journal
-Article. \emph{Journal of Contemporary Criminal Justice} 31 (1):
-90--106. \url{https://doi.org/10.1177/1043986214552619}.
-
-\leavevmode\hypertarget{ref-He2017built}{}%
-He, Li, Antonio Páez, and Desheng Liu. 2017. ``Built Environment and
-Violent Crime: An Environmental Audit Approach Using Google Street
-View.'' Journal Article. \emph{Computers, Environment and Urban Systems}
-66: 83--95.
-\url{https://doi.org/http://dx.doi.org/10.1016/j.compenvurbsys.2017.08.001}.
-
-\leavevmode\hypertarget{ref-Hipp2007block}{}%
-Hipp, J. R. 2007. ``Block, Tract, and Levels of Aggregation:
-Neighborhood Structure and Crime and Disorder as a Case in Point.''
-Journal Article. \emph{American Sociological Review} 72 (5): 659--80.
-\url{https://doi.org/10.1177/000312240707200501}.
-
-\leavevmode\hypertarget{ref-Hirschi1969hellfire}{}%
-Hirschi, Travis, and Rodney Stark. 1969. ``Hellfire and Delinquency.''
-\emph{Social Problems} 17 (2): 202--13.
-
-\leavevmode\hypertarget{ref-Hoffmann2015religion}{}%
-Hoffmann, John P. 2015. ``Religion and Crime.'' \emph{The Encyclopedia
-of Crime and Punishment}, 1--5.
-
-\leavevmode\hypertarget{ref-Johnson2011more}{}%
-Johnson, Byron. 2011. \emph{More God, Less Crime: Why Faith Matters and
-How It Could Matter More}. Templeton Foundation Press.
-
-\leavevmode\hypertarget{ref-Lee2004love}{}%
-Lee, M. R., and J. P. Bartkowski. 2004. ``Love Thy Neighbor? Moral
-Communities, Civic Engagement, and Juvenile Homicide in Rural Areas.''
-Journal Article. \emph{Social Forces} 82 (3): 1001--35.
-\url{https://doi.org/10.1353/sof.2004.0044}.
-
-\leavevmode\hypertarget{ref-Lipton2008spatial}{}%
-Lipton, R., A. Banerjee, D. Levy, N. Manzanilla, and M. Cochrane. 2008.
-``The Spatial Distribution of Underage Tobacco Sales in Los Angeles.''
-Journal Article. \emph{Substance Use \& Misuse} 43 (11): 1597--1617.
-\url{https://doi.org/10.1080/10826080802241110}.
-
-\leavevmode\hypertarget{ref-Loukaitou2001measuring}{}%
-Loukaitou-Sideris, A., R. Liggett, H. Iseki, and W. Thurlow. 2001.
-``Measuring the Effects of Built Environment on Bus Stop Crime.''
-Journal Article. \emph{Environment and Planning B-Planning \& Design} 28
-(2): 255--80. \url{https://doi.org/10.1068/b2642r}.
-
-\leavevmode\hypertarget{ref-Menezes2013spatial}{}%
-Menezes, T., R. Silveira-Neto, C. Monteiro, and J. L. Ratton. 2013.
-``Spatial Correlation Between Homicide Rates and Inequality: Evidence
-from Urban Neighborhoods.'' Journal Article. \emph{Economics Letters}
-120 (1): 97--99. \url{https://doi.org/10.1016/j.econlet.2013.03.040}.
-
-\leavevmode\hypertarget{ref-Murray2013crime}{}%
-Murray, J., D. R. D. Cerqueira, and T. Kahn. 2013. ``Crime and Violence
-in Brazil: Systematic Review of Time Trends, Prevalence Rates and Risk
-Factors.'' Journal Article. \emph{Aggression and Violent Behavior} 18
-(5): 471--83. \url{https://doi.org/10.1016/j.avb.2013.07.003}.
-
-\leavevmode\hypertarget{ref-Nie2019smoking}{}%
-Nie, F. H., and X. Z. Y. Yang. 2019. ``Smoking in the Temple of the Holy
-Spirit? Geographic Location Matters.'' Journal Article. \emph{Health \&
-Place} 58. \url{https://doi.org/10.1016/j.healthplace.2019.05.017}.
-
-\leavevmode\hypertarget{ref-Quick2017influence}{}%
-Quick, M., J. Law, and H. Luan. 2017. ``The Influence of on-Premise and
-Off-Premise Alcohol Outlets on Reported Violent Crime in the Region of
-Waterloo, Ontario: Applying Bayesian Spatial Modeling to Inform Land Use
-Planning and Policy.'' Journal Article. \emph{Applied Spatial Analysis
-and Policy} 10 (3): 435--54.
-\url{https://doi.org/10.1007/s12061-016-9191-5}.
-
-\leavevmode\hypertarget{ref-Regnerus2003moral}{}%
-Regnerus, M. D. 2003. ``Moral Communities and Adolescent Delinquency:
-Religious Contexts and Community Social Control.'' Journal Article.
-\emph{Sociological Quarterly} 44 (4): 523--54.
-\href{\%3CGo\%20to\%20ISI\%3E://WOS:000188319800001}{\textless{}Go to ISI\textgreater{}://WOS:000188319800001}.
-
-\leavevmode\hypertarget{ref-Reiss1951delinquency}{}%
-Reiss, Albert J. 1951. ``Delinquency as the Failure of Personal and
-Social Controls.'' Journal Article. \emph{American Sociological Review}
-16 (2): 196--207. \url{https://doi.org/10.2307/2087693}.
-
-\leavevmode\hypertarget{ref-Rohrbaugh1975religiosity}{}%
-Rohrbaugh, John, and Richard Jessor. 1975. ``Religiosity in Youth: A
-Personal Control Against Deviant Behavior.'' \emph{Journal of
-Personality} 43 (1): 136--55.
-
-\leavevmode\hypertarget{ref-Stark1996religion}{}%
-Stark, R. 1996. ``Religion as Context: Hellfire and Delinquency One More
-Time.'' Journal Article. \emph{Sociology of Religion} 57 (2): 163--73.
-\url{https://doi.org/10.2307/3711948}.
-
-\leavevmode\hypertarget{ref-Stroope2018moral}{}%
-Stroope, S., and J. O. Baker. 2018. ``Whose Moral Community?
-Religiosity, Secularity, and Self-Rated Health Across Communal Religious
-Contexts.'' Journal Article. \emph{Journal of Health and Social
-Behavior} 59 (2): 185--99.
-\url{https://doi.org/10.1177/0022146518755698}.
-
-\leavevmode\hypertarget{ref-Topalli2013god}{}%
-Topalli, V., T. Brezina, and M. Bernhardt. 2013. ``With God on My Side:
-The Paradoxical Relationship Between Religious Belief and Criminality
-Among Hardcore Street Offenders.'' Journal Article. \emph{Theoretical
-Criminology} 17 (1): 49--69.
-\url{https://doi.org/10.1177/1362480612463114}.
-
-\leavevmode\hypertarget{ref-Traunmuller2011moral}{}%
-Traunmuller, R. 2011. ``Moral Communities? Religion as a Source of
-Social Trust in a Multilevel Analysis of 97 German Regions.'' Journal
-Article. \emph{European Sociological Review} 27 (3): 346--63.
-\url{https://doi.org/10.1093/esr/jcq011}.
-
-\leavevmode\hypertarget{ref-Unodc2019executive}{}%
-United Nations Office on Drugs and Crime. 2019a. ``Global Study on
-Homicide 2019: Executive Summary.'' UNODC Vienna.
-
-\leavevmode\hypertarget{ref-Unodc2019development}{}%
----------. 2019b. ``Global Study on Homicide 2019: Homicide, Development
-and the Sustainable Development Goals.'' UNODC Vienna.
-
-\leavevmode\hypertarget{ref-Warner2019neighborhood}{}%
-Warner, B. D., and R. H. Konkel. 2019. ``Neighborhood Churches and Their
-Relationship to Neighborhood Processes Important for Crime Prevention.''
-Journal Article. \emph{Journal of Urban Affairs}.
-\url{https://doi.org/10.1080/07352166.2019.1581030}.
-
 \bibliographystyle{spbasic}

### 2019-10-15 Moral_Communities_and_Crime_v1.Rmd
First complete draft of paper and title page

@@
 
-Words go here.
+Theories of moral communities and social disorganization suggest that the presence of institutions that can provide a measure of social control can help to improve crime prevention, and this has been borne to some extent by previous empirical research.
+However, as noted by Warner and Konkel [-@Warner2019neighborhood], it is important to distinguish between different religious traditions.
+
+In this respect, it has been argued that Catholics and some Protestant groups are more adept at generating bridging social capital with stron intergroup relations.
+This can result in greater collective efficacy - which in turn can decrease the likelihood of crime [see @Hoffmann2015religion, p.
+2].
+Along these lines, Beyerlein and Hipp [-@Beyerlein2005social], examined data for 3,157 counties in the United States in 2000, and found that counties with greater percentages of people affiliated with mainline Protestant and Catholic traditions tended to have lower crime rates.
+Similarly, Warner and Konkel [-@Warner2019neighborhood] found that mainline Protestant and briding churches (but not Catholic churches) correlated positively with informal social control at the neighborhood level in two cities in Kentucky.
+Nie and Yang [-@Nie2019smoking], in a study of smoking, also found that share of Catholic population in counties did not reduce the rate of smoking, thus bringing into question the efficacy of this type of moral community in generating social control - albeit at a fairly aggregated scale (counties).
+Our findings with respect to Catholic places of worship echo these results at a much higher level of disaggregation, as we fail to find evidence that Catholic churches project a moral community in space more than any of various common types of business establishments.
+
+Conservative Protestant and Evangelic traditions, unlike mainline Protestants and Catholics, appear to focus less on bridging and more on bonding social capital with strong within-group relations.
+Associated with this, Conservative Protestant doctrine emphasizes individual responsibility, and therefore social ills tend to be seen as primarily personal ills in need of religious redemption instead of secular interventions [@Nie2019smoking, pp.
+2-3].
+Under this light, a moral community with such views of social problems may not be effective to deter crime [@Ellison2003enduring]; instead, conservative Protestant doctrine may be more tolerant of violent behavior when associated with defense of honor, family, property, or women, or unfortunate events may be seen as representing legitimate celestial retribution for moral torpitude.
+Empirical research by Beyerlein and Hipp [-@Beyerlein2005social] found that counties with greater percentages of people affiliated with Evangelical traditions tended to have higher crime rates.
+More recent research by Desmond et al.
+[-@Desmond2010congregations] in more than 400 block groups in Indianapolis found that neighborhoods with a greater presence of Evangelical congregations had higher rates of both violent and property crimes.
+In the specific context of Brazil, it is important to consider the role that religion has had in interpreting and conferring meaning to urban violence, or what Brazilian anthropologist Patricia Birman has termed "The Violence of the Just" [see @Birman2012violencia; @Birman2019narrativas].
+
+Our analysis suggests that the role of Evangelical churches may be more complex than thought: on the one hand, our results indicate that Lethal and Intentional Violent Crime tends to be more intense in the proximity of Evangelical places of worship, up to a distance of approximately 100 m, but then declines even below the global intensity and the background intenstiy at distances between 300 m and 100 m.
+It is at the moment unclear what could cause this, and candidate explanations could include the way police respond to crime incidents in different neighborhoods [e.g., Warner, 1997], the likelihood of residents reporting crime to the police (e.g., Goudriaan, Witterbrood, & Niewbeerta, 2006), and the level of opportunity for crimes (e.g., Rountree, Land, & Miethe, 1994).
+Could Evangelicals churches provide greater levels of opportunity for crime, if adherents to this tradition are seen as meek?
+And is it possible that at a different scale, unlike reports by other researchers, Evangelical churches are more effective at keeping social control?
+While these speculations cannot be answered in the context of the present study, we suggest that these might be fruitful areas for future research.
+
+Finally, we would like to remark on the results regarding Spiritist churches, which in some way are the mirror image of those for Evangelical churches.
+An extensive scan of the literature finds much less information about this religious tradition, in particular in the context of Brazil where it adapted in ways that made it very different from its original French predecessor [see @daGraca2011espiritismo].
+Could it be that Spiritism is seen by potential criminals as spiritually or materially risky, if supernatural retribution is feared?
+Again, we cannot do more than speculate about this at the moment, but this might be another worthwhle avenue for future investigation.
 

### 2019-10-15 Moral_Communities_and_Crime_v1.tex
First complete draft of paper and title page

@@
 
-Words go here.
+Theories of moral communities and social disorganization suggest that
+the presence of institutions that can provide a measure of social
+control can help to improve crime prevention, and this has been borne to
+some extent by previous empirical research. However, as noted by Warner
+and Konkel (2019), it is important to distinguish between different
+religious traditions.
+
+In this respect, it has been argued that Catholics and some Protestant
+groups are more adept at generating bridging social capital with stron
+intergroup relations. This can result in greater collective efficacy -
+which in turn can decrease the likelihood of crime (see Hoffmann 2015,
+2). Along these lines, Beyerlein and Hipp (2005), examined data for
+3,157 counties in the United States in 2000, and found that counties
+with greater percentages of people affiliated with mainline Protestant
+and Catholic traditions tended to have lower crime rates. Similarly,
+Warner and Konkel (2019) found that mainline Protestant and briding
+churches (but not Catholic churches) correlated positively with informal
+social control at the neighborhood level in two cities in Kentucky. Nie
+and Yang (2019), in a study of smoking, also found that share of
+Catholic population in counties did not reduce the rate of smoking, thus
+bringing into question the efficacy of this type of moral community in
+generating social control - albeit at a fairly aggregated scale
+(counties). Our findings with respect to Catholic places of worship echo
+these results at a much higher level of disaggregation, as we fail to
+find evidence that Catholic churches project a moral community in space
+more than any of various common types of business establishments.
+
+Conservative Protestant and Evangelic traditions, unlike mainline
+Protestants and Catholics, appear to focus less on bridging and more on
+bonding social capital with strong within-group relations. Associated
+with this, Conservative Protestant doctrine emphasizes individual
+responsibility, and therefore social ills tend to be seen as primarily
+personal ills in need of religious redemption instead of secular
+interventions (Nie and Yang 2019, 2--3). Under this light, a moral
+community with such views of social problems may not be effective to
+deter crime (Ellison, Burr, and McCall 2003); instead, conservative
+Protestant doctrine may be more tolerant of violent behavior when
+associated with defense of honor, family, property, or women, or
+unfortunate events may be seen as representing legitimate celestial
+retribution for moral torpitude. Empirical research by Beyerlein and
+Hipp (2005) found that counties with greater percentages of people
+affiliated with Evangelical traditions tended to have higher crime
+rates. More recent research by Desmond et al.~(2010) in more than 400
+block groups in Indianapolis found that neighborhoods with a greater
+presence of Evangelical congregations had higher rates of both violent
+and property crimes. In the specific context of Brazil, it is important
+to consider the role that religion has had in interpreting and
+conferring meaning to urban violence, or what Brazilian anthropologist
+Patricia Birman has termed ``The Violence of the Just'' (see Birman and
+Machado 2012; Birman 2019).
+
+Our analysis suggests that the role of Evangelical churches may be more
+complex than thought: on the one hand, our results indicate that Lethal
+and Intentional Violent Crime tends to be more intense in the proximity
+of Evangelical places of worship, up to a distance of approximately 100
+m, but then declines even below the global intensity and the background
+intenstiy at distances between 300 m and 100 m. It is at the moment
+unclear what could cause this, and candidate explanations could include
+the way police respond to crime incidents in different neighborhoods
+{[}e.g., Warner, 1997{]}, the likelihood of residents reporting crime to
+the police (e.g., Goudriaan, Witterbrood, \& Niewbeerta, 2006), and the
+level of opportunity for crimes (e.g., Rountree, Land, \& Miethe, 1994).
+Could Evangelicals churches provide greater levels of opportunity for
+crime, if adherents to this tradition are seen as meek? And is it
+possible that at a different scale, unlike reports by other researchers,
+Evangelical churches are more effective at keeping social control? While
+these speculations cannot be answered in the context of the present
+study, we suggest that these might be fruitful areas for future
+research.
+
+Finally, we would like to remark on the results regarding Spiritist
+churches, which in some way are the mirror image of those for
+Evangelical churches. An extensive scan of the literature finds much
+less information about this religious tradition, in particular in the
+context of Brazil where it adapted in ways that made it very different
+from its original French predecessor (see Graca Arribas 2011). Could it
+be that Spiritism is seen by potential criminals as spiritually or
+materially risky, if supernatural retribution is feared? Again, we
+cannot do more than speculate about this at the moment, but this might
+be another worthwhle avenue for future investigation.
 
@@
 
+\leavevmode\hypertarget{ref-Beyerlein2005social}{}%
+Beyerlein, Kraig, and John R Hipp. 2005. ``Social Capital, Too Much of a
+Good Thing? American Religious Traditions and Community Crime.'' Journal
+Article. \emph{Social Forces} 84 (2): 995--1013.
+
+\leavevmode\hypertarget{ref-Birman2012violencia}{}%
+Birman, Patricia, and Carly Machado. 2012. ``A Violencia Dos Justos:
+Evangelicos, Midia E Periferias Da Metropole.'' \emph{Revista Brasileira
+de Ciências Sociais} 27 (80): 55--69.
+
+\leavevmode\hypertarget{ref-Birman2019narrativas}{}%
+Birman, Patrocia. 2019. ``NARRATIVAS Seculares E Religiosas Sobre a
+Violencia: AS Fronteiras Do Humano No Governo Dos Pobres.''
+\emph{Sociologia \& Antropologia} 9 (1).
+
 \leavevmode\hypertarget{ref-Brower2007spatial}{}%
@@
 
+\leavevmode\hypertarget{ref-Desmond2010congregations}{}%
+Desmond, S. A., G. Kikuchi, and K. H. Morgan. 2010. ``Congregations and
+Crime: Is the Spatial Distribution of Congregations Associated with
+Neighborhood Crime Rates?'' Journal Article. \emph{Journal for the
+Scientific Study of Religion} 49 (1): 37--55.
+\url{https://doi.org/10.1111/j.1468-5906.2009.01491.x}.
+
 \leavevmode\hypertarget{ref-Donovan1986different}{}%
@@
 61--81. \url{https://doi.org/10.1111/j.1468-5906.2010.01552.x}.
+
+\leavevmode\hypertarget{ref-Ellison2003enduring}{}%
+Ellison, Christopher G, Jeffrey A Burr, and Patricia L McCall. 2003.
+``The Enduring Puzzle of Southern Homicide: Is Regional Religious
+Culture the Missing Piece?'' \emph{Homicide Studies} 7 (4): 326--52.
 
@@
 
+\leavevmode\hypertarget{ref-daGraca2011espiritismo}{}%
+Graca Arribas, Celia da. 2011. ``Espiritismo: Entre Crime E Religiao.''
+\emph{Mneme-Revista de Humanidades} 12 (29).
+
 \leavevmode\hypertarget{ref-Groff2015informal}{}%

### 2019-10-16 Moral_Communities_and_Crime_Title_Page_v1.Rmd
Initial submission to Applied Spatial Analysis and Policy

@@
 - name: Edivaldo Constantino das Neves Júnior 
-  address: A Department, University of XXX
+  address: University of Sao Paulo - Brazil
   email: edivaldoconstantino@gmail.com
@@
 - name: Breno Caldas de Araujo
-  address: A Department, University of XXX
-  email: def@example.edu
+  address: Federal University of Pernambuco - Brazil
+  email: brenocaldasdearaujo@gmail.com
 
 - name: Tatiane Almeida de Menezes
-  address: Programa de Pós-Graduação em Gestão e Economia da Saúde, Universidade Federal de Pernambuco
+  address: Federal University of Pernambuco - Brazil
   email: tatianedemenezes@gmail.com
@@
 - name: Antonio Paez
-  address: School of Geography and Earth Sciences, McMaster University
+  address: McMaster University - Canada
   email: paezha@mcmaster.ca
@@
 With the advent of the scientific study of religion, there has been a growth of interest in the role of religions to operate as moral communities.
-Moral communities, a concept closely related to informal social controls, are of interest in countries and regions where formal controls are weak and ineffective.
+Moral communities, a concept closely related to informal social control, are of interest in countries and regions where formal controls are weak and ineffective.
 The objective of this paper is to present a spatial analysis of Violent and Intentional Crime in the city of Recife in Brazil, with a focus on the possible interactions between criminal events and places of worship.
 Previous research into moral communities has advocated the need for analysis at different scales, and this analysis contributes to the literature by using micro-level data and appropriate spatial analytical tools for spatial point patterns.
-Analysis is condcuted using three different types of places of worship (Catholic, Evangelical, and Spiritist) and three types of business establishments as controls (ice cream shops, pharmacies, and supermarkets).
-The results suggest that Catholic and Spiritist places of worship do not project moral communities geographically more than, say, ice cream shops.
-The intensity of criminal events in the proximity of Evangelical places of worship, in contrast, is markedly higher than for any of our other referential events.
+Analysis is conducted using three different types of places of worship (Catholic, Evangelical, and Spiritist) and three types of business establishments as controls (ice cream shops, pharmacies, and supermarkets).
+The results suggest that Catholic places of worship do not project moral communities geographically more than, say, ice cream shops.
+The intensity of criminal events in the proximity of Evangelical places of worship, in contrast, is markedly higher at short distances than for any of our other referential events; however, at distances between 300 and 500 m the intensity is lower.
+Finally, the intensity of crime with respect to Spiritist churches, albeit lower at short distances, tends to be significantly higher at distances between 300 and 500 m.
 

### 2019-10-16 Moral_Communities_and_Crime_Title_Page_v1.tex
Initial submission to Applied Spatial Analysis and Policy

@@
         Edivaldo Constantino das Neves Júnior \at
-     A Department, University of XXX \\
+     University of Sao Paulo - Brazil \\
      \email{\href{mailto:edivaldoconstantino@gmail.com}{\nolinkurl{edivaldoconstantino@gmail.com}}}  %  \\
@@
         Breno Caldas de Araujo \at
-     A Department, University of XXX \\
-     \email{\href{mailto:def@example.edu}{\nolinkurl{def@example.edu}}}  %  \\
+     Federal University of Pernambuco - Brazil \\
+     \email{\href{mailto:brenocaldasdearaujo@gmail.com}{\nolinkurl{brenocaldasdearaujo@gmail.com}}}  %  \\
 %             \emph{Present address:} of F. Author  %  if needed
@@
         Tatiane Almeida de Menezes \at
-     Programa de Pós-Graduação em Gestão e Economia da Saúde, Universidade
- Federal de Pernambuco \\
+     Federal University of Pernambuco - Brazil \\
      \email{\href{mailto:tatianedemenezes@gmail.com}{\nolinkurl{tatianedemenezes@gmail.com}}}  %  \\
@@
         Antonio Paez \at
-     School of Geography and Earth Sciences, McMaster University \\
+     McMaster University - Canada \\
      \email{\href{mailto:paezha@mcmaster.ca}{\nolinkurl{paezha@mcmaster.ca}}}  %  \\
@@
 as moral communities. Moral communities, a concept closely related to
-informal social controls, are of interest in countries and regions where
+informal social control, are of interest in countries and regions where
 formal controls are weak and ineffective. The objective of this paper is
@@
 micro-level data and appropriate spatial analytical tools for spatial
-point patterns. Analysis is condcuted using three different types of
+point patterns. Analysis is conducted using three different types of
 places of worship (Catholic, Evangelical, and Spiritist) and three types
 of business establishments as controls (ice cream shops, pharmacies, and
-supermarkets). The results suggest that Catholic and Spiritist places of
-worship do not project moral communities geographically more than, say,
-ice cream shops. The intensity of criminal events in the proximity of
-Evangelical places of worship, in contrast, is markedly higher than for
-any of our other referential events.
+supermarkets). The results suggest that Catholic places of worship do
+not project moral communities geographically more than, say, ice cream
+shops. The intensity of criminal events in the proximity of Evangelical
+places of worship, in contrast, is markedly higher at short distances
+than for any of our other referential events; however, at distances
+between 300 and 500 m the intensity is lower. Finally, the intensity of
+crime with respect to Spiritist churches, albeit lower at short
+distances, tends to be significantly higher at distances between 300 and
+500 m.
 \\

### 2019-10-16 Moral_Communities_and_Crime_v1.Rmd
Initial submission to Applied Spatial Analysis and Policy

@@
 
-<!-- This paper is an R markdown document that includes self-contained, reproducible analysis.
-To reproduce the paper the following file is needed: `Moral_Communities_and_Crime.RData`.
+<!-- This paper is an R markdown document that includes self-contained, reproducible analysis. To reproduce the paper the following files are needed:
+
+1. Moral_Communities_and_Crime_v1.Rmd (this R markdown document)
+
+2.
+`Moral_Communities_and_Crime_Recife.RData`.
 For reference: this data file was created using R notebook `00.
@@
 Neither of these two files are needed for this document.
+
+3. For compiling into a pdf in the style of a Springer document: svjour3.cls, svglov3.clo, spphys.bst, spmpsci.bst, spbasic.bst
+
 -->
@@
 
-This concern is not iddle: the endemic malady of violent crime represents an important factor that threatens to derail sustainable development goals.
-Violent crime, furthermore, tends to be more accute in those regions most in need of development.
+This concern is not idle: the endemic malady of violent crime represents an important factor that threatens to derail sustainable development goals.
+Violent crime, furthermore, tends to be more acute in those regions most in need of development.
 On the one hand, violent crime represents an economic and social drag [@Becker2017analise].
@@
 In turn, these negative effects combine to create the unfortunate conditions that tend to breed criminal behavior, thus creating a vicious cycle of economic disadvantage and violence [@Unodc2019development].
-It is not surprising, given these pecularities, that researchers have joined calls for increased attention to the study of the patterns of violent crime in Low and Middle Income Countries [LMIC; @Murray2013crime].
+It is not surprising, given these peculiarities, that researchers have joined calls for increased attention to the study of the patterns of violent crime in Low and Middle Income Countries [LMIC; @Murray2013crime].
 
@@
 Serious institutional problems include the inefficiency of the police, the lack of national legislation, the glacial pace of judicial processes, and the weak situation of the prison system in the country [@Menezes2013spatial].
-This makes it even more urgent to understand the role of _informal social controls_, i.e., the ability of community organizations such as schools, clubs, and neighborbood associations to suppress crime by strengthening the capabilities of neighbors to control inappropriate behavior [@Groff2015informal].
+This makes it even more urgent to understand the role of _informal social controls_, i.e., the ability of community organizations such as schools, clubs, and neighborhood associations to suppress crime by strengthening the capabilities of neighbors to control inappropriate behavior [@Groff2015informal].
 
 In order to understand the factors that can potentially help to deter crime, beyond formal control, it is important to identify empirical regularities.
-Criminological factors include concrete elements (such as te presence of arms or drugs, or elements of the built environment), and also figurative factors, which include the soci costs of deviant behavior as well as family supervision.
+Criminological factors include concrete elements (such as the presence of arms or drugs, or elements of the built environment), and also figurative factors, which include the social cost of deviant behavior.
 Accordingly, a number of studies have investigated various aspects of the environment and neighborhood design [e.g., @Foster2010neighbourhood; @He2017built; @Loukaitou2001measuring], whereas other studies have focused on exposure to environmental attributes that signal weakened norms, such as liquor and tobacco outlets [e.g., @Brower2007spatial; @Deryol2016crime; @Lipton2008spatial; @Quick2017influence].
@@
 
-With the above considerations in mind, the objective of this paper is to investigate whether and how places of workship correlate spatially with criminal events, or in other words, to investigate whether their signals as moral communities have a discernible spatial dimension.
+With the above considerations in mind, the objective of this paper is to investigate whether and how places of worship correlate spatially with criminal events, or in other words, to investigate whether their signals as moral communities have a discernible spatial dimension.
 The case study is the city of Recife in the state of Pernambuco, in Brazil's Northeast.
@@
 An early effort to theorize the effect of religion on criminal behavior was the _hellfire hypothesis_ of Hirschi and Stark [-@Hirschi1969hellfire].
-The focus of this hypothesis is the threat of extratemporal (and possibly eternal) punishment, and how this threat can deter believers from commandment-breaking (also see Pascal's wager).
+The focus of this hypothesis is the threat of extra-temporal (and possibly eternal) punishment, and how this threat can deter believers from commandment-breaking (also see Pascal's wager).
 Accordingly, Hirsch and Stark [-@Hirschi1969hellfire] posit that negative correlations between crime and religion are a consequence of a sense of commitment with normative values, a commitment that is ritually reinforced in a regular fashion (e.g., the liturgical rite of peace in Catholic Mass).
@@
 The hypothesis of moral communities as a form of informal social control, on the other hand, has been less studied from a geographical perspective, and it is only recently that has attracted the attention of researchers.
-Groff [-@Groff2015informal], for instance, discusses informal social control as a phenomenon that can plaussibly operate at different geographical scales, from the level of the home and family, through the street block and neighborhood, possibly to other scales such as the county.
+Groff [-@Groff2015informal], for instance, discusses informal social control as a phenomenon that can plausibly operate at different geographical scales, from the level of the home and family, through the street block and neighborhood, possibly to other scales such as the county.
 In this way, Nie and Yang [-@Nie2019smoking] remark on a recent paper on the lack of research conducted to study how the religious context of a geographical area (e.g., a county) may influence (youth smoking) behavior (p.
@@
 
-Using quadrats of equal size is a convenient simplication, but in principle the areas could be different - in which case the count of events would be proportional to the area of the quadrat under homogeneity.
+Using quadrats of equal size is a convenient simplification, but in principle the areas could be different - in which case the count of events would be proportional to the area of the quadrat under homogeneity.
 An interesting variation of this technique is conditional quadrat counts, whereby the partition of regions $\mathcal{A}_1,\cdots,\mathcal{A}_m$ is done to reflect an underlying covariate of interest, say $Z$.
@@
                                               "Body injury followed by death",
-                                              "Black and White",
+                                              "Black and Brown",
                                               "Yellow and White",
@@
 These three types of establishments comply with the criteria of being morally neutral and having a spatial distribution commensurate with places of worship.
-Figure 2 shows the spatial distribution of the control establishments, namely pharmacies, bakeries, and ice cream shops in the city of Recife.
+
+Figure \ref{fig:plot-events} shows the spatial distribution of the places of worhsip and control establishments in the city of Recife.
 Note the similarity between the maps.
@@
 The figure shows a veritable mosaic of affluence and deprivation, with high deprivation areas directly in contact, and in some cases almost completely surrounded, by low deprivation areas.
-This geographical pattern of unequality, on the other hand, seems to be characteristic of Brazilian metropolitan regions, where enclaves of wealth and _favelas_ (i.e., urban slums) can be found in close proximity [see for example @Feitosa2007global].
+This geographical pattern of inequality, on the other hand, seems to be characteristic of Brazilian metropolitan regions, where enclaves of wealth and _favelas_ (i.e., urban slums) can be found in close proximity [see for example @Feitosa2007global].
 
@@
 The dotted line indicates the global intensity, simply the total number of criminal events divided by the area of the region.
-It can be seen that the intensity of crime clearly increases with Socio-Economic Deprivation, and the intensity in areas with the highest levels of Socio-Economic Deprivation is `r crime5_intensity[5]/crime5_intensity[1] * 100` % higher than in places with the lowest levels of Socio-Economic Deprivation.
+It can be seen that the intensity of crime clearly increases with Socio-Economic Deprivation, and the intensity in areas with the highest levels of Socio-Economic Deprivation is `r round(crime5_intensity[5]/crime5_intensity[1] * 100, 2)`% higher than in places with the lowest levels of Socio-Economic Deprivation.
 This is as expected, as research has consistently found correlations between Socio-Economic Deprivation and a variety of negative health outcomes, including homicide [@Ichihara2018area].
@@
 
-Three types of commercial establishments also appear to display inhomogeneous locational patterns (Table \ref{tab:quadrat-test-statistics}), with the $\chi^2$ test of independence emphaticaly rejecting the null hypothesis.
+Three types of commercial establishments also appear to display inhomogeneous locational patterns (Table \ref{tab:quadrat-test-statistics}), with the $\chi^2$ test of independence emphatically rejecting the null hypothesis.
 As seen in Figure \ref{fig:plot-events-quadrat}, ice cream shops tend to be found more frequently in places with middle-low Socio-Economic Deprivation, and less frequently in places with middle-high Socio-Economic Deprivation.
@@
                                    "Spiritist",
-                                   "Ice Crem Shop",
+                                   "Ice Cream Shop",
                                    "Pharmacy",
@@
 The general trends shown by the relative distribution functions are similar to those seen above, before the introduction of a baseline function.
-We notice that at short distances the intensity of crime as a function of distance to the various places of worship and commercial establishments is higher than the baseline intesity (i.e., the intensity according to Socio-Ecomomic Deprivation).
+We notice that at short distances the intensity of crime as a function of distance to the various places of worship and commercial establishments is higher than the baseline intensity (i.e., the intensity according to Socio-Economic Deprivation).
 The only exception is intensity as a function of distance to a Spiritist church, which at short distances is not significantly different from the baseline intensity (see the confidence bands).
 
-The peak intensity as a function of distance to Evangelical churches is `r round(max_int_evangelical_bl, 2)` times the baseline intensity; for Catholic places of worship, the highest intesity is `r round(max_int_catholic_bl, 2)` times the baseline intensity; and the intensity reaches a peak of `r round(max_int_spiritist_bl, 2)` times the baseline intensity when the covariate is distance to Spiritist churches.
+The peak intensity as a function of distance to Evangelical churches is `r round(max_int_evangelical_bl, 2)` times the baseline intensity; for Catholic places of worship, the highest intensity is `r round(max_int_catholic_bl, 2)` times the baseline intensity; and the intensity reaches a peak of `r round(max_int_spiritist_bl, 2)` times the baseline intensity when the covariate is distance to Spiritist churches.
 Turning now to inspect the three types of commercial establishments, we see that the highest intensities are `r round(max_int_ice_cream_bl, 2)`, `r round(max_int_pharmacy_bl, 2)`, and `r round(max_int_supermarket_bl, 2)` times the intensity according to Socio-Economic Deprivation for ice cream shops, pharmacies, and supermarkets, respectively.
@@
 
-In this respect, it has been argued that Catholics and some Protestant groups are more adept at generating bridging social capital with stron intergroup relations.
+In this respect, it has been argued that Catholics and some Protestant groups are more adept at generating bridging social capital with strong intergroup relations.
 This can result in greater collective efficacy - which in turn can decrease the likelihood of crime [see @Hoffmann2015religion, p.
@@
 Along these lines, Beyerlein and Hipp [-@Beyerlein2005social], examined data for 3,157 counties in the United States in 2000, and found that counties with greater percentages of people affiliated with mainline Protestant and Catholic traditions tended to have lower crime rates.
-Similarly, Warner and Konkel [-@Warner2019neighborhood] found that mainline Protestant and briding churches (but not Catholic churches) correlated positively with informal social control at the neighborhood level in two cities in Kentucky.
+Similarly, Warner and Konkel [-@Warner2019neighborhood] found that mainline Protestant and bridging churches (but not Catholic churches) correlated positively with informal social control at the neighborhood level in two cities in Kentucky.
 Nie and Yang [-@Nie2019smoking], in a study of smoking, also found that share of Catholic population in counties did not reduce the rate of smoking, thus bringing into question the efficacy of this type of moral community in generating social control - albeit at a fairly aggregated scale (counties).
@@
 2-3].
-Under this light, a moral community with such views of social problems may not be effective to deter crime [@Ellison2003enduring]; instead, conservative Protestant doctrine may be more tolerant of violent behavior when associated with defense of honor, family, property, or women, or unfortunate events may be seen as representing legitimate celestial retribution for moral torpitude.
+Under this light, a moral community with such views of social problems may not be effective to deter crime [@Ellison2003enduring]; instead, conservative Protestant doctrine may be more tolerant of violent behavior when associated with defense of honor, family, property, or women, or unfortunate events may be seen as representing legitimate celestial retribution for moral turpitude.
 Empirical research by Beyerlein and Hipp [-@Beyerlein2005social] found that counties with greater percentages of people affiliated with Evangelical traditions tended to have higher crime rates.
@@
 
-Our analysis suggests that the role of Evangelical churches may be more complex than thought: on the one hand, our results indicate that Lethal and Intentional Violent Crime tends to be more intense in the proximity of Evangelical places of worship, up to a distance of approximately 100 m, but then declines even below the global intensity and the background intenstiy at distances between 300 m and 100 m.
-It is at the moment unclear what could cause this, and candidate explanations could include the way police respond to crime incidents in different neighborhoods [e.g., Warner, 1997], the likelihood of residents reporting crime to the police (e.g., Goudriaan, Witterbrood, & Niewbeerta, 2006), and the level of opportunity for crimes (e.g., Rountree, Land, & Miethe, 1994).
+Our analysis suggests that the role of Evangelical churches may be more complex than thought: on the one hand, our results indicate that Lethal and Intentional Violent Crime tends to be more intense in the proximity of Evangelical places of worship, up to a distance of approximately 100 m, but then declines even below the global intensity and the background intensity at distances between 300 m and 100 m.
+It is at the moment unclear what could cause this, and candidate explanations could include the way police respond to crime incidents in different neighborhoods, the likelihood of residents reporting crime to the police, and the level of opportunity for crimes [@Warner2019neighborhood].
 Could Evangelicals churches provide greater levels of opportunity for crime, if adherents to this tradition are seen as meek?
@@
 Could it be that Spiritism is seen by potential criminals as spiritually or materially risky, if supernatural retribution is feared?
-Again, we cannot do more than speculate about this at the moment, but this might be another worthwhle avenue for future investigation.
+Again, we cannot do more than speculate about this at the moment, but this might be another worthwhile avenue for future investigation.
 
@@
 1.
-Catholic places of worhsip do not seem to geographically project more or less of a moral community effect than, say, ice cream shops, pharmacies, or supermarkets: the intensity of crime as a function of distance to Catholic churches is not significantly different that what we observe when we consider distance to any of these types of commercial establishments, at any distance.
+Catholic places of worship do not seem to geographically project more or less of a moral community effect than, say, ice cream shops, pharmacies, or supermarkets: the intensity of crime as a function of distance to Catholic churches is not significantly different that what we observe when we consider distance to any of these types of commercial establishments, at any distance.
 
 2.
-The intensity of crime with respect to distance to Evangelical churches is more complex than previously known: whereas the intensity of crime is higher at relativey short distance from Evangelical places of worship, at longer distances it tends to decline even below the global intensity of crime and the background level of crime according to Socio-Economic Deprivation.
+The intensity of crime with respect to distance to Evangelical churches is more complex than previously known: whereas the intensity of crime is higher at relatively short distance from Evangelical places of worship, at longer distances it tends to decline even below the global intensity of crime and the background level of crime according to Socio-Economic Deprivation.
 
@@
 
-In summary, the research presented on this paper provides information about the potential of different types of moral communities to reduce crime, and should be of interest to policy makers as they assess whether formal or informal forms of social control can be effective to deter criminal behavior.
+In summary, the research presented on this paper provides information about the potential of different types of moral communities to reduce crime, and should be of interest to policy makers as they assess whether formal or informal forms of social control can be effective to deter criminal behavior, in order to achieve development goals.
+
+# Acknowledgments
+
+The following `R` packages were used in the course of this investigation and the authors wish to acknowledge their developers: `faterize` [@Ross2018fasterize], `ggthemes` [@Arnold2018], `kableExtra` [@Zhu2018], `knitr` [@Xie2015; @Xie2018], `maptools` [@Bivand2019maptools], `spatstat` [@Baddeley2005spatstat], `raster` [@Hijmans2018raster], `rticles` [@Allaire2018rticles], `sf` [@Pebesma2018simple], and `tidyverse` [@Wickham2017].
 

### 2019-10-16 Moral_Communities_and_Crime_v1.tex
Initial submission to Applied Spatial Analysis and Policy

@@
 
-This concern is not iddle: the endemic malady of violent crime
-represents an important factor that threatens to derail sustainable
-development goals. Violent crime, furthermore, tends to be more accute
-in those regions most in need of development. On the one hand, violent
-crime represents an economic and social drag (Becker and Kassouf 2017).
-The impacts are multifaceted, since they affect people's well-being,
-either through the loss of human life, mental health, and limitations in
-the right to public spaces (Doran and Burgess 2011), or through
-disturbances in schooling and academic achievement (United Nations
-Office on Drugs and Crime 2019b). In turn, these negative effects
-combine to create the unfortunate conditions that tend to breed criminal
-behavior, thus creating a vicious cycle of economic disadvantage and
-violence (United Nations Office on Drugs and Crime 2019b). It is not
-surprising, given these pecularities, that researchers have joined calls
-for increased attention to the study of the patterns of violent crime in
-Low and Middle Income Countries (LMIC; Murray, Cerqueira, and Kahn
-2013).
+This concern is not idle: the endemic malady of violent crime represents
+an important factor that threatens to derail sustainable development
+goals. Violent crime, furthermore, tends to be more acute in those
+regions most in need of development. On the one hand, violent crime
+represents an economic and social drag (Becker and Kassouf 2017). The
+impacts are multifaceted, since they affect people's well-being, either
+through the loss of human life, mental health, and limitations in the
+right to public spaces (Doran and Burgess 2011), or through disturbances
+in schooling and academic achievement (United Nations Office on Drugs
+and Crime 2019b). In turn, these negative effects combine to create the
+unfortunate conditions that tend to breed criminal behavior, thus
+creating a vicious cycle of economic disadvantage and violence (United
+Nations Office on Drugs and Crime 2019b). It is not surprising, given
+these peculiarities, that researchers have joined calls for increased
+attention to the study of the patterns of violent crime in Low and
+Middle Income Countries (LMIC; Murray, Cerqueira, and Kahn 2013).
 
@@
 role of \emph{informal social controls}, i.e., the ability of community
-organizations such as schools, clubs, and neighborbood associations to
+organizations such as schools, clubs, and neighborhood associations to
 suppress crime by strengthening the capabilities of neighbors to control
@@
 regularities. Criminological factors include concrete elements (such as
-te presence of arms or drugs, or elements of the built environment), and
-also figurative factors, which include the soci costs of deviant
-behavior as well as family supervision. Accordingly, a number of studies
-have investigated various aspects of the environment and neighborhood
-design (e.g., Foster, Giles-Corti, and Knuiman 2010; He, Páez, and Liu
-2017; Loukaitou-Sideris et al. 2001), whereas other studies have focused
-on exposure to environmental attributes that signal weakened norms, such
-as liquor and tobacco outlets (e.g., Brower and Carroll 2007; Deryol et
-al. 2016; Lipton et al. 2008; Quick, Law, and Luan 2017).
+the presence of arms or drugs, or elements of the built environment),
+and also figurative factors, which include the social cost of deviant
+behavior. Accordingly, a number of studies have investigated various
+aspects of the environment and neighborhood design (e.g., Foster,
+Giles-Corti, and Knuiman 2010; He, Páez, and Liu 2017; Loukaitou-Sideris
+et al. 2001), whereas other studies have focused on exposure to
+environmental attributes that signal weakened norms, such as liquor and
+tobacco outlets (e.g., Brower and Carroll 2007; Deryol et al. 2016;
+Lipton et al. 2008; Quick, Law, and Luan 2017).
 
@@
 With the above considerations in mind, the objective of this paper is to
-investigate whether and how places of workship correlate spatially with
+investigate whether and how places of worship correlate spatially with
 criminal events, or in other words, to investigate whether their signals
@@
 was the \emph{hellfire hypothesis} of Hirschi and Stark (1969). The
-focus of this hypothesis is the threat of extratemporal (and possibly
+focus of this hypothesis is the threat of extra-temporal (and possibly
 eternal) punishment, and how this threat can deter believers from
@@
 Groff (2015), for instance, discusses informal social control as a
-phenomenon that can plaussibly operate at different geographical scales,
+phenomenon that can plausibly operate at different geographical scales,
 from the level of the home and family, through the street block and
@@
 
-Using quadrats of equal size is a convenient simplication, but in
+Using quadrats of equal size is a convenient simplification, but in
 principle the areas could be different - in which case the count of
@@
 \multicolumn{2}{l}{\textbf{Ethnicity of Victim}}\\
-\hspace{1em}Black and White & 95.11\\
+\hspace{1em}Black and Brown & 95.11\\
 \hspace{1em}Yellow and White & 1.27\\
@@
 neutral and having a spatial distribution commensurate with places of
-worship. Figure 2 shows the spatial distribution of the control
-establishments, namely pharmacies, bakeries, and ice cream shops in the
-city of Recife. Note the similarity between the maps. As expected, there
-are differences in the number of points, but the locations of cases and
-controls are quite similar.
+worship.
+
+Figure \ref{fig:plot-events} shows the spatial distribution of the
+places of worhsip and control establishments in the city of Recife. Note
+the similarity between the maps. As expected, there are differences in
+the number of points, but the locations of cases and controls are quite
+similar.
 
@@
 directly in contact, and in some cases almost completely surrounded, by
-low deprivation areas. This geographical pattern of unequality, on the
+low deprivation areas. This geographical pattern of inequality, on the
 other hand, seems to be characteristic of Brazilian metropolitan
@@
 and the intensity in areas with the highest levels of Socio-Economic
-Deprivation is 208.7065292 \% higher than in places with the lowest
-levels of Socio-Economic Deprivation. This is as expected, as research
-has consistently found correlations between Socio-Economic Deprivation
-and a variety of negative health outcomes, including homicide (Ichihara
-et al. 2018). A \(\chi^2\) test of independence on the quadrat counts
-yields a value of \(p\leq 0.0001\) (four degrees of freedom), which
-comfortably rejects the null hypothesis of homogeneity of intensity of
-crime by level of Socio-Economic Deprivation.
+Deprivation is 208.71\% higher than in places with the lowest levels of
+Socio-Economic Deprivation. This is as expected, as research has
+consistently found correlations between Socio-Economic Deprivation and a
+variety of negative health outcomes, including homicide (Ichihara et al.
+2018). A \(\chi^2\) test of independence on the quadrat counts yields a
+value of \(p\leq 0.0001\) (four degrees of freedom), which comfortably
+rejects the null hypothesis of homogeneity of intensity of crime by
+level of Socio-Economic Deprivation.
 
@@
 \ref{tab:quadrat-test-statistics}), with the \(\chi^2\) test of
-independence emphaticaly rejecting the null hypothesis. As seen in
+independence emphatically rejecting the null hypothesis. As seen in
 Figure \ref{fig:plot-events-quadrat}, ice cream shops tend to be found
@@
 Spiritist & 9.108 & 4 & 0.1169\\
-Ice Crem Shop & 15.108 & 4 & 0.009\\
+Ice Cream Shop & 15.108 & 4 & 0.009\\
 Pharmacy & 51.869 & 4 & $<0.001$\\
@@
 various places of worship and commercial establishments is higher than
-the baseline intesity (i.e., the intensity according to Socio-Ecomomic
+the baseline intensity (i.e., the intensity according to Socio-Economic
 Deprivation). The only exception is intensity as a function of distance
@@
 3.33 times the baseline intensity; for Catholic places of worship, the
-highest intesity is 2.12 times the baseline intensity; and the intensity
-reaches a peak of 2.67 times the baseline intensity when the covariate
-is distance to Spiritist churches. Turning now to inspect the three
-types of commercial establishments, we see that the highest intensities
-are 2.28, 2.46, and 2.09 times the intensity according to Socio-Economic
-Deprivation for ice cream shops, pharmacies, and supermarkets,
-respectively.
+highest intensity is 2.12 times the baseline intensity; and the
+intensity reaches a peak of 2.67 times the baseline intensity when the
+covariate is distance to Spiritist churches. Turning now to inspect the
+three types of commercial establishments, we see that the highest
+intensities are 2.28, 2.46, and 2.09 times the intensity according to
+Socio-Economic Deprivation for ice cream shops, pharmacies, and
+supermarkets, respectively.
 
@@
 In this respect, it has been argued that Catholics and some Protestant
-groups are more adept at generating bridging social capital with stron
+groups are more adept at generating bridging social capital with strong
 intergroup relations. This can result in greater collective efficacy -
@@
 and Catholic traditions tended to have lower crime rates. Similarly,
-Warner and Konkel (2019) found that mainline Protestant and briding
+Warner and Konkel (2019) found that mainline Protestant and bridging
 churches (but not Catholic churches) correlated positively with informal
@@
 unfortunate events may be seen as representing legitimate celestial
-retribution for moral torpitude. Empirical research by Beyerlein and
+retribution for moral turpitude. Empirical research by Beyerlein and
 Hipp (2005) found that counties with greater percentages of people
@@
 m, but then declines even below the global intensity and the background
-intenstiy at distances between 300 m and 100 m. It is at the moment
+intensity at distances between 300 m and 100 m. It is at the moment
 unclear what could cause this, and candidate explanations could include
-the way police respond to crime incidents in different neighborhoods
-{[}e.g., Warner, 1997{]}, the likelihood of residents reporting crime to
-the police (e.g., Goudriaan, Witterbrood, \& Niewbeerta, 2006), and the
-level of opportunity for crimes (e.g., Rountree, Land, \& Miethe, 1994).
-Could Evangelicals churches provide greater levels of opportunity for
-crime, if adherents to this tradition are seen as meek? And is it
-possible that at a different scale, unlike reports by other researchers,
-Evangelical churches are more effective at keeping social control? While
-these speculations cannot be answered in the context of the present
-study, we suggest that these might be fruitful areas for future
-research.
+the way police respond to crime incidents in different neighborhoods,
+the likelihood of residents reporting crime to the police, and the level
+of opportunity for crimes (Warner and Konkel 2019). Could Evangelicals
+churches provide greater levels of opportunity for crime, if adherents
+to this tradition are seen as meek? And is it possible that at a
+different scale, unlike reports by other researchers, Evangelical
+churches are more effective at keeping social control? While these
+speculations cannot be answered in the context of the present study, we
+suggest that these might be fruitful areas for future research.
 
@@
 cannot do more than speculate about this at the moment, but this might
-be another worthwhle avenue for future investigation.
+be another worthwhile avenue for future investigation.
 
@@
 \item
-  Catholic places of worhsip do not seem to geographically project more
+  Catholic places of worship do not seem to geographically project more
   or less of a moral community effect than, say, ice cream shops,
@@
   churches is more complex than previously known: whereas the intensity
-  of crime is higher at relativey short distance from Evangelical places
-  of worship, at longer distances it tends to decline even below the
-  global intensity of crime and the background level of crime according
-  to Socio-Economic Deprivation.
+  of crime is higher at relatively short distance from Evangelical
+  places of worship, at longer distances it tends to decline even below
+  the global intensity of crime and the background level of crime
+  according to Socio-Economic Deprivation.
 \item
@@
 formal or informal forms of social control can be effective to deter
-criminal behavior.
+criminal behavior, in order to achieve development goals.
+
+\hypertarget{acknowledgments}{%
+\section{Acknowledgments}\label{acknowledgments}}
+
+The following \texttt{R} packages were used in the course of this
+investigation and the authors wish to acknowledge their developers:
+\texttt{faterize} (Ross 2018), \texttt{ggthemes} (Arnold 2018),
+\texttt{kableExtra} (Zhu 2018), \texttt{knitr} (Xie 2015, 2018),
+\texttt{maptools} (Bivand and Lewin-Koh 2019), \texttt{spatstat}
+(Baddeley and Turner 2005), \texttt{raster} (Hijmans 2018),
+\texttt{rticles} (Allaire et al. 2018), \texttt{sf} (Pebesma 2018), and
+\texttt{tidyverse} (Wickham 2017).
 
@@
 
+\leavevmode\hypertarget{ref-Allaire2018rticles}{}%
+Allaire, JJ, Yihui Xie, R Foundation, Hadley Wickham, Journal of
+Statistical Software, Ramnath Vaidyanathan, Association for Computing
+Machinery, et al. 2018. \emph{Rticles: Article Formats for R Markdown}.
+\url{https://CRAN.R-project.org/package=rticles}.
+
+\leavevmode\hypertarget{ref-Arnold2018}{}%
+Arnold, Jeffrey B. 2018. \emph{Ggthemes: Extra Themes, Scales and Geoms
+for 'Ggplot2'}. \url{https://CRAN.R-project.org/package=ggthemes}.
+
 \leavevmode\hypertarget{ref-Baddeley2015spatial}{}%
@@
 Patterns: Methodology and Applications with R}. Book. Chapman; Hall/CRC.
+
+\leavevmode\hypertarget{ref-Baddeley2005spatstat}{}%
+Baddeley, Adrian, and Rolf Turner. 2005. ``spatstat: An R Package for
+Analyzing Spatial Point Patterns.'' \emph{Journal of Statistical
+Software} 12 (6): 1--42. \url{http://www.jstatsoft.org/v12/i06/}.
 
@@
 \emph{Sociologia \& Antropologia} 9 (1).
+
+\leavevmode\hypertarget{ref-Bivand2019maptools}{}%
+Bivand, Roger, and Nicholas Lewin-Koh. 2019. \emph{Maptools: Tools for
+Handling Spatial Objects}.
+\url{https://CRAN.R-project.org/package=maptools}.
 
@@
 
+\leavevmode\hypertarget{ref-Hijmans2018raster}{}%
+Hijmans, Robert J. 2018. \emph{Raster: Geographic Data Analysis and
+Modeling}. \url{https://CRAN.R-project.org/package=raster}.
+
 \leavevmode\hypertarget{ref-Hipp2007block}{}%
@@
 Sciences}, edited by N. Wrigley, 127--44. London: Pion.
+
+\leavevmode\hypertarget{ref-Pebesma2018simple}{}%
+Pebesma, Edzer. 2018. ``Simple Features for R: Standardized Support for
+Spatial Vector Data.'' \emph{The R Journal} 10 (1): 439--46.
+\url{https://doi.org/10.32614/RJ-2018-009}.
 
@@
 
+\leavevmode\hypertarget{ref-Ross2018fasterize}{}%
+Ross, Noam. 2018. \emph{Fasterize: Fast Polygon to Raster Conversion}.
+\url{https://CRAN.R-project.org/package=fasterize}.
+
 \leavevmode\hypertarget{ref-Stark1996religion}{}%
@@
 
+\leavevmode\hypertarget{ref-Wickham2017}{}%
+Wickham, Hadley. 2017. \emph{Tidyverse: Easily Install and Load the
+'Tidyverse'}. \url{https://CRAN.R-project.org/package=tidyverse}.
+
+\leavevmode\hypertarget{ref-Xie2015}{}%
+Xie, Yihui. 2015. \emph{Dynamic Documents with R and Knitr}. 2nd ed.
+Boca Raton, Florida: Chapman; Hall/CRC. \url{https://yihui.name/knitr/}.
+
+\leavevmode\hypertarget{ref-Xie2018}{}%
+---------. 2018. \emph{Knitr: A General-Purpose Package for Dynamic
+Report Generation in R}. \url{https://yihui.name/knitr/}.
+
+\leavevmode\hypertarget{ref-Zhu2018}{}%
+Zhu, Hao. 2018. \emph{KableExtra: Construct Complex Table with 'Kable'
+and Pipe Syntax}. \url{https://CRAN.R-project.org/package=kableExtra}.
+
 \bibliographystyle{spbasic}

### 2019-10-16 Moral_Communities_and_Crime_Title_Page_v1.tex
Repeated initial submission after including statements re. funding information and conflicts of interest

@@
 analysis of places of worship and violent crime in the city of Recife,
-Brazil \thanks{Grants or other notes about the article that should go on the front page
-should be placed here. General acknowledgments should be placed at the
-end of the article.} }
+Brazil \thanks{\textbf{Funding:} Edivaldo Constantino das Neves Júnior received support
+from the Government of Canada, through the program Emerging Leaders in
+the Americas. \textbf{Conflicts of Interest:} The authors declare that
+they have no conflict of interest.} }
 

### 2019-10-17 Moral_Communities_and_Crime_v1.tex
Knitted the pdf file with the text again to remove the fictitious author names that the editorial assistant did not find amusing.

@@
 
-\author{  Aye B. Cee \and  Dee E. Eff \and  Gee H. Eye \and  Jay K. Elle \and  }
+\author{  Author One \and  Author Two \and  Author Three \and  Author Four \and  }
 
 \institute{
-        Aye B. Cee \at
+        Author One \at
      A Department, University of XXX \\
@@
     \and
-        Dee E. Eff \at
+        Author Two \at
      A Department, University of XXX \\
@@
     \and
-        Gee H. Eye \at
+        Author Three \at
      A Department, University of XXX \\
@@
     \and
-        Jay K. Elle \at
+        Author Four \at
      Some School, University of WWW \\

### 2020-04-07 Moral_Communities_and_Crime_v1_Word.Rmd
Create Word version to work on revisions after receiving feedback from reviewers.

--- 2019-10-16 Moral_Communities_and_Crime_v1.Rmd
@@
 ---
-title: Do moral communities have a spatial dimension? A spatial exploratory analysis of places of worship and violent crime in the city of Recife, Brazil
-subtitle: 
-titlerunning: Moral communities and crime
-authorrunning: 
-thanks: | 
-    Grants or other notes about the article that should go on the front 
-    page should be placed here. General acknowledgments should be placed at the
-    end of the article.
-
-authors: 
-- name: Aye B. Cee
-  address: A Department, University of XXX
-  email: abc@example.edu
-
-- name: Dee E. Eff
-  address: A Department, University of XXX
-  email: def@example.edu
-
-- name: Gee H. Eye
-  address: A Department, University of XXX
-  email: ghi@example.edu
-
-- name: Jay K. Elle
-  address: Some School, University of WWW
-  email: jkl@school.edu
-
-keywords:
-- Moral communities
-- Crime
-- Places of worship
-- Point pattern analysis
-- Intensity
-
-abstract: |
-  Religious tenets of the type "thou shalt not kill" and their equivalents in many world religions have functioned as _de facto_ social policy for determining appropriate and acceptable behavior over the centuries.
-With the advent of the scientific study of religion, there has been a growth of interest in the role of religions to operate as moral communities.
-Moral communities, a concept closely related to informal social control, are of interest in countries and regions where formal controls are weak and ineffective.
-The objective of this paper is to present a spatial analysis of Violent and Intentional Crime in the city of Recife in Brazil, with a focus on the possible interactions between criminal events and places of worship.
-Previous research into moral communities has advocated the need for analysis at different scales, and this analysis contributes to the literature by using micro-level data and appropriate spatial analytical tools for spatial point patterns.
-Analysis is conducted using three different types of places of worship (Catholic, Evangelical, and Spiritist) and three types of business establishments as controls (ice cream shops, pharmacies, and supermarkets).
-The results suggest that Catholic places of worship do not project moral communities geographically more than, say, ice cream shops.
-The intensity of criminal events in the proximity of Evangelical places of worship, in contrast, is markedly higher at short distances than for any of our other referential events; however, at distances between 300 and 500 m the intensity is lower.
-Finally, the intensity of crime with respect to Spiritist churches, albeit lower at short distances, tends to be significantly higher at distances between 300 and 500 m.
-
-bibliography: bibliography.bib
-output: rticles::springer_article
+title: "Moral_Communities_and_Crime_v1"
+author: "Edivaldo Constantino das Neves Júnior et al."
+date: "4/7/2020"
+output: word_document
 ---
-
-<!-- This paper is an R markdown document that includes self-contained, reproducible analysis. To reproduce the paper the following files are needed:
-
-1. Moral_Communities_and_Crime_v1.Rmd (this R markdown document)
-
-2.
-`Moral_Communities_and_Crime_Recife.RData`.
-For reference: this data file was created using R notebook `00.
-Data Preprocessing.Rmd`.
-The initial analysis was done in R notebook `01.
-Analysis.Rmd`.
-Neither of these two files are needed for this document.
-
-3. For compiling into a pdf in the style of a Springer document: svjour3.cls, svglov3.clo, spphys.bst, spmpsci.bst, spbasic.bst
-
--->
 

### 2020-05-18 Moral-Communities-and-Crime/Moral_Communities_and_Crime.Rmd
Tidied up folders before revisions for version 2

--- 2019-10-16 Moral_Communities_and_Crime_v1.Rmd
@@
 authors: 
-- name: Aye B. Cee
+- name: Author One
   address: A Department, University of XXX
@@
 
-- name: Dee E. Eff
+- name: Author Two
   address: A Department, University of XXX
@@
 
-- name: Gee H. Eye
+- name: Author Three
   address: A Department, University of XXX
@@
 
-- name: Jay K. Elle
+- name: Author Four
   address: Some School, University of WWW

### 2020-05-18 Moral-Communities-and-Crime/Moral_Communities_and_Crime_Title_Page.Rmd
Tidied up folders before revisions for version 2

--- 2019-10-16 Moral_Communities_and_Crime_Title_Page_v1.Rmd
@@
 thanks: | 
-    Grants or other notes about the article that should go on the front 
-    page should be placed here. General acknowledgments should be placed at the
-    end of the article.
+    **Funding:** Edivaldo Constantino das Neves Júnior received support from the Government of Canada, through the program Emerging Leaders in the Americas.
+**Conflicts of Interest:** The authors declare that they have no conflict of interest.
 

### 2020-06-10 Response-to-Reviewers-v2.Rmd
v2 Working on revisions after Edivaldo N.

---
title: "Response to Reviewers"
output: pdf_document
---

We would like to thank the editor and two anonymous reviewers for their comments and suggestions to revise the paper. In this letter we respond to the feedback received, and describe the actions taken to revise the paper.

\newpage

#Reviewer 1

This article uses spatial point pattern methods to explore the distribution of violent crime as it relates to the location of three types of churches and other land use types in Recife, Brazil. The authors suggest that the intensity of violent crimes is higher near to evangelical churches than other types of churches.

>\textcolor{blue}{Thank you for reading the paper and for your thoughtful suggestions to improve it.}

The article as currently written is lacking a thorough and logical argument as to why the location of churches can be hypothesized to influence violent crime at the point-level (i.e., where distance is the measure of interest rather than location within an area).

>\textcolor{blue}{Thank you for this comment. We believe this question has two parts: can we hypothesize that the location of churches influences violent crime? And if so, does this influence happen at the point level?}

>\textcolor{blue}{We believe that the first question is critical: what are the theoretical arguments for the influence of places of worship on crime? We can sketch this argument in the following way:}

>>\textcolor{blue}{1. Informal social controls can suppress crime by strengthening the capabilities of neighbors to control inappropriate behavior.}

>>\textcolor{blue}{2. Several types of community organizations, including schools, clubs, and neighborhood associations, can strengthen the capabilities of communities to exert informal social control.}

>>\textcolor{blue}{3. Of several types of community organizations, churches are among the most overtly moral, since they often have prescriptive codes of behavior.}

>>\textcolor{blue}{4. Why would the location of places of worship matter? Groff (2015) reviews several theoretical perspectives that address the role of informal social control in micro-places: these include "eyes on the street" and human territorial functioning. By affecting informal surveillance, movement governing rules, direct intervention, and the social climate at places, places of worship are hypothesized to contribute to informal social controls.}

>\textcolor{blue}{We believe that this argument is logical and sound, and churches, in fact, have been studied in the past for their potential to supress crime (Abdullah, 2018; Davignon, 2015; Furr, 2010; Traunmuller, 2011). A more novel aspect is the study of places of worship, as distinct entities from the members of the congregations: as Warner and Konkel (2019) note, this is an aspect that has received less attention in empirical and theoretical research. }

>\textcolor{blue}{The second question is related to the above: if places of worship somehow help to strengthen social control, does this happen at the level of points or for areas? In our view, Groff (2015) already explains why these investigations should matter at the micro-level. But there is also an analytical aspect to this question: what is an appropriate unit of support for the analysis? In this sense, we would note that aggregating point data to areas inevitably involves loss of information. Sometimes this situation is unavoidable (say, statistics are recorded for areas, so locational information is lost from the origin.) But when we have access to detailed geographical information, converting to area data is an unecessary loss of information. The question that we are asking is actually quite geographical: if there is evidence of a moral community effect, what is its distance decay? This question is best answered by considering the distance from criminal events to places of worship, instead of simple propinquity of criminal events and places of worship in areas.}

>\textcolor{blue}{We have edited the paper to increase its clarity.}

The current framing of the article also does not convincingly situate 'moral communities' within a geographical framework and does not make clear the hypothesized mechanisms that link the presence/absence of a church at a specific location with criminal behavior and, more specifically, the presence/absence of churches of specific religions with violent crimes. The authors should make sure not to conflate the processes linking crime with population characteristics (percent of individuals who are religious in an area), with religion more broadly (hellfire hypothesis), and with the location of a church in geographic space. This theoretical work is necessary to justify why this study is relevant and make clear how it contributes to the existing literature.

>\textcolor{blue}{Thank you for this comment. As pointed out above, the theoretical work is already discussed by Groff, among others. Although conceptually the study draws from much earlier research (e.g., Appleyard, Jacobs, Taylor), empirical work has lagged, as noted by Warner and Konkel (2019). This is where the current paper aims to make a contribution.}

The introduction of the paper does a really good job setting out the importance of reducing violent crime for achieving policy goals, but it is not clear how this research will contribute to developing, implementing, or evaluating policy. For this journal, I think it is necessary for the research to contribute to policy; so, how can police, religious officials, and other policymakers use the results of this study?

>\textcolor{blue}{The analysis of elements that can influence the occurrence of the criminal act is able to increase the range of tools that public authorities can use to face the problem, particularly in a reality as difficult as that of Brazil. The notion of informal social control allows new strategies to combat crime to be explored. In particular, we find that Catholic churches do not "project" any more moral community than pharmacies, which means that there is probably not much that can be done in terms of leveraging them for informal social control purposes. Spiritis churches have a lower relative risk at relatively short distances of less than approximately 50 m (which would be the side of a small block). Why is this? We cannot posit a causality relationship, but could this be due to superstitious beliefs associated by laypeople to this religion? And if so, could this be leveraged for policy purposes? Evangelical churches, on the other hand, have a higher relative risk than any other establishment or place of worship at distance between 20-100 m. A relevant question is whether this is crime by Evangelicals or crime towards Evangelicals. This is not something that the analysis can throw light on, but that deserves further investigation to determine whether it can lead to actionable policy recommendations.}

I currently find the logic of the analysis and the results to be confusing and I think that adding the results of more exploratory work before showing results of very specific analyses (i.e., those with SES quintiles and different church types) would help to provide a convincing argument that the results of this research are solid. For example, the authors may first examine the intensity of crime as it relates to all churches (aggregated), then to adjust for ses with all churches, and then to adjust for land use. Then, after presenting the results of the initial analysis for the aggregated churches, the results of the specific church types can be presented.

>\textcolor{blue}{Words.}

The authors should add a review of social disorganization theory and informal social control as these concepts are mentioned a number of times but not discussed in any detail. The similarities and differences between these concepts and the geographically-situated moral community concept should also be discussed.

>\textcolor{blue}{The referee's suggestion has been incorporated into the text. Two footnotes have been included that detail theories of social disorganization and the theory of social control. The focus was to present the main characteristics of these theoretical views and to relate them to the geographical aspect. In general, both views highlight the relationship between local characteristics (neighborhoods, for example), social networks and criminal behavior.}

The authors should justify their choice of method over other approaches, such as regression models, which provide a framework for quantifying the relationship between crime and churches after accounting for SES factors, land use types, and population at risk. There are a number of regression models that allow  coefficients to vary over space, which would quantify how the impact of churches on crime changes by distance (see: generalized additive models, spatially-varying coefficients, geographically weighted regression). Without this type of analysis, or at least a more thorough discussion of why this point pattern method was chosen, I am not convinced that crime is, in fact, associated with evangelical churches at close distances and that this isn't due to other factors, like evangelical churches being located in high population density areas (and so crime is due to higher population density, not the presence of a church), for example.

>\textcolor{blue}{Thank you for this comment. We agree, there are many methods that could be applied. In our analysis we use two tools:}

>\textcolor{blue}{1. The relative distribution, which is akin to a smooth estimate of the intensity as a function of distance.}

>\textcolor{blue}{2. The relative distribution function with a baseline function, which adjusts the relative distribution using other risk factors. In our case, those other factors are summarized by the SED score. The interpretation of the relative distribution with a baseline is as follows: values of rho greater than 1 indicate that the intensity of crime is higher over what the baseline explains, and values less than 1 indicate that the intensity of crime is lower than what the baseline explains.}

>\textcolor{blue}{In the previous version of the paper, our SED score was calculated using several potential confounders of crime: median income, proportion of popultion unemployed, proportion of households with a single mother, proportion of households with a young single mother, proportion of population in poverty. In the current version of the paper, we also include population density in the calculation of the SED score (thank you for the suggestion). Accordingly, the relative distribution with a baseline function is the smooth estimate of the intensity of crime after controlling for potential confounders, including population density.}

>\textcolor{blue}{We believe that this approach is informative, and since we are interested in the distribution of crime in the proximity of places of worship, it is also appropriate. Furthermore, by considering other underlying risk factors (in the form of the SED score), we believe that we have dilligently covered most reasonable confounders. We have edited the text to explain this.}

Please define the term 'neutral morality profile' and how it relates to the spatial distribution of crime around land uses. Please also clarify what 'having a spatial distribution commensurate with places of worship' means and provide evidence to support this claim.

>\textcolor{blue}{Following the referee suggestion, a footnote was added to the data section explaining clearly the highlighted terms. The idea of building a control group is to compare similar establishments about their spatial distribution, but different in relation to their neutrality profile in relation to crime, seeking to simulate the existence of a case and a control group.}

Did this study investigate the location of crime events, offender locations, or victim locations? I assume crime events, but it would be good to clarify nonetheless.

>\textcolor{blue}{In the data section, we changed the description. In the third paragraph, we use the term crime event location, which is more precise.}

For the intensity measures, what is the unit area for the results? Would it be possible to provide an intensity unit such as per mile^2 or per km^2, for example? I think that this would improve the interpretation of the results and Figures 6 and 7 in particular.

>\textcolor{blue}{Words.}

When discussing the quantitative results throughout Sections 5.3 and 5.4, please add uncertainty intervals so that readers can interpret the degree to which two quantities are, in fact, different? For example, p.16-line 49: "reaching a peak of 2.4916 * 10e-5 (95% CI: 2.1 * 10e-5, 3.2 * 10e-5)." (ANTÔNIO)

>\textcolor{blue}{Words.}

Please provide some additional text describing the equations in Section 3.2. It is not immediately clear to me how to interpret equations and how they are reflected in the results. For example, what is B(s), how is it calculated, and what does it mean for understanding the distribution of crime in relation to churches? (ANTÔNIO)

>\textcolor{blue}{Words.}

For readers (like me) who are unfamiliar with the geography of Recife, Brazil, it would be helpful if a detailed map was included in the paper that shows where the downtown is, major roads, different neighborhoods, etc. This could be included in the background of Figure 1 or in an Appendix. It would also be helpful if the authors provided some description of the patterns observed in Figure 1 and Figure 2 identifying where violent crimes (or churches, supermarkets, etc.) were clustered and where they weren't clustered.

>\textcolor{blue}{The appendix included information that allows a more in-depth characterization of the city of Recife: population distribution, income, inequality, urban facilities (education and health) and location of residential areas. We also have included a brief explanation of the requested clustering pattern in the text.}

How is the SES data, which I assume is distributed at the area-level, translated to points so it can be used in the point pattern analysis? Please describe how this was done and the limitations of doing so, if any. Related, why did the authors choose to break up SES into quintiles rather than keep it as a continuous measure or break up into other classifications?

>\textcolor{blue}{Thank you for this comment. When using quadrats in the earlier part of the analysis, the region is partitioned using quintiles; this is a requirement, since quadrats function as "containers" and need to be partitioned in discrete segments. However, when calculating the distribution functions, the continuous value of the SED score is used. We have clarified this in the text.}

What are the counts of the different types of churches, ice cream shops, pharmacies, and supermarkets? I am wondering if the frequency of each of these 'events' would influence the chi-square statistics and associated p-values shown in Table 3. Is this possible?

>\textcolor{blue}{Thank you for this comment. The chi-square statistic is affected if there are small numbers if events in some quadrats, but that was not the case for any of our events. We have included a new table with the quadrat counts.}

\newpage

#Reviewer 2

This was an interesting paper to review, challenging even.

>\textcolor{blue}{Thank you for reading the paper and for your thoughtful suggestions to improve it.}

We’ll start with a couple of minor issues

– in the quadrat analysis are the results conditioned on the quadrat size. If I understand correctly this might not be an issue with conditional quadrat counts, through the use of a covariate.  It does remind me of Openshaw early work on optimal zoning systems.  Perhaps a simple example to guide the reader might help.

>\textcolor{blue}{Words.}

- the symbolism in Figure 1, 2, and 3 are rather large.  I’m at a loss to interpret Figure 3, because as the image is built up (I have a rather slow graphics card) I see symbols overlapping.

>\textcolor{blue}{Words.}

There’s a cultural issue which I find challenging.  I think I know what Catholic and Evangelical churches may be, but this is first encounter with Spiritism. I assume that the first two are differentiated along some transubstantiation/consubstantiation divide, with the Catholics being in the former group. As for Spiritist churches I’m at a loss.  Do these follow the teaching of Allan Kardec; I presume it’s related to the Spiritualist churches found in the Anglophone countries, which go back to Swedenborg and Mesmer, and which seem to be originally motivated by trying to make contact with the dead. If so, then it’s very much a minority sect.  Perhaps the authors might like to expand on the details further for those readers who are not familiar with religious practices, and how they are undertaken in Brazil. Would I recognise the same in Cape Town, to take a city at random. Would this be a typical Spiritist church: https://goo.gl/maps/kwWe2B7didMk8vfDA

>\textcolor{blue}{We have included in the text an analysis of the religions studied in the article.}

Some description of the geography of Recife would also help. Where are residential areas, where are industrial zones, is there a university, and so on.

>\textcolor{blue}{The appendix included information that allows a more in-depth characterization of the city of Recife: population distribution, income, inequality, urban facilities (education and health) and location of residential areas.}

What’s the reason for the choice of controls – what if we take another three, such as take-away food outlets, garden centres, and clothing/footwear shops? Would we get similar of different results? You could use simulation to build a distribution of randomly chosen groups of different land uses. I’m not proposing that this should happen, but it’s a possibility.

>\textcolor{blue}{Words.}

I am reminded of Openshaw’s early work on cluster detection – work was based around trying to link clusters of leukaemia to electricity substations, and overhead transmission lines. Alan Craft was rumoured to have asked something along the lines of “how do we know it’s substations, or fish chip shops.  What if it’s telephone booths?”  Openshaw turned his approach on its head to ask where the clusters were occurring. This lead him to a nuclear waste reprocessing plant, and an urban domestic waste incineration plant. Does the possibility of turning the question to look at this from another point of view?

>\textcolor{blue}{Words.}

I’m left wondering whether this paper tells me something about spatial relationships in Recife.  The author(s) assert that the research is exploratory.  But to what extent might they apply elsewhere? Would these relationships translate to Vienna, for example?

>\textcolor{blue}{Words.}

I think that ASAP readers might be find this paper rarified.  They certainly need help with some of the cultural issues.

>\textcolor{blue}{Words.}

### 2020-06-10 Moral-Communities-and-Crime/Moral_Communities_and_Crime.Rmd
v2 Working on revisions after Edivaldo N.

@@
   st_drop_geometry() %>%
-  dplyr::select(median_income, prop_unemployed, prop_single_mother, prop_young_single_mother, prop_poverty) %>%
+  dplyr::select(pop_density, median_income, prop_unemployed, prop_single_mother, prop_young_single_mother, prop_poverty) %>%
   prcomp(center = TRUE, scale = TRUE)
@@
 
-pr_var[1:5]
+pr_var[1:6]
 
@@
 
-# Once that there is principal component (call it SED for socio-economic deprivation), it can be joined to the `sf` object before it can be rasterized and converted to an `im` object:
+# Once that there is a principal component (call it SED for socio-economic deprivation), it can be joined to the `sf` object before it can be rasterized and converted to an `im` object:
 
@@
 
->> High levels of homicidal violence are concentrated in geographic and demographic “pockets”, so
-achieving target 16.1 of the Sustainable Development Goals requires interventions within the specific
-regions, countries, communities and population groups that are most at risk [@Unodc2019executive, p. 35]
-
-This concern is not idle: the endemic malady of violent crime represents an important factor that threatens to derail sustainable development goals.
-Violent crime, furthermore, tends to be more acute in those regions most in need of development.
+>> High levels of homicidal violence are concentrated in geographic and demographic “pockets”, so achieving target 16.1 of the Sustainable Development Goals requires interventions within the specific regions, countries, communities and population groups that are most at risk [@Unodc2019executive, p.
+35]
+
+This concern is not idle: violent crime tends to be more acute in those regions most in need of development.
 On the one hand, violent crime represents an economic and social drag [@Becker2017analise].
@@
 
-Added to the scenario, formal social control institutions in countries like Brazil leave something (or much) to be desired: while in practice they should inhibit criminal behavior, they suffer from deep deficiencies that end up reducing the deterrent power of the justice system.
+Added to the scenario, formal social control institutions in countries like Brazil often leave something (or much) to be desired: while in practice they should inhibit criminal behavior, they suffer from deep deficiencies that end up eroding the deterrent power of the justice system.
 Serious institutional problems include the inefficiency of the police, the lack of national legislation, the glacial pace of judicial processes, and the weak situation of the prison system in the country [@Menezes2013spatial].
@@
 With the above considerations in mind, the objective of this paper is to investigate whether and how places of worship correlate spatially with criminal events, or in other words, to investigate whether their signals as moral communities have a discernible spatial dimension.
+Our argument is as follows:
+
+1.
+Since places of worship are associated with normative codes of behavior, can they by their presence contribute to create moral communities?
+The mechanisms for this are explained by Groff [@Groff2015informal] in terms of "Eyes on the Street" [see @Jacobs1961death; @Jacobs1968community; @Appleyard1980livable] and Human Territorial Functioning [@Taylor1988human; @Taylor1998crime].
+
+2. If so, what is the distance decay of said moral community effect? Is it uniform over a certain distance or does it vary?
+
 The case study is the city of Recife in the state of Pernambuco, in Brazil's Northeast.
@@
 
-# Background: Theoretical Perspectives on Religion and Crime {#background}
+# Background {#background}
+
+## Theoretical Perspectives on Religion and Crime
 
@@
 In this way, the hypothesis of _moral communities_ [@Stark1996religion] recognizes that social integration is essential for increasing social control, thus reducing the practice of behavior that is not in accordance with current norms.
-Rohrbaugh and Jessor [-@Rohrbaugh1975religiosity] also emphasize that religion acts as social control since it defines what is an appropriate attitude according to moral values, thus making moral communities a close relative of the concept of _informal social control_, defined as “the ability of social groups or institutions to make norms or rules effective” [@Reiss1951delinquency, p.
+Rohrbaugh and Jessor [-@Rohrbaugh1975religiosity] also emphasize that religion acts as social control since it defines what is an appropriate attitude according to moral values, thus making moral communities a close relative of the concept of _informal social control_, defined as “the ability of social groups or institutions to make norms or rules effective”\footnote{The discussion about disorder is directly related to the idea of social control.
+According to this perspective, control mechanisms to contain deviant actions are considered indispensable [@Rodrigues2012medo].
+This theory assumes that the deviant act is seen as normal in social organization, and coercive forces are needed to limit such impulses.
+This control can be exercised through formal institutions, as well as through informal mechanisms.
+It would be a set of norms as well as positive and negative sanctions, which are specified throughout the socialization process, which aims to establish values and build the individual's morals, disciplining society and subjecting individuals to standards and principles predefined.
+As the individual finds himself more and more inserted in the social context, less will be the chances of criminal behavior.
+From this perspective, the existence of bonds of solidarity and trust would be relevant for the effective exercise of supervision over crime in the neighborhood.
+Informal control is then linked to the ability of a neighborhood to establish communication and surveillance networks, interfering in cases of criminal actions and increasing security in the region.} [@Reiss1951delinquency, p.
 196; cited in @Groff2015informal, p.
@@
 2).
-This point is echoed by Warner and Konkel [-@Warner2019neighborhood], who moreover recommend the use of even smaller units of analysis (e.g., Census Block Groups), in an effort to reduce aggregation bias [see @Hipp2007block], but more importantly because the processes defined by social disorganization theory are thought to occur at relatively small geographical scales.
+This point is echoed by Warner and Konkel [-@Warner2019neighborhood], who moreover recommend the use of even smaller units of analysis (e.g., Census Block Groups), in an effort to reduce aggregation bias [see @Hipp2007block], but more importantly because the processes defined by social disorganization theory\footnote{The social disorganization theory (Shaw e McKay, 1942)  presents the social structure as an indispensable factor for the elaboration of theoretical and methodological models related to crime in urban neighborhoods.
+According to Villareal e Silva [-@Villarreal2006social], “neighborhoods characterized by dense social networks will experience greater trust among residents and cooperation in the enforcement of social norms against crime.
+Disadvantaged neighborhoods are thought to have fewer social ties among residents and lower rates of organizational participation”.
+The argument is that community characteristics such as poverty, ethnic heterogeneity and residential mobility disrupt the social organization of urban neighborhoods in such a way as to reduce their capacity to exercise social control [@Rodrigues2012medo].
+Thus, aspects such as residential instability or low socioeconomic status would contribute to the weakening of the community organization, increasing crime rates.
+This theory highlights the importance of the geography of crime and the characteristics of disorder in its analysis.} are thought to occur at relatively small geographical scales.
 
@@
 This requires micro-level data both of crimes and places of worship, and the use of appropriate spatial analytical tools for point patterns, as discussed next.
+
+## Religion in Brazil
+
+Brazil is a country with a Christian predominance.
+In the first national census in which religion was considered, in 1872, practically the entire population (99.72%) identified themselves as Catholic.
+In the 1940 census, when the religion question was researched again, Catholicism was still the religion with which the vast majority of the population identified: 95.2% of Brazilians declared themselves as Catholics that year.
+
+The dominance of Catholicism in the religious life of Brazil is unsurprising given the colonial past of the country.
+However, the religious makeup of the population started was not completely homogeneous.
+Among other offerings in the market place of spiritual ideas, Spiritism stands out.
+The Spiritist Doctrine was founded in nineteenth-century France by Allan Kardec as a way of understanding the world and its relations with the "beyond": a curious mix of pseudo-scientific language, philosophy, and religion [@Arribas2011doutrina].
+Spiritism arrived in Brazil in the middle of the previous century and since then it slowly spread in large and medium-sized cities.
+Today, Spiritism is the third largest religious segment in Brazil, after Catholicism and Evangelical Christianism, and shares with these religions the Christian principle of charity through pious assistance.
+
+Another signal event in the developmento of religion in Brazil is the arrival of Pentecostal Protestant churches, such as the Christian Congregation of Brazil in 1910 and the Assembly of God in 1911.
+Initially, these religious denominations did not make important inroads in Brazilian society (as evidenced by the 1940 census).
+However, in the context of rapid industrialization and major urbanization, they contributed to the change in the religious landscape along with other religions [@deSouza2019pluralidade]: according to the 2010 census the configuration of the religious scene in Brazil had changed considerably from the previous century: 64.6% of the population identified as Catholic and 22.2% as Evangelical Christians.
+The acceleration of this demographic change was directly related to the inroads made by Evangelical churches: The Assembly of God, for instance, grew considerably to total more than 12.3 million adherents.
+This church is represented extensively in many regions of the country, following a system of loosely allied ministries.
+With a charismatic profile and the lack of a centralized authority, growth of this church has occurred mainly through subdivision into numerous small and autonomous churches.
+These churches have in common the Prosperity Gospel and an emphasis on economic entrepreneurship  [e.g., @Mora2008marketing].
+They also do not shy away from partisan political action.
+
+In general, there have been two major changes in Brazil regarding religious issues: the slow decline of Catholicism and the rapid growth of Evangelical Christianism.
+These changes are evident in the daily life of the average Brazilian, either through the radio or through TV programs, which give more and more space to religious doctrines.
+At the same time, there has been an increase in cases of religious intolerance in Brazil, generating violence acts against temples, objects of worship, and people.
 
@@
 With a population of 1,550,390 million inhabitants in 2010, Recife is one of the main Brazilian metropolises, exerting a great economic influence in neighboring regions.
-However, the city experiences a serious problem with violent crime, and has the dubious honor of being one of the five capital cities in Brazil with the highest homicide rates in the period under study [@Menezes2013spatial].
+However, the city experiences a serious problem with violent crime, and has the dubious honor of being one of the five capital cities in Brazil with the highest homicide rates in the period under study\footnote{The appendix included information that allows a more in-depth characterization of the city of Recife: population distribution, income, inequality, urban facilities (education and health) and location of residential areas.} [@Menezes2013spatial].
 
@@
 
-The data are organized at the individual level and it is possible to obtain information about the location, day of the week, day of the month, period of the day, as well as gender, age, and race of the victim.
+The data are organized at the individual level and it is possible to obtain information about the crime event location, day of the week, day of the month, period of the day, as well as gender, age, and race of the victim.
 The database used in this study comprises the period from July 1, 2008 to June 30, 2010.
@@
 In addition to places of worship, the National Register of Addresses for Statistical Purposes was queried to extract facilities other than places of worship.
+To study the spatial relationship between churches and crime, a control group was built and then a comparison of the empirical distribution was carried out.
+Gilberston and Hayes [-@Gilbertson2012engineering] highlight that different configurations of urban space affect how people perceive the environment around them.
+In this perspective, churches are continually affecting the impression that people have of a region.
+Thus, we assume that churches have unique characteristics that may or may not inhibit the occurrence of a criminal act: they are urban constructions that are not neutral in relation to crime.
+On the other hand, there are urban facilities that do not provide any specific reason to justify the fact that the crime occurs closer or further from these buildings.
+Thus, it can be thought that any urban construction belongs to one of the following distinct situations: neutral or non-neutral to crime.
+Pharmacies, ice cream parlors and bakeries will be used in the construction of the control group and will be considered neutral.
+The comparison between the distributions of non-neutral and crime-neutral establishments forms the empirical basis of the study.
+The choice of establishments that make up the control was based on two criteria: (a) being neutral to crime and (b) having a spatial distribution similar to non-neutral establishments.
+The strategy used is to compare similar establishments about their spatial distribution, but different in relation to their neutrality profile in relation to crime, seeking to simulate the existence of a case and a control group.
+The case will be composed of churches, while the control will be pharmacies, ice cream parlors and bakeries.
+
 As discussed above, the idea is to identify points of reference that can be used as controls, having a neutral morality profile.
@@
 Note the similarity between the maps.
-As expected, there are differences in the number of points, but the locations of cases and controls are quite similar.
+As expected, there are differences in the number of points, but the locations of cases and controls are quite similar: the distribution of the churches, as well as of the commercial establishments follow the same pattern, being correlated with the spatial distribution of the population (see in the appendix).
+Since such urban constructions respond to society's demand, they tend to be located close to population agglomerations.
 
@@
 
-sed_descriptives <- data.frame(variable = c("Median Income", 
+sed_descriptives <- data.frame(variable = c("Population Density",
+                                            "Median Income", 
                     "Proportion Unemployed", 
@@
                     "Proportion Young Single Mother"),
-           min = c(min(setores.sf$median_income),
+           min = c(min(setores.sf$pop_density),
+                   min(setores.sf$median_income),
                    min(setores.sf$prop_unemployed),
@@
                    min(setores.sf$prop_young_single_mother)),
-           second_q = c(quantile(setores.sf$median_income, c(0, 0.25, 0.5, 0.75, 1))[2],
+           second_q = c(quantile(setores.sf$pop_density, c(0, 0.25, 0.5, 0.75, 1))[2],
+                        quantile(setores.sf$median_income, c(0, 0.25, 0.5, 0.75, 1))[2],
                         quantile(setores.sf$prop_unemployed, c(0, 0.25, 0.5, 0.75, 1))[2],
@@
                         quantile(setores.sf$prop_young_single_mother, c(0, 0.25, 0.5, 0.75, 1))[2]),
-           median = c(median(setores.sf$median_income),
+           median = c(median(setores.sf$pop_density),
+                      median(setores.sf$median_income),
                       median(setores.sf$prop_unemployed),
@@
                       median(setores.sf$prop_young_single_mother)),
-           mean = c(mean(setores.sf$median_income),
+           mean = c(mean(setores.sf$pop_density),
+                    mean(setores.sf$median_income),
                     mean(setores.sf$prop_unemployed),
@@
                     mean(setores.sf$prop_young_single_mother)),
-           third_q = c(quantile(setores.sf$median_income, c(0, 0.25, 0.5, 0.75, 1))[4],
+           third_q = c(quantile(setores.sf$pop_density, c(0, 0.25, 0.5, 0.75, 1))[4],
+             quantile(setores.sf$median_income, c(0, 0.25, 0.5, 0.75, 1))[4],
                        quantile(setores.sf$prop_unemployed, c(0, 0.25, 0.5, 0.75, 1))[4],
@@
                        quantile(setores.sf$prop_young_single_mother, c(0, 0.25, 0.5, 0.75, 1))[4]),
-           max = c(max(setores.sf$median_income),
+           max = c(max(setores.sf$pop_density),
+                   max(setores.sf$median_income),
                    max(setores.sf$prop_unemployed),
@@
                    max(setores.sf$prop_young_single_mother)),
-           pca_loadings = as.numeric(setores.pca$rotation[c(1, 2, 5, 3, 4), 1]))
+           pca_loadings = as.numeric(setores.pca$rotation[c(1, 2, 3, 6, 4, 5), 1]))
 ```
@@
   footnote(general = c("The variable for young single mothers is for women aged 15-25",
-                       "The first principal component accounts for 53.91% of the variance"))
-
+                       "The first principal component accounts for 44.9% of the variance"))
 ```
@@
 In essence, this entails counting the number of events at each level of the covariate, and then calculating an estimator of the intensity.
+The global intensity of each type of event (i.e., places of worship and businesses), as well as quadrant counts and intensity at each level of SED can be consulted in Table \ref{tab:quadrat-count-results}.
 
@@
 
-```{r instensity-dataframe, include=FALSE}
+```{r counts-intensity-dataframe, include=FALSE}
 
@@
                                        each = 5),
+                            Quadrat_Count = c(crime5,
+                                          catholic5,
+                                          evangelic5,
+                                          spiritist5,
+                                          ice_cream5,
+                                          pharmacies5,
+                                          super_markets5),
                             Intensity = c(crime5_intensity,
@@
 
+```{r table-counts-intensity-summary, echo=FALSE}
+kable(intensity_all %>% dplyr::select(Type, Global_Intensity, SED, Quadrat_Count, Intensity),
+      "latex",
+      booktabs = TRUE,
+      digits = 3,
+      escape = FALSE,
+      align = c("l", "c", "c", "c"),
+      caption = "\\label{tab:quadrat-count-results}Conditional quadrat counts results",
+      col.names = linebreak(c("Event", 
+                              "Global Intensity", 
+                              "SED Quintile", 
+                              "Quadrat Count",
+                              "Intensity"))) %>%
+  kable_styling(bootstrap_options = c("striped", "scaled_down")) %>% 
+  collapse_rows(1:2, latex_hline = "none")
+```
+
 Figure \ref{fig:plot-crime-quadrat} shows the results of conditional quadrat counts for Lethal and Intentional Violent Crime.
@@
 Finally, we would like to remark on the results regarding Spiritist churches, which in some way are the mirror image of those for Evangelical churches.
-An extensive scan of the literature finds much less information about this religious tradition, in particular in the context of Brazil where it adapted in ways that made it very different from its original French predecessor [see @daGraca2011espiritismo].
+An extensive scan of the literature finds much less information about this religious tradition, in particular in the context of Brazil where it adapted in ways that made it very different from its original French predecessor [see @Arribas2011espiritismo].
 Could it be that Spiritism is seen by potential criminals as spiritually or materially risky, if supernatural retribution is feared?

### 2020-06-10 Moral-Communities-and-Crime/Moral_Communities_and_Crime.tex
v2 Working on revisions after Edivaldo N.

--- 2019-10-17 Moral_Communities_and_Crime_v1.tex
@@
 
-This concern is not idle: the endemic malady of violent crime represents
-an important factor that threatens to derail sustainable development
-goals. Violent crime, furthermore, tends to be more acute in those
+This concern is not idle: violent crime tends to be more acute in those
 regions most in need of development. On the one hand, violent crime
@@
 Added to the scenario, formal social control institutions in countries
-like Brazil leave something (or much) to be desired: while in practice
-they should inhibit criminal behavior, they suffer from deep
-deficiencies that end up reducing the deterrent power of the justice
+like Brazil often leave something (or much) to be desired: while in
+practice they should inhibit criminal behavior, they suffer from deep
+deficiencies that end up eroding the deterrent power of the justice
 system. Serious institutional problems include the inefficiency of the
@@
 criminal events, or in other words, to investigate whether their signals
-as moral communities have a discernible spatial dimension. The case
-study is the city of Recife in the state of Pernambuco, in Brazil's
-Northeast. Recife is a large and important metropolitan area in a
-historically poor region, and afflicted by high levels of violent crime,
-having the dubious honor of being one of five state capitals with the
-highest rates of homicide in the period under study (Menezes et al.
+as moral communities have a discernible spatial dimension. Our argument
+is as follows:
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\item
+  Since places of worship are associated with normative codes of
+  behavior, can they by their presence contribute to create moral
+  communities? The mechanisms for this are explained by Groff (Groff
+  2015) in terms of ``Eyes on the Street'' (see Jacobs 1961, 1968;
+  Appleyard 1980) and Human Territorial Functioning (Taylor 1988, 1998).
+\item
+  If so, what is the distance decay of said moral community effect? Is
+  it uniform over a certain distance or does it vary?
+\end{enumerate}
+
+The case study is the city of Recife in the state of Pernambuco, in
+Brazil's Northeast. Recife is a large and important metropolitan area in
+a historically poor region, and afflicted by high levels of violent
+crime, having the dubious honor of being one of five state capitals with
+the highest rates of homicide in the period under study (Menezes et al.
 2013). The empirical strategy is to use spatial analysis to explore the
@@
 \hypertarget{background}{%
-\section{Background: Theoretical Perspectives on Religion and
-Crime}\label{background}}
+\section{Background}\label{background}}
+
+\hypertarget{theoretical-perspectives-on-religion-and-crime}{%
+\subsection{Theoretical Perspectives on Religion and
+Crime}\label{theoretical-perspectives-on-religion-and-crime}}
 
@@
 \emph{informal social control}, defined as ``the ability of social
-groups or institutions to make norms or rules effective'' (Reiss 1951,
-196; cited in Groff 2015, 91).
+groups or institutions to make norms or rules
+effective''\footnote{The discussion about disorder is directly related to the idea of social control.
+According to this perspective, control mechanisms to contain deviant actions are considered indispensable [@Rodrigues2012medo].
+This theory assumes that the deviant act is seen as normal in social organization, and coercive forces are needed to limit such impulses.
+This control can be exercised through formal institutions, as well as through informal mechanisms.
+It would be a set of norms as well as positive and negative sanctions, which are specified throughout the socialization process, which aims to establish values and build the individual's morals, disciplining society and subjecting individuals to standards and principles predefined.
+As the individual finds himself more and more inserted in the social context, less will be the chances of criminal behavior.
+From this perspective, the existence of bonds of solidarity and trust would be relevant for the effective exercise of supervision over crime in the neighborhood.
+Informal control is then linked to the ability of a neighborhood to establish communication and surveillance networks, interfering in cases of criminal actions and increasing security in the region.}
+(Reiss 1951, 196; cited in Groff 2015, 91).
 
@@
 effort to reduce aggregation bias (see Hipp 2007), but more importantly
-because the processes defined by social disorganization theory are
-thought to occur at relatively small geographical scales.
+because the processes defined by social disorganization
+theory\footnote{The social disorganization theory (Shaw e McKay, 1942)  presents the social structure as an indispensable factor for the elaboration of theoretical and methodological models related to crime in urban neighborhoods.
+According to Villareal e Silva [-@Villarreal2006social], “neighborhoods characterized by dense social networks will experience greater trust among residents and cooperation in the enforcement of social norms against crime.
+Disadvantaged neighborhoods are thought to have fewer social ties among residents and lower rates of organizational participation”.
+The argument is that community characteristics such as poverty, ethnic heterogeneity and residential mobility disrupt the social organization of urban neighborhoods in such a way as to reduce their capacity to exercise social control [@Rodrigues2012medo].
+Thus, aspects such as residential instability or low socioeconomic status would contribute to the weakening of the community organization, increasing crime rates.
+This theory highlights the importance of the geography of crime and the characteristics of disorder in its analysis.}
+are thought to occur at relatively small geographical scales.
 
@@
 analytical tools for point patterns, as discussed next.
+
+\hypertarget{religion-in-brazil}{%
+\subsection{Religion in Brazil}\label{religion-in-brazil}}
+
+Brazil is a country with a Christian predominance. In the first national
+census in which religion was considered, in 1872, practically the entire
+population (99.72\%) identified themselves as Catholic. In the 1940
+census, when the religion question was researched again, Catholicism was
+still the religion with which the vast majority of the population
+identified: 95.2\% of Brazilians declared themselves as Catholics that
+year.
+
+The dominance of Catholicism in the religious life of Brazil is
+unsurprising given the colonial past of the country. However, the
+religious makeup of the population started was not completely
+homogeneous. Among other offerings in the market place of spiritual
+ideas, Spiritism stands out. The Spiritist Doctrine was founded in
+nineteenth-century France by Allan Kardec as a way of understanding the
+world and its relations with the ``beyond'': a curious mix of
+pseudo-scientific language, philosophy, and religion (Arribas 2011a).
+Spiritism arrived in Brazil in the middle of the previous century and
+since then it slowly spread in large and medium-sized cities. Today,
+Spiritism is the third largest religious segment in Brazil, after
+Catholicism and Evangelical Christianism, and shares with these
+religions the Christian principle of charity through pious assistance.
+
+Another signal event in the developmento of religion in Brazil is the
+arrival of Pentecostal Protestant churches, such as the Christian
+Congregation of Brazil in 1910 and the Assembly of God in 1911.
+Initially, these religious denominations did not make important inroads
+in Brazilian society (as evidenced by the 1940 census). However, in the
+context of rapid industrialization and major urbanization, they
+contributed to the change in the religious landscape along with other
+religions (Souza 2019): according to the 2010 census the configuration
+of the religious scene in Brazil had changed considerably from the
+previous century: 64.6\% of the population identified as Catholic and
+22.2\% as Evangelical Christians. The acceleration of this demographic
+change was directly related to the inroads made by Evangelical churches:
+The Assembly of God, for instance, grew considerably to total more than
+12.3 million adherents. This church is represented extensively in many
+regions of the country, following a system of loosely allied ministries.
+With a charismatic profile and the lack of a centralized authority,
+growth of this church has occurred mainly through subdivision into
+numerous small and autonomous churches. These churches have in common
+the Prosperity Gospel and an emphasis on economic entrepreneurship
+(e.g., Mora 2008). They also do not shy away from partisan political
+action.
+
+In general, there have been two major changes in Brazil regarding
+religious issues: the slow decline of Catholicism and the rapid growth
+of Evangelical Christianism. These changes are evident in the daily life
+of the average Brazilian, either through the radio or through TV
+programs, which give more and more space to religious doctrines. At the
+same time, there has been an increase in cases of religious intolerance
+in Brazil, generating violence acts against temples, objects of worship,
+and people.
 
@@
 capital cities in Brazil with the highest homicide rates in the period
-under study (Menezes et al. 2013).
+under
+study\footnote{The appendix included information that allows a more in-depth characterization of the city of Recife: population distribution, income, inequality, urban facilities (education and health) and location of residential areas.}
+(Menezes et al. 2013).
 
@@
 The data are organized at the individual level and it is possible to
-obtain information about the location, day of the week, day of the
-month, period of the day, as well as gender, age, and race of the
+obtain information about the crime event location, day of the week, day
+of the month, period of the day, as well as gender, age, and race of the
 victim. The database used in this study comprises the period from July
@@
 
-\includegraphics{Moral_Communities_and_Crime_v1_files/figure-latex/fig-plot-crime-1} \hfill{}
+\includegraphics{Moral_Communities_and_Crime_files/figure-latex/fig-plot-crime-1} \hfill{}
 
@@
 Statistical Purposes was queried to extract facilities other than places
-of worship. As discussed above, the idea is to identify points of
-reference that can be used as controls, having a neutral morality
-profile. For the sake of the present study, we selected pharmacies, ice
-cream shops, and bakeries to construct our control group. These three
-types of establishments comply with the criteria of being morally
-neutral and having a spatial distribution commensurate with places of
-worship.
+of worship. To study the spatial relationship between churches and
+crime, a control group was built and then a comparison of the empirical
+distribution was carried out. Gilberston and Hayes (2012) highlight that
+different configurations of urban space affect how people perceive the
+environment around them. In this perspective, churches are continually
+affecting the impression that people have of a region. Thus, we assume
+that churches have unique characteristics that may or may not inhibit
+the occurrence of a criminal act: they are urban constructions that are
+not neutral in relation to crime. On the other hand, there are urban
+facilities that do not provide any specific reason to justify the fact
+that the crime occurs closer or further from these buildings. Thus, it
+can be thought that any urban construction belongs to one of the
+following distinct situations: neutral or non-neutral to crime.
+Pharmacies, ice cream parlors and bakeries will be used in the
+construction of the control group and will be considered neutral. The
+comparison between the distributions of non-neutral and crime-neutral
+establishments forms the empirical basis of the study. The choice of
+establishments that make up the control was based on two criteria: (a)
+being neutral to crime and (b) having a spatial distribution similar to
+non-neutral establishments. The strategy used is to compare similar
+establishments about their spatial distribution, but different in
+relation to their neutrality profile in relation to crime, seeking to
+simulate the existence of a case and a control group. The case will be
+composed of churches, while the control will be pharmacies, ice cream
+parlors and bakeries.
+
+As discussed above, the idea is to identify points of reference that can
+be used as controls, having a neutral morality profile. For the sake of
+the present study, we selected pharmacies, ice cream shops, and bakeries
+to construct our control group. These three types of establishments
+comply with the criteria of being morally neutral and having a spatial
+distribution commensurate with places of worship.
 
@@
 the number of points, but the locations of cases and controls are quite
-similar.
+similar: the distribution of the churches, as well as of the commercial
+establishments follow the same pattern, being correlated with the
+spatial distribution of the population (see in the appendix). Since such
+urban constructions respond to society's demand, they tend to be located
+close to population agglomerations.
 
@@
 
-\includegraphics{Moral_Communities_and_Crime_v1_files/figure-latex/fig-plot-events-1} \hfill{}
+\includegraphics{Moral_Communities_and_Crime_files/figure-latex/fig-plot-events-1} \hfill{}
 
@@
 \midrule
+Population Density & 0.00 & 10564.16 & 16333.62 & 19230.23 & 25106.07 & 181642.55 & 0.03\\
 Median Income & 122.05 & 337.40 & 581.67 & 1031.09 & 1283.79 & 7954.19 & -0.42\\
@@
 Proportion Single Mother & 0.00 & 0.33 & 0.37 & 0.37 & 0.41 & 0.70 & 0.17\\
+\addlinespace
 Proportion Young Single Mother & 0.00 & 0.01 & 0.02 & 0.02 & 0.02 & 0.21 & 0.38\\
@@
 \centering
-\includegraphics{Moral_Communities_and_Crime_v1_files/figure-latex/plot-sed-as-quintiles-1.pdf}
+\includegraphics{Moral_Communities_and_Crime_files/figure-latex/plot-sed-as-quintiles-1.pdf}
 \caption{\label{fig:plot-sed-as-quintiles}Socio-Economic Deprivation in
@@
 events at each level of the covariate, and then calculating an estimator
-of the intensity.
+of the intensity. The global intensity of each type of event (i.e.,
+places of worship and businesses), as well as quadrant counts and
+intensity at each level of SED can be consulted in Table
+\ref{tab:quadrat-count-results}.
+
+\begin{table}
+
+\caption{\label{tab:table-counts-intensity-summary}\label{tab:quadrat-count-results}Conditional quadrat counts results}
+\centering
+\begin{tabular}[t]{lcccl}
+\toprule
+Event & Global Intensity & SED Quintile & Quadrat Count & Intensity\\
+\midrule
+ &  & Low & 201 & 0\\
+
+ &  & Middle Low & 295 & 0\\
+
+ &  & Middle & 364 & 0\\
+
+ &  & Middle High & 479 & 0\\
+
+\multirow{-5}{*}{\raggedright\arraybackslash Violent Crime} &  & High & 318 & 0\\
+
+ &  & Low & 32 & 0\\
+
+ &  & Middle Low & 26 & 0\\
+
+ &  & Middle & 33 & 0\\
+
+ &  & Middle High & 29 & 0\\
+
+\multirow{-5}{*}{\raggedright\arraybackslash Catholic} &  & High & 11 & 0\\
+
+ &  & Low & 209 & 0\\
+
+ &  & Middle Low & 304 & 0\\
+
+ &  & Middle & 334 & 0\\
+
+ &  & Middle High & 345 & 0\\
+
+\multirow{-5}{*}{\raggedright\arraybackslash Evangelical} &  & High & 185 & 0\\
+
+ &  & Low & 37 & 0\\
+
+ &  & Middle Low & 50 & 0\\
+
+ &  & Middle & 40 & 0\\
+
+ &  & Middle High & 40 & 0\\
+
+\multirow{-5}{*}{\raggedright\arraybackslash Spiritist} &  & High & 20 & 0\\
+
+ &  & Low & 45 & 0\\
+
+ &  & Middle Low & 60 & 0\\
+
+ &  & Middle & 39 & 0\\
+
+ &  & Middle High & 37 & 0\\
+
+\multirow{-5}{*}{\raggedright\arraybackslash Ice Cream} &  & High & 26 & 0\\
+
+ &  & Low & 116 & 0\\
+
+ &  & Middle Low & 122 & 0\\
+
+ &  & Middle & 91 & 0\\
+
+ &  & Middle High & 84 & 0\\
+
+\multirow{-5}{*}{\raggedright\arraybackslash Pharmacies} &  & High & 25 & 0\\
+
+ &  & Low & 27 & 0\\
+
+ &  & Middle Low & 32 & 0\\
+
+ &  & Middle & 38 & 0\\
+
+ &  & Middle High & 16 & 0\\
+
+\multirow{-5}{*}{\raggedright\arraybackslash Super Markets} & \multirow{-35}{*}{\centering\arraybackslash 0} & High & 12 & 0\\
+\bottomrule
+\end{tabular}
+\end{table}
 
@@
 and the intensity in areas with the highest levels of Socio-Economic
-Deprivation is 208.71\% higher than in places with the lowest levels of
+Deprivation is 216.2\% higher than in places with the lowest levels of
 Socio-Economic Deprivation. This is as expected, as research has
@@
 \centering
-\includegraphics{Moral_Communities_and_Crime_v1_files/figure-latex/plot-crime-quadrat-1.pdf}
+\includegraphics{Moral_Communities_and_Crime_files/figure-latex/plot-crime-quadrat-1.pdf}
 \caption{\label{fig:plot-crime-quadrat}Intensity of crime by level of
@@
 \centering
-\includegraphics{Moral_Communities_and_Crime_v1_files/figure-latex/plot-events-quadrat-1.pdf}
+\includegraphics{Moral_Communities_and_Crime_files/figure-latex/plot-events-quadrat-1.pdf}
 \caption{\label{fig:plot-events-quadrat}Intensity of places of worship
@@
 \midrule
-Catholic & 6.913 & 4 & 0.2811\\
-Evangelical & 38.669 & 4 & $<0.001$\\
-Spiritist & 9.108 & 4 & 0.1169\\
-Ice Cream Shop & 15.108 & 4 & 0.009\\
-Pharmacy & 51.869 & 4 & $<0.001$\\
+Catholic & 7.074 & 4 & 0.264\\
+Evangelical & 32.976 & 4 & $<0.001$\\
+Spiritist & 7.561 & 4 & 0.2181\\
+Ice Cream Shop & 13.946 & 4 & 0.0149\\
+Pharmacy & 53.220 & 4 & $<0.001$\\
 \addlinespace
-Supermarket & 23.826 & 4 & $<0.001$\\
+Supermarket & 18.819 & 4 & 0.0017\\
 \bottomrule
@@
 \centering
-\includegraphics{Moral_Communities_and_Crime_v1_files/figure-latex/figure-plot-relative-distribution-1.pdf}
+\includegraphics{Moral_Communities_and_Crime_files/figure-latex/figure-plot-relative-distribution-1.pdf}
 \caption{\label{fig:plot-relative-distribution}Intensity of crime as a
@@
 The peak intensity as a function of distance to Evangelical churches is
-3.33 times the baseline intensity; for Catholic places of worship, the
+3.32 times the baseline intensity; for Catholic places of worship, the
 highest intensity is 2.12 times the baseline intensity; and the
@@
 three types of commercial establishments, we see that the highest
-intensities are 2.28, 2.46, and 2.09 times the intensity according to
+intensities are 2.28, 2.45, and 2.09 times the intensity according to
 Socio-Economic Deprivation for ice cream shops, pharmacies, and
@@
 by means of a baseline function, the peak relative intensity as a
-function of distance to Evangelical churches declines to 176.53\% higher
+function of distance to Evangelical churches declines to 176.48\% higher
 than the intensity as a function of distance to Catholic places of
@@
 comparing Evangelical and Spiritist churches at the distance of the peak
-intensity of Evangelical churches, this difference declines to 186.45\%
+intensity of Evangelical churches, this difference declines to 186.43\%
 after accounting for Socio-Economic Deprivation. Similarly, this changes
-from 174.19\% to 168.33\% in the case of ice cream shops, from 151.52\%
-to 140.97\% in the case of pharmacies, and finally from 177.63\% to
-164.33\% in the case of supermarkets.
+from 174.19\% to 168.37\% in the case of ice cream shops, from 151.52\%
+to 140.98\% in the case of pharmacies, and finally from 177.63\% to
+164.31\% in the case of supermarkets.
 
@@
 \centering
-\includegraphics{Moral_Communities_and_Crime_v1_files/figure-latex/figure-plot-relative-distribution-with-baseline-1.pdf}
+\includegraphics{Moral_Communities_and_Crime_files/figure-latex/figure-plot-relative-distribution-with-baseline-1.pdf}
 \caption{\label{fig:plot-relative-distribution-with-baseline}Intensity
@@
 context of Brazil where it adapted in ways that made it very different
-from its original French predecessor (see Graca Arribas 2011). Could it
-be that Spiritism is seen by potential criminals as spiritually or
+from its original French predecessor (see Arribas 2011b). Could it be
+that Spiritism is seen by potential criminals as spiritually or
 materially risky, if supernatural retribution is feared? Again, we
@@
 Clustering.'' Journal Article. \emph{Jurnal Teknologi} 80 (4): 43--53.
-\href{\%3CGo\%20to\%20ISI\%3E://WOS:000437011000005}{\textless{}Go to ISI\textgreater{}://WOS:000437011000005}.
+\url{\%3CGo\%20to\%20ISI\%3E://WOS:000437011000005}.
 
@@
 
+\leavevmode\hypertarget{ref-Appleyard1980livable}{}%
+Appleyard, D. 1980. ``Livable Streets - Protected Neighborhoods.''
+Journal Article. \emph{Annals of the American Academy of Political and
+Social Science} 451 (SEP): 106--17.
+\url{\%3CGo\%20to\%20ISI\%3E://A1980KK14700011}.
+
 \leavevmode\hypertarget{ref-Arnold2018}{}%
@@
 for 'Ggplot2'}. \url{https://CRAN.R-project.org/package=ggthemes}.
+
+\leavevmode\hypertarget{ref-Arribas2011doutrina}{}%
+Arribas, Celia da Graca. 2011a. ``A Doutrina Espirita Na Formacao Da
+Diversidade Religiosa Brasileira.'' \emph{Anais Do XXVI Simpósio
+Nacional de História - ANPUH, Sao Paulo}.
+
+\leavevmode\hypertarget{ref-Arribas2011espiritismo}{}%
+---------. 2011b. ``Espiritismo: Entre Crime E Religiao.''
+\emph{Mneme-Revista de Humanidades} 12 (29).
 
@@
 Geographical Information Science} 21 (3): 299--323.
-\href{ISI:000244944300004\%0AC:/Papers/International\%20Journal\%20of\%20Geographical\%20Information\%20Science/IJGIS\%20(2007)\%2021\%20(3)\%20299-323.pdf}{ISI:000244944300004
-C:/Papers/International Journal of Geographical Information Science/IJGIS (2007) 21 (3) 299-323.pdf}.
+\url{ISI:000244944300004\%0AC:/Papers/International\%20Journal\%20of\%20Geographical\%20Information\%20Science/IJGIS\%20(2007)\%2021\%20(3)\%20299-323.pdf}.
 
@@
 
-\leavevmode\hypertarget{ref-daGraca2011espiritismo}{}%
-Graca Arribas, Celia da. 2011. ``Espiritismo: Entre Crime E Religiao.''
-\emph{Mneme-Revista de Humanidades} 12 (29).
+\leavevmode\hypertarget{ref-Gilbertson2012engineering}{}%
+Gilbertson, Alan, and Alan Hayes. 2012. ``Engineering to Reduce Crime
+and Disorder in Public Places.'' In \emph{Proceedings of the Institution
+of Civil Engineers-Municipal Engineer}, 165:175--83. 3. Thomas Telford
+Ltd.
 
@@
 
+\leavevmode\hypertarget{ref-Jacobs1961death}{}%
+Jacobs, Jane. 1961. \emph{The Death and Life of Great American Cities}.
+Book. New York: Vintage.
+
+\leavevmode\hypertarget{ref-Jacobs1968community}{}%
+---------. 1968. ``Community on the City Streets.'' Book Section. In
+\emph{The Search for Community in Modern America}, edited by D.
+Baltzell, 74--93. New York: Harper \& Row.
+
 \leavevmode\hypertarget{ref-Johnson2011more}{}%
@@
 120 (1): 97--99. \url{https://doi.org/10.1016/j.econlet.2013.03.040}.
+
+\leavevmode\hypertarget{ref-Mora2008marketing}{}%
+Mora, G. Cristina. 2008. ``Marketing the `Health and Wealth Gospel'
+Across National Borders; Evidence from Brazil and the United States.''
+\emph{Poetics} 36 (5): 404--20.
+\url{https://doi.org/https://doi.org/10.1016/j.poetic.2008.06.008}.
 
@@
 \emph{Sociological Quarterly} 44 (4): 523--54.
-\href{\%3CGo\%20to\%20ISI\%3E://WOS:000188319800001}{\textless{}Go to ISI\textgreater{}://WOS:000188319800001}.
+\url{\%3CGo\%20to\%20ISI\%3E://WOS:000188319800001}.
 
@@
 \url{https://CRAN.R-project.org/package=fasterize}.
+
+\leavevmode\hypertarget{ref-deSouza2019pluralidade}{}%
+Souza, André Ricardo de. 2019. ``Pluralidade Cristã E Algumas Questões
+Do Cenário Religioso Brasileiro.'' \emph{Revista USP}, no. 120: 13--22.
 
@@
 
+\leavevmode\hypertarget{ref-Taylor1988human}{}%
+Taylor, Ralph B. 1988. \emph{Human Territorial Functioning: An
+Empirical, Evolutionary Perspective on Individual and Small Group
+Territorial Cognitions, Behaviors, and Consequences}. Book. Cambridge
+University Press.
+
+\leavevmode\hypertarget{ref-Taylor1998crime}{}%
+Taylor, Ralph B. 1998. ``Crime and Small-Scale Places: What We Know,
+What We Can Prevent, and What Else We Need to Know.'' Book Section. In
+\emph{Crime and Place: Plenary Papers of the 1997 Conference on Criminal
+Justice Research and Evaluation}, edited by Ralph B. Taylor, 1--22.
+Washington, D.C.: National Institute of Justice.
+
 \leavevmode\hypertarget{ref-Topalli2013god}{}%

### 2020-06-11 Moral-Communities-and-Crime/Response-to-Reviewers-v2.Rmd
v2 Finished revisions, sent back to Edivaldo and Tatiane for comments.

--- 2020-06-10 Response-to-Reviewers-v2.Rmd
@@
 
-We would like to thank the editor and two anonymous reviewers for their comments and suggestions to revise the paper.
+We would like to thank the editor and two anonymous reviewers for their thoughtful comments and suggestions to revise the paper.
+We believe that the paper has been substantially improved by their input.
 In this letter we respond to the feedback received, and describe the actions taken to revise the paper.
@@
 
->\textcolor{blue}{We believe that this argument is logical and sound, and churches, in fact, have been studied in the past for their potential to supress crime (Abdullah, 2018; Davignon, 2015; Furr, 2010; Traunmuller, 2011).
+>\textcolor{blue}{We believe that this argument is logical and sound, and churches, in fact, have been studied in the past for their potential to suppress crime (Abdullah, 2018; Davignon, 2015; Furr, 2010; Traunmuller, 2011).
 A more novel aspect is the study of places of worship, as distinct entities from the members of the congregations: as Warner and Konkel (2019) note, this is an aspect that has received less attention in empirical and theoretical research.
@@
 In this sense, we would note that aggregating point data to areas inevitably involves loss of information.
-Sometimes this situation is unavoidable (say, statistics are recorded for areas, so locational information is lost from the origin.) But when we have access to detailed geographical information, converting to area data is an unecessary loss of information.
+Sometimes this situation is unavoidable (say, statistics are recorded for areas, so locational information is lost from the origin.) But when we have access to detailed geographical information, converting to area data is an unnecessary loss of information.
 The question that we are asking is actually quite geographical: if there is evidence of a moral community effect, what is its distance decay?
@@
 >\textcolor{blue}{Thank you for this comment.
-As pointed out above, the theoretical work is already discussed by Groff, among others.
-Although conceptually the study draws from much earlier research (e.g., Appleyard, Jacobs, Taylor), empirical work has lagged, as noted by Warner and Konkel (2019).
+As pointed out above, the theoretical work was discussed by Groff, among others.
+Although conceptually the study draws from much earlier research (e.g., Appleyard, Jacobs, Taylor), empirical work has lagged, as recently noted by Warner and Konkel (2019).
 This is where the current paper aims to make a contribution.}
@@
 In particular, we find that Catholic churches do not "project" any more moral community than pharmacies, which means that there is probably not much that can be done in terms of leveraging them for informal social control purposes.
-Spiritis churches have a lower relative risk at relatively short distances of less than approximately 50 m (which would be the side of a small block).
+Spiritist churches have a lower relative risk at relatively short distances of less than approximately 50 m (which would be the side of a small block).
 Why is this?
@@
 
->\textcolor{blue}{Words.}
+>\textcolor{blue}{We hope that the revisions elsewhere (e.g., an improved presentation of the argument behind the analysis) have resolved this issue.}
 
@@
 
->\textcolor{blue}{Thank you for this comment. We agree, there are many methods that could be applied. In our analysis we use two tools:}
+>\textcolor{blue}{Thank you for this comment. We agree, there are many methods that could be applied. In our analysis we use three tools:}
 
->\textcolor{blue}{1. The relative distribution, which is akin to a smooth estimate of the intensity as a function of distance.}
+>\textcolor{blue}{1. Quadrat counts, which are discrete estimates of the intensity of a point process in a region.}
+
+>\textcolor{blue}{1. The relative distribution, which is a smooth estimate of the intensity of a process as a function of distance to a signal event.}
 
@@
 
->\textcolor{blue}{In the previous version of the paper, our SED score was calculated using several potential confounders of crime: median income, proportion of popultion unemployed, proportion of households with a single mother, proportion of households with a young single mother, proportion of population in poverty.
+>\textcolor{blue}{In the previous version of the paper, our SED score was calculated using several potential confounders of crime: median income, proportion of population unemployed, proportion of households with a single mother, proportion of households with a young single mother, proportion of population in poverty.
 In the current version of the paper, we also include population density in the calculation of the SED score (thank you for the suggestion).
@@
 >\textcolor{blue}{We believe that this approach is informative, and since we are interested in the distribution of crime in the proximity of places of worship, it is also appropriate.
-Furthermore, by considering other underlying risk factors (in the form of the SED score), we believe that we have dilligently covered most reasonable confounders.
+Furthermore, by considering other underlying risk factors (in the form of the SED score), we believe that we have diligently covered most reasonable confounders.
 We have edited the text to explain this.}
@@
 
->\textcolor{blue}{Words.}
+>\textcolor{blue}{Thank you for this comment. The intensity is in events/square meter, except in the case of the relative distribution with a baseline function, which is dimensionless (see below).}
 
 When discussing the quantitative results throughout Sections 5.3 and 5.4, please add uncertainty intervals so that readers can interpret the degree to which two quantities are, in fact, different?
-For example, p.16-line 49: "reaching a peak of 2.4916 * 10e-5 (95% CI: 2.1 * 10e-5, 3.2 * 10e-5)." (ANTÔNIO)
+For example, p.16-line 49: "reaching a peak of 2.4916 * 10e-5 (95% CI: 2.1 * 10e-5, 3.2 * 10e-5)."
 
->\textcolor{blue}{Words.}
+>\textcolor{blue}{Thank you for this comment. The intervals of confidence are now reported in the figures.}
 
@@
 For example, what is B(s), how is it calculated, and what does it mean for understanding the distribution of crime in relation to churches?
-(ANTÔNIO)
 
->\textcolor{blue}{Words.}
+>\textcolor{blue}{Both quadrats and the relative distribution function are estimators of intensity.
+In the case of quadrats, the estimator is the number of events in the quadrat divided by area of the quadrat (conditional quadrat counts use irregular quadrats that are defined based on some control variable, in our case SED).
+In the case of the relative distribution, the output is an estimate of intensity conditioned on distance from a signal event, i.e., a place of worship.
+The relative distribution can be adjusted by means of a baseline function B(s), which is also an estimate of intensity but conditioned on some control covariate (in our cases SED).
+Therefore, the interpretation of the relative distribution with a baseline function is as a ratio of the intensity with respect to distance to the baseline intensity according to the control covariate.
+If the ratio, which is dimensionless, is 1, then intensity as a function of distance is equal to the baseline intensity.}
 
@@
 
->\textcolor{blue}{The appendix included information that allows a more in-depth characterization of the city of Recife: population distribution, income, inequality, urban facilities (education and health) and location of residential areas.
-We also have included a brief explanation of the requested clustering pattern in the text.}
+>\textcolor{blue}{We now include a figure with maps for socio-economic and demographic variables which help to provide context to SED.
+We also have included a brief explanation of the requested clustering pattern in the text and recreated the figures after adding some transparecy to compensate for overplotting.
+This should make it easier to appreciate the patterns.}
 
@@
 When using quadrats in the earlier part of the analysis, the region is partitioned using quintiles; this is a requirement, since quadrats function as "containers" and need to be partitioned in discrete segments.
+In Figure 4 each color is a "container".
 However, when calculating the distribution functions, the continuous value of the SED score is used.
@@
 
->\textcolor{blue}{Words.}
+>\textcolor{blue}{The results are conditioned by the size of the quadrat, but in this case since the quadrats are defined based on a covariate, we have a conceptual reason to prefer them to other partitions (each quintile in the SED is one quadrat).
+We have explained in the text how the quadrats function as containers for counting events.}
 
@@
 
->\textcolor{blue}{Words.}
+>\textcolor{blue}{Thank you for this comment. We have recreated the figures with some transparency to compensate for overplotting.}
 
@@
 
->\textcolor{blue}{We have included in the text an analysis of the religions studied in the article.}
+>\textcolor{blue}{We have included in the text some additional information of the religions studied in the article.}
 
@@
 
->\textcolor{blue}{Words.}
+>\textcolor{blue}{The results for the various controls are qualitatively similar both for quadrat analysis and relative distribution.
+After choosing these three types of establishments semi-randomly and verifying their similarities, we think it unlikely that other types of common establishments would change the results.
+In this case simulation of a random variable is not useful, because it would detect clustering of all the events.
+Rather, what we are interested in is clustering of one type of event with respect to another.}
 
@@
 
->\textcolor{blue}{Words.}
+>\textcolor{blue}{This is a very perceptive comment.
+In fact, the approach that Openshaw pioneered was later converted into several forms of spatial scan statistics.
+Its legacy is still felt here: we can explore co-location patterns by considering signal events that we are interested in for theoretical/conceptual reasons (i.e., places of worship).}
 
@@
 
->\textcolor{blue}{Words.}
+>\textcolor{blue}{The specific results are applicable to Recife only.
+However, the general principle could be more general, and it aligns well with one to two ideas: the Violence of the Just, or the possibility that violence is a consequence of religious intolerance.
+Irrespective of which of these two propositions is true, the general principle could be in operation in areas where conservative Evangelicals are present (similar results, although at the aggregated level, have been reported by Desmond et al., 2010).
+For this reason, Vienna might be a stretch, but other cities in Latin America and North America not so much.}
 
@@
 
->\textcolor{blue}{Words.}
+>\textcolor{blue}{We have made an effort to provide additional context, which we hope will make the paper more approachable to readers who are less familiar with the Latin American context.}

### 2020-06-11 Moral-Communities-and-Crime/Moral_Communities_and_Crime.Rmd
v2 Finished revisions, sent back to Edivaldo and Tatiane for comments.

@@
 # Load packages used in this document
-
-library(tidyverse)
+library(fasterize)
+library(ggthemes)
+library(gridExtra)
 library(kableExtra)
+library(maptools)
+library(raster)
 library(sf)
 library(spatstat)
-library(maptools)
-library(ggthemes)
-library(raster)
-library(fasterize)
+library(tidyverse)
 ```
@@
 
-2. If so, what is the distance decay of said moral community effect? Is it uniform over a certain distance or does it vary?
+2. If so, what is the distance decay of said moral community effect? Is it uniform over a certain distance, such that it could be well approximated by areas, or does it vary?
 
@@
 The empirical strategy is to use spatial analysis to explore the potential geographical relationships between violent criminal events, on the one hand, and places of worship and a selection of commercial establishments that serve as controls, on the other.
-Dissagregated data allows us to analyze the relevant point patterns at a very high level of resolution.
-
-After this introductory section, the rest of this paper is structured as follows: Section \ref{background} reviews some relevant theoretical perspectives on religion and crime; then Section \ref{methods} presents the empirical strategy used in the work.
-Section \ref{case} presents the context as well as the data used in the study; Section \ref{results} presents the analysis, whereas the results are discussed in Section \ref{discussion}.
-Finally, some concluding remarks are offered in Section \ref{conclusions}.
+
+Note that this is a reproducible research document.
+The analysis is implemented using the `R` statistical computing language, and documents to replicate the analysis, in addition to all data necessary, are available from a repository\footnote{\url{https://drive.google.com/open?id=1tuJM4Mhi0Ftq3ZEwjv6RP9vO3veGjIR_}}.
 
@@
 2).
-This point is echoed by Warner and Konkel [-@Warner2019neighborhood], who moreover recommend the use of even smaller units of analysis (e.g., Census Block Groups), in an effort to reduce aggregation bias [see @Hipp2007block], but more importantly because the processes defined by social disorganization theory\footnote{The social disorganization theory (Shaw e McKay, 1942)  presents the social structure as an indispensable factor for the elaboration of theoretical and methodological models related to crime in urban neighborhoods.
+This point is echoed by Warner and Konkel [-@Warner2019neighborhood], who moreover recommend the use of smaller units of analysis (e.g., Census Block Groups), in an effort to reduce aggregation bias [see @Hipp2007block], but more importantly because the processes defined by social disorganization theory\footnote{The social disorganization theory (Shaw e McKay, 1942)  presents the social structure as an indispensable factor for the elaboration of theoretical and methodological models related to crime in urban neighborhoods.
 According to Villareal e Silva [-@Villarreal2006social], “neighborhoods characterized by dense social networks will experience greater trust among residents and cooperation in the enforcement of social norms against crime.
@@
 
+Groff [@Groff2015informal], in particular, explains the small-scale mechanisms that could be in operation in terms of "Eyes on the Street" and Human Territorial Functioning.
+"Eyes on the Street" [see @Jacobs1961death; @Jacobs1968community; @Appleyard1980livable] refers to involvment in small civic acts by users of public spaces, which signal social organization to an observer.
+Community engagement, moreover, is engendered at small scales, as shown by authors that include Appleyard [-@Appleyard1980livable] and more recently Grannis [-@Grannis2009from].
+Human Territorial Functioning [HTF; @Taylor1988human; @Taylor1998crime] also provides a valuable perspective on social informal control This concept posits that territorial functioning is a product of the shared perceptions of individuals in regards to places, and how these perceptions are shaped by the physical, social, and cultural environments.
+For example, Paez [-@Paez2013mapping] discusses how perceptions of safety are the outcome of a socially interactive process, and shows how this process leads to spatial patterning at very small-scales.
 The preceding review, as well as other research [-@Malleson2019identifying], makes clear the need for research at various geographical scales.
 From a theoretical perspective, the social mechanisms that underpin the hypothesis of moral communities and informal social control can happen at various geographical scales, some of which previous research has already addressed: the work of Nie and Yang [-@Nie2019smoking], for one, highlights the role of processes detectable at a relatively high level of aggregation, whereas Warner and Konkel [-@Warner2019neighborhood], looking specifically at the effect of places of worship, bring the study closer to the level of the neighborhood.
-In what follows, we propose to adopt a much more disaggregated approach by considering the intensity of criminal events with respect to places of worship.
-This requires micro-level data both of crimes and places of worship, and the use of appropriate spatial analytical tools for point patterns, as discussed next.
 
@@
 
-Brazil is a country with a Christian predominance.
-In the first national census in which religion was considered, in 1872, practically the entire population (99.72%) identified themselves as Catholic.
-In the 1940 census, when the religion question was researched again, Catholicism was still the religion with which the vast majority of the population identified: 95.2% of Brazilians declared themselves as Catholics that year.
-
-The dominance of Catholicism in the religious life of Brazil is unsurprising given the colonial past of the country.
-However, the religious makeup of the population started was not completely homogeneous.
+Brazil is a predominantly Christian country.
+The first national census that collected information on religion was in 1872, a year when practically the entire population (99.72%) identified themselves as Catholic.
+In the 1940 census, when the religion question was asked again, Catholicism was still the dominant religion with 95.2% of Brazilians declaring themselves as Catholics.
+The dominance of Catholicism in the religious life of Brazil is unsurprising given the country's past as a former colony of a Catholic power (Portugal).
+However, the religious makeup of the population started to become less homogeneous in the second half of the twentieth century.
 Among other offerings in the market place of spiritual ideas, Spiritism stands out.
-The Spiritist Doctrine was founded in nineteenth-century France by Allan Kardec as a way of understanding the world and its relations with the "beyond": a curious mix of pseudo-scientific language, philosophy, and religion [@Arribas2011doutrina].
-Spiritism arrived in Brazil in the middle of the previous century and since then it slowly spread in large and medium-sized cities.
+The Spiritist Doctrine was founded in nineteenth-century France by Allan Kardec as a way of understanding the world and its relations with the "beyond", in a curious mix of pseudo-scientific language, philosophy, and religion [@Arribas2011doutrina].
+Spiritism arrived in Brazil in the middle of the nineteenth century, and since then it slowly spread in large and medium-sized cities.
 Today, Spiritism is the third largest religious segment in Brazil, after Catholicism and Evangelical Christianism, and shares with these religions the Christian principle of charity through pious assistance.
 
-Another signal event in the developmento of religion in Brazil is the arrival of Pentecostal Protestant churches, such as the Christian Congregation of Brazil in 1910 and the Assembly of God in 1911.
+Another signal event in the developmento of religion in Brazil was the arrival of Pentecostal Protestant churches, such as the Christian Congregation of Brazil in 1910 and the Assembly of God in 1911.
 Initially, these religious denominations did not make important inroads in Brazilian society (as evidenced by the 1940 census).
 However, in the context of rapid industrialization and major urbanization, they contributed to the change in the religious landscape along with other religions [@deSouza2019pluralidade]: according to the 2010 census the configuration of the religious scene in Brazil had changed considerably from the previous century: 64.6% of the population identified as Catholic and 22.2% as Evangelical Christians.
-The acceleration of this demographic change was directly related to the inroads made by Evangelical churches: The Assembly of God, for instance, grew considerably to total more than 12.3 million adherents.
+The acceleration of this demographic change was directly related to successful proselitism by Evangelical churches: The Assembly of God, for instance, is a church that grew considerably to total more than 12.3 million adherents.
 This church is represented extensively in many regions of the country, following a system of loosely allied ministries.
@@
 These changes are evident in the daily life of the average Brazilian, either through the radio or through TV programs, which give more and more space to religious doctrines.
-At the same time, there has been an increase in cases of religious intolerance in Brazil, generating violence acts against temples, objects of worship, and people.
+At the same time, there has been an increase in cases of religious intolerance in Brazil, generating violence acts against temples, objects of worship, and people [e.g., @deSouza2019pluralidade; @Fonseca2018primeiras].
+
+In what follows, we propose to adopt a disaggregated approach by considering the intensity of criminal events with respect to places of worship.
+This requires micro-level data both of crimes and places of worship, and the use of appropriate spatial analytical tools for point patterns, as discussed next.
 
@@
 
-where $\lambda(s)$ is the intensity of the process at point $s$, given by the expected number of events in a small area $\text{d}s$ around $s$, as the area becomes arbitrarily small.
+\noindent where $\lambda(s)$ is the intensity of the process at point $s$, given by the expected number of events in a small area $\text{d}s$ around $s$, as the area becomes arbitrarily small.
 If the process is homogeneous (i.e., the intensity is a constant over space), an apt estimator of the intensity is the global intensity, simply the number of events divided by the area of the region under analysis:
@@
 It can be seen that $\rho(z)=1$ is the baseline intensity; values of $\rho(z)>1$ correspond to intensities higher than the baseline _as a function of_ $Z(u)$, whereas values of $\rho(z)<1$ correspond to intensities lower than the baseline, again as a function of $Z(u)$.
+
+## Interpretation
+
+Both quadrats and the relative distribution function are estimators of intensity.
+In the case of quadrats, the estimator is the number of events in the quadrat divided by area of the quadrat (conditional quadrat counts use irregular quadrats that are defined based on some control variable, in our case SED).
+In the case of the relative distribution, the output is an estimate of intensity conditioned on distance from a signal event, e.g., a place of worship.
+The relative distribution can be adjusted by means of a baseline function $B(s)$, which is also an estimate of intensity but conditioned on some control covariate (in our cases SED).
+Therefore, the interpretation of the relative distribution with a baseline function is as a ratio of the intensity with respect to distance, to the baseline intensity according to the control covariate.
+If the ratio, which is dimensionless, is 1, then intensity as a function of distance is equal to the baseline intensity.
 
@@
 ggplot(data = events_recife.sf %>% filter(Type == "Violent Crime")) + 
-  geom_sf(color = "red") +
+  geom_sf(color = "red", alpha = 0.1, shape = 16) +
   geom_sf(data = recife.sf, fill = NA) + 
@@
 ggplot(data = events_recife.sf %>% filter(Type != "Violent Crime")) + 
-  geom_sf(aes(color = Type, fill = Type)) +
+  geom_sf(aes(color = Type, fill = Type), alpha = 0.1, shape = 16) +
   geom_sf(data = recife.sf, fill = NA) + 
@@
 
-The analysis is implemented using the `R` statistical computing language, and documents to replicate the analysis, in addition to all data necessary, can be obtained from this anonymous shared folder:
-
-https://drive.google.com/open?id=1tuJM4Mhi0Ftq3ZEwjv6RP9vO3veGjIR_
-
 ## Socio-Economic Deprivation
@@
 
-sed_descriptives <- data.frame(variable = c("Population Density",
-                                            "Median Income", 
+sed_descriptives <- data.frame(variable = c("Median Income", 
                                             "Proportion Unemployed", 
@@
                                             "Proportion Single Mother", 
-                    "Proportion Young Single Mother"),
-           min = c(min(setores.sf$pop_density),
-                   min(setores.sf$median_income),
+                                            "Proportion Young Single Mother",
+                                            "Population Density"),
+                               min = c(min(setores.sf$median_income),
                                        min(setores.sf$prop_unemployed),
@@
                                        min(setores.sf$prop_single_mother),
-                   min(setores.sf$prop_young_single_mother)),
-           second_q = c(quantile(setores.sf$pop_density, c(0, 0.25, 0.5, 0.75, 1))[2],
-                        quantile(setores.sf$median_income, c(0, 0.25, 0.5, 0.75, 1))[2],
+                                       min(setores.sf$prop_young_single_mother),
+                                       min(setores.sf$pop_density)),
+                               second_q = c(quantile(setores.sf$median_income, c(0, 0.25, 0.5, 0.75, 1))[2],
                                             quantile(setores.sf$prop_unemployed, c(0, 0.25, 0.5, 0.75, 1))[2],
@@
                                             quantile(setores.sf$prop_single_mother, c(0, 0.25, 0.5, 0.75, 1))[2],
-                        quantile(setores.sf$prop_young_single_mother, c(0, 0.25, 0.5, 0.75, 1))[2]),
-           median = c(median(setores.sf$pop_density),
-                      median(setores.sf$median_income),
+                                            quantile(setores.sf$prop_young_single_mother, c(0, 0.25, 0.5, 0.75, 1))[2],
+                                            quantile(setores.sf$pop_density, c(0, 0.25, 0.5, 0.75, 1))[2]),
+                               median = c(median(setores.sf$median_income),
                                           median(setores.sf$prop_unemployed),
@@
                                           median(setores.sf$prop_single_mother),
-                      median(setores.sf$prop_young_single_mother)),
-           mean = c(mean(setores.sf$pop_density),
-                    mean(setores.sf$median_income),
+                                          median(setores.sf$prop_young_single_mother),
+                                          median(setores.sf$pop_density)),
+                               mean = c(mean(setores.sf$median_income),
                                         mean(setores.sf$prop_unemployed),
@@
                                         mean(setores.sf$prop_single_mother),
-                    mean(setores.sf$prop_young_single_mother)),
-           third_q = c(quantile(setores.sf$pop_density, c(0, 0.25, 0.5, 0.75, 1))[4],
-             quantile(setores.sf$median_income, c(0, 0.25, 0.5, 0.75, 1))[4],
+                                        mean(setores.sf$prop_young_single_mother),
+                                        mean(setores.sf$pop_density)),
+                               third_q = c(quantile(setores.sf$median_income, c(0, 0.25, 0.5, 0.75, 1))[4],
                                            quantile(setores.sf$prop_unemployed, c(0, 0.25, 0.5, 0.75, 1))[4],
@@
                                            quantile(setores.sf$prop_single_mother, c(0, 0.25, 0.5, 0.75, 1))[4],
-                       quantile(setores.sf$prop_young_single_mother, c(0, 0.25, 0.5, 0.75, 1))[4]),
-           max = c(max(setores.sf$pop_density),
-                   max(setores.sf$median_income),
+                                           quantile(setores.sf$prop_young_single_mother, c(0, 0.25, 0.5, 0.75, 1))[4],
+                                           quantile(setores.sf$pop_density, c(0, 0.25, 0.5, 0.75, 1))[4]),
+                               max = c(max(max(setores.sf$median_income),
                                            max(setores.sf$prop_unemployed),
@@
                                            max(setores.sf$prop_single_mother),
-                   max(setores.sf$prop_young_single_mother)),
-           pca_loadings = as.numeric(setores.pca$rotation[c(1, 2, 3, 6, 4, 5), 1]))
+                                           max(setores.sf$prop_young_single_mother),
+                                           setores.sf$pop_density)),
+                               pca_loadings = as.numeric(setores.pca$rotation[c(2, 3, 6, 4, 5, 1), 1]))
 ```
@@
 In the present case, we have for the city of Recife information about the variables shown in Table \ref{tab:sed-descriptive-statistics} at the level of _setores_, a small Brazilian census geography.
-
-As seen in the table, Recife is a city with large socio-economic and demographic disparities; for example, the _setor_ with the highest median income has a median income that is `r prettyNum(filter(sed_descriptives, variable == "Median Income")["max"]/filter(sed_descriptives, variable == "Median Income")["min"] * 100, digits = 2, big.mark = ",")`% higher than the median income in the _setor_ with the lowest median income.
-Whereas there are _setores_ where the proportion of population who are unemployed is zero, there are _setores_ where almost 60% of the population are unemployed.
-Likewise, there are _setores_ where almost 70% of the population lives in poverty.
-In addition to these economic indicators, two variables are used to represent family disruption, the proportion of families whose head is a single mother, and the proportion of families whose head is a _young_ single mother, that is, a woman between the ages of 15 and 25.
-As can be seen, there are _setores_ with approximately 70% of households led by single mothers, and of these, over 21% are led by younger women, indicating a high degree of family disruption.
-
-For the analysis, we use Principal Component Analysis, a data reduction technique, to obtain an indicator of Socio-Economic Deprivation, essentially the first principal component in the output.
-The loadings of this factor (Table \ref{tab:sed-descriptive-statistics}) indicate that high Socio-Economic Deprivation is a combination of (in terms of importance): high levels of poverty, high levels of unemployment, and low median income, followed by high levels of family disruption, in particular proportion of families led by young single mothers.
-This factor accounts for almost 54% of the variance..
 
@@
 
+As seen in the table, Recife is a city with large socio-economic and demographic disparities; for example, the _setor_ with the highest median income has a median income that is `r prettyNum(filter(sed_descriptives, variable == "Median Income")["max"]/filter(sed_descriptives, variable == "Median Income")["min"] * 100, digits = 2, big.mark = ",")`% higher than the median income in the _setor_ with the lowest median income.
+Whereas there are _setores_ where the proportion of population who are unemployed is zero, there are _setores_ where almost 60% of the population are unemployed.
+Likewise, there are _setores_ where almost 70% of the population lives in poverty.
+In addition to these economic indicators, two variables are used to represent family disruption, the proportion of families whose head is a single mother, and the proportion of families whose head is a _young_ single mother, that is, a woman between the ages of 15 and 25.
+As can be seen, there are _setores_ with approximately 70% of households led by single mothers, and of these, over 21% are led by younger women, indicating a high degree of family disruption.
+Unlike other places, where higher population density is associated with more poverty, the distribution of this variable in Recife is more mixed: middle income and higher income households favor high density development in many parts of the city.
+The spatial distribution of these variables can be seen in Figure \ref{fig:plot-sed-variables}.
+
+```{r fig-plot-sed-variables, echo=FALSE, fig.height= 9, fig.align = "left", fig.cap="\\label{fig:plot-sed-variables}Spatial distribution of socio-economic and demographic variables in Recife"}
+
+median_income.plot <- ggplot(setores.sf) + 
+  geom_sf(aes(fill = median_income), color = NA) +
+  geom_sf(data = recife.sf, fill = NA) + 
+  scale_fill_viridis_c(name = "Median Income", option = "viridis", direction = -1) +
+  theme_tufte()  +
+  theme(axis.text.x = element_text(angle = 90),
+        legend.position = "bottom")
+
+prop_unemployed.plot <- ggplot(setores.sf) + 
+  geom_sf(aes(fill = prop_unemployed), color = NA) +
+  geom_sf(data = recife.sf, fill = NA) + 
+  scale_fill_viridis_c(name = "Prop. Unemployed", option = "viridis", direction = -1) +
+  theme_tufte()  +
+  theme(axis.text.x = element_text(angle = 90),
+        legend.position = "bottom")
+
+prop_poverty.plot <- ggplot(setores.sf) + 
+  geom_sf(aes(fill = prop_poverty), color = NA) +
+  geom_sf(data = recife.sf, fill = NA) + 
+  scale_fill_viridis_c(name = "Prop. Poverty", option = "viridis", direction = -1) +
+  theme_tufte()  +
+  theme(axis.text.x = element_text(angle = 90),
+        legend.position = "bottom")
+
+prop_unemployed.plot <- ggplot(setores.sf) + 
+  geom_sf(aes(fill = prop_unemployed), color = NA) +
+  geom_sf(data = recife.sf, fill = NA) + 
+  scale_fill_viridis_c(name = "Prop. Unemployed", option = "viridis", direction = -1) +
+  theme_tufte()  +
+  theme(axis.text.x = element_text(angle = 90),
+        legend.position = "bottom")
+
+prop_single_mother.plot <- ggplot(setores.sf) + 
+  geom_sf(aes(fill = prop_single_mother), color = NA) +
+  geom_sf(data = recife.sf, fill = NA) + 
+  scale_fill_viridis_c(name = "Prop. Single Mother", option = "viridis", direction = -1) +
+  theme_tufte()  +
+  theme(axis.text.x = element_text(angle = 90),
+        legend.position = "bottom")
+
+prop_young_single_mother.plot <- ggplot(setores.sf) + 
+  geom_sf(aes(fill = prop_young_single_mother), color = NA) +
+  geom_sf(data = recife.sf, fill = NA) + 
+  scale_fill_viridis_c(name = "Prop. Young Single Mother", option = "viridis", direction = -1) +
+  theme_tufte()  +
+  theme(axis.text.x = element_text(angle = 90),
+        legend.position = "bottom")
+
+pop_density.plot <- ggplot(setores.sf) + 
+  geom_sf(aes(fill = pop_density), color = NA) +
+  geom_sf(data = recife.sf, fill = NA) + 
+  scale_fill_viridis_c(name = "Pop. Density", option = "viridis", direction = -1) +
+  theme_tufte()  +
+  theme(axis.text.x = element_text(angle = 90),
+        legend.position = "bottom")
+
+grid.arrange(median_income.plot, 
+             prop_poverty.plot, 
+             prop_unemployed.plot, 
+             prop_single_mother.plot, 
+             prop_young_single_mother.plot, 
+             pop_density.plot,
+             nrow = 3)
+```
+
+For the analysis, we use Principal Component Analysis, a data reduction technique, to obtain an indicator of Socio-Economic Deprivation, essentially the first principal component in the output.
+The loadings of this factor (Table \ref{tab:sed-descriptive-statistics}) indicate that high Socio-Economic Deprivation is a combination of (in terms of importance): high levels of poverty, high levels of unemployment, and low median income, followed by high levels of family disruption, in particular proportion of families led by young single mothers.
+This factor accounts for almost 45% of the variance.
+
 Figure \ref{fig:plot-sed-as-quintiles} displays the geography of Socio-Economic Deprivation in Recife, after classifying _setores_ by quintiles, whereby the "High" class corresponds to _setores_ in the top 20% of the Socio-Economic Deprivation indicator, and the "Low" class corresponds to _setores_ in the bottom 20% of the Socio-Economic Deprivation indicator.
@@
 As previously described, the quadrats are irregular areas that are defined based on a covariate, in this case the quintiles of the Socio-Economic Deprivation indicator.
+Each color in Figure \ref{fig:plot-sed-as-quintiles} is a "container" for counting events.
 In essence, this entails counting the number of events at each level of the covariate, and then calculating an estimator of the intensity.
@@
   xlab("Socio-Economic Deprivation") +
+  ylab(expression(Intensity~(events/m^{2}))) +
   geom_hline(aes(yintercept = Global_Intensity), linetype = "dotted", size = 1) +
@@
   geom_abline(intercept = 7.575416e-06, slope = 0, linetype = "dotted") +
+  xlab("z (distance in m)") +
+  ylab(expression(rho~(events/m^{2}))) +
   #xlim(c(0, 500)) +
@@
   geom_abline(intercept = 1, slope = 0, linetype = "dotted") +
+  xlab("z (distance in m)") +
+  ylab(expression(rho)) +
   #xlim(c(0, 500)) +
@@
 The objective of this research was to investigate the covariation of incidence of crime to proximity of places of worship.
-The hypothesis of moral communities (closely related to social disorganization theory) posits that a negative correlation between the presence of churches and crime.
+The hypothesis of moral communities (closely related to social informal controls) posits a negative correlation between the presence of churches and crime.
 Recent papers have argued that it is important to study this phenomenon from a geographical perspective, paying attention to the level of aggregation.
@@
 2.
-Following the suggestion of Warner and Konkel [-@Warner2019neighborhood], we also disaggregate our data by type of place of worship, and consider the three following denominations: Catholic, Evangelical, and Spiritist.
-
-2. Use of relative distribution functions is effectively a multi-scale technique, since the intensity of crime can be estimated at different distances from various types of places of worship.
+Following the suggestion of Warner and Konkel [-@Warner2019neighborhood], we also disaggregated our data by type of place of worship, and considered the three following denominations: Catholic, Evangelical, and Spiritist.
+
+2. We used relative distribution functions, a multi-scale technique, that allowed us to estimate the intensity of crime at different distances from various types of places of worship.
 
@@
 3.
-Spiritist churches are associated with lower intensity of crime at very short distance; however, at longer distances(approximately between 200 m and 500 m), the intensity of crime is in fact higher than the intensity of crime at comparable distances from Catholic and Evangelical places of worship, or any of the three kinds of commercial establishments.
-
-These results are interesting of and in themselves as they help to clarify the potential of various types of moral communities to act as restraints on crime or, on the contrary, as criminogenic factors.
-They also suggest several avenues for further research.
+Spiritist churches are associated with lower intensity of crime at very short distance; however, at longer distances (approximately between 200 m and 500 m), the intensity of crime is in fact higher than the intensity of crime at comparable distances from Catholic and Evangelical places of worship, or any of the three kinds of commercial establishments.
+
+These results are interesting in and of themselves as they help to clarify the potential of various types of moral communities to act as restraints on crime or, on the contrary, as criminogenic factors.
+They also suggest several recommendations.
 
@@
 And finally, in relation to the latter point, more in-depth research, perhaps observational, ethnographic, or participatory studies, could help to develop a more refined understanding of the kinds of processes that operate at different scales and how they can influence moral behavior and crime.
+For example, an important question is whether the higher intensity of crime in the neighborhood of Evangelical churches is related to "The Violence of the Just" [see @Birman2012violencia; @Birman2019narrativas], or contrariwise, crime committed _against_ Evangelical Christians as a result of religious intolerance [see @deSouza2019pluralidade; @Fonseca2018primeiras].
 

### 2020-06-11 Moral-Communities-and-Crime/Moral_Communities_and_Crime.tex
v2 Finished revisions, sent back to Edivaldo and Tatiane for comments.

@@
   If so, what is the distance decay of said moral community effect? Is
-  it uniform over a certain distance or does it vary?
+  it uniform over a certain distance, such that it could be well
+  approximated by areas, or does it vary?
 \end{enumerate}
@@
 the one hand, and places of worship and a selection of commercial
-establishments that serve as controls, on the other. Dissagregated data
-allows us to analyze the relevant point patterns at a very high level of
-resolution.
-
-After this introductory section, the rest of this paper is structured as
-follows: Section \ref{background} reviews some relevant theoretical
-perspectives on religion and crime; then Section \ref{methods} presents
-the empirical strategy used in the work. Section \ref{case} presents the
-context as well as the data used in the study; Section \ref{results}
-presents the analysis, whereas the results are discussed in Section
-\ref{discussion}. Finally, some concluding remarks are offered in
-Section \ref{conclusions}.
+establishments that serve as controls, on the other.
+
+Note that this is a reproducible research document. The analysis is
+implemented using the \texttt{R} statistical computing language, and
+documents to replicate the analysis, in addition to all data necessary,
+are available from a
+repository\footnote{\url{https://drive.google.com/open?id=1tuJM4Mhi0Ftq3ZEwjv6RP9vO3veGjIR_}}.
 
@@
 point is echoed by Warner and Konkel (2019), who moreover recommend the
-use of even smaller units of analysis (e.g., Census Block Groups), in an
+use of smaller units of analysis (e.g., Census Block Groups), in an
 effort to reduce aggregation bias (see Hipp 2007), but more importantly
@@
 
-The preceding review, as well as other research (2019), makes clear the
-need for research at various geographical scales. From a theoretical
+Groff (Groff 2015), in particular, explains the small-scale mechanisms
+that could be in operation in terms of ``Eyes on the Street'' and Human
+Territorial Functioning. ``Eyes on the Street'' (see Jacobs 1961, 1968;
+Appleyard 1980) refers to involvment in small civic acts by users of
+public spaces, which signal social organization to an observer.
+Community engagement, moreover, is engendered at small scales, as shown
+by authors that include Appleyard (1980) and more recently Grannis
+(2009). Human Territorial Functioning (HTF; Taylor 1988, 1998) also
+provides a valuable perspective on social informal control This concept
+posits that territorial functioning is a product of the shared
+perceptions of individuals in regards to places, and how these
+perceptions are shaped by the physical, social, and cultural
+environments. For example, Paez (2013) discusses how perceptions of
+safety are the outcome of a socially interactive process, and shows how
+this process leads to spatial patterning at very small-scales. The
+preceding review, as well as other research (2019), makes clear the need
+for research at various geographical scales. From a theoretical
 perspective, the social mechanisms that underpin the hypothesis of moral
@@
 places of worship, bring the study closer to the level of the
-neighborhood. In what follows, we propose to adopt a much more
-disaggregated approach by considering the intensity of criminal events
-with respect to places of worship. This requires micro-level data both
-of crimes and places of worship, and the use of appropriate spatial
-analytical tools for point patterns, as discussed next.
+neighborhood.
 
@@
 
-Brazil is a country with a Christian predominance. In the first national
-census in which religion was considered, in 1872, practically the entire
-population (99.72\%) identified themselves as Catholic. In the 1940
-census, when the religion question was researched again, Catholicism was
-still the religion with which the vast majority of the population
-identified: 95.2\% of Brazilians declared themselves as Catholics that
-year.
-
-The dominance of Catholicism in the religious life of Brazil is
-unsurprising given the colonial past of the country. However, the
-religious makeup of the population started was not completely
-homogeneous. Among other offerings in the market place of spiritual
-ideas, Spiritism stands out. The Spiritist Doctrine was founded in
-nineteenth-century France by Allan Kardec as a way of understanding the
-world and its relations with the ``beyond'': a curious mix of
-pseudo-scientific language, philosophy, and religion (Arribas 2011a).
-Spiritism arrived in Brazil in the middle of the previous century and
-since then it slowly spread in large and medium-sized cities. Today,
-Spiritism is the third largest religious segment in Brazil, after
-Catholicism and Evangelical Christianism, and shares with these
-religions the Christian principle of charity through pious assistance.
-
-Another signal event in the developmento of religion in Brazil is the
+Brazil is a predominantly Christian country. The first national census
+that collected information on religion was in 1872, a year when
+practically the entire population (99.72\%) identified themselves as
+Catholic. In the 1940 census, when the religion question was asked
+again, Catholicism was still the dominant religion with 95.2\% of
+Brazilians declaring themselves as Catholics. The dominance of
+Catholicism in the religious life of Brazil is unsurprising given the
+country's past as a former colony of a Catholic power (Portugal).
+However, the religious makeup of the population started to become less
+homogeneous in the second half of the twentieth century. Among other
+offerings in the market place of spiritual ideas, Spiritism stands out.
+The Spiritist Doctrine was founded in nineteenth-century France by Allan
+Kardec as a way of understanding the world and its relations with the
+``beyond'', in a curious mix of pseudo-scientific language, philosophy,
+and religion (Arribas 2011a). Spiritism arrived in Brazil in the middle
+of the nineteenth century, and since then it slowly spread in large and
+medium-sized cities. Today, Spiritism is the third largest religious
+segment in Brazil, after Catholicism and Evangelical Christianism, and
+shares with these religions the Christian principle of charity through
+pious assistance.
+
+Another signal event in the developmento of religion in Brazil was the
 arrival of Pentecostal Protestant churches, such as the Christian
@@
 22.2\% as Evangelical Christians. The acceleration of this demographic
-change was directly related to the inroads made by Evangelical churches:
-The Assembly of God, for instance, grew considerably to total more than
-12.3 million adherents. This church is represented extensively in many
-regions of the country, following a system of loosely allied ministries.
-With a charismatic profile and the lack of a centralized authority,
-growth of this church has occurred mainly through subdivision into
-numerous small and autonomous churches. These churches have in common
-the Prosperity Gospel and an emphasis on economic entrepreneurship
-(e.g., Mora 2008). They also do not shy away from partisan political
-action.
+change was directly related to successful proselitism by Evangelical
+churches: The Assembly of God, for instance, is a church that grew
+considerably to total more than 12.3 million adherents. This church is
+represented extensively in many regions of the country, following a
+system of loosely allied ministries. With a charismatic profile and the
+lack of a centralized authority, growth of this church has occurred
+mainly through subdivision into numerous small and autonomous churches.
+These churches have in common the Prosperity Gospel and an emphasis on
+economic entrepreneurship (e.g., Mora 2008). They also do not shy away
+from partisan political action.
 
@@
 in Brazil, generating violence acts against temples, objects of worship,
-and people.
+and people (e.g., Souza 2019; Fonseca 2018).
+
+In what follows, we propose to adopt a disaggregated approach by
+considering the intensity of criminal events with respect to places of
+worship. This requires micro-level data both of crimes and places of
+worship, and the use of appropriate spatial analytical tools for point
+patterns, as discussed next.
 
@@
 
-where \(\lambda(s)\) is the intensity of the process at point \(s\),
-given by the expected number of events in a small area \(\text{d}s\)
-around \(s\), as the area becomes arbitrarily small. If the process is
-homogeneous (i.e., the intensity is a constant over space), an apt
-estimator of the intensity is the global intensity, simply the number of
-events divided by the area of the region under analysis: \[
+\noindent where \(\lambda(s)\) is the intensity of the process at point
+\(s\), given by the expected number of events in a small area
+\(\text{d}s\) around \(s\), as the area becomes arbitrarily small. If
+the process is homogeneous (i.e., the intensity is a constant over
+space), an apt estimator of the intensity is the global intensity,
+simply the number of events divided by the area of the region under
+analysis: \[
 \hat{\lambda} = \frac{n}{|\mathcal{A}|}
@@
 lower than the baseline, again as a function of \(Z(u)\).
+
+\hypertarget{interpretation}{%
+\subsection{Interpretation}\label{interpretation}}
+
+Both quadrats and the relative distribution function are estimators of
+intensity. In the case of quadrats, the estimator is the number of
+events in the quadrat divided by area of the quadrat (conditional
+quadrat counts use irregular quadrats that are defined based on some
+control variable, in our case SED). In the case of the relative
+distribution, the output is an estimate of intensity conditioned on
+distance from a signal event, e.g., a place of worship. The relative
+distribution can be adjusted by means of a baseline function \(B(s)\),
+which is also an estimate of intensity but conditioned on some control
+covariate (in our cases SED). Therefore, the interpretation of the
+relative distribution with a baseline function is as a ratio of the
+intensity with respect to distance, to the baseline intensity according
+to the control covariate. If the ratio, which is dimensionless, is 1,
+then intensity as a function of distance is equal to the baseline
+intensity.
 
@@
 
-The analysis is implemented using the \texttt{R} statistical computing
-language, and documents to replicate the analysis, in addition to all
-data necessary, can be obtained from this anonymous shared folder:
-
-\url{https://drive.google.com/open?id=1tuJM4Mhi0Ftq3ZEwjv6RP9vO3veGjIR_}
-
 \hypertarget{socio-economic-deprivation}{%
@@
 
+\begin{table}
+
+\caption{\label{tab:table-sed-descriptive-statistics}\label{tab:sed-descriptive-statistics}Descriptive statistics of some key socio-economic and demographic variables in Recife}
+\centering
+\begin{tabular}[t]{lccccccc}
+\toprule
+Variable & Min & 2nd Quartile & Median & Mean & 3rd Quartile & Max & \makecell[l]{PC Factor 1\\ Loadings}\\
+\midrule
+Median Income & 122.05 & 337.40 & 581.67 & 1031.09 & 1283.79 & 181642.5 & -0.42\\
+Proportion Unemployed & 0.00 & 0.05 & 0.10 & 0.10 & 0.14 & 181642.5 & 0.56\\
+Proportion Poverty & 0.00 & 0.05 & 0.11 & 0.12 & 0.17 & 181642.5 & 0.58\\
+Proportion Single Mother & 0.00 & 0.33 & 0.37 & 0.37 & 0.41 & 181642.5 & 0.17\\
+Proportion Young Single Mother & 0.00 & 0.01 & 0.02 & 0.02 & 0.02 & 181642.5 & 0.38\\
+\addlinespace
+Population Density & 0.00 & 10564.16 & 16333.62 & 19230.23 & 25106.07 & 181642.5 & 0.03\\
+\bottomrule
+\multicolumn{8}{l}{\textit{Note: }}\\
+\multicolumn{8}{l}{The variable for young single mothers is for women aged 15-25}\\
+\multicolumn{8}{l}{The first principal component accounts for 44.9\% of the variance}\\
+\end{tabular}
+\end{table}
+
 As seen in the table, Recife is a city with large socio-economic and
 demographic disparities; for example, the \emph{setor} with the highest
-median income has a median income that is 6,517\% higher than the median
-income in the \emph{setor} with the lowest median income. Whereas there
-are \emph{setores} where the proportion of population who are unemployed
-is zero, there are \emph{setores} where almost 60\% of the population
-are unemployed. Likewise, there are \emph{setores} where almost 70\% of
-the population lives in poverty. In addition to these economic
-indicators, two variables are used to represent family disruption, the
-proportion of families whose head is a single mother, and the proportion
-of families whose head is a \emph{young} single mother, that is, a woman
-between the ages of 15 and 25. As can be seen, there are \emph{setores}
-with approximately 70\% of households led by single mothers, and of
-these, over 21\% are led by younger women, indicating a high degree of
-family disruption.
+median income has a median income that is 148,826\% higher than the
+median income in the \emph{setor} with the lowest median income. Whereas
+there are \emph{setores} where the proportion of population who are
+unemployed is zero, there are \emph{setores} where almost 60\% of the
+population are unemployed. Likewise, there are \emph{setores} where
+almost 70\% of the population lives in poverty. In addition to these
+economic indicators, two variables are used to represent family
+disruption, the proportion of families whose head is a single mother,
+and the proportion of families whose head is a \emph{young} single
+mother, that is, a woman between the ages of 15 and 25. As can be seen,
+there are \emph{setores} with approximately 70\% of households led by
+single mothers, and of these, over 21\% are led by younger women,
+indicating a high degree of family disruption. Unlike other places,
+where higher population density is associated with more poverty, the
+distribution of this variable in Recife is more mixed: middle income and
+higher income households favor high density development in many parts of
+the city. The spatial distribution of these variables can be seen in
+Figure \ref{fig:plot-sed-variables}.
+
+\begin{figure}
+
+\includegraphics{Moral_Communities_and_Crime_files/figure-latex/fig-plot-sed-variables-1} \hfill{}
+
+\caption{\label{fig:plot-sed-variables}Spatial distribution of socio-economic and demographic variables in Recife}\label{fig:fig-plot-sed-variables}
+\end{figure}
 
@@
 particular proportion of families led by young single mothers. This
-factor accounts for almost 54\% of the variance..
-
-\begin{table}
-
-\caption{\label{tab:table-sed-descriptive-statistics}\label{tab:sed-descriptive-statistics}Descriptive statistics of some key socio-economic and demographic variables in Recife}
-\centering
-\begin{tabular}[t]{lccccccc}
-\toprule
-Variable & Min & 2nd Quartile & Median & Mean & 3rd Quartile & Max & \makecell[l]{PC Factor 1\\ Loadings}\\
-\midrule
-Population Density & 0.00 & 10564.16 & 16333.62 & 19230.23 & 25106.07 & 181642.55 & 0.03\\
-Median Income & 122.05 & 337.40 & 581.67 & 1031.09 & 1283.79 & 7954.19 & -0.42\\
-Proportion Unemployed & 0.00 & 0.05 & 0.10 & 0.10 & 0.14 & 0.59 & 0.56\\
-Proportion Poverty & 0.00 & 0.05 & 0.11 & 0.12 & 0.17 & 0.63 & 0.58\\
-Proportion Single Mother & 0.00 & 0.33 & 0.37 & 0.37 & 0.41 & 0.70 & 0.17\\
-\addlinespace
-Proportion Young Single Mother & 0.00 & 0.01 & 0.02 & 0.02 & 0.02 & 0.21 & 0.38\\
-\bottomrule
-\multicolumn{8}{l}{\textit{Note: }}\\
-\multicolumn{8}{l}{The variable for young single mothers is for women aged 15-25}\\
-\multicolumn{8}{l}{The first principal component accounts for 53.91\% of the variance}\\
-\end{tabular}
-\end{table}
+factor accounts for almost 45\% of the variance.
 
@@
 based on a covariate, in this case the quintiles of the Socio-Economic
-Deprivation indicator. In essence, this entails counting the number of
-events at each level of the covariate, and then calculating an estimator
-of the intensity. The global intensity of each type of event (i.e.,
-places of worship and businesses), as well as quadrant counts and
-intensity at each level of SED can be consulted in Table
-\ref{tab:quadrat-count-results}.
+Deprivation indicator. Each color in Figure
+\ref{fig:plot-sed-as-quintiles} is a ``container'' for counting events.
+In essence, this entails counting the number of events at each level of
+the covariate, and then calculating an estimator of the intensity. The
+global intensity of each type of event (i.e., places of worship and
+businesses), as well as quadrant counts and intensity at each level of
+SED can be consulted in Table \ref{tab:quadrat-count-results}.
 
@@
 incidence of crime to proximity of places of worship. The hypothesis of
-moral communities (closely related to social disorganization theory)
-posits that a negative correlation between the presence of churches and
-crime. Recent papers have argued that it is important to study this
-phenomenon from a geographical perspective, paying attention to the
-level of aggregation. Accordingly, our research for the city of Recife
-in Brazil makes the following contributions to the literature:
+moral communities (closely related to social informal controls) posits a
+negative correlation between the presence of churches and crime. Recent
+papers have argued that it is important to study this phenomenon from a
+geographical perspective, paying attention to the level of aggregation.
+Accordingly, our research for the city of Recife in Brazil makes the
+following contributions to the literature:
 
@@
   Following the suggestion of Warner and Konkel (2019), we also
-  disaggregate our data by type of place of worship, and consider the
+  disaggregated our data by type of place of worship, and considered the
   three following denominations: Catholic, Evangelical, and Spiritist.
 \item
-  Use of relative distribution functions is effectively a multi-scale
-  technique, since the intensity of crime can be estimated at different
-  distances from various types of places of worship.
+  We used relative distribution functions, a multi-scale technique, that
+  allowed us to estimate the intensity of crime at different distances
+  from various types of places of worship.
 \item
@@
   Spiritist churches are associated with lower intensity of crime at
-  very short distance; however, at longer distances(approximately
+  very short distance; however, at longer distances (approximately
   between 200 m and 500 m), the intensity of crime is in fact higher
@@
 
-These results are interesting of and in themselves as they help to
+These results are interesting in and of themselves as they help to
 clarify the potential of various types of moral communities to act as
 restraints on crime or, on the contrary, as criminogenic factors. They
-also suggest several avenues for further research.
+also suggest several recommendations.
 
@@
   that operate at different scales and how they can influence moral
-  behavior and crime.
+  behavior and crime. For example, an important question is whether the
+  higher intensity of crime in the neighborhood of Evangelical churches
+  is related to ``The Violence of the Just'' (see Birman and Machado
+  2012; Birman 2019), or contrariwise, crime committed \emph{against}
+  Evangelical Christians as a result of religious intolerance (see Souza
+  2019; Fonseca 2018).
 \end{enumerate}
@@
 
+\leavevmode\hypertarget{ref-Fonseca2018primeiras}{}%
+Fonseca, Alexandre Brasil. 2018. ``Primeiras Análises Dos Dados Do
+Relatório Sobre Intolerância E Violência Religiosa No Brasil
+(2011-2015).'' Journal Article. \emph{Estado Laico, Intolerância E
+Diversidade Religiosa No Brasil. Brasília, Ministério Dos Direitos
+Humanos}, 22--47.
+
 \leavevmode\hypertarget{ref-Foster2010neighbourhood}{}%
@@
 
+\leavevmode\hypertarget{ref-Grannis2009from}{}%
+Grannis, R. 2009. \emph{From the Ground up: Translating Geography into
+Community Through Neighbor Networks}. Book. Princeton: Princeton
+University Press.
+
 \leavevmode\hypertarget{ref-Groff2015informal}{}%
@@
 Sciences}, edited by N. Wrigley, 127--44. London: Pion.
+
+\leavevmode\hypertarget{ref-Paez2013mapping}{}%
+Paez, A. 2013. ``Mapping Travelers' Attitudes: Does Space Matter?''
+Journal Article. \emph{Journal of Transport Geography} 26: 117--25.
+\url{https://doi.org/10.1016/j.jtrangeo.2012.09.002}.
 

### 2020-06-13 Moral-Communities-and-Crime/Response-to-Reviewers-v2.Rmd
v2 Second submission after comments from Edivaldo and Tatiane

@@
 >\textcolor{blue}{We believe that this argument is logical and sound, and churches, in fact, have been studied in the past for their potential to suppress crime (Abdullah, 2018; Davignon, 2015; Furr, 2010; Traunmuller, 2011).
-A more novel aspect is the study of places of worship, as distinct entities from the members of the congregations: as Warner and Konkel (2019) note, this is an aspect that has received less attention in empirical and theoretical research.
-}
+A more novel aspect is the study of places of worship, as distinct entities from the members of the congregations: as Warner and Konkel (2019) note, this is an aspect that has received less attention in empirical and theoretical research.}
 
@@
 >\textcolor{blue}{Thank you for this comment.
-As pointed out above, the theoretical work was discussed by Groff, among others.
-Although conceptually the study draws from much earlier research (e.g., Appleyard, Jacobs, Taylor), empirical work has lagged, as recently noted by Warner and Konkel (2019).
+As pointed out above, the theoretical work was discussed by Groff (2015), among others.
+Although conceptually the study draws from much earlier research (e.g., Appleyard, 1980; Jacobs, 1961, 1968; Taylor, 1988, 1998), empirical work has lagged, as recently noted by Warner and Konkel (2019).
 This is where the current paper aims to make a contribution.}
@@
 
->\textcolor{blue}{The appendix included information that allows a more in-depth characterization of the city of Recife: population distribution, income, inequality, urban facilities (education and health) and location of residential areas.}
+>\textcolor{blue}{We now include a figure with maps for socio-economic and demographic variables which help to provide additional context.}
 

### 2020-06-13 Moral-Communities-and-Crime/Moral_Communities_and_Crime.Rmd
v2 Second submission after comments from Edivaldo and Tatiane

@@
 
-The case study is the city of Recife in the state of Pernambuco, in Brazil's Northeast.
+The empirical analysis uses geographically detailed data from the city of Recife in the state of Pernambuco, in Brazil's Northeast.
 Recife is a large and important metropolitan area in a historically poor region, and afflicted by high levels of violent crime, having the dubious honor of being one of five state capitals with the highest rates of homicide in the period under study [@Menezes2013spatial].
@@
 The database used in this study comprises the period from July 1, 2008 to June 30, 2010.
-Some descriptive statistics regarding this dataset are reported in Table 1, where it can be seen that overall, about 91% of the victims of LIVC in the period analyzed were men, while approximately 98% of violent crimes were homicides.
+Some descriptive statistics regarding this dataset are reported in Table 1, where it can be seen that overall, about 91% of the victims of LIVC in the period analyzed were men.
 In addition, most of the victims were black or brown, and the youth population between the ages of 18 and 30 is the most affected by violent crime.
@@
                                            quantile(setores.sf$pop_density, c(0, 0.25, 0.5, 0.75, 1))[4]),
-                               max = c(max(max(setores.sf$median_income),
+                               max = c(max(setores.sf$median_income),
                                            max(setores.sf$prop_unemployed),
@@
                                            max(setores.sf$prop_young_single_mother),
-                                           setores.sf$pop_density)),
+                                           max(setores.sf$pop_density)),
                                pca_loadings = as.numeric(setores.pca$rotation[c(2, 3, 6, 4, 5, 1), 1]))
@@
 As seen in the table, Recife is a city with large socio-economic and demographic disparities; for example, the _setor_ with the highest median income has a median income that is `r prettyNum(filter(sed_descriptives, variable == "Median Income")["max"]/filter(sed_descriptives, variable == "Median Income")["min"] * 100, digits = 2, big.mark = ",")`% higher than the median income in the _setor_ with the lowest median income.
-Whereas there are _setores_ where the proportion of population who are unemployed is zero, there are _setores_ where almost 60% of the population are unemployed.
+Whereas there are _setores_ where the proportion of population who are unemployed is zero, there are _setores_ where almost 60% of the population are unemployed (which for some may mean they are part of the informal econom).
 Likewise, there are _setores_ where almost 70% of the population lives in poverty.
@@
 In essence, this entails counting the number of events at each level of the covariate, and then calculating an estimator of the intensity.
-The global intensity of each type of event (i.e., places of worship and businesses), as well as quadrant counts and intensity at each level of SED can be consulted in Table \ref{tab:quadrat-count-results}.
+The global intensity of each type of event, as well as the quadrant counts and intensity at each level of SED can be consulted in Table \ref{tab:quadrat-count-results}.
 
@@
                                                      summary_evangelic$intensity,
-                                                     summary_spiritist$intensity,
+                                                     summary_spiritist$intensity + 1e-8, #small correction to distinguish from ice cream in table
                                                      summary_ice_cream$intensity,
@@
 ```{r table-counts-intensity-summary, echo=FALSE}
-kable(intensity_all %>% dplyr::select(Type, Global_Intensity, SED, Quadrat_Count, Intensity),
+kable(intensity_all %>% 
+        dplyr::select(Type, Global_Intensity, SED, Quadrat_Count, Intensity) %>% 
+        mutate(Global_Intensity = prettyNum(Global_Intensity, digits = 3),
+               Intensity = prettyNum(Intensity, digits = 3)),
       "latex",
       booktabs = TRUE,
-      digits = 3,
+      #digits = 3,
       escape = FALSE,
@@
   kable_styling(bootstrap_options = c("striped", "scaled_down")) %>% 
-  collapse_rows(1:2, latex_hline = "none")
+  collapse_rows(1:2, latex_hline = "major")
 ```

### 2020-06-13 Moral-Communities-and-Crime/Moral_Communities_and_Crime.tex
v2 Second submission after comments from Edivaldo and Tatiane

@@
 
-The case study is the city of Recife in the state of Pernambuco, in
-Brazil's Northeast. Recife is a large and important metropolitan area in
-a historically poor region, and afflicted by high levels of violent
-crime, having the dubious honor of being one of five state capitals with
-the highest rates of homicide in the period under study (Menezes et al.
-2013). The empirical strategy is to use spatial analysis to explore the
-potential geographical relationships between violent criminal events, on
-the one hand, and places of worship and a selection of commercial
-establishments that serve as controls, on the other.
+The empirical analysis uses geographically detailed data from the city
+of Recife in the state of Pernambuco, in Brazil's Northeast. Recife is a
+large and important metropolitan area in a historically poor region, and
+afflicted by high levels of violent crime, having the dubious honor of
+being one of five state capitals with the highest rates of homicide in
+the period under study (Menezes et al. 2013). The empirical strategy is
+to use spatial analysis to explore the potential geographical
+relationships between violent criminal events, on the one hand, and
+places of worship and a selection of commercial establishments that
+serve as controls, on the other.
 
@@
 dataset are reported in Table 1, where it can be seen that overall,
-about 91\% of the victims of LIVC in the period analyzed were men, while
-approximately 98\% of violent crimes were homicides. In addition, most
-of the victims were black or brown, and the youth population between the
-ages of 18 and 30 is the most affected by violent crime. Lastly, it
-should be noted that about 88\% of the criminal events under analysis
-involved firearms. Figure \ref{fig:plot-crime} shows the spatial
-distribution of the 1,657 LIVC crimes that occurred in the city of
-Recife between July, 2008 and June, 2010.
+about 91\% of the victims of LIVC in the period analyzed were men. In
+addition, most of the victims were black or brown, and the youth
+population between the ages of 18 and 30 is the most affected by violent
+crime. Lastly, it should be noted that about 88\% of the criminal events
+under analysis involved firearms. Figure \ref{fig:plot-crime} shows the
+spatial distribution of the 1,657 LIVC crimes that occurred in the city
+of Recife between July, 2008 and June, 2010.
 
@@
 \midrule
-Median Income & 122.05 & 337.40 & 581.67 & 1031.09 & 1283.79 & 181642.5 & -0.42\\
-Proportion Unemployed & 0.00 & 0.05 & 0.10 & 0.10 & 0.14 & 181642.5 & 0.56\\
-Proportion Poverty & 0.00 & 0.05 & 0.11 & 0.12 & 0.17 & 181642.5 & 0.58\\
-Proportion Single Mother & 0.00 & 0.33 & 0.37 & 0.37 & 0.41 & 181642.5 & 0.17\\
-Proportion Young Single Mother & 0.00 & 0.01 & 0.02 & 0.02 & 0.02 & 181642.5 & 0.38\\
+Median Income & 122.05 & 337.40 & 581.67 & 1031.09 & 1283.79 & 7954.19 & -0.42\\
+Proportion Unemployed & 0.00 & 0.05 & 0.10 & 0.10 & 0.14 & 0.59 & 0.56\\
+Proportion Poverty & 0.00 & 0.05 & 0.11 & 0.12 & 0.17 & 0.63 & 0.58\\
+Proportion Single Mother & 0.00 & 0.33 & 0.37 & 0.37 & 0.41 & 0.70 & 0.17\\
+Proportion Young Single Mother & 0.00 & 0.01 & 0.02 & 0.02 & 0.02 & 0.21 & 0.38\\
 \addlinespace
-Population Density & 0.00 & 10564.16 & 16333.62 & 19230.23 & 25106.07 & 181642.5 & 0.03\\
+Population Density & 0.00 & 10564.16 & 16333.62 & 19230.23 & 25106.07 & 181642.55 & 0.03\\
 \bottomrule
@@
 demographic disparities; for example, the \emph{setor} with the highest
-median income has a median income that is 148,826\% higher than the
-median income in the \emph{setor} with the lowest median income. Whereas
-there are \emph{setores} where the proportion of population who are
-unemployed is zero, there are \emph{setores} where almost 60\% of the
-population are unemployed. Likewise, there are \emph{setores} where
-almost 70\% of the population lives in poverty. In addition to these
-economic indicators, two variables are used to represent family
-disruption, the proportion of families whose head is a single mother,
-and the proportion of families whose head is a \emph{young} single
-mother, that is, a woman between the ages of 15 and 25. As can be seen,
-there are \emph{setores} with approximately 70\% of households led by
-single mothers, and of these, over 21\% are led by younger women,
-indicating a high degree of family disruption. Unlike other places,
-where higher population density is associated with more poverty, the
-distribution of this variable in Recife is more mixed: middle income and
-higher income households favor high density development in many parts of
-the city. The spatial distribution of these variables can be seen in
-Figure \ref{fig:plot-sed-variables}.
+median income has a median income that is 6,517\% higher than the median
+income in the \emph{setor} with the lowest median income. Whereas there
+are \emph{setores} where the proportion of population who are unemployed
+is zero, there are \emph{setores} where almost 60\% of the population
+are unemployed (which for some may mean they are part of the informal
+econom). Likewise, there are \emph{setores} where almost 70\% of the
+population lives in poverty. In addition to these economic indicators,
+two variables are used to represent family disruption, the proportion of
+families whose head is a single mother, and the proportion of families
+whose head is a \emph{young} single mother, that is, a woman between the
+ages of 15 and 25. As can be seen, there are \emph{setores} with
+approximately 70\% of households led by single mothers, and of these,
+over 21\% are led by younger women, indicating a high degree of family
+disruption. Unlike other places, where higher population density is
+associated with more poverty, the distribution of this variable in
+Recife is more mixed: middle income and higher income households favor
+high density development in many parts of the city. The spatial
+distribution of these variables can be seen in Figure
+\ref{fig:plot-sed-variables}.
 
@@
 the covariate, and then calculating an estimator of the intensity. The
-global intensity of each type of event (i.e., places of worship and
-businesses), as well as quadrant counts and intensity at each level of
-SED can be consulted in Table \ref{tab:quadrat-count-results}.
+global intensity of each type of event, as well as the quadrant counts
+and intensity at each level of SED can be consulted in Table
+\ref{tab:quadrat-count-results}.
 
@@
 \midrule
- &  & Low & 201 & 0\\
-
- &  & Middle Low & 295 & 0\\
-
- &  & Middle & 364 & 0\\
-
- &  & Middle High & 479 & 0\\
-
-\multirow{-5}{*}{\raggedright\arraybackslash Violent Crime} &  & High & 318 & 0\\
-
- &  & Low & 32 & 0\\
-
- &  & Middle Low & 26 & 0\\
-
- &  & Middle & 33 & 0\\
-
- &  & Middle High & 29 & 0\\
-
-\multirow{-5}{*}{\raggedright\arraybackslash Catholic} &  & High & 11 & 0\\
-
- &  & Low & 209 & 0\\
-
- &  & Middle Low & 304 & 0\\
-
- &  & Middle & 334 & 0\\
-
- &  & Middle High & 345 & 0\\
-
-\multirow{-5}{*}{\raggedright\arraybackslash Evangelical} &  & High & 185 & 0\\
-
- &  & Low & 37 & 0\\
-
- &  & Middle Low & 50 & 0\\
-
- &  & Middle & 40 & 0\\
-
- &  & Middle High & 40 & 0\\
-
-\multirow{-5}{*}{\raggedright\arraybackslash Spiritist} &  & High & 20 & 0\\
-
- &  & Low & 45 & 0\\
-
- &  & Middle Low & 60 & 0\\
-
- &  & Middle & 39 & 0\\
-
- &  & Middle High & 37 & 0\\
-
-\multirow{-5}{*}{\raggedright\arraybackslash Ice Cream} &  & High & 26 & 0\\
-
- &  & Low & 116 & 0\\
-
- &  & Middle Low & 122 & 0\\
-
- &  & Middle & 91 & 0\\
-
- &  & Middle High & 84 & 0\\
-
-\multirow{-5}{*}{\raggedright\arraybackslash Pharmacies} &  & High & 25 & 0\\
-
- &  & Low & 27 & 0\\
-
- &  & Middle Low & 32 & 0\\
-
- &  & Middle & 38 & 0\\
-
- &  & Middle High & 16 & 0\\
-
-\multirow{-5}{*}{\raggedright\arraybackslash Super Markets} & \multirow{-35}{*}{\centering\arraybackslash 0} & High & 12 & 0\\
+ &  & Low & 201 & 4.6e-06\\
+
+ &  & Middle Low & 295 & 6.75e-06\\
+
+ &  & Middle & 364 & 8.33e-06\\
+
+ &  & Middle High & 479 & 8.65e-06\\
+
+\multirow{-5}{*}{\raggedright\arraybackslash Violent Crime} & \multirow{-5}{*}{\centering\arraybackslash 7.58e-06} & High & 318 & 9.94e-06\\
+\cmidrule{1-5}
+ &  & Low & 32 & 7.32e-07\\
+
+ &  & Middle Low & 26 & 5.95e-07\\
+
+ &  & Middle & 33 & 7.55e-07\\
+
+ &  & Middle High & 29 & 5.24e-07\\
+
+\multirow{-5}{*}{\raggedright\arraybackslash Catholic} & \multirow{-5}{*}{\centering\arraybackslash 5.99e-07} & High & 11 & 3.44e-07\\
+\cmidrule{1-5}
+ &  & Low & 209 & 4.78e-06\\
+
+ &  & Middle Low & 304 & 6.95e-06\\
+
+ &  & Middle & 334 & 7.65e-06\\
+
+ &  & Middle High & 345 & 6.23e-06\\
+
+\multirow{-5}{*}{\raggedright\arraybackslash Evangelical} & \multirow{-5}{*}{\centering\arraybackslash 6.3e-06} & High & 185 & 5.78e-06\\
+\cmidrule{1-5}
+ &  & Low & 37 & 8.46e-07\\
+
+ &  & Middle Low & 50 & 1.14e-06\\
+
+ &  & Middle & 40 & 9.16e-07\\
+
+ &  & Middle High & 40 & 7.22e-07\\
+
+\multirow{-5}{*}{\raggedright\arraybackslash Spiritist} & \multirow{-5}{*}{\centering\arraybackslash 8.65e-07} & High & 20 & 6.25e-07\\
+\cmidrule{1-5}
+ &  & Low & 45 & 1.03e-06\\
+
+ &  & Middle Low & 60 & 1.37e-06\\
+
+ &  & Middle & 39 & 8.93e-07\\
+
+ &  & Middle High & 37 & 6.68e-07\\
+
+\multirow{-5}{*}{\raggedright\arraybackslash Ice Cream} & \multirow{-5}{*}{\centering\arraybackslash 9.47e-07} & High & 26 & 8.12e-07\\
+\cmidrule{1-5}
+ &  & Low & 116 & 2.65e-06\\
+
+ &  & Middle Low & 122 & 2.79e-06\\
+
+ &  & Middle & 91 & 2.08e-06\\
+
+ &  & Middle High & 84 & 1.52e-06\\
+
+\multirow{-5}{*}{\raggedright\arraybackslash Pharmacies} & \multirow{-5}{*}{\centering\arraybackslash 2e-06} & High & 25 & 7.81e-07\\
+\cmidrule{1-5}
+ &  & Low & 27 & 6.17e-07\\
+
+ &  & Middle Low & 32 & 7.32e-07\\
+
+ &  & Middle & 38 & 8.7e-07\\
+
+ &  & Middle High & 16 & 2.89e-07\\
+
+\multirow{-5}{*}{\raggedright\arraybackslash Super Markets} & \multirow{-5}{*}{\centering\arraybackslash 5.72e-07} & High & 12 & 3.75e-07\\
 \bottomrule

### 2020-08-07 Moral-Communities-and-Crime/Moral_Communities_and_Crime.Rmd
Revisions before submission to Journal of Quantitative Criminology

@@
 ---
-title: Do moral communities have a spatial dimension? A spatial exploratory analysis of places of worship and violent crime in the city of Recife, Brazil
+title: Spatial clustering and violent crime near places of worship in Recife, Brazil. Do moral communities have a spatial dimension?
 subtitle: 
@@
 It is thus that in a recent paper, Warner and Konkel [-@Warner2019neighborhood] note that the role of places of worship, as distinct entities from the members of the congregations, have received less attention in empirical and theoretical research for their deterrence potential.
-The role of these institutions might be particularly important in places where formal state institutions lack the means or the will to enforce norms in a consistent way - as is the case in Brazil [@Garmany2014space].
+The role of these institutions might be particularly important in places where formal state institutions lack the means or the will to enforce norms in a consistent way, such as disadvanteged Black communities in the American South [e.g., @Thomson2020religion] and in regions in South America such as Brazil [e.g., @Garmany2014space].
 
@@
 In what follows, we propose to adopt a disaggregated approach by considering the intensity of criminal events with respect to places of worship.
+An advantage of a disaggregated approach is the use of micro-spatial units to investigate the spatial concentration (or clustering) of events [see @Andresen2017crime].
 This requires micro-level data both of crimes and places of worship, and the use of appropriate spatial analytical tools for point patterns, as discussed next.
@@
 More recent studies of crime as point patterns are Nakaya and Yano [-@Nakaya2010visualizing], Kiani et al.
-[-@Kiani2015analysis], and Malleson and Andresen [-@Malleson2015spatio].
+[-@Kiani2015analysis], and Malleson and Andresen [-@Malleson2015spatio], .
 
@@
 Some descriptive statistics regarding this dataset are reported in Table 1, where it can be seen that overall, about 91% of the victims of LIVC in the period analyzed were men.
-In addition, most of the victims were black or brown, and the youth population between the ages of 18 and 30 is the most affected by violent crime.
+In addition, most of the victims were Black or brown, and the youth population between the ages of 18 and 30 is the most affected by violent crime.
 Lastly, it should be noted that about 88% of the criminal events under analysis involved firearms.

### 2020-08-07 Moral-Communities-and-Crime/Moral_Communities_and_Crime.tex
Revisions before submission to Journal of Quantitative Criminology

@@
 
-\title{Do moral communities have a spatial dimension? A spatial exploratory
-analysis of places of worship and violent crime in the city of Recife,
-Brazil \thanks{Grants or other notes about the article that should go on the front page
+\title{Spatial clustering and violent crime near places of worship in Recife,
+Brazil. Do moral communities have a spatial dimension? \thanks{Grants or other notes about the article that should go on the front page
 should be placed here. General acknowledgments should be placed at the
@@
 important in places where formal state institutions lack the means or
-the will to enforce norms in a consistent way - as is the case in Brazil
-(Garmany 2014).
+the will to enforce norms in a consistent way, such as disadvanteged
+Black communities in the American South (e.g., Thomson, n.d.) and in
+regions in South America such as Brazil (e.g., Garmany 2014).
 
@@
 considering the intensity of criminal events with respect to places of
-worship. This requires micro-level data both of crimes and places of
-worship, and the use of appropriate spatial analytical tools for point
-patterns, as discussed next.
+worship. An advantage of a disaggregated approach is the use of
+micro-spatial units to investigate the spatial concentration (or
+clustering) of events (see Andresen, Linning, and Malleson 2017). This
+requires micro-level data both of crimes and places of worship, and the
+use of appropriate spatial analytical tools for point patterns, as
+discussed next.
 
@@
 Nakaya and Yano (2010), Kiani et al.~(2015), and Malleson and Andresen
-(2015).
+(2015), .
 
@@
 about 91\% of the victims of LIVC in the period analyzed were men. In
-addition, most of the victims were black or brown, and the youth
+addition, most of the victims were Black or brown, and the youth
 population between the ages of 18 and 30 is the most affected by violent
@@
 
+\leavevmode\hypertarget{ref-Andresen2017crime}{}%
+Andresen, M. A., S. J. Linning, and N. Malleson. 2017. ``Crime at Places
+and Spatial Concentrations: Exploring the Spatial Stability of Property
+Crime in Vancouver Bc, 2003-2013.'' Journal Article. \emph{Journal of
+Quantitative Criminology} 33 (2): 255--75.
+\url{https://doi.org/10.1007/s10940-016-9295-8}.
+
 \leavevmode\hypertarget{ref-Appleyard1980livable}{}%
@@
 
+\leavevmode\hypertarget{ref-Thomson2020religion}{}%
+Thomson, R. A. n.d. ``Religion in Civil Society: The Influence of Black
+Religious Ecology on Crime in the South.'' Journal Article.
+\emph{Journal of Quantitative Criminology}, 27.
+\url{https://doi.org/10.1007/s10940-019-09444-7}.
+
 \leavevmode\hypertarget{ref-Topalli2013god}{}%

### 2020-08-07 Moral-Communities-and-Crime/Moral_Communities_and_Crime_Title_Page.tex
Revisions before submission to Journal of Quantitative Criminology

--- 2019-10-16 Moral_Communities_and_Crime_Title_Page_v1.tex
@@
 
-\title{Do moral communities have a spatial dimension? A spatial exploratory
-analysis of places of worship and violent crime in the city of Recife,
-Brazil \thanks{\textbf{Funding:} Edivaldo Constantino das Neves Júnior received support
+\title{Spatial clustering and violent crime near places of worship in Recife,
+Brazil. Do moral communities have a spatial dimension? \thanks{\textbf{Funding:} Edivaldo Constantino das Neves Júnior received support
 from the Government of Canada, through the program Emerging Leaders in
